### 2018-02-06 PAPER/paper.tex
Paper template.

% $Id: template.tex 11 2007-04-03 22:25:53Z jpeltier $

\documentclass{vgtc}                          % final (conference style)
%\documentclass[review]{vgtc}                 % review
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint]{vgtc}               % preprint
%\documentclass[electronic]{vgtc}             % electronic version

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Figures should be in CMYK or Grey scale format, otherwise, colour 
%% shifting may occur during the printing process.

%% These few lines make a distinction between latex and pdflatex calls and they
%% bring in essential packages for graphics and font handling.
%% Note that due to the \DeclareGraphicsExtensions{} call it is no longer necessary
%% to provide the the path and extension of a graphics file:
%% \includegraphics{diamondrule} is completely sufficient.
%%
\ifpdf%                                % if we use pdflatex
  \pdfoutput=1\relax                   % create PDFs from pdfLaTeX
  \pdfcompresslevel=9                  % PDF Compression
  \pdfoptionpdfminorversion=7          % create PDF 1.7
  \ExecuteOptions{pdftex}
  \usepackage{graphicx}                % allow us to embed graphics files
  \DeclareGraphicsExtensions{.pdf,.png,.jpg,.jpeg} % for pdflatex we expect .pdf, .png, or .jpg files
\else%                                 % else we use pure latex
  \ExecuteOptions{dvips}
  \usepackage{graphicx}                % allow us to embed graphics files
  \DeclareGraphicsExtensions{.eps}     % for pure latex we expect eps files
\fi%

%% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
\graphicspath{{figures/}{pictures/}{images/}{./}} % where to search for the images

\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
\usepackage{textcomp}                  % use better special symbols
\usepackage{mathptmx}                  % use matching math font
\usepackage{times}                     % we use Times as the main font
\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
\usepackage{cite}                      % needed to automatically sort the references
\usepackage{tabu}                      % only used for the table example
\usepackage{booktabs}                  % only used for the table example
%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{Machine Perception?}

%% This is how authors are specified in the conference style

%% Author and Affiliation (single author).
%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}}
%%\affiliation{\scriptsize Allied Widgets Research}

%% Author and Affiliation (multiple authors with single affiliations).
%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com} %
%%\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com} %
%%\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}}
%%\affiliation{\scriptsize Martha Stewart Enterprises \\ Microsoft Research}

%% Author and Affiliation (multiple authors with multiple affiliations)
\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ %
        \scriptsize Starbucks Research %
\and Ed Grimley\thanks{e-mail: ed.grimley@aol.com}\\ %
     \scriptsize Grimley Widgets, Inc. %
\and Martha Stewart\thanks{e-mail: martha.stewart@marthastewart.com}\\ %
     \parbox{1.4in}{\scriptsize \centering Martha Stewart Enterprises \\ Microsoft Research}}

%% A teaser figure can be included as follows, but is not recommended since
%% the space is now taken up by a full width abstract.
%\teaser{
%  \includegraphics[width=1.5in]{sample.eps}
%  \caption{Lookit! Lookit!}
%}

%% Abstract section.
\abstract{TODO%
} % end of abstract

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/about/class> for details.
%% We recommend the 2012 system <http://www.acm.org/about/class/class/2012>
%% For the 2012 system use the ``\CCScatTwelve'' which command takes four arguments.
%% The 1998 system <http://www.acm.org/about/class/class/2012> is still possible
%% For the 1998 system use the ``\CCScat'' which command takes four arguments.
%% In both cases the last two arguments (1998) or last three (2012) can be empty.

\CCScatlist{
  \CCScatTwelve{Human-centered computing}{Visu\-al\-iza\-tion}{Visu\-al\-iza\-tion techniques}{Treemaps};
  \CCScatTwelve{Human-centered computing}{Visu\-al\-iza\-tion}{Visualization design and evaluation methods}{}
}

%\CCScatlist{
  %\CCScat{H.5.2}{User Interfaces}{User Interfaces}{Graphical user interfaces (GUI)}{};
  %\CCScat{H.5.m}{Information Interfaces and Presentation}{Miscellaneous}{}{}
%}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

%% if specified like this the section will be committed in review mode
%\acknowledgments{
%The authors wish to thank A, B, and C. This work was supported in part by
%a grant from XYZ.}

%\bibliographystyle{abbrv}
\bibliographystyle{abbrv-doi}
%\bibliographystyle{abbrv-doi-narrow}
%\bibliographystyle{abbrv-doi-hyperref}
%\bibliographystyle{abbrv-doi-hyperref-narrow}

\bibliography{paper.bib}
\end{document}

### 2018-02-07 PAPER/01_related_work.tex

\section{Related Work}
\cite{linegraph_vs_scatterplot} \cite{cleveland_mcgill}

### 2018-02-07 PAPER/paper.tex
Paper

@@
-% $Id: template.tex 11 2007-04-03 22:25:53Z jpeltier $
-
-\documentclass{vgtc}                          % final (conference style)
-%\documentclass[review]{vgtc}                 % review
+\documentclass[journal]{vgtc}                % final (journal style)
+%\documentclass[review,journal]{vgtc}         % review (journal style)
 %\documentclass[widereview]{vgtc}             % wide-spaced review
-%\documentclass[preprint]{vgtc}               % preprint
-%\documentclass[electronic]{vgtc}             % electronic version
+%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
 
@@
 %% submission, ``preprint'' is for pre-publication, and the final version
-%% doesn't use a specific qualifier. Further, ``electronic'' includes
-%% hyperreferences for more convenient online viewing.
+%% doesn't use a specific qualifier.
 
@@
 
-%% Figures should be in CMYK or Grey scale format, otherwise, colour 
-%% shifting may occur during the printing process.
+%% Please note that the use of figures other than the optional teaser is not permitted on the first page
+%% of the journal version.  Figures should begin on the second page and be
+%% in CMYK or Grey scale format, otherwise, colour shifting may occur
+%% during the printing process.  Papers submitted with figures other than the optional teaser on the
+%% first page will be refused. Also, the teaser figure should only have the
+%% width of the abstract as the template enforces it.
 
@@
 %% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
-\graphicspath{{figures/}{pictures/}{images/}{./}} % where to search for the images
+\graphicspath{{gfx/}{./}} % where to search for the images
 
@@
 
+%% In preprint mode you may define your own headline.
+%\preprinttext{To appear in IEEE Transactions on Visualization and Computer Graphics.}
+
 %% If you are submitting a paper to a conference for review with a double
@@
 \vgtccategory{Research}
-
-%% allow for this line if you want the electronic option to work properly
-\vgtcinsertpkg
-
-%% In preprint mode you may define your own headline.
-%\preprinttext{To appear in an IEEE VGTC sponsored conference.}
+%% please declare the paper type of your paper to help reviewers, only shown in review mode
+%% choices:
+%% * algorithm/technique
+%% * application/design study
+%% * evaluation
+%% * system
+%% * theory/model
+\vgtcpapertype{theory/model}
 
 %% Paper title.
+\title{Experiments of Cleveland and McGill for Machine Perception}
 
-\title{Machine Perception?}
+%% This is how authors are specified in the journal style
 
-%% This is how authors are specified in the conference style
+%% indicate IEEE Member or Student Member in form indicated below
+\author{Daniel Haehn, \textit{Member, IEEE}, James Tompkin, and Hanspeter Pfister}
+\authorfooter{
+%% insert punctuation at end of each item
+\item Daniel Haehn, and Hanspeter Pfister are with the Paulson School of Engineering and Applied Sciences at Harvard University. \\
+E-mail: \{haehn,pfister\}@seas.harvard.edu.
+%
+\item James Tompkin is with the Thomas J. Watson Sr. Center for Information Technology at Brown University. \\E-mail: james\_tompkin@brown.edu.
+}
 
-%% Author and Affiliation (single author).
-%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}}
-%%\affiliation{\scriptsize Allied Widgets Research}
-
-%% Author and Affiliation (multiple authors with single affiliations).
-%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com} %
-%%\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com} %
-%%\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}}
-%%\affiliation{\scriptsize Martha Stewart Enterprises \\ Microsoft Research}
-
-%% Author and Affiliation (multiple authors with multiple affiliations)
-\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ %
-        \scriptsize Starbucks Research %
-\and Ed Grimley\thanks{e-mail: ed.grimley@aol.com}\\ %
-     \scriptsize Grimley Widgets, Inc. %
-\and Martha Stewart\thanks{e-mail: martha.stewart@marthastewart.com}\\ %
-     \parbox{1.4in}{\scriptsize \centering Martha Stewart Enterprises \\ Microsoft Research}}
-
-%% A teaser figure can be included as follows, but is not recommended since
-%% the space is now taken up by a full width abstract.
-%\teaser{
-%  \includegraphics[width=1.5in]{sample.eps}
-%  \caption{Lookit! Lookit!}
-%}
+%other entries to be set up for journal
+\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Experiments of Cleveland and McGill applied to Machine Perception}
+%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
 
@@
 
+%% Keywords that describe your work. Will show as 'Index Terms' in journal
+%% please capitalize first letter and insert punctuation after last keyword
+\keywords{Machine Perception, Deep Learning}
+
 %% ACM Computing Classification System (CCS). 
-%% See <http://www.acm.org/about/class> for details.
-%% We recommend the 2012 system <http://www.acm.org/about/class/class/2012>
-%% For the 2012 system use the ``\CCScatTwelve'' which command takes four arguments.
-%% The 1998 system <http://www.acm.org/about/class/class/2012> is still possible
-%% For the 1998 system use the ``\CCScat'' which command takes four arguments.
-%% In both cases the last two arguments (1998) or last three (2012) can be empty.
+%% See <http://www.acm.org/class/1998/> for details.
+%% The ``\CCScat'' command takes four arguments.
 
-\CCScatlist{
-  \CCScatTwelve{Human-centered computing}{Visu\-al\-iza\-tion}{Visu\-al\-iza\-tion techniques}{Treemaps};
-  \CCScatTwelve{Human-centered computing}{Visu\-al\-iza\-tion}{Visualization design and evaluation methods}{}
+%\CCScatlist{ % not used in journal version
+% \CCScat{K.6.1}{Management of Computing and Information Systems}%
+%{Project and People Management}{Life Cycle};
+% \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
+%}
+
+%% Uncomment below to include a teaser figure.
+\teaser{
+  \centering
+  \includegraphics[width=\linewidth]{fish.jpg}
+  \caption{Here is a fish.}
+	\label{fig:teaser}
 }
 
-%\CCScatlist{
-  %\CCScat{H.5.2}{User Interfaces}{User Interfaces}{Graphical user interfaces (GUI)}{};
-  %\CCScat{H.5.m}{Information Interfaces and Presentation}{Miscellaneous}{}{}
-%}
+%% Uncomment below to disable the manuscript note
+%\renewcommand{\manuscriptnotetxt}{}
 
@@
 % \nocopyrightspace
+
+\vgtcinsertpkg
 
@@
 %% the only exception to this rule is the \firstsection command
-\firstsection{Introduction}
+
+\input{00_introduction.tex}
 
 \maketitle
+
+\input{01_related_work.tex}
 
@@
 %The authors wish to thank A, B, and C. This work was supported in part by
-%a grant from XYZ.}
+%a grant from XYZ (\# 12345-67890).}
 

### 2018-02-08 PAPER/01_related_work.tex
Re-store James edits.
Update some related work.

@@
 \section{Related Work}
-\cite{linegraph_vs_scatterplot} \cite{cleveland_mcgill}
+
+\cite{cleveland_mcgill} Clevenland McGill experiments
+
+\cite{Wang_linegraph_vs_scatterplot} Comparing linegraph vs. scatterplot
+
+\cite{mckenzie_2d_3d} \cite{forsberg2009comparing_3d_vector} \cite{laidlaw_2d_vector} 2D vs. 3D Evaluation (vector fields)
+
+\cite{kindlmann2002color} \cite{rheingans1992color} \cite{ware1988color} \cite{Rogowitz2001_colormaps} color maps
+
+2d vs. 3d + color \cite{borkin2011arteries}
+
+black hat vis \cite{heer2017blackhat}
+
+timeseries \cite{herr2009timeseries}
+
+munzner \cite{munzner2015visualization}
+
+open vs. closed shaped \cite{open_vs_closed_shapes}
+
+visualization ranking \cite{harrison2014_webers_law_rank}
+
+Pineo et al.~\cite{Pineo2012_computational_perception} present a method to automatically evaluating and optimizing visualizations using a computational model of human vision, based on a neural network simulation of the early perceptual processing in the retina and primary visual cortex.
+[JT] Copied from their abstract.

### 2018-03-06 PAPER/00_introduction.tex
Rough paper outline.

\firstsection{Introduction}

\maketitle

Artificial intelligence has taken the world of technology by storm. Deep multilayer neural networks are being successfully applied in a wide range of applications that are regularly outperforming humans in object recognition \cite{krizhevsky_imagenet2012, simonyan_very_deep2014,szegedy2015}. 
Originally inspired by neuroscientific discoveries, the recent advances in deep learning have been the direct results of engineering efforts, more specifically in convolutional neural networks (CNNs). 
While there has been significant advancement, this does not mean we understand what CNNs are doing and we are in fact treating them as blackboxes without detracting from their success \cite{goodfellow_book, deeplearning_blackbox2017}. 
Our current knowledge of biological vision suggests that modern machine learning models indeed mimic the underlying biology by abstracting the many details of biological neural networks \cite{yamins2016using, hassabis2017neuroscience}.

Despite tremendous research efforts and generating massive datasets, we are far from fully understanding biological vision. 
Similarly, we are constantly developing inventive features for deep artificial networks without truly understanding it in its entirety. As a result, a double discrepancy is observed. 
Advances in neuroscience can revolutionize machine learning by reverse-engineering neural circuits, yielding new classifiers which, in turn, can help process the massive biological data. 
There are many existing questions which must be answered in order to fill this gap of our understanding.

We focus on perception... experiments from cleveland mcgill..

%\emph{Can we leverage decades of visualization research to understand the way convolutional neural networks process data?}
%
%Information visualization has been an established research field for decades which has resulted in numerous insightful findings in regards to how human beings can best process information visually.
%Unpublished preliminary experiments have shown that data representations customized for the human eye also can improve the performance of an automatic classifier. 
%While the reasons for this are still unknown, specified research can most certainly advance the understanding of deep neural networks.

\subsection{Biological Vision}

\begin{figure}[t]
	  \includegraphics[width=\linewidth]{biology_vs_cnn.png}
  \caption{The Biological Vision (schematic)}
	\label{fig:vision}
\end{figure}

Biological vision is an extremely powerful system which allows humans the ability, and seemingly without effort, to recognize an enormous amount of distinct objects in the world. 
Object detection is extremely difficult and therefore is especially impressive as light intensities can change by levels of magnitude and contrast between foreground and background is so often low. 
In addition, the visual scene changes every time the human body or human eyes move. 
This visual system exhibits a very noisy structure but because it is organized by layers it has inspired the mathematical theory of multilayer neural networks. 
What is remarkable is that even though current machine learning models do not resemble the complexity of its biological pendant, they inherently generalize extremely well. 
Neural networks trained on one specific task can be used to perform detection or segmentation of, seemingly, unrelated objects with relatively minor retraining. 
The reported classification performance is superior to that of humans and the question in regards to their functionality opens an interesting research topic.

GOALS
...reduce the gap between neurobiology and data science to advance the understanding of visual cortex inspired machine learning

CONTRIBUTIONS

- experiments of cleveland mcgill with systematic parametrization and evaluation

- ranking like cleveland mcgill for machine perception?

- many other insights?

- framework

### 2018-03-06 PAPER/paper.tex
Update some related work.
Rough paper outline.

@@
 %% Paper title.
-\title{Experiments of Cleveland and McGill for Machine Perception}
+\title{Visualization for Machine Perception}
 
@@
 %other entries to be set up for journal
-\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Experiments of Cleveland and McGill applied to Machine Perception}
+\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Visualization for Machine Perception}
 %\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
@@
 
-\maketitle
+\input{01_related_work.tex}
 
-\input{01_related_work.tex}
+\section{Experimental Setup}
+
+\subsection{Measures}
+
+\subsection{Classifiers}
+
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{fish.jpg}
+  \caption{The Classifiers...}
+	\label{fig:classifiers}
+\end{figure}
+
+\section{Elementary Perceptual Tasks}
+
+THe Figure 1 of Cleveland McGill
+
+\subsection{Parametrization}
+
+Here we show each of Figure 1 with its parametrization.
+
+\section{Position-Angle Experiment}
+
+This is pie chart vs bar chart
+
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{fish.jpg}
+  \caption{The Position-Angle Experiment}
+	\label{fig:position_angle_experiment}
+\end{figure}
+
+\section{Position-Length Experiment}
+
+This is the one where we estimate two selected bars compared to the longest one
+
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{fish.jpg}
+  \caption{The Position-Length Experiment}
+	\label{fig:position_length_experiment}
+\end{figure}
+
+\section{Bars and Framed Rectangles Experiment}
+
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{fish.jpg}
+  \caption{The Bars and Framed Rectangles Experiment}
+	\label{fig:bars_and_framed_rectangles_experiment}
+\end{figure}
+
+\subsection{Weber's Law}
+
+\section{Results and Discussion}
+
+Cleveland McGills Ranking - can we observe something similar?
+
+1. Position along a common scale e.g. scatter plot
+
+2. Position on identical but nonaligned scales e.g. multiple scatter plots
+
+3. Length e.g. bar chart
+
+4.Angle \& Slope (tie) e.g. pie chart
+
+5. Area e.g. bubbles
+
+6. Volume, density, and color saturation (tie) e.g. heatmap
+
+7. Color hue e.g. newsmap
+
+\section{Conclusions}
+
+Future work: allow insights for infovis for machines
 

### 2018-03-07 EXP/ClevelandMcGill/__init__.py
Nice way of plotting our images.

@@ -1,2 +1,3 @@
+from figure1_position_nonaligned_scale import Figure1_Position_Nonaligned_Scale
 from figure1_position_common_scale import Figure1_Position_Common_Scale
 from figure12 import Figure12

### 2018-03-07 EXP/ClevelandMcGill/figure1_position_nonaligned_scale.py

@@ -1,0 +1,214 @@
+import cPickle as pickle
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+import sys
+
+sys.path.append('../')
+from util import Util
+
+class Figure1_Position_Nonaligned_Scale:
+
+  def __init__(self, variable_x=False, \
+                     variable_spot_size=False, \
+                     constrain = True):
+
+    '''
+    '''
+    self.X = None
+    self.Y = None
+    self.ORIGIN = [0, 0]
+    self.SPOT_SIZE = None
+
+    self.label = None
+
+    self._delta = 20 # we use a bit larger delta here, it needs to be bigger
+    # than the max. spot size / 2. for sure
+
+    self._size = (100,100)
+    self._spot_sizes = [1, 3, 5, 7, 9, 11]
+
+    #
+    # variability switches
+    #
+    #  stage 1: var. Y
+    #  stage 2: var. WIDTH
+    #  stage 3: var. X
+    #    
+    self._variable_x = variable_x
+    self._variable_spot_size = variable_spot_size
+    self._constrain = constrain
+
+
+  def create(self, verbose=False):
+    '''
+    Create a valid datapoint with variability settings in mind.
+    '''
+
+    parameters = 1
+
+    #
+    # Y is always variable since it encodes our value
+    #
+    Y, p = Util.parameter(self._delta, self._size[0]-self._delta + 1)
+    parameters *= p
+
+    X = self._size[1] / 2
+    if self._variable_x:
+      X, p = Util.parameter(self._delta, self._size[1]-self._delta + 1)
+      parameters *= p
+
+    spot_size = 5
+    if self._variable_spot_size:
+      spot_size = np.random.choice(self._spot_sizes)
+      parameters *= len(self._spot_sizes)
+
+    origin = [0, 0]
+
+    # update the label
+    self.X = X
+    self.Y = Y
+    self.ORIGIN = origin
+    self.SPOT_SIZE = spot_size
+    self.label = Y # label is the same as Y
+
+
+  def to_sparse(self):
+    '''
+    Convert to sparse representation
+      [X, Y, origin, spot_size], label
+    '''
+    return [self.X, self.Y, self.ORIGIN, self.SPOT_SIZE], self.label
+
+
+  @staticmethod
+  def from_sparse(sparse_representation, \
+                  variable_x=False, \
+                  variable_spot_size=False, \
+                  constrain = True):
+    '''
[…]

### 2018-03-07 EXP/util.py

@@ -1,3 +1,4 @@
+import matplotlib.pyplot as plt
 import numpy as np
 
 class Util:
@@ -9,3 +10,21 @@ class Util:
     parameters = len(range(start-delta, end+delta))
 
     return value, parameters
+
+  @staticmethod
+  def imshow_nicely(image, filename=None):
+
+    plt.figure()
+    plt.imshow(image, cmap='Greys', interpolation='nearest')
+    ax = plt.gca()
+    ax.set_xticklabels('')
+    ax.set_yticklabels('')
+    ax.set_xticks(np.arange(-.5, 100, 10), minor=False);
+    ax.set_yticks(np.arange(-.5, 100, 10), minor=False);
+    ax.grid(which='major', color='gray', linestyle=':', linewidth='0.5')
+    ax.set_axisbelow(True)
+
+    if filename:
+
+      plt.savefig(filename)
+    

### 2018-03-08 EXP/ClevelandMcGill/figure1_position_common_scale.py
Moving on with Figure 1 (finally).

@@ -7,7 +7,7 @@ import sys
 sys.path.append('../')
 from util import Util
 
-class Figure1_Position_Common_Scale:
+class Figure1_Position_Common_Scale(object):
 
   def __init__(self, variable_x=False, \
                      variable_spot_size=False, \
@@ -70,7 +70,10 @@ class Figure1_Position_Common_Scale:
     self.Y = Y
     self.ORIGIN = origin
     self.SPOT_SIZE = spot_size
-    self.label = Y # label is the same as Y
+    self.label = Y - self.ORIGIN[0]# label is the same as Y
+
+    if verbose:
+      print '# Parameters', parameters
 
 
   def to_sparse(self):
@@ -101,7 +104,7 @@ class Figure1_Position_Common_Scale:
     fig.SPOT_SIZE = sparse_representation[3]
 
     # update the label
-    fig.label = fig.Y
+    fig.label = fig.Y - fig.ORIGIN[0]
 
     return fig
 

### 2018-03-08 EXP/ClevelandMcGill/figure1_position_nonaligned_scale.py

@@ -7,37 +7,22 @@ import sys
 sys.path.append('../')
 from util import Util
 
-class Figure1_Position_Nonaligned_Scale:
+from figure1_position_common_scale import Figure1_Position_Common_Scale
 
-  def __init__(self, variable_x=False, \
+class Figure1_Position_Nonaligned_Scale(Figure1_Position_Common_Scale):
+
+  def __init__(self, variable_y=False, \
+                     variable_x=False, \
                      variable_spot_size=False, \
+                     variable_origin=True, \
                      constrain = True):
 
-    '''
-    '''
-    self.X = None
-    self.Y = None
-    self.ORIGIN = [0, 0]
-    self.SPOT_SIZE = None
+    self._variable_y = variable_y
+    self._variable_origin = True
 
-    self.label = None
-
-    self._delta = 20 # we use a bit larger delta here, it needs to be bigger
-    # than the max. spot size / 2. for sure
-
-    self._size = (100,100)
-    self._spot_sizes = [1, 3, 5, 7, 9, 11]
-
-    #
-    # variability switches
-    #
-    #  stage 1: var. Y
-    #  stage 2: var. WIDTH
-    #  stage 3: var. X
-    #    
-    self._variable_x = variable_x
-    self._variable_spot_size = variable_spot_size
-    self._constrain = constrain
+    super(Figure1_Position_Nonaligned_Scale, self).__init__(variable_x=variable_x, \
+                                                            variable_spot_size=variable_spot_size, \
+                                                            constrain=constrain)
 
 
   def create(self, verbose=False):
@@ -50,8 +35,10 @@ class Figure1_Position_Nonaligned_Scale:
     #
     # Y is always variable since it encodes our value
     #
-    Y, p = Util.parameter(self._delta, self._size[0]-self._delta + 1)
-    parameters *= p
+    Y = self._size[0] / 2
+    if self._variable_y:
+      Y, p = Util.parameter(self._delta, self._size[0]-self._delta + 1)
+      parameters *= p
 
     X = self._size[1] / 2
     if self._variable_x:
@@ -64,35 +51,42 @@ class Figure1_Position_Nonaligned_Scale:
       parameters *= len(self._spot_sizes)
 
     origin = [0, 0]
+    if self._variable_origin:
+      origin_y, p = Util.parameter(0, self._delta/2)
+      parameters *= p
+
+      origin[0] = origin_y
+      #
[…]

### 2018-03-09 EXP/ClevelandMcGill/__init__.py
Draw Direction, Angle, Area, Curvature.

@@ -1,3 +1,4 @@
+from figure1_length import Figure1_Length
 from figure1_position_nonaligned_scale import Figure1_Position_Nonaligned_Scale
 from figure1_position_common_scale import Figure1_Position_Common_Scale
 from figure12 import Figure12

### 2018-03-09 EXP/ClevelandMcGill/figure1_length.py

@@ -1,0 +1,225 @@
+import cPickle as pickle
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+import sys
+
+sys.path.append('../')
+from util import Util
+
+class Figure1_Length(object):
+
+  def __init__(self, variable_x=False, \
+                     variable_y=False, \
+                     variable_width=False, \
+                     constrain = True):
+
+    '''
+    '''
+    self.X = None
+    self.Y = None
+    self.LENGTH = None
+    self.WIDTH = None
+
+    self.label = None
+
+    self._delta = 20 # we use a bit larger delta here, it needs to be bigger
+    # than the max. spot size / 2. for sure
+
+    self._size = (100,100)
+    self._widths = [1, 3, 5, 7, 9, 11]
+
+    #
+    # variability switches
+    #
+    #  stage 1: var. X
+    #  stage 2: var. Y
+    #  stage 3: var. WIDTH
+    #    
+    self._variable_x = variable_x
+    self._variable_y = variable_y
+    self._variable_width = variable_width
+    self._constrain = constrain
+
+
+  def create(self, verbose=False):
+    '''
+    Create a valid datapoint with variability settings in mind.
+    '''
+
+    parameters = 1
+
+    length, p = Util.parameter(1,41)
+    parameters *= p
+
+    Y = self._size[0] / 2
+    if self._variable_y:
+      Y, p = Util.parameter(self._delta, self._size[0]-self._delta + 1)
+      parameters *= p
+
+    X = self._size[1] / 2
+    if self._variable_x:
+      X, p = Util.parameter(self._delta, self._size[1]-self._delta + 1)
+      parameters *= p
+
+    width = 1
+    if self._variable_width:
+      width = np.random.choice(self._widths)
+      parameters *= len(self._widths)
+
+    # update the label
+    self.X = X
+    self.Y = Y
+    self.LENGTH = length
+    self.WIDTH = width
+    self.label = length
+
+    if verbose:
+      print '# Parameters', parameters
+
+
+  def to_sparse(self):
+    '''
+    Convert to sparse representation
+      [X, Y, origin, spot_size], label
+    '''
+    return [self.X, self.Y, self.LENGTH, self.WIDTH], self.label
+
+
+  @staticmethod
+  def from_sparse(sparse_representation, \
+                  variable_x=False, \
+                  variable_y=False, \
+                  variable_width=False, \
+                  constrain = True):
+    '''
+    From sparse:
[…]

### 2018-03-10 EXP/ClevelandMcGill/__init__.py
New stuff.

@@ -1,3 +1,5 @@
+from figure1 import Figure1
+
 from figure1_length import Figure1_Length
 from figure1_position_nonaligned_scale import Figure1_Position_Nonaligned_Scale
 from figure1_position_common_scale import Figure1_Position_Common_Scale

### 2018-03-10 EXP/ClevelandMcGill/figure1.py

@@ -1,0 +1,226 @@
+import numpy as np
+import os
+import sys
+
+sys.path.append('../')
+from util import Util
+
+class Figure1:
+
+  DELTA_MIN = 20
+  DELTA_MAX = 80
+  SIZE = (100, 100)
+
+  @staticmethod
+  def position_common_scale(flags=[False, False]):
+    '''
+    '''
+    var_x = flags[0]
+    var_spot = flags[1]
+
+
+    sparse = None
+    image = None
+    label = None
+    parameters = 1
+
+
+    Y_RANGE = (Figure1.DELTA_MIN, Figure1.DELTA_MAX)
+    X_RANGE = (Figure1.DELTA_MIN, Figure1.DELTA_MAX)
+
+    Y, p = Util.parameter(Y_RANGE[0], Y_RANGE[1])
+    parameters *= p
+
+    X = Figure1.SIZE[1] / 2
+    if var_x:
+      X, p = Util.parameter(X_RANGE[0], X_RANGE[1])
+      parameters *= p
+
+    SPOT_SIZE = 5
+    if var_spot:
+      sizes = [1, 3, 5, 7, 9, 11]
+      SPOT_SIZE = np.random.choice(sizes)
+      parameters *= len(sizes)
+
+    ORIGIN = 10
+
+    sparse = [X, Y, SPOT_SIZE]
+
+    image = np.zeros(Figure1.SIZE, dtype=np.bool)
+
+    # draw axis
+    image[Y_RANGE[0]:Y_RANGE[1], ORIGIN] = 1
+
+    # draw spot
+    half_spot_size = SPOT_SIZE / 2
+    image[Y-half_spot_size:Y+half_spot_size+1, X-half_spot_size:X+half_spot_size+1] = 1
+
+    label = Y - Figure1.DELTA_MIN
+
+    return sparse, image, label, parameters
+
+
+
+  @staticmethod
+  def position_non_aligned_scale(flags=[False, False, False]):
+
+    var_x = flags[0]
+    var_y = flags[1]
+    var_spot = flags[2]
+
+    sparse = None
+    image = None
+    label = None
+    parameters = 1
+
+    OFFSET, p = Util.parameter(1, 11)
+    # print 'OFFSET', OFFSET
+    parameters *= p
+
+    Y_RANGE = (Figure1.DELTA_MIN-OFFSET, Figure1.DELTA_MAX-OFFSET)
[…]

### 2018-03-10 EXP/ClevelandMcGill/figure1.py
Add cleaner Figure1 file with all variations and some noise experiments.

@@ -1,5 +1,6 @@
 import numpy as np
 import os
+import skimage.draw
 import sys
 
 sys.path.append('../')
@@ -44,7 +45,7 @@ class Figure1:
 
     ORIGIN = 10
 
-    sparse = [X, Y, SPOT_SIZE]
+    sparse = [Y, X, SPOT_SIZE]
 
     image = np.zeros(Figure1.SIZE, dtype=np.bool)
 
@@ -64,8 +65,8 @@ class Figure1:
   @staticmethod
   def position_non_aligned_scale(flags=[False, False, False]):
 
-    var_x = flags[0]
-    var_y = flags[1]
+    var_y = flags[0]
+    var_x = flags[1]
     var_spot = flags[2]
 
     sparse = None
@@ -100,7 +101,7 @@ class Figure1:
 
     ORIGIN = 10
 
-    sparse = [X, Y, SPOT_SIZE]
+    sparse = [Y, X, SPOT_SIZE]
 
     image = np.zeros(Figure1.SIZE, dtype=np.bool)
 
@@ -120,8 +121,8 @@ class Figure1:
   @staticmethod
   def length(flags=[False, False, False]):
 
-    var_x = flags[0]
-    var_y = flags[1]
+    var_y = flags[0]
+    var_x = flags[1]
     var_width = flags[2]
 
 
@@ -130,15 +131,106 @@ class Figure1:
     label = None
     parameters = 1
 
-    length, p = Util.parameter(1,41)
+    Y_RANGE = (Figure1.DELTA_MIN, Figure1.DELTA_MAX)
+    X_RANGE = (Figure1.DELTA_MIN, Figure1.DELTA_MAX)
+
+    LENGTH, p = Util.parameter(1, Y_RANGE[1]-Y_RANGE[0]+1) # 1..60
     parameters *= p
 
+    MAX_LENGTH = Y_RANGE[1]-Y_RANGE[0]
+    # print 'Max length', MAX_LENGTH
+
+    X = Figure1.SIZE[1] / 2
+    if var_x:
+      X, p = Util.parameter(X_RANGE[0], X_RANGE[1])
+      parameters *= p
+
+    Y = Y_RANGE[0]
+    if var_y:
+      
+      Y, p = Util.parameter(0, Figure1.SIZE[0]-MAX_LENGTH)
+      # print 'Y',Y
+      parameters *= p
+
+    WIDTH = 1
+    if var_width:
[…]

### 2018-03-12 EXP/run_regression.py
Brand new stuff.

@@ -1,0 +1,338 @@
+from keras import models
+from keras import layers
+from keras import optimizers
+import keras.applications
+import keras.callbacks
+from keras import backend as K
+from keras.utils.np_utils import to_categorical
+import sklearn.metrics
+
+import cPickle as pickle
+import numpy as np
+import os
+import sys
+import time
+
+import ClevelandMcGill as C
+
+
+EXPERIMENT = sys.argv[1] # f.e. Angle
+DATASET = int(sys.argv[2]) # How many TRUE flags to pass to data generation
+CLASSIFIER = sys.argv[3] # 'LeNet'
+NOISE = sys.argv[4] # True
+JOB_INDEX = int(sys.argv[5])
+
+#
+#
+#
+print 'Running', EXPERIMENT, 'on dataset', DATASET, 'with', CLASSIFIER, 'Noise:', NOISE, 'Job Index', JOB_INDEX
+
+#
+#
+# PROCESS SOME FLAGS
+#
+#
+SUFFIX = '.'
+if NOISE == 'True':
+  NOISE = True
+  SUFFIX = '_noise.'
+else:
+  NOISE = False
+
+DATATYPE = eval(EXPERIMENT)
+
+FLAGS = [False] * 10 # never more than 10 flags
+for f in range(DATASET):
+  FLAGS[f] = True
+
+
+if os.path.abspath('~').startswith('/n/'):
+  # we are on the cluster
+  PREFIX = '/n/regal/pfister_lab/PERCEPTION/'
+else:
+  PREFIX = '/home/d/PERCEPTION/'
+RESULTS_DIR = PREFIX + 'RESULTS/'
+
+OUTPUT_DIR = RESULTS_DIR + EXPERIMENT + '/' + str(DATASET) + '/' + CLASSIFIER + '/'
+if not os.path.exists(OUTPUT_DIR):
+  os.makedirs(OUTPUT_DIR)
+
+STATSFILE = OUTPUT_DIR + str(JOB_INDEX).zfill(2) + SUFFIX + 'p'
+MODELFILE = OUTPUT_DIR + str(JOB_INDEX).zfill(2) + SUFFIX + 'h5'
+
+print 'Working in', OUTPUT_DIR
+print 'Storing', STATSFILE
+print 'Storing', MODELFILE
+
+if os.path.exists(STATSFILE) and os.path.exists(MODELFILE):
[…]

### 2018-03-12 EXP/run_regression.py
Fix bug

@@ -263,6 +263,8 @@ elif CLASSIFIER == 'MLP':
   X_train = X_train.reshape(len(X_train), 100*100)
   X_val = X_val.reshape(len(X_val), 100*100)
   X_test = X_test.reshape(len(X_test), 100*100)
+
+  feature_shape = 100*100
   ##
 
 #

### 2018-03-12 EXP/run_regression.py
Fix race condition.

@@ -55,7 +55,11 @@ RESULTS_DIR = PREFIX + 'RESULTS/'
 
 OUTPUT_DIR = RESULTS_DIR + EXPERIMENT + '/' + str(DATASET) + '/' + CLASSIFIER + '/'
 if not os.path.exists(OUTPUT_DIR):
-  os.makedirs(OUTPUT_DIR)
+  # here can be a race condition
+  try:
+    os.makedirs(OUTPUT_DIR)
+  except:
+    print 'Race condition!', os.path.exists(OUTPUT_DIR)
 
 STATSFILE = OUTPUT_DIR + str(JOB_INDEX).zfill(2) + SUFFIX + 'p'
 MODELFILE = OUTPUT_DIR + str(JOB_INDEX).zfill(2) + SUFFIX + 'h5'

### 2018-03-12 EXP/run_regression.py
Another bug fix aaaaahhhh

@@ -211,7 +211,7 @@ if CLASSIFIER == 'VGG19' or CLASSIFIER == 'XCEPTION':
 
   if CLASSIFIER == 'VGG19':  
     feature_generator = keras.applications.VGG19(include_top=False, weights='imagenet', input_shape=(100,100,3))
-  elif CLASSIFIER == 'Xception':
+  elif CLASSIFIER == 'XCEPTION':
     feature_generator = keras.applications.Xception(include_top=False, weights='imagenet', input_shape=(100,100,3))
   elif CLASSIFIER == 'RESNET50':
     print 'Not yet - we need some padding and so on!!!'

### 2018-03-13 EXP/ClevelandMcGill/__init__.py
Avoid some leakage between train/val/test and also work on Figure3.

@@ -1,4 +1,5 @@
 from figure1 import Figure1
+from figure3 import Figure3
 
 from figure1_length import Figure1_Length
 from figure1_position_nonaligned_scale import Figure1_Position_Nonaligned_Scale

### 2018-03-13 EXP/ClevelandMcGill/figure3.py

@@ -1,0 +1,148 @@
+import numpy as np
+import os
+import skimage.draw
+import sys
+
+sys.path.append('../')
+from util import Util
+
+class Figure3:
+
+  SIZE = (100, 100)
+
+
+  @staticmethod
+  def generate_datapoint():
+    '''
+    '''
+
+
+    # from codementum.org/cleveland-mcgill/
+    #
+    #
+    # randomize data according to Cleveland84
+    # specifically:
+    #  - 5 numbers
+    #  - add to 100
+    #  - none less than 3
+    #  - none greater than 39
+    #  - differences greater than .1
+    #
+    def randomize_data():
+      max = 36;
+      min = 3;
+      diff = 0.1;
+
+      d = []
+
+      while len(d) < 5:
+        randomnumber=np.ceil(np.random.random()*36 + 3);
+        found=False;
+        for i in range(len(d)):
+          if not ensure_difference(d, randomnumber):
+            found = True
+            break
+            
+        if not found:
+          d.append(randomnumber)
+
+
+      return d;
+
+
+    def ensure_difference(A, c):
+      result = True;
+      for i in range(len(A)):
+        if c > (A[i] - 3) and c < (A[i] + 3):
+          result = False
+
+      return result
+
+    sum = -1
+    while(sum != 100):
+      data = randomize_data()
+      sum = data[0] + data[1] + data[2] + data[3] + data[4]
+
+    labels = np.zeros((5), dtype=np.float32)
+    for i,d in enumerate(data):
+      labels[i] = d/float(np.max(data))
+
+    #
+    #
+    # ATTENTION, HERE WE NEED TO ORDER THE LABELS ACCORDING
+    # OUR CONVENTION
+    #
+    # NOW, THE MARKED ELEMENT IS AT POSITION 0 BUT ONLY IN THE
+    # LABELS. THIS MEANS WE CAN NOW GO LEFT TO RIGHT (ROLLING) IN THE BARCHART
+    # STARTING FROM THE MARKED ONE. AND SIMILARLY, IN THE PIE CHART WE CAN GO
+    # COUNTER-CLOCKWISE STARTING FROM THE MARKED ONE.
+    #
+    labels = np.roll(labels, 5-np.where(labels==1)[0])
+
+    return data, labels
+
+
+  @staticmethod
+  def data_to_barchart(data):
+    '''
+    '''
+    barchart = np.zeros((100,100), dtype=np.bool)
+
[…]

### 2018-03-13 EXP/run_regression.py

@@ -164,25 +164,27 @@ print 'Done', time.time()-t0, 'seconds (', all_counter, 'iterations)'
 
 #
 #
-# NORMALIZE DATA IN-PLACE
+# NORMALIZE DATA IN-PLACE (BUT SEPERATELY)
 #
 #
-X_min = min(X_train.min(), X_val.min(), X_test.min())
-X_max = max(X_train.max(), X_val.max(), X_test.max())
-y_min = min(y_train.min(), y_val.min(), y_test.min())
-y_max = max(y_train.max(), y_val.max(), y_test.max())
+X_min = X_train.min()
+X_max = X_train.max()
+y_min = y_train.min()
+y_max = y_train.max()
 
 # scale in place
 X_train -= X_min
 X_train /= (X_max - X_min)
-X_val -= X_min
-X_val /= (X_max - X_min)
-X_test -= X_min
-X_test /= (X_max - X_min)
 y_train -= y_min
 y_train /= (y_max - y_min)
+
+X_val -= X_min
+X_val /= (X_max - X_min)
 y_val -= y_min
 y_val /= (y_max - y_min)
+
+X_test -= X_min
+X_test /= (X_max - X_min)
 y_test -= y_min
 y_test /= (y_max - y_min)
 

### 2018-03-13 EXP/run_position_angle.py
Ready for figure 3 experiments.

@@ -1,0 +1,360 @@
+from keras import models
+from keras import layers
+from keras import optimizers
+import keras.applications
+import keras.callbacks
+from keras import backend as K
+from keras.utils.np_utils import to_categorical
+import sklearn.metrics
+
+import cPickle as pickle
+import numpy as np
+import os
+import sys
+import time
+
+import ClevelandMcGill as C
+
+
+EXPERIMENT = sys.argv[1] # f.e. Angle
+CLASSIFIER = sys.argv[3] # 'LeNet'
+NOISE = sys.argv[4] # True
+JOB_INDEX = int(sys.argv[5])
+
+#
+#
+#
+print 'Running', EXPERIMENT, 'with', CLASSIFIER, 'Noise:', NOISE, 'Job Index', JOB_INDEX
+
+#
+#
+# PROCESS SOME FLAGS
+#
+#
+SUFFIX = '.'
+if NOISE == 'True':
+  NOISE = True
+  SUFFIX = '_noise.'
+else:
+  NOISE = False
+
+DATATYPE = eval(EXPERIMENT)
+
+
+
+if os.path.abspath('~').startswith('/n/'):
+  # we are on the cluster
+  PREFIX = '/n/regal/pfister_lab/PERCEPTION/'
+else:
+  PREFIX = '/home/d/PERCEPTION/'
+RESULTS_DIR = PREFIX + 'RESULTS/'
+
+OUTPUT_DIR = RESULTS_DIR + EXPERIMENT + '/' + CLASSIFIER + '/'
+if not os.path.exists(OUTPUT_DIR):
+  # here can be a race condition
+  try:
+    os.makedirs(OUTPUT_DIR)
+  except:
+    print 'Race condition!', os.path.exists(OUTPUT_DIR)
+
+STATSFILE = OUTPUT_DIR + str(JOB_INDEX).zfill(2) + SUFFIX + 'p'
+MODELFILE = OUTPUT_DIR + str(JOB_INDEX).zfill(2) + SUFFIX + 'h5'
+
+print 'Working in', OUTPUT_DIR
+print 'Storing', STATSFILE
+print 'Storing', MODELFILE
+
+if os.path.exists(STATSFILE) and os.path.exists(MODELFILE):
+  print 'WAIT A MINUTE!! WE HAVE DONE THIS ONE BEFORE!'
+  sys.exit(0)
+
+
+#
+#
+# DATA GENERATION
+#
+#
+def in_array(haystack, needle):
+  for h in haystack:
[…]

### 2018-03-13 EXP/run_position_angle.py
Update Fig 3 experiments.

@@ -17,9 +17,9 @@ import ClevelandMcGill as C
 
 
 EXPERIMENT = sys.argv[1] # f.e. Angle
-CLASSIFIER = sys.argv[3] # 'LeNet'
-NOISE = sys.argv[4] # True
-JOB_INDEX = int(sys.argv[5])
+CLASSIFIER = sys.argv[2] # 'LeNet'
+NOISE = sys.argv[3] # True
+JOB_INDEX = int(sys.argv[4])
 
 #
 #

### 2018-03-13 EXP/util.py
Add new figure option for nice plots.

@@ -12,9 +12,11 @@ class Util:
     return value, parameters
 
   @staticmethod
-  def imshow_nicely(image, filename=None):
+  def imshow_nicely(image, filename=None, new_figure=True):
 
-    plt.figure()
+    if new_figure:
+      plt.figure()
+      
     plt.imshow(image, cmap='Greys', interpolation='nearest')
     ax = plt.gca()
     ax.set_xticklabels('')

### 2018-03-14 PAPER/paper.tex
Update title.
Update stuffffff...
Work on the paper.

@@
 %% Paper title.
-\title{Visualization for Machine Perception}
+\title{Computing Graphical Perception}
 
@@
 %other entries to be set up for journal
-\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Visualization for Machine Perception}
+\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Computing Graphical Perception}
 %\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
@@
 
+The experiments shown in this paper are either regression or classification tasks.
+We formulate any estimation of quantities (e.g.
+angles, positions, lengths etc.) as a regression problem between $0$ and $1$.
+The output indicates the percentage in regards to the degrees of freedom of the individual experiment.
+If the experiment involves a choice, we formulate it as a classification problem.
+
 \subsection{Measures}
 
+\textbf{Accuracy.} We use the same metric as Cleveland McGill to measure accuracy.
+\begin{equation}
+	\log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
+\end{equation}
+
+\noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
+For regression tasks the loss is defined as mean squared error (MSE) and for classification tasks the loss is categorical cross-entropy.
+
 \subsection{Classifiers}
+
+Our classifiers are built upon a multilayer perceptron (MLP) which is a feedforward artificial neural network.
+We combine this MLP with different convolutional neural networks (CNNs) for preprocessing and feature generation.
+These include the traditional LeNet trained from scratch, as well as VGG19 and Xception trained using ImageNet.
+\\~\\
+\noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units.
+We then add a dropout layer to prevent overfitting and compute linear regression or classification (softmax).
+\\~\\
+\noindent{\textbf{Convolutional Neural Networks.}} We use CNNs to generate additional features as input to the MLP.
+For this, we train a \emph{LeNet} classifier with different samples for each experiment.
+
+All networks are trained with the same parameters.
+
+%\begin{table}[t]
+%\caption{Training parameters, cost, and results of our guided proofreading classifier versus focused proofreading by Plaza~\cite{focused_proofreading}.
+Both methods were trained on the same mouse brain dataset using the same hardware (Tesla X GPU).}%While the training of our classifier is more expensive, testing accuracy is superior.
+}
+%
+%\small{
+%\begin{tabular}{lll}
+%	\toprule
+%	Classifier & Parameters & Layers
+%	\begin{tabular}{l}
+%		\vspace{0.2mm} \\
+%		\midrule
+%		\emph{Results---Test Set} \\ \midrule Cost [m]: 383 \\ Val. loss: 0.0845 \\ Val. acc.: 0.969 \\ Test. acc.: 0.94 \\ Prec./Recall: 0.94/0.94 \\ F1 Score: 0.94 \\ ~ \\
+%	\end{tabular}
+%\end{tabular}
+%
+%\vspace{0.5mm}
+%\begin{tabular}{ll}
+%	\toprule
+%	\begin{tabular}{l}
+%		\textbf{Focused Proofreading}\\ \midrule
+%		\emph{Parameters} \\ \midrule
+%		Iterations: 3 \\
+%		Learning strategy: 2\\
+%		Mito agglomeration: Off~~~~~~ \\  % Crappy alignment spacing
+%		Threshold: 0.0\\~\\
+%	\end{tabular}
+%	&
+%	\begin{tabular}{l}
+%		\vspace{0.2mm} \\
+%		\midrule
+%		\emph{Results---Test Set} \\ \midrule Cost [m]: 217 \\ Val. acc.: 0.99 \\ Test. acc.: 0.68 \\ Prec./Recall: 0.58/0.56 \\ F1 Score: 0.54 \\
+%	\end{tabular}
+%%	\bottomrule
+%\end{tabular}
+%\hrule
+%}
+%\label{tab:parameters}
+%\vspace{-4mm}
+%\end{table}
 
@@
 
+\begin{figure*}[h]
+	  \includegraphics[width=\linewidth]{figure1.pdf}
+  \caption{Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
+We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
+	\label{fig:figure1_results}
+\end{figure*}
+
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
+  \caption{Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts.
+We report the MSE measure for both encodings of four different classifier on previously unseen validation data.}
+	\label{fig:position_angle_results}
+\end{figure}
+
+\begin{figure*}[t]
+	  \includegraphics[width=\linewidth]{figure3_mlae.pdf}
+  \caption{Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
+We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
+	\label{fig:position_angle_results}
+\end{figure*}
+
 Cleveland McGills Ranking - can we observe something similar?

### 2018-03-14 PAPER/01_related_work.tex

@@
-\section{Related Work}
+\section{Previous Work}
 
-\cite{cleveland_mcgill} Clevenland McGill experiments
+\textbf{Graphical Perception.} Cleveland and McGill~\cite{cleveland_mcgill} introduced the concept of \emph{graphical perception} and investigated how different visual attributes and encodings are perceivable by humans.
+They define \emph{elementary perceptual tasks} as mental-visual stimuli to understand encodings in visualizations.
+
+\cite{HeerBostock2010} Heer and Bostock, CHI 2010 --- Crowdsourcing Graphical Perception: Using Mechanical Turk to Assess Visualization Design
 
@@
 
-Pineo et al.~\cite{Pineo2012_computational_perception} present a method to automatically evaluating and optimizing visualizations using a computational model of human vision, based on a neural network simulation of the early perceptual processing in the retina and primary visual cortex.
+\textbf{Visual Cortex Inspired Machine Learning.} LeNet etc and work from THomas Serre
+
+\textbf{Computing Perception.}
+
+Pineo et al.~\cite{Pineo2012_computational_perception} present a method to automatically evaluate and optimize visualizations using a computational model of human vision, based on a neural network simulation of the early perceptual processing in the retina and primary visual cortex.
 [JT] Copied from their abstract.

### 2018-03-16 PAPER/paper.tex

@@
 \usepackage{booktabs}                  % only used for the table example
+\usepackage{makecell}
+\usepackage{multirow}
+
 %% We encourage the use of mathptmx for consistent usage of times font
@@
 \\~\\
-\noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units.
+\noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units~(Fig.~\ref{fig:classifiers}).
 We then add a dropout layer to prevent overfitting and compute linear regression or classification (softmax).
@@
 \noindent{\textbf{Convolutional Neural Networks.}} We use CNNs to generate additional features as input to the MLP.
-For this, we train a \emph{LeNet} classifier with different samples for each experiment.
+We train the \emph{LeNet} classifier with tune it specifically towards each visualization.
+For \emph{VGG19} and \emph{Xception}, we generate features using previously trained weights on ImageNet.
+\\~\\
+\noindent{\textbf{Optimization.}} All networks are optimized using stochastic gradient descent with Nesterov momentum using fixed parameters (Table \ref{tab:parameters}).
+We train for $1000$ epochs but stop early if the loss does not decrease for ten epochs.
 
-All networks are trained with the same parameters.
-
-%\begin{table}[t]
-%\caption{Training parameters, cost, and results of our guided proofreading classifier versus focused proofreading by Plaza~\cite{focused_proofreading}.
-Both methods were trained on the same mouse brain dataset using the same hardware (Tesla X GPU).}%While the training of our classifier is more expensive, testing accuracy is superior.
+\begin{table}[t]
+\centering
+\caption{We use different feature generators as input to a multilayer perceptron which performs linear regression or the classification task.
+This yields different sets of trainable parameters.
+We also train the MLP directly on the visualizations without any additional feature generation.}
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lrl}
+%	\toprule
+%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
+%	\midrule
+%	MLP & $0$ & $2,560,513$ \\
+%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
+%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
+%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
+%	\bottomrule
+	\toprule
+	Classifier & \makecell{Trainable\\Parameters} & Optimization \\
+	\midrule
+	MLP & $2,560,513$ & SGD (Nesterov momentum)\\
+	\emph{LeNet} + MLP & $8,026,083$ & Learning rate: $0.0001$\\
+	\emph{VGG19} + MLP & $21,204,545$ & Momentum: $0.9$ \\
+	\emph{Xception} + MLP & $25,580,585$ & \makecell[tl]{Batchsize: 32\\Epochs: $1000$ (Early Stopping)}\\
+	\bottomrule
+\end{tabular}
 }
-%
-%\small{
-%\begin{tabular}{lll}
-%	\toprule
-%	Classifier & Parameters & Layers
-%	\begin{tabular}{l}
-%		\vspace{0.2mm} \\
-%		\midrule
-%		\emph{Results---Test Set} \\ \midrule Cost [m]: 383 \\ Val. loss: 0.0845 \\ Val. acc.: 0.969 \\ Test. acc.: 0.94 \\ Prec./Recall: 0.94/0.94 \\ F1 Score: 0.94 \\ ~ \\
-%	\end{tabular}
-%\end{tabular}
-%
-%\vspace{0.5mm}
-%\begin{tabular}{ll}
-%	\toprule
-%	\begin{tabular}{l}
-%		\textbf{Focused Proofreading}\\ \midrule
-%		\emph{Parameters} \\ \midrule
-%		Iterations: 3 \\
-%		Learning strategy: 2\\
-%		Mito agglomeration: Off~~~~~~ \\  % Crappy alignment spacing
-%		Threshold: 0.0\\~\\
-%	\end{tabular}
-%	&
-%	\begin{tabular}{l}
-%		\vspace{0.2mm} \\
-%		\midrule
-%		\emph{Results---Test Set} \\ \midrule Cost [m]: 217 \\ Val. acc.: 0.99 \\ Test. acc.: 0.68 \\ Prec./Recall: 0.58/0.56 \\ F1 Score: 0.54 \\
-%	\end{tabular}
-%%	\bottomrule
-%\end{tabular}
-%\hrule
-%}
-%\label{tab:parameters}
-%\vspace{-4mm}
-%\end{table}
+\label{tab:parameters}
+\vspace{-4mm}
+\end{table}
 
 \begin{figure}[t]
-	  \includegraphics[width=\linewidth]{fish.jpg}
-  \caption{The Classifiers...}
+	\centering
+	  \includegraphics[width=.8\linewidth]{classifiers.pdf}
+  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
+We use Dropout regularization to prevent overfitting.
+We learn categorical and unordered dependent variables using the softmax function and perform linear regression for continuous variables.
+The MLP can learn the visualizations directly but we also learn features generated by LeNet (2 conv.
+layers, filter size $5\time5$), VGG19 trained on ImageNet (16 conv.
+layers, filter size $3\times3$), or Xception trained on ImageNet (36 conv.
+layers, filter size $3\times3$) to increase the number of trainable parameters.}
 	\label{fig:classifiers}

### 2018-03-16 PAPER/00_introduction.tex
Paper draft in the works.

@@
 GOALS
-...reduce the gap between neurobiology and data science to advance the understanding of visual cortex inspired machine learning
+\begin{itemize}
+	\item reduce the gap between neurobiology and data science to advance the understanding of visual cortex inspired machine learning
+	\item insights for infovis for machines
+\end{itemize}
 
 CONTRIBUTIONS
-
-- experiments of cleveland mcgill with systematic parametrization and evaluation
-
-- ranking like cleveland mcgill for machine perception?
-
-- many other insights?
-
-- framework
+\begin{itemize}
+	\item experiments of cleveland mcgill with systematic parametrization and evaluation for computational perception
+	\item ranking of elementary perceptual tasks for machine perception
+	\item validation of weber's law for feedforward neural networks
+	\item many insights for infovis for machines
+\end{itemize}

### 2018-03-16 PAPER/paper.tex
Paper draft in the works.

@@
   \centering
-  \includegraphics[width=\linewidth]{fish.jpg}
-  \caption{Here is a fish.}
+  \includegraphics[width=\linewidth]{teaser.pdf}
+  \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking visual cortex inspired machine learning classifiers to assess the relationship between values encoded in pie charts and bar charts.
+Similar to the findings of Clevenland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs predict more accurately when working with bar charts (mean squared error, MSE in green).}
 	\label{fig:teaser}
@@
 
-The experiments shown in this paper are either regression or classification tasks.
+The experiments shown in this paper are either supervised regression or classification tasks.
 We formulate any estimation of quantities (e.g.
@@
 
-\textbf{Accuracy.} We use the same metric as Cleveland McGill to measure accuracy.
+\textbf{Accuracy.} We use the same metric as Cleveland and McGill to measure accuracy.
 \begin{equation}
@@
 \end{equation}
-
+\\~\\
+\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the confidence intervals.
+\\~\\
 \noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
@@
 
+\subsection{Data}
+
+We create all visualizations as parametrized rasterized images without interpolation.
+The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}.
+We add subtle random noise ($0.05$) to each pixel to introduce additional variation.
+\\~\\
+\noindent\textbf{Training/Validation/Test Splits.} We specify the size of each split set as follows: 60,000 training images, 20,000 validation images, and 20,000 test images.
+We then randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables.
+This eliminates leakage during training and evaluation.
+We also scale each set independently: images to the range of $-.5$ to $.5$ and labels to the range of $0.0$ to $1.0$.
+\\~\\
+\noindent\textbf{Cross-classifier variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.
+
 \section{Elementary Perceptual Tasks}
 
-THe Figure 1 of Cleveland McGill
+Cleveland and McGill describe the mapping of graphical elements to quantitative variables as \emph{elementary perceptual tasks} and introduce a list of ten different encodings in their paper~\cite{cleveland_mcgill}.
+We create visualizations of these tasks as rasterized images (Fig.~\ref{fig:elementary_perceptual_tasks}).
 
-\subsection{Parametrization}
+\subsection{Parametrizations}
+\label{sec:parametrizations}
+We generate multiple parameterizations for each elementary perceptual task (Fig.~\ref{fig:elementary_perceptual_tasks}) and sequentially increase the number of parameters.
+For instance, for \emph{position non-aligned scale} we first only vary the origin of the coordinate system which yields just 10 different parameters.
+We then include translation along the y-axis with a significant increase in variability.
+We then also add x-movement and a variable spot size.
+This results in more complex datasets depending on the variability setting.
+Table~\ref{tab:encoding_parameters} shows the different settings.
+It is important to consider this variability when evaluating different classifiers with individual trainable parameters (Table~\ref{tab:parameters}).
 
-Here we show each of Figure 1 with its parametrization.
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{figure1_overview.pdf}
+  \caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of the elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
+We vary the parameters of each perceptual task and then assess the interpretability of feed-forward neural networks.}
+	\label{fig:elementary_perceptual_tasks}
+\end{figure}
+
+\begin{table}[t]
+\centering
+\caption{\textbf{Variability of Elementary Perceptual Tasks.} We sequentially increase the number of parameters for every visual encoding of the elementary perceptual tasks.
+This introduces variability and increasingly more complex datasets.}
+\resizebox{\linewidth}{!}{
+\begin{tabular}{llr}
+%	\toprule
+%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
+%	\midrule
+%	MLP & $0$ & $2,560,513$ \\
+%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
+%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
+%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
+%	\bottomrule
+	\toprule
+	Elementary Perceptual Task & Variability & Parameters\\
+	\midrule
+	\makecell[tl]{\emph{Position Common Scale}} & \makecell[tl]{Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $60$ \\ $3600$ \\ $21600$}\\
+
+	\midrule
+	\makecell[tl]{\emph{Position Non-Aligned Scale}} & \makecell[tl]{Origin\\+ Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $10$ \\$600$ \\ $36000$ \\ $216000$}\\
+
+	\midrule
+	\makecell[tl]{\emph{Length}} & \makecell[tl]{Length\\+ Position Y \\+ Position X \\+ Width} & \makecell[tr]{ $60$ \\ $2400$ \\ $144000$\\$864000$}\\
+
+	\midrule
+	\makecell[tl]{\emph{Direction}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $360$ \\ $21600$ \\ $1296000$}\\
+
+	\midrule
+	\makecell[tl]{\emph{Angle}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $90$ \\ $5400$ \\ $324000$}\\
+
+	\midrule
+	\makecell[tl]{\emph{Area}} & \makecell[tl]{Radius\\+ Position Y \\+ Position X} & \makecell[tr]{ $40$ \\ $800$ \\ $16000$}\\
+
+	\midrule
+	\makecell[tl]{\emph{Volume}} & \makecell[tl]{Cube Sidelength\\+ Position Y \\+ Position X} & \makecell[tr]{ $20$ \\ $400$ \\ $8000$}\\
+
+	\midrule
+	\makecell[tl]{\emph{Curvature}} & \makecell[tl]{Midpoint Curvature\\+ Position Y \\+ Position X\\+ Width} & \makecell[tr]{ $80$ \\ $1600$ \\ $64000$ \\ $2560000$}\\
+
+	\midrule
+	\makecell[tl]{\emph{Shading}} & \makecell[tl]{Density\\+ Position Y \\+ Position X} & \makecell[tr]{ $100$ \\ $1600$ \\ $2000$ \\ $40000$}\\
+
+	\bottomrule
+\end{tabular}
+}
+\label{tab:encoding_parameters}
+\end{table}
+
+\subsection{Hypotheses}
+
+We proposed four hypotheses entering the elementary perceptual task experiment:
+
+\begin{itemize}
+	\item \textbf{H1.1} \textbf{Visual cortex inspired classifiers are able to connect graphical elements to their quantitative variables.} While much simpler models than their biological pendant, convolutional neural networks are heavily influenced by our biological knowledge of the visual system.
+Such classifiers therefor follow the same principles as human perception.
+	\item \textbf{H1.2} \textbf{Computed perceptual performance is dependent on classifier complexity.} We evaluate multiple classifiers with different numbers of trainable parameters.
+A more complex classifier (with higher number of parameters) will perform better on elementary perceptual tasks.
+	\item \textbf{H1.3} \textbf{Some visual encodings are better than others for computations.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
+We investigate whether this order is also relevant for computing graphical perception.
+	\item \textbf{H1.4} \textbf{Classifiers trained on perceptual tasks can generalize to more or less complex variations of the same task.} Recent research suggests that convolutional neural networks generalize extremely well.
+While the underlying reasons are mainly yet unknown, this property allows them to perform on variations of a similar perceptual task.
+\end{itemize}
 
@@
 
-This is pie chart vs bar chart
+The position-angle experiment was originally performed by Cleveland and McGill to measure whether humans can better perceive quantities encoded as positions or as angles~\cite{cleveland_mcgill}.
+The actual experiment then compares pie charts versus bar charts since these map down to elementary position and angle judgement.
+We create rasterized images mimicking Cleveland and McGill's proposed encoding and investigate computational perception of our four classifiers.
 
 \begin{figure}[t]
-	  \includegraphics[width=\linewidth]{fish.jpg}
-  \caption{The Position-Angle Experiment}
+	  \includegraphics[width=\linewidth]{figure3_overview}
+  \caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
+The experimental task involves the judgement of different encoded values in comparison to the largest encoded values.
+The pie chart (left) and the bar chart (right) visualize the same data point.
+In their paper, Cleveland and McGill report less errors using bar charts.}
 	\label{fig:position_angle_experiment}
@@
 
+\subsection{Hypotheses}
+
+We proposed four hypotheses entering the elementary perceptual task experiment:
+
+\begin{itemize}
+	\item \textbf{H2.1} \textbf{Computed perceptual performance is better using bar charts than pie charts.} Cleveland and McGill report that position judgements are almost twice as accurate as angle judgements.
+This renders bar charts superior to pie charts and should also be the case for convolutional neural networks.
+	\item \textbf{H2.2} \textbf{Classifiers can learn position faster than angles.} We assume that understanding bar charts is easier than understanding pie charts.
+We suspect that our classifiers learn encodings of positions faster than of angles resulting in more efficient training and faster convergence.
+\end{itemize}
+
 \section{Position-Length Experiment}
 
-This is the one where we estimate two selected bars compared to the longest one
+This is the one where we estimate two selected bars compared to the longest one - very similar to the previous one but, not yet done.
+We basically test divided versus grouped barchart and we estimate one relation between two marked quanitities: what percent the smaller is of the larger.
+
+There are five types: type 1-3 this is a position judgement along a common scale.
+(btw all classifiers seem to do that extremely well in the elementary tasks so we assume this will work well here too).
+Types 4-5 are length judgements and we know that the classifiers struggle with that quite a bit.
+
+The setup from Cleveland McGill is first a classification task: which one is smaller? and then a regression task: how much smaller. So we have to see how to encode this.
+
+\subsection{Hypotheses}
+
+We proposed two hypotheses entering the elementary perceptual task experiment:
+
+\begin{itemize}
+	\item \textbf{H3.1} \textbf{Grouped bar charts are better computational perceivable than divided bar charts.} A grouped bar chart involves judging a position while a divided bar chart most likely (if not the bottom is looked at) requires length judgements.
+Classifiers are better at judging position than at judging length so grouped bar charts are easier to grasp in terms of computational perception.
+	\item \textbf{H3.2} \textbf{not yet} Any ideas?
+\end{itemize}
 
 \begin{figure}[t]
-	  \includegraphics[width=\linewidth]{fish.jpg}
-  \caption{The Position-Length Experiment}
+	  \includegraphics[width=\linewidth]{position_length_original1.png}
+	  \includegraphics[width=\linewidth]{position_length_original2.png}
+
+  \caption{\textbf{Position-Length Experiment.} (Not yet) Rasterized versions of the graphs of Cleveland and McGill's position-length experiment.
+The perceptual task involves comparing.
+the two dot-marked quantities across five different visual encodings of either grouped or divided bar charts.
+We evaluate which type of bar chart performs better with our neural networks as a combined classification and regression problem.
+The first task is to select which of the marked quantities is smaller (classification) and the second task is to specify how much smaller it is (regression).}
 	\label{fig:position_length_experiment}
@@
 
+Visual cues can help converting graphical elements back to their real world variables.
+Cleveland and McGill introduced the bars and framed rectangles experiment which judges the elementary perceptual task of position along non-aligned scales~\cite{cleveland_mcgill}.
+
+\subsection{Hypotheses}
+
+We proposed two hypotheses entering the elementary perceptual task experiment:
+
+\begin{itemize}
+	\item \textbf{H4.1} \textbf{Classifiers can leverage additional visual cues.} The original bar and framed rectangle experiment shows how visual cues aid humans in mapping graphical elements to quantitative variables.
+This should be the same for feed-forward neural networks since they are inspired by the visual system.
+	\item \textbf{H4.2} \textbf{Weber's law can be transferred to computational perception.} Cleveland and McGill confirmed Weber's law based on the bar and framed rectangle experiment.
+For humans, the ability to perceive change within a distribution is proportional to the size of the initial distribution.
+\end{itemize}
+
 \begin{figure}[t]
-	  \includegraphics[width=\linewidth]{fish.jpg}
-  \caption{The Bars and Framed Rectangles Experiment}
+	  \includegraphics[width=\linewidth]{figure12_overview}
+  \caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales.
+For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles (right).
+In this case, the right bar is heigher as visible with less free space when adding the frame.
+We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
 	\label{fig:bars_and_framed_rectangles_experiment}
@@
 
-\subsection{Weber's Law}
+\subsection{Weber-Fechner's Law}
+
+As identified by Cleveland and McGill, the bar and framed rectangle experiment is closely related to Weber's law.
+This psychophysics law states that perceivable difference within a distribution is proportional to the initial size of the distribution.
+Weber's law goes hand-in-hand with Fechner's law.
+We conduct an additional experiment based on the original illustrations of the Weber-Fechner law to investigate whether this law can be applied to computational perception of our classifiers (Fig.~\ref{fig:webers_law}).
+
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{webers_law}
+  \caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution.
+The lower square contains 10 more dots than the upper one on both sides.
+However, the difference is easily perceivable on the left while the squares on the right almost look the same.
+We generate rasterized visualizations similar to this setup and evaluate our classifiers.}
+	\label{fig:webers_law}
+\end{figure}
 
@@
 
+\subsection{Elementary Perceptual Tasks}
+
+some are good and some are bad.. why?
+
 \begin{figure*}[h]
 	  \includegraphics[width=\linewidth]{figure1.pdf}
-  \caption{Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
+  \caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
 We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
@@
 
+\textbf{Computational Perception Ranking.}
+
+Cleveland McGills Ranking - can we observe something similar?
+
+\begin{enumerate}
+	\item Position along a common scale e.g. scatter plot
+	\item Position on identical but nonaligned scales e.g. multiple scatter plots
+	\item Length e.g. bar chart
+	\item Angle \& Slope (tie) e.g. pie chart
+	\item Area e.g. bubbles
+	\item Volume, density, and color saturation (tie) e.g. heatmap
+	\item Color hue e.g. newsmap
+\end{enumerate}
+
+\noindent{\textbf{Cross-classifier variability.}}
+
+Can a neural network generalize on simple perceptual tasks?
+
+\subsection{Position-Angle Experiment}
+
+Bar charts are more accurate (Fig.~\ref{fig:figure3_mlae}) and networks converge faster (Fig.~\ref{fig:figure3_val_loss}). This is great.
+
 \begin{figure}[t]
 	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
-  \caption{Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts.
+  \caption{\textbf{Classifier Efficiency of the Position-Angle experiment.} Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts.
 We report the MSE measure for both encodings of four different classifier on previously unseen validation data.}
-	\label{fig:position_angle_results}
+	\label{fig:figure3_val_loss}
 \end{figure}
@@
 	  \includegraphics[width=\linewidth]{figure3_mlae.pdf}
-  \caption{Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
+  \caption{\textbf{Computational results of the Position-Angle experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
 We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:position_angle_results}
+	\label{fig:figure3_mlae}
 \end{figure*}
 
-Cleveland McGills Ranking - can we observe something similar?
+\subsection{Position-Length Experiment}
 
-1. Position along a common scale e.g. scatter plot
+We have nothing yet here...
 
-2. Position on identical but nonaligned scales e.g. multiple scatter plots
+\subsection{Bars and Framed Rectangles Experiment}
 
-3. Length e.g. bar chart
+First run indicates that framed rectangles perform better but we dont really know it yet.
 
-4.Angle \& Slope (tie) e.g. pie chart
-
-5. Area e.g. bubbles
-
-6. Volume, density, and color saturation (tie) e.g. heatmap
-
-7. Color hue e.g. newsmap
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{figure12_val_loss.png}
+  \caption{\textbf{Classifier Efficiency of the Bars and Framed Rectangles experiment.} Categorical Cross-Entropy loss for the \emph{bars and framed rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
+The frame around the bars adds an additional visual cue enables faster network convergence.
+This is not yet reproducible!}
+	\label{fig:figure12_val_loss}
+\end{figure}
 

### 2018-03-20 PAPER/paper.tex
Latest draft.

@@
 %% Paper title.
+% James' title suggestions
+% - Reproducing 'Graphical Perception' with CNNs
+% - 
 \title{Computing Graphical Perception}
@@
 %% Abstract section.
-\abstract{TODO%
+\abstract{
+Convolutional neural networks can successfully perform many computer vision tasks on images, and their early-layer learned representations are often said to mimic the early stages of the visual cortex.
+But can CNNs solve graphical perception tasks for visualization?
+We investigate this question by reproducing Cleveland and McGill's seminal 1984 paper, which defined elementary perceptual tasks for visualization, and measured human perception efficiency of different visual encodings with user studies.
+We measure the graphical perceptual capabilities of four classifiers on a) elementary perceptual tasks with increasing parametric complexity, b) the position-angle experiment which compares pie and bar charts, c) the position-length experiment which compares grouped and divided bar charts, and d) the bars and framed rectangles experiment where visual cues help to perceive information.
+We also study how feed-forward neural networks obey Weber's law, which defines the proportional relation between perceivable information and distribution density.
+We present the results of these experiments to foster the understanding of how such classifiers succeed and fail when applied to data visualizations.
 } % end of abstract
@@
 We train for $1000$ epochs but stop early if the loss does not decrease for ten epochs.
+\\~\\
+\noindent{\textbf{Environment.}} We run all experiments on an NVIDIA DGX1 machine with Tesla V100 graphical processing units. We use the KERAS framework with tensorflow.
 

### 2018-03-20 PAPER/paper.tex
Update paper.

@@
 %% Abstract section.
-\abstract{
+\abstract{%Convolutional neural networks are being successfully used for image understanding and object recognition tasks while regularly outperforming humans.
+Despite their tremendous success, little is known perceptional capabilities.
+In there Graphical Perception paper from 1984, Cleveland and McGill define elementary perceptual tasks that let people extract quantitative information from visualizations and measure human perception of different visual encodings with in depth user studies.
+%We replicate the experimental setup of Cleveland and McGill and evaluate the perceptional capabilities of four modern classifiers based on neural networks.
+We systematically test how the classifers perform on a) elementary perceptual tasks with increasing parametric complexity, b) the position-angle experiment which compares pie and bar charts, c) the position-length experiment which compares grouped and divided bar charts, and d) the bars and framed rectangles experiment where visual cues help to perceive information.
+We present the results of these experiments to foster the understanding of how such classifiers can be successfully applied to data visualizations.
+On this journey, we also study how the feed-forward neural networks obey Weber's law which defines the proportional relation between perceivable information and distribution size.
+We further introduce a ranking of elementary visual encodings targeted towards modern deep neural networks and derive practical evidence for properties of convolutional neural networks such as the translation invariance.
 Convolutional neural networks can successfully perform many computer vision tasks on images, and their early-layer learned representations are often said to mimic the early stages of the visual cortex.
 But can CNNs solve graphical perception tasks for visualization?
-We investigate this question by reproducing Cleveland and McGill's seminal 1984 paper, which defined elementary perceptual tasks for visualization, and measured human perception efficiency of different visual encodings with user studies.
+We investigate this question by reproducing Cleveland and McGill's seminal 1984 experiments, which measured human perception efficiency of different visual encodings and defined elementary perceptual tasks for visualization.
 We measure the graphical perceptual capabilities of four classifiers on a) elementary perceptual tasks with increasing parametric complexity, b) the position-angle experiment which compares pie and bar charts, c) the position-length experiment which compares grouped and divided bar charts, and d) the bars and framed rectangles experiment where visual cues help to perceive information.
@@
 
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{crossnetwork.pdf}
+  \caption{\textbf{Cross-classifier variability for the perceptual task of measuring curvature.} We use predictions of LeNet classifiers trained on different parametrizations of the \emph{curvature} elementary perceptual task and measure the mean logistic absolute error (MLAE).
+The lower score, the better.
+Classifiers trained on curves with variable position can generalize even if the axis of translation varies.
+However, classifiers trained on fixed positions of curves are not able to measure translated curves.}
+	\label{fig:cross_network}
+\end{figure}
+
 \subsection{Position-Angle Experiment}

### 2018-03-20 PAPER/03_elementary_perceptual_tasks.tex
Add subfiles to the tex and some updates from VCG.

--- 2018-03-16 PAPER/paper.tex
@@
-\documentclass[journal]{vgtc}                % final (journal style)
-%\documentclass[review,journal]{vgtc}         % review (journal style)
-%\documentclass[widereview]{vgtc}             % wide-spaced review
-%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
-
-%% Uncomment one of the lines above depending on where your paper is
-%% in the conference process. ``review'' and ``widereview'' are for review
-%% submission, ``preprint'' is for pre-publication, and the final version
-%% doesn't use a specific qualifier.
-
-%% Please use one of the ``review'' options in combination with the
-%% assigned online id (see below) ONLY if your paper uses a double blind
-%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
-%% in the past.
-
-%% Please note that the use of figures other than the optional teaser is not permitted on the first page
-%% of the journal version.  Figures should begin on the second page and be
-%% in CMYK or Grey scale format, otherwise, colour shifting may occur
-%% during the printing process.  Papers submitted with figures other than the optional teaser on the
-%% first page will be refused. Also, the teaser figure should only have the
-%% width of the abstract as the template enforces it.
-
-%% These few lines make a distinction between latex and pdflatex calls and they
-%% bring in essential packages for graphics and font handling.
-%% Note that due to the \DeclareGraphicsExtensions{} call it is no longer necessary
-%% to provide the the path and extension of a graphics file:
-%% \includegraphics{diamondrule} is completely sufficient.
-%%
-\ifpdf%                                % if we use pdflatex
-  \pdfoutput=1\relax                   % create PDFs from pdfLaTeX
-  \pdfcompresslevel=9                  % PDF Compression
-  \pdfoptionpdfminorversion=7          % create PDF 1.7
-  \ExecuteOptions{pdftex}
-  \usepackage{graphicx}                % allow us to embed graphics files
-  \DeclareGraphicsExtensions{.pdf,.png,.jpg,.jpeg} % for pdflatex we expect .pdf, .png, or .jpg files
-\else%                                 % else we use pure latex
-  \ExecuteOptions{dvips}
-  \usepackage{graphicx}                % allow us to embed graphics files
-  \DeclareGraphicsExtensions{.eps}     % for pure latex we expect eps files
-\fi%
-
-%% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
-\graphicspath{{gfx/}{./}} % where to search for the images
-
-\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
-\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
-\usepackage{textcomp}                  % use better special symbols
-\usepackage{mathptmx}                  % use matching math font
-\usepackage{times}                     % we use Times as the main font
-\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
-\usepackage{cite}                      % needed to automatically sort the references
-\usepackage{tabu}                      % only used for the table example
-\usepackage{booktabs}                  % only used for the table example
-\usepackage{makecell}
-\usepackage{multirow}
-
-%% We encourage the use of mathptmx for consistent usage of times font
-%% throughout the proceedings. However, if you encounter conflicts
-%% with other math-related packages, you may want to disable it.
-
-%% In preprint mode you may define your own headline.
-%\preprinttext{To appear in IEEE Transactions on Visualization and Computer Graphics.}
-
-%% If you are submitting a paper to a conference for review with a double
-%% blind reviewing process, please replace the value ``0'' below with your
-%% OnlineID. Otherwise, you may safely leave it at ``0''.
-\onlineid{0}
-
-%% declare the category of your paper, only shown in review mode
-\vgtccategory{Research}
-%% please declare the paper type of your paper to help reviewers, only shown in review mode
-%% choices:
-%% * algorithm/technique
-%% * application/design study
-%% * evaluation
-%% * system
-%% * theory/model
-\vgtcpapertype{theory/model}
-
-%% Paper title.
-\title{Computing Graphical Perception}
-
-%% This is how authors are specified in the journal style
-
-%% indicate IEEE Member or Student Member in form indicated below
-\author{Daniel Haehn, \textit{Member, IEEE}, James Tompkin, and Hanspeter Pfister}
-\authorfooter{
-%% insert punctuation at end of each item
-\item Daniel Haehn, and Hanspeter Pfister are with the Paulson School of Engineering and Applied Sciences at Harvard University. \\
-E-mail: \{haehn,pfister\}@seas.harvard.edu.
-%
-\item James Tompkin is with the Thomas J. Watson Sr. Center for Information Technology at Brown University. \\E-mail: james\_tompkin@brown.edu.
-}
-
-%other entries to be set up for journal
-\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Computing Graphical Perception}
-%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
-
-%% Abstract section.
-\abstract{TODO%
-} % end of abstract
-
-%% Keywords that describe your work. Will show as 'Index Terms' in journal
-%% please capitalize first letter and insert punctuation after last keyword
-\keywords{Machine Perception, Deep Learning}
-
-%% ACM Computing Classification System (CCS). 
-%% See <http://www.acm.org/class/1998/> for details.
-%% The ``\CCScat'' command takes four arguments.
-
-%\CCScatlist{ % not used in journal version
-% \CCScat{K.6.1}{Management of Computing and Information Systems}%
-%{Project and People Management}{Life Cycle};
-% \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
-%}
-
-%% Uncomment below to include a teaser figure.
-\teaser{
-  \centering
-  \includegraphics[width=\linewidth]{teaser.pdf}
-  \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking visual cortex inspired machine learning classifiers to assess the relationship between values encoded in pie charts and bar charts.
-Similar to the findings of Clevenland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs predict more accurately when working with bar charts (mean squared error, MSE in green).}
-	\label{fig:teaser}
-}
-
-%% Uncomment below to disable the manuscript note
-%\renewcommand{\manuscriptnotetxt}{}
-
-%% Copyright space is enabled by default as required by guidelines.
-%% It is disabled by the 'review' option or via the following command:
-% \nocopyrightspace
-
-\vgtcinsertpkg
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-%% The ``\maketitle'' command must be the first command after the
-%% ``\begin{document}'' command. It prepares and prints the title block.
-
-%% the only exception to this rule is the \firstsection command
-
-\input{00_introduction.tex}
-
-\input{01_related_work.tex}
-
-\section{Experimental Setup}
-
-The experiments shown in this paper are either supervised regression or classification tasks.
-We formulate any estimation of quantities (e.g.
-angles, positions, lengths etc.) as a regression problem between $0$ and $1$.
-The output indicates the percentage in regards to the degrees of freedom of the individual experiment.
-If the experiment involves a choice, we formulate it as a classification problem.
-
-\subsection{Measures}
-
-\textbf{Accuracy.} We use the same metric as Cleveland and McGill to measure accuracy.
-\begin{equation}
-	\log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
-\end{equation}
-\\~\\
-\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the confidence intervals.
-\\~\\
-\noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
-For regression tasks the loss is defined as mean squared error (MSE) and for classification tasks the loss is categorical cross-entropy.
-
-\subsection{Classifiers}
-
-Our classifiers are built upon a multilayer perceptron (MLP) which is a feedforward artificial neural network.
-We combine this MLP with different convolutional neural networks (CNNs) for preprocessing and feature generation.
-These include the traditional LeNet trained from scratch, as well as VGG19 and Xception trained using ImageNet.
-\\~\\
-\noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units~(Fig.~\ref{fig:classifiers}).
-We then add a dropout layer to prevent overfitting and compute linear regression or classification (softmax).
-\\~\\
-\noindent{\textbf{Convolutional Neural Networks.}} We use CNNs to generate additional features as input to the MLP.
-We train the \emph{LeNet} classifier with tune it specifically towards each visualization.
-For \emph{VGG19} and \emph{Xception}, we generate features using previously trained weights on ImageNet.
-\\~\\
-\noindent{\textbf{Optimization.}} All networks are optimized using stochastic gradient descent with Nesterov momentum using fixed parameters (Table \ref{tab:parameters}).
-We train for $1000$ epochs but stop early if the loss does not decrease for ten epochs.
-
-\begin{table}[t]
-\centering
-\caption{We use different feature generators as input to a multilayer perceptron which performs linear regression or the classification task.
-This yields different sets of trainable parameters.
-We also train the MLP directly on the visualizations without any additional feature generation.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lrl}
-%	\toprule
-%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
-%	\midrule
-%	MLP & $0$ & $2,560,513$ \\
-%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
-%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
-%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
-%	\bottomrule
-	\toprule
-	Classifier & \makecell{Trainable\\Parameters} & Optimization \\
-	\midrule
-	MLP & $2,560,513$ & SGD (Nesterov momentum)\\
-	\emph{LeNet} + MLP & $8,026,083$ & Learning rate: $0.0001$\\
-	\emph{VGG19} + MLP & $21,204,545$ & Momentum: $0.9$ \\
-	\emph{Xception} + MLP & $25,580,585$ & \makecell[tl]{Batchsize: 32\\Epochs: $1000$ (Early Stopping)}\\
-	\bottomrule
-\end{tabular}
-}
-\label{tab:parameters}
-\vspace{-4mm}
-\end{table}
-
-\begin{figure}[t]
-	\centering
-	  \includegraphics[width=.8\linewidth]{classifiers.pdf}
-  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
-We use Dropout regularization to prevent overfitting.
-We learn categorical and unordered dependent variables using the softmax function and perform linear regression for continuous variables.
-The MLP can learn the visualizations directly but we also learn features generated by LeNet (2 conv.
-layers, filter size $5\time5$), VGG19 trained on ImageNet (16 conv.
-layers, filter size $3\times3$), or Xception trained on ImageNet (36 conv.
-layers, filter size $3\times3$) to increase the number of trainable parameters.}
-	\label{fig:classifiers}
-\end{figure}
-
-\subsection{Data}
-
-We create all visualizations as parametrized rasterized images without interpolation.
-The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}.
-We add subtle random noise ($0.05$) to each pixel to introduce additional variation.
-\\~\\
-\noindent\textbf{Training/Validation/Test Splits.} We specify the size of each split set as follows: 60,000 training images, 20,000 validation images, and 20,000 test images.
-We then randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables.
-This eliminates leakage during training and evaluation.
-We also scale each set independently: images to the range of $-.5$ to $.5$ and labels to the range of $0.0$ to $1.0$.
-\\~\\
-\noindent\textbf{Cross-classifier variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.
-
 \section{Elementary Perceptual Tasks}
@@
 \end{itemize}
-
-\section{Position-Angle Experiment}
-
-The position-angle experiment was originally performed by Cleveland and McGill to measure whether humans can better perceive quantities encoded as positions or as angles~\cite{cleveland_mcgill}.
-The actual experiment then compares pie charts versus bar charts since these map down to elementary position and angle judgement.
-We create rasterized images mimicking Cleveland and McGill's proposed encoding and investigate computational perception of our four classifiers.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_overview}
-  \caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
-The experimental task involves the judgement of different encoded values in comparison to the largest encoded values.
-The pie chart (left) and the bar chart (right) visualize the same data point.
-In their paper, Cleveland and McGill report less errors using bar charts.}
-	\label{fig:position_angle_experiment}
-\end{figure}
-
-\subsection{Hypotheses}
-
-We proposed four hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H2.1} \textbf{Computed perceptual performance is better using bar charts than pie charts.} Cleveland and McGill report that position judgements are almost twice as accurate as angle judgements.
-This renders bar charts superior to pie charts and should also be the case for convolutional neural networks.
-	\item \textbf{H2.2} \textbf{Classifiers can learn position faster than angles.} We assume that understanding bar charts is easier than understanding pie charts.
-We suspect that our classifiers learn encodings of positions faster than of angles resulting in more efficient training and faster convergence.
-\end{itemize}
-
-\section{Position-Length Experiment}
-
-This is the one where we estimate two selected bars compared to the longest one - very similar to the previous one but, not yet done.
-We basically test divided versus grouped barchart and we estimate one relation between two marked quanitities: what percent the smaller is of the larger.
-
-There are five types: type 1-3 this is a position judgement along a common scale.
-(btw all classifiers seem to do that extremely well in the elementary tasks so we assume this will work well here too).
-Types 4-5 are length judgements and we know that the classifiers struggle with that quite a bit.
-
-The setup from Cleveland McGill is first a classification task: which one is smaller? and then a regression task: how much smaller. So we have to see how to encode this.
-
-\subsection{Hypotheses}
-
-We proposed two hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H3.1} \textbf{Grouped bar charts are better computational perceivable than divided bar charts.} A grouped bar chart involves judging a position while a divided bar chart most likely (if not the bottom is looked at) requires length judgements.
-Classifiers are better at judging position than at judging length so grouped bar charts are easier to grasp in terms of computational perception.
-	\item \textbf{H3.2} \textbf{not yet} Any ideas?
-\end{itemize}
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{position_length_original1.png}
-	  \includegraphics[width=\linewidth]{position_length_original2.png}
-
-  \caption{\textbf{Position-Length Experiment.} (Not yet) Rasterized versions of the graphs of Cleveland and McGill's position-length experiment.
-The perceptual task involves comparing.
-the two dot-marked quantities across five different visual encodings of either grouped or divided bar charts.
-We evaluate which type of bar chart performs better with our neural networks as a combined classification and regression problem.
-The first task is to select which of the marked quantities is smaller (classification) and the second task is to specify how much smaller it is (regression).}
-	\label{fig:position_length_experiment}
-\end{figure}
-
-\section{Bars and Framed Rectangles Experiment}
-
-Visual cues can help converting graphical elements back to their real world variables.
-Cleveland and McGill introduced the bars and framed rectangles experiment which judges the elementary perceptual task of position along non-aligned scales~\cite{cleveland_mcgill}.
-
-\subsection{Hypotheses}
-
-We proposed two hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H4.1} \textbf{Classifiers can leverage additional visual cues.} The original bar and framed rectangle experiment shows how visual cues aid humans in mapping graphical elements to quantitative variables.
-This should be the same for feed-forward neural networks since they are inspired by the visual system.
-	\item \textbf{H4.2} \textbf{Weber's law can be transferred to computational perception.} Cleveland and McGill confirmed Weber's law based on the bar and framed rectangle experiment.
-For humans, the ability to perceive change within a distribution is proportional to the size of the initial distribution.
-\end{itemize}
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_overview}
-  \caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales.
-For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles (right).
-In this case, the right bar is heigher as visible with less free space when adding the frame.
-We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
-	\label{fig:bars_and_framed_rectangles_experiment}
-\end{figure}
-
-\subsection{Weber-Fechner's Law}
-
-As identified by Cleveland and McGill, the bar and framed rectangle experiment is closely related to Weber's law.
-This psychophysics law states that perceivable difference within a distribution is proportional to the initial size of the distribution.
-Weber's law goes hand-in-hand with Fechner's law.
-We conduct an additional experiment based on the original illustrations of the Weber-Fechner law to investigate whether this law can be applied to computational perception of our classifiers (Fig.~\ref{fig:webers_law}).
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{webers_law}
-  \caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution.
-The lower square contains 10 more dots than the upper one on both sides.
-However, the difference is easily perceivable on the left while the squares on the right almost look the same.
-We generate rasterized visualizations similar to this setup and evaluate our classifiers.}
-	\label{fig:webers_law}
-\end{figure}
-
-\section{Results and Discussion}
-
-\subsection{Elementary Perceptual Tasks}
-
-some are good and some are bad.. why?
-
-\begin{figure*}[h]
-	  \includegraphics[width=\linewidth]{figure1.pdf}
-  \caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure1_results}
-\end{figure*}
-
-\textbf{Computational Perception Ranking.}
-
-Cleveland McGills Ranking - can we observe something similar?
-
-\begin{enumerate}
-	\item Position along a common scale e.g. scatter plot
-	\item Position on identical but nonaligned scales e.g. multiple scatter plots
-	\item Length e.g. bar chart
-	\item Angle \& Slope (tie) e.g. pie chart
-	\item Area e.g. bubbles
-	\item Volume, density, and color saturation (tie) e.g. heatmap
-	\item Color hue e.g. newsmap
-\end{enumerate}
-
-\noindent{\textbf{Cross-classifier variability.}}
-
-Can a neural network generalize on simple perceptual tasks?
-
-\subsection{Position-Angle Experiment}
-
-Bar charts are more accurate (Fig.~\ref{fig:figure3_mlae}) and networks converge faster (Fig.~\ref{fig:figure3_val_loss}). This is great.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
-  \caption{\textbf{Classifier Efficiency of the Position-Angle experiment.} Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts.
-We report the MSE measure for both encodings of four different classifier on previously unseen validation data.}
-	\label{fig:figure3_val_loss}
-\end{figure}
-
-\begin{figure*}[t]
-	  \includegraphics[width=\linewidth]{figure3_mlae.pdf}
-  \caption{\textbf{Computational results of the Position-Angle experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure3_mlae}
-\end{figure*}
-
-\subsection{Position-Length Experiment}
-
-We have nothing yet here...
-
-\subsection{Bars and Framed Rectangles Experiment}
-
-First run indicates that framed rectangles perform better but we dont really know it yet.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_val_loss.png}
-  \caption{\textbf{Classifier Efficiency of the Bars and Framed Rectangles experiment.} Categorical Cross-Entropy loss for the \emph{bars and framed rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-The frame around the bars adds an additional visual cue enables faster network convergence.
-This is not yet reproducible!}
-	\label{fig:figure12_val_loss}
-\end{figure}
-
-\section{Conclusions}
-
-Future work: allow insights for infovis for machines
-
-%% if specified like this the section will be committed in review mode
-%\acknowledgments{
-%The authors wish to thank A, B, and C. This work was supported in part by
-%a grant from XYZ (\# 12345-67890).}
-
-%\bibliographystyle{abbrv}
-\bibliographystyle{abbrv-doi}
-%\bibliographystyle{abbrv-doi-narrow}
-%\bibliographystyle{abbrv-doi-hyperref}
-%\bibliographystyle{abbrv-doi-hyperref-narrow}
-
-\bibliography{paper.bib}
-\end{document}

### 2018-03-20 PAPER/05_position_length_experiment.tex
Add subfiles to the tex and some updates from VCG.

--- 2018-03-16 PAPER/paper.tex
@@
-\documentclass[journal]{vgtc}                % final (journal style)
-%\documentclass[review,journal]{vgtc}         % review (journal style)
-%\documentclass[widereview]{vgtc}             % wide-spaced review
-%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
-
-%% Uncomment one of the lines above depending on where your paper is
-%% in the conference process. ``review'' and ``widereview'' are for review
-%% submission, ``preprint'' is for pre-publication, and the final version
-%% doesn't use a specific qualifier.
-
-%% Please use one of the ``review'' options in combination with the
-%% assigned online id (see below) ONLY if your paper uses a double blind
-%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
-%% in the past.
-
-%% Please note that the use of figures other than the optional teaser is not permitted on the first page
-%% of the journal version.  Figures should begin on the second page and be
-%% in CMYK or Grey scale format, otherwise, colour shifting may occur
-%% during the printing process.  Papers submitted with figures other than the optional teaser on the
-%% first page will be refused. Also, the teaser figure should only have the
-%% width of the abstract as the template enforces it.
-
-%% These few lines make a distinction between latex and pdflatex calls and they
-%% bring in essential packages for graphics and font handling.
-%% Note that due to the \DeclareGraphicsExtensions{} call it is no longer necessary
-%% to provide the the path and extension of a graphics file:
-%% \includegraphics{diamondrule} is completely sufficient.
-%%
-\ifpdf%                                % if we use pdflatex
-  \pdfoutput=1\relax                   % create PDFs from pdfLaTeX
-  \pdfcompresslevel=9                  % PDF Compression
-  \pdfoptionpdfminorversion=7          % create PDF 1.7
-  \ExecuteOptions{pdftex}
-  \usepackage{graphicx}                % allow us to embed graphics files
-  \DeclareGraphicsExtensions{.pdf,.png,.jpg,.jpeg} % for pdflatex we expect .pdf, .png, or .jpg files
-\else%                                 % else we use pure latex
-  \ExecuteOptions{dvips}
-  \usepackage{graphicx}                % allow us to embed graphics files
-  \DeclareGraphicsExtensions{.eps}     % for pure latex we expect eps files
-\fi%
-
-%% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
-\graphicspath{{gfx/}{./}} % where to search for the images
-
-\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
-\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
-\usepackage{textcomp}                  % use better special symbols
-\usepackage{mathptmx}                  % use matching math font
-\usepackage{times}                     % we use Times as the main font
-\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
-\usepackage{cite}                      % needed to automatically sort the references
-\usepackage{tabu}                      % only used for the table example
-\usepackage{booktabs}                  % only used for the table example
-\usepackage{makecell}
-\usepackage{multirow}
-
-%% We encourage the use of mathptmx for consistent usage of times font
-%% throughout the proceedings. However, if you encounter conflicts
-%% with other math-related packages, you may want to disable it.
-
-%% In preprint mode you may define your own headline.
-%\preprinttext{To appear in IEEE Transactions on Visualization and Computer Graphics.}
-
-%% If you are submitting a paper to a conference for review with a double
-%% blind reviewing process, please replace the value ``0'' below with your
-%% OnlineID. Otherwise, you may safely leave it at ``0''.
-\onlineid{0}
-
-%% declare the category of your paper, only shown in review mode
-\vgtccategory{Research}
-%% please declare the paper type of your paper to help reviewers, only shown in review mode
-%% choices:
-%% * algorithm/technique
-%% * application/design study
-%% * evaluation
-%% * system
-%% * theory/model
-\vgtcpapertype{theory/model}
-
-%% Paper title.
-\title{Computing Graphical Perception}
-
-%% This is how authors are specified in the journal style
-
-%% indicate IEEE Member or Student Member in form indicated below
-\author{Daniel Haehn, \textit{Member, IEEE}, James Tompkin, and Hanspeter Pfister}
-\authorfooter{
-%% insert punctuation at end of each item
-\item Daniel Haehn, and Hanspeter Pfister are with the Paulson School of Engineering and Applied Sciences at Harvard University. \\
-E-mail: \{haehn,pfister\}@seas.harvard.edu.
-%
-\item James Tompkin is with the Thomas J. Watson Sr. Center for Information Technology at Brown University. \\E-mail: james\_tompkin@brown.edu.
-}
-
-%other entries to be set up for journal
-\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Computing Graphical Perception}
-%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
-
-%% Abstract section.
-\abstract{TODO%
-} % end of abstract
-
-%% Keywords that describe your work. Will show as 'Index Terms' in journal
-%% please capitalize first letter and insert punctuation after last keyword
-\keywords{Machine Perception, Deep Learning}
-
-%% ACM Computing Classification System (CCS). 
-%% See <http://www.acm.org/class/1998/> for details.
-%% The ``\CCScat'' command takes four arguments.
-
-%\CCScatlist{ % not used in journal version
-% \CCScat{K.6.1}{Management of Computing and Information Systems}%
-%{Project and People Management}{Life Cycle};
-% \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
-%}
-
-%% Uncomment below to include a teaser figure.
-\teaser{
-  \centering
-  \includegraphics[width=\linewidth]{teaser.pdf}
-  \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking visual cortex inspired machine learning classifiers to assess the relationship between values encoded in pie charts and bar charts.
-Similar to the findings of Clevenland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs predict more accurately when working with bar charts (mean squared error, MSE in green).}
-	\label{fig:teaser}
-}
-
-%% Uncomment below to disable the manuscript note
-%\renewcommand{\manuscriptnotetxt}{}
-
-%% Copyright space is enabled by default as required by guidelines.
-%% It is disabled by the 'review' option or via the following command:
-% \nocopyrightspace
-
-\vgtcinsertpkg
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-%% The ``\maketitle'' command must be the first command after the
-%% ``\begin{document}'' command. It prepares and prints the title block.
-
-%% the only exception to this rule is the \firstsection command
-
-\input{00_introduction.tex}
-
-\input{01_related_work.tex}
-
-\section{Experimental Setup}
-
-The experiments shown in this paper are either supervised regression or classification tasks.
-We formulate any estimation of quantities (e.g.
-angles, positions, lengths etc.) as a regression problem between $0$ and $1$.
-The output indicates the percentage in regards to the degrees of freedom of the individual experiment.
-If the experiment involves a choice, we formulate it as a classification problem.
-
-\subsection{Measures}
-
-\textbf{Accuracy.} We use the same metric as Cleveland and McGill to measure accuracy.
-\begin{equation}
-	\log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
-\end{equation}
-\\~\\
-\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the confidence intervals.
-\\~\\
-\noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
-For regression tasks the loss is defined as mean squared error (MSE) and for classification tasks the loss is categorical cross-entropy.
-
-\subsection{Classifiers}
-
-Our classifiers are built upon a multilayer perceptron (MLP) which is a feedforward artificial neural network.
-We combine this MLP with different convolutional neural networks (CNNs) for preprocessing and feature generation.
-These include the traditional LeNet trained from scratch, as well as VGG19 and Xception trained using ImageNet.
-\\~\\
-\noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units~(Fig.~\ref{fig:classifiers}).
-We then add a dropout layer to prevent overfitting and compute linear regression or classification (softmax).
-\\~\\
-\noindent{\textbf{Convolutional Neural Networks.}} We use CNNs to generate additional features as input to the MLP.
-We train the \emph{LeNet} classifier with tune it specifically towards each visualization.
-For \emph{VGG19} and \emph{Xception}, we generate features using previously trained weights on ImageNet.
-\\~\\
-\noindent{\textbf{Optimization.}} All networks are optimized using stochastic gradient descent with Nesterov momentum using fixed parameters (Table \ref{tab:parameters}).
-We train for $1000$ epochs but stop early if the loss does not decrease for ten epochs.
-
-\begin{table}[t]
-\centering
-\caption{We use different feature generators as input to a multilayer perceptron which performs linear regression or the classification task.
-This yields different sets of trainable parameters.
-We also train the MLP directly on the visualizations without any additional feature generation.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lrl}
-%	\toprule
-%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
-%	\midrule
-%	MLP & $0$ & $2,560,513$ \\
-%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
-%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
-%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
-%	\bottomrule
-	\toprule
-	Classifier & \makecell{Trainable\\Parameters} & Optimization \\
-	\midrule
-	MLP & $2,560,513$ & SGD (Nesterov momentum)\\
-	\emph{LeNet} + MLP & $8,026,083$ & Learning rate: $0.0001$\\
-	\emph{VGG19} + MLP & $21,204,545$ & Momentum: $0.9$ \\
-	\emph{Xception} + MLP & $25,580,585$ & \makecell[tl]{Batchsize: 32\\Epochs: $1000$ (Early Stopping)}\\
-	\bottomrule
-\end{tabular}
-}
-\label{tab:parameters}
-\vspace{-4mm}
-\end{table}
-
-\begin{figure}[t]
-	\centering
-	  \includegraphics[width=.8\linewidth]{classifiers.pdf}
-  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
-We use Dropout regularization to prevent overfitting.
-We learn categorical and unordered dependent variables using the softmax function and perform linear regression for continuous variables.
-The MLP can learn the visualizations directly but we also learn features generated by LeNet (2 conv.
-layers, filter size $5\time5$), VGG19 trained on ImageNet (16 conv.
-layers, filter size $3\times3$), or Xception trained on ImageNet (36 conv.
-layers, filter size $3\times3$) to increase the number of trainable parameters.}
-	\label{fig:classifiers}
-\end{figure}
-
-\subsection{Data}
-
-We create all visualizations as parametrized rasterized images without interpolation.
-The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}.
-We add subtle random noise ($0.05$) to each pixel to introduce additional variation.
-\\~\\
-\noindent\textbf{Training/Validation/Test Splits.} We specify the size of each split set as follows: 60,000 training images, 20,000 validation images, and 20,000 test images.
-We then randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables.
-This eliminates leakage during training and evaluation.
-We also scale each set independently: images to the range of $-.5$ to $.5$ and labels to the range of $0.0$ to $1.0$.
-\\~\\
-\noindent\textbf{Cross-classifier variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.
-
-\section{Elementary Perceptual Tasks}
-
-Cleveland and McGill describe the mapping of graphical elements to quantitative variables as \emph{elementary perceptual tasks} and introduce a list of ten different encodings in their paper~\cite{cleveland_mcgill}.
-We create visualizations of these tasks as rasterized images (Fig.~\ref{fig:elementary_perceptual_tasks}).
-
-\subsection{Parametrizations}
-\label{sec:parametrizations}
-We generate multiple parameterizations for each elementary perceptual task (Fig.~\ref{fig:elementary_perceptual_tasks}) and sequentially increase the number of parameters.
-For instance, for \emph{position non-aligned scale} we first only vary the origin of the coordinate system which yields just 10 different parameters.
-We then include translation along the y-axis with a significant increase in variability.
-We then also add x-movement and a variable spot size.
-This results in more complex datasets depending on the variability setting.
-Table~\ref{tab:encoding_parameters} shows the different settings.
-It is important to consider this variability when evaluating different classifiers with individual trainable parameters (Table~\ref{tab:parameters}).
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure1_overview.pdf}
-  \caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of the elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
-We vary the parameters of each perceptual task and then assess the interpretability of feed-forward neural networks.}
-	\label{fig:elementary_perceptual_tasks}
-\end{figure}
-
-\begin{table}[t]
-\centering
-\caption{\textbf{Variability of Elementary Perceptual Tasks.} We sequentially increase the number of parameters for every visual encoding of the elementary perceptual tasks.
-This introduces variability and increasingly more complex datasets.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{llr}
-%	\toprule
-%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
-%	\midrule
-%	MLP & $0$ & $2,560,513$ \\
-%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
-%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
-%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
-%	\bottomrule
-	\toprule
-	Elementary Perceptual Task & Variability & Parameters\\
-	\midrule
-	\makecell[tl]{\emph{Position Common Scale}} & \makecell[tl]{Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $60$ \\ $3600$ \\ $21600$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Position Non-Aligned Scale}} & \makecell[tl]{Origin\\+ Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $10$ \\$600$ \\ $36000$ \\ $216000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Length}} & \makecell[tl]{Length\\+ Position Y \\+ Position X \\+ Width} & \makecell[tr]{ $60$ \\ $2400$ \\ $144000$\\$864000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Direction}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $360$ \\ $21600$ \\ $1296000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Angle}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $90$ \\ $5400$ \\ $324000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Area}} & \makecell[tl]{Radius\\+ Position Y \\+ Position X} & \makecell[tr]{ $40$ \\ $800$ \\ $16000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Volume}} & \makecell[tl]{Cube Sidelength\\+ Position Y \\+ Position X} & \makecell[tr]{ $20$ \\ $400$ \\ $8000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Curvature}} & \makecell[tl]{Midpoint Curvature\\+ Position Y \\+ Position X\\+ Width} & \makecell[tr]{ $80$ \\ $1600$ \\ $64000$ \\ $2560000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Shading}} & \makecell[tl]{Density\\+ Position Y \\+ Position X} & \makecell[tr]{ $100$ \\ $1600$ \\ $2000$ \\ $40000$}\\
-
-	\bottomrule
-\end{tabular}
-}
-\label{tab:encoding_parameters}
-\end{table}
-
-\subsection{Hypotheses}
-
-We proposed four hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H1.1} \textbf{Visual cortex inspired classifiers are able to connect graphical elements to their quantitative variables.} While much simpler models than their biological pendant, convolutional neural networks are heavily influenced by our biological knowledge of the visual system.
-Such classifiers therefor follow the same principles as human perception.
-	\item \textbf{H1.2} \textbf{Computed perceptual performance is dependent on classifier complexity.} We evaluate multiple classifiers with different numbers of trainable parameters.
-A more complex classifier (with higher number of parameters) will perform better on elementary perceptual tasks.
-	\item \textbf{H1.3} \textbf{Some visual encodings are better than others for computations.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
-We investigate whether this order is also relevant for computing graphical perception.
-	\item \textbf{H1.4} \textbf{Classifiers trained on perceptual tasks can generalize to more or less complex variations of the same task.} Recent research suggests that convolutional neural networks generalize extremely well.
-While the underlying reasons are mainly yet unknown, this property allows them to perform on variations of a similar perceptual task.
-\end{itemize}
-
-\section{Position-Angle Experiment}
-
-The position-angle experiment was originally performed by Cleveland and McGill to measure whether humans can better perceive quantities encoded as positions or as angles~\cite{cleveland_mcgill}.
-The actual experiment then compares pie charts versus bar charts since these map down to elementary position and angle judgement.
-We create rasterized images mimicking Cleveland and McGill's proposed encoding and investigate computational perception of our four classifiers.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_overview}
-  \caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
-The experimental task involves the judgement of different encoded values in comparison to the largest encoded values.
-The pie chart (left) and the bar chart (right) visualize the same data point.
-In their paper, Cleveland and McGill report less errors using bar charts.}
-	\label{fig:position_angle_experiment}
-\end{figure}
-
-\subsection{Hypotheses}
-
-We proposed four hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H2.1} \textbf{Computed perceptual performance is better using bar charts than pie charts.} Cleveland and McGill report that position judgements are almost twice as accurate as angle judgements.
-This renders bar charts superior to pie charts and should also be the case for convolutional neural networks.
-	\item \textbf{H2.2} \textbf{Classifiers can learn position faster than angles.} We assume that understanding bar charts is easier than understanding pie charts.
-We suspect that our classifiers learn encodings of positions faster than of angles resulting in more efficient training and faster convergence.
-\end{itemize}
-
 \section{Position-Length Experiment}
@@
 \end{figure}
-
-\section{Bars and Framed Rectangles Experiment}
-
-Visual cues can help converting graphical elements back to their real world variables.
-Cleveland and McGill introduced the bars and framed rectangles experiment which judges the elementary perceptual task of position along non-aligned scales~\cite{cleveland_mcgill}.
-
-\subsection{Hypotheses}
-
-We proposed two hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H4.1} \textbf{Classifiers can leverage additional visual cues.} The original bar and framed rectangle experiment shows how visual cues aid humans in mapping graphical elements to quantitative variables.
-This should be the same for feed-forward neural networks since they are inspired by the visual system.
-	\item \textbf{H4.2} \textbf{Weber's law can be transferred to computational perception.} Cleveland and McGill confirmed Weber's law based on the bar and framed rectangle experiment.
-For humans, the ability to perceive change within a distribution is proportional to the size of the initial distribution.
-\end{itemize}
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_overview}
-  \caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales.
-For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles (right).
-In this case, the right bar is heigher as visible with less free space when adding the frame.
-We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
-	\label{fig:bars_and_framed_rectangles_experiment}
-\end{figure}
-
-\subsection{Weber-Fechner's Law}
-
-As identified by Cleveland and McGill, the bar and framed rectangle experiment is closely related to Weber's law.
-This psychophysics law states that perceivable difference within a distribution is proportional to the initial size of the distribution.
-Weber's law goes hand-in-hand with Fechner's law.
-We conduct an additional experiment based on the original illustrations of the Weber-Fechner law to investigate whether this law can be applied to computational perception of our classifiers (Fig.~\ref{fig:webers_law}).
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{webers_law}
-  \caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution.
-The lower square contains 10 more dots than the upper one on both sides.
-However, the difference is easily perceivable on the left while the squares on the right almost look the same.
-We generate rasterized visualizations similar to this setup and evaluate our classifiers.}
-	\label{fig:webers_law}
-\end{figure}
-
-\section{Results and Discussion}
-
-\subsection{Elementary Perceptual Tasks}
-
-some are good and some are bad.. why?
-
-\begin{figure*}[h]
-	  \includegraphics[width=\linewidth]{figure1.pdf}
-  \caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure1_results}
-\end{figure*}
-
-\textbf{Computational Perception Ranking.}
-
-Cleveland McGills Ranking - can we observe something similar?
-
-\begin{enumerate}
-	\item Position along a common scale e.g. scatter plot
-	\item Position on identical but nonaligned scales e.g. multiple scatter plots
-	\item Length e.g. bar chart
-	\item Angle \& Slope (tie) e.g. pie chart
-	\item Area e.g. bubbles
-	\item Volume, density, and color saturation (tie) e.g. heatmap
-	\item Color hue e.g. newsmap
-\end{enumerate}
-
-\noindent{\textbf{Cross-classifier variability.}}
-
-Can a neural network generalize on simple perceptual tasks?
-
-\subsection{Position-Angle Experiment}
-
-Bar charts are more accurate (Fig.~\ref{fig:figure3_mlae}) and networks converge faster (Fig.~\ref{fig:figure3_val_loss}). This is great.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
-  \caption{\textbf{Classifier Efficiency of the Position-Angle experiment.} Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts.
-We report the MSE measure for both encodings of four different classifier on previously unseen validation data.}
-	\label{fig:figure3_val_loss}
-\end{figure}
-
-\begin{figure*}[t]
-	  \includegraphics[width=\linewidth]{figure3_mlae.pdf}
-  \caption{\textbf{Computational results of the Position-Angle experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure3_mlae}
-\end{figure*}
-
-\subsection{Position-Length Experiment}
-
-We have nothing yet here...
-
-\subsection{Bars and Framed Rectangles Experiment}
-
-First run indicates that framed rectangles perform better but we dont really know it yet.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_val_loss.png}
-  \caption{\textbf{Classifier Efficiency of the Bars and Framed Rectangles experiment.} Categorical Cross-Entropy loss for the \emph{bars and framed rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-The frame around the bars adds an additional visual cue enables faster network convergence.
-This is not yet reproducible!}
-	\label{fig:figure12_val_loss}
-\end{figure}
-
-\section{Conclusions}
-
-Future work: allow insights for infovis for machines
-
-%% if specified like this the section will be committed in review mode
-%\acknowledgments{
-%The authors wish to thank A, B, and C. This work was supported in part by
-%a grant from XYZ (\# 12345-67890).}
-
-%\bibliographystyle{abbrv}
-\bibliographystyle{abbrv-doi}
-%\bibliographystyle{abbrv-doi-narrow}
-%\bibliographystyle{abbrv-doi-hyperref}
-%\bibliographystyle{abbrv-doi-hyperref-narrow}
-
-\bibliography{paper.bib}
-\end{document}

### 2018-03-21 PAPER/00_introduction.tex
Update paper.

@@
 
+In 1962 Hubel and Wiesel were the first to begin studying the visual cortex from the standpoint of a neuroscientist.
+Their experimental findings on cats and macaque monkeys suggested a hierarchy of cells with increasing complexity which was then later transferred to the hierarchical model of different layers.
+Twenty years later, this insight was translated to the Neocognitron quantitative model, by Fukushima and Miyake, which ultimately led to the important work of Hinton, Bengio, and LeCun in the 1980s.
+Their work on stochastic gradient descent approximation, and the availability of faster computer hardware then led to today’s breakthrough of deep learning networks.
+In the last decade, this field has exhibited rapid growth, constant evolution, and new applications in various domains.
+
 GOALS

### 2018-03-21 PAPER/02_experimental_setup.tex

--- 2018-03-16 PAPER/paper.tex
@@
-\documentclass[journal]{vgtc}                % final (journal style)
-%\documentclass[review,journal]{vgtc}         % review (journal style)
-%\documentclass[widereview]{vgtc}             % wide-spaced review
-%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
-
-%% Uncomment one of the lines above depending on where your paper is
-%% in the conference process. ``review'' and ``widereview'' are for review
-%% submission, ``preprint'' is for pre-publication, and the final version
-%% doesn't use a specific qualifier.
-
-%% Please use one of the ``review'' options in combination with the
-%% assigned online id (see below) ONLY if your paper uses a double blind
-%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
-%% in the past.
-
-%% Please note that the use of figures other than the optional teaser is not permitted on the first page
-%% of the journal version.  Figures should begin on the second page and be
-%% in CMYK or Grey scale format, otherwise, colour shifting may occur
-%% during the printing process.  Papers submitted with figures other than the optional teaser on the
-%% first page will be refused. Also, the teaser figure should only have the
-%% width of the abstract as the template enforces it.
-
-%% These few lines make a distinction between latex and pdflatex calls and they
-%% bring in essential packages for graphics and font handling.
-%% Note that due to the \DeclareGraphicsExtensions{} call it is no longer necessary
-%% to provide the the path and extension of a graphics file:
-%% \includegraphics{diamondrule} is completely sufficient.
-%%
-\ifpdf%                                % if we use pdflatex
-  \pdfoutput=1\relax                   % create PDFs from pdfLaTeX
-  \pdfcompresslevel=9                  % PDF Compression
-  \pdfoptionpdfminorversion=7          % create PDF 1.7
-  \ExecuteOptions{pdftex}
-  \usepackage{graphicx}                % allow us to embed graphics files
-  \DeclareGraphicsExtensions{.pdf,.png,.jpg,.jpeg} % for pdflatex we expect .pdf, .png, or .jpg files
-\else%                                 % else we use pure latex
-  \ExecuteOptions{dvips}
-  \usepackage{graphicx}                % allow us to embed graphics files
-  \DeclareGraphicsExtensions{.eps}     % for pure latex we expect eps files
-\fi%
-
-%% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
-\graphicspath{{gfx/}{./}} % where to search for the images
-
-\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
-\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
-\usepackage{textcomp}                  % use better special symbols
-\usepackage{mathptmx}                  % use matching math font
-\usepackage{times}                     % we use Times as the main font
-\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
-\usepackage{cite}                      % needed to automatically sort the references
-\usepackage{tabu}                      % only used for the table example
-\usepackage{booktabs}                  % only used for the table example
-\usepackage{makecell}
-\usepackage{multirow}
-
-%% We encourage the use of mathptmx for consistent usage of times font
-%% throughout the proceedings. However, if you encounter conflicts
-%% with other math-related packages, you may want to disable it.
-
-%% In preprint mode you may define your own headline.
-%\preprinttext{To appear in IEEE Transactions on Visualization and Computer Graphics.}
-
-%% If you are submitting a paper to a conference for review with a double
-%% blind reviewing process, please replace the value ``0'' below with your
-%% OnlineID. Otherwise, you may safely leave it at ``0''.
-\onlineid{0}
-
-%% declare the category of your paper, only shown in review mode
-\vgtccategory{Research}
-%% please declare the paper type of your paper to help reviewers, only shown in review mode
-%% choices:
-%% * algorithm/technique
-%% * application/design study
-%% * evaluation
-%% * system
-%% * theory/model
-\vgtcpapertype{theory/model}
-
-%% Paper title.
-\title{Computing Graphical Perception}
-
-%% This is how authors are specified in the journal style
-
-%% indicate IEEE Member or Student Member in form indicated below
-\author{Daniel Haehn, \textit{Member, IEEE}, James Tompkin, and Hanspeter Pfister}
-\authorfooter{
-%% insert punctuation at end of each item
-\item Daniel Haehn, and Hanspeter Pfister are with the Paulson School of Engineering and Applied Sciences at Harvard University. \\
-E-mail: \{haehn,pfister\}@seas.harvard.edu.
-%
-\item James Tompkin is with the Thomas J. Watson Sr. Center for Information Technology at Brown University. \\E-mail: james\_tompkin@brown.edu.
-}
-
-%other entries to be set up for journal
-\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Computing Graphical Perception}
-%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
-
-%% Abstract section.
-\abstract{TODO%
-} % end of abstract
-
-%% Keywords that describe your work. Will show as 'Index Terms' in journal
-%% please capitalize first letter and insert punctuation after last keyword
-\keywords{Machine Perception, Deep Learning}
-
-%% ACM Computing Classification System (CCS). 
-%% See <http://www.acm.org/class/1998/> for details.
-%% The ``\CCScat'' command takes four arguments.
-
-%\CCScatlist{ % not used in journal version
-% \CCScat{K.6.1}{Management of Computing and Information Systems}%
-%{Project and People Management}{Life Cycle};
-% \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
-%}
-
-%% Uncomment below to include a teaser figure.
-\teaser{
-  \centering
-  \includegraphics[width=\linewidth]{teaser.pdf}
-  \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking visual cortex inspired machine learning classifiers to assess the relationship between values encoded in pie charts and bar charts.
-Similar to the findings of Clevenland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs predict more accurately when working with bar charts (mean squared error, MSE in green).}
-	\label{fig:teaser}
-}
-
-%% Uncomment below to disable the manuscript note
-%\renewcommand{\manuscriptnotetxt}{}
-
-%% Copyright space is enabled by default as required by guidelines.
-%% It is disabled by the 'review' option or via the following command:
-% \nocopyrightspace
-
-\vgtcinsertpkg
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-%% The ``\maketitle'' command must be the first command after the
-%% ``\begin{document}'' command. It prepares and prints the title block.
-
-%% the only exception to this rule is the \firstsection command
-
-\input{00_introduction.tex}
-
-\input{01_related_work.tex}
-
 \section{Experimental Setup}
@@
 \noindent\textbf{Cross-classifier variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.
-
-\section{Elementary Perceptual Tasks}
-
-Cleveland and McGill describe the mapping of graphical elements to quantitative variables as \emph{elementary perceptual tasks} and introduce a list of ten different encodings in their paper~\cite{cleveland_mcgill}.
-We create visualizations of these tasks as rasterized images (Fig.~\ref{fig:elementary_perceptual_tasks}).
-
-\subsection{Parametrizations}
-\label{sec:parametrizations}
-We generate multiple parameterizations for each elementary perceptual task (Fig.~\ref{fig:elementary_perceptual_tasks}) and sequentially increase the number of parameters.
-For instance, for \emph{position non-aligned scale} we first only vary the origin of the coordinate system which yields just 10 different parameters.
-We then include translation along the y-axis with a significant increase in variability.
-We then also add x-movement and a variable spot size.
-This results in more complex datasets depending on the variability setting.
-Table~\ref{tab:encoding_parameters} shows the different settings.
-It is important to consider this variability when evaluating different classifiers with individual trainable parameters (Table~\ref{tab:parameters}).
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure1_overview.pdf}
-  \caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of the elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
-We vary the parameters of each perceptual task and then assess the interpretability of feed-forward neural networks.}
-	\label{fig:elementary_perceptual_tasks}
-\end{figure}
-
-\begin{table}[t]
-\centering
-\caption{\textbf{Variability of Elementary Perceptual Tasks.} We sequentially increase the number of parameters for every visual encoding of the elementary perceptual tasks.
-This introduces variability and increasingly more complex datasets.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{llr}
-%	\toprule
-%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
-%	\midrule
-%	MLP & $0$ & $2,560,513$ \\
-%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
-%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
-%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
-%	\bottomrule
-	\toprule
-	Elementary Perceptual Task & Variability & Parameters\\
-	\midrule
-	\makecell[tl]{\emph{Position Common Scale}} & \makecell[tl]{Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $60$ \\ $3600$ \\ $21600$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Position Non-Aligned Scale}} & \makecell[tl]{Origin\\+ Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $10$ \\$600$ \\ $36000$ \\ $216000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Length}} & \makecell[tl]{Length\\+ Position Y \\+ Position X \\+ Width} & \makecell[tr]{ $60$ \\ $2400$ \\ $144000$\\$864000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Direction}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $360$ \\ $21600$ \\ $1296000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Angle}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $90$ \\ $5400$ \\ $324000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Area}} & \makecell[tl]{Radius\\+ Position Y \\+ Position X} & \makecell[tr]{ $40$ \\ $800$ \\ $16000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Volume}} & \makecell[tl]{Cube Sidelength\\+ Position Y \\+ Position X} & \makecell[tr]{ $20$ \\ $400$ \\ $8000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Curvature}} & \makecell[tl]{Midpoint Curvature\\+ Position Y \\+ Position X\\+ Width} & \makecell[tr]{ $80$ \\ $1600$ \\ $64000$ \\ $2560000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Shading}} & \makecell[tl]{Density\\+ Position Y \\+ Position X} & \makecell[tr]{ $100$ \\ $1600$ \\ $2000$ \\ $40000$}\\
-
-	\bottomrule
-\end{tabular}
-}
-\label{tab:encoding_parameters}
-\end{table}
-
-\subsection{Hypotheses}
-
-We proposed four hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H1.1} \textbf{Visual cortex inspired classifiers are able to connect graphical elements to their quantitative variables.} While much simpler models than their biological pendant, convolutional neural networks are heavily influenced by our biological knowledge of the visual system.
-Such classifiers therefor follow the same principles as human perception.
-	\item \textbf{H1.2} \textbf{Computed perceptual performance is dependent on classifier complexity.} We evaluate multiple classifiers with different numbers of trainable parameters.
-A more complex classifier (with higher number of parameters) will perform better on elementary perceptual tasks.
-	\item \textbf{H1.3} \textbf{Some visual encodings are better than others for computations.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
-We investigate whether this order is also relevant for computing graphical perception.
-	\item \textbf{H1.4} \textbf{Classifiers trained on perceptual tasks can generalize to more or less complex variations of the same task.} Recent research suggests that convolutional neural networks generalize extremely well.
-While the underlying reasons are mainly yet unknown, this property allows them to perform on variations of a similar perceptual task.
-\end{itemize}
-
-\section{Position-Angle Experiment}
-
-The position-angle experiment was originally performed by Cleveland and McGill to measure whether humans can better perceive quantities encoded as positions or as angles~\cite{cleveland_mcgill}.
-The actual experiment then compares pie charts versus bar charts since these map down to elementary position and angle judgement.
-We create rasterized images mimicking Cleveland and McGill's proposed encoding and investigate computational perception of our four classifiers.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_overview}
-  \caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
-The experimental task involves the judgement of different encoded values in comparison to the largest encoded values.
-The pie chart (left) and the bar chart (right) visualize the same data point.
-In their paper, Cleveland and McGill report less errors using bar charts.}
-	\label{fig:position_angle_experiment}
-\end{figure}
-
-\subsection{Hypotheses}
-
-We proposed four hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H2.1} \textbf{Computed perceptual performance is better using bar charts than pie charts.} Cleveland and McGill report that position judgements are almost twice as accurate as angle judgements.
-This renders bar charts superior to pie charts and should also be the case for convolutional neural networks.
-	\item \textbf{H2.2} \textbf{Classifiers can learn position faster than angles.} We assume that understanding bar charts is easier than understanding pie charts.
-We suspect that our classifiers learn encodings of positions faster than of angles resulting in more efficient training and faster convergence.
-\end{itemize}
-
-\section{Position-Length Experiment}
-
-This is the one where we estimate two selected bars compared to the longest one - very similar to the previous one but, not yet done.
-We basically test divided versus grouped barchart and we estimate one relation between two marked quanitities: what percent the smaller is of the larger.
-
-There are five types: type 1-3 this is a position judgement along a common scale.
-(btw all classifiers seem to do that extremely well in the elementary tasks so we assume this will work well here too).
-Types 4-5 are length judgements and we know that the classifiers struggle with that quite a bit.
-
-The setup from Cleveland McGill is first a classification task: which one is smaller? and then a regression task: how much smaller. So we have to see how to encode this.
-
-\subsection{Hypotheses}
-
-We proposed two hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H3.1} \textbf{Grouped bar charts are better computational perceivable than divided bar charts.} A grouped bar chart involves judging a position while a divided bar chart most likely (if not the bottom is looked at) requires length judgements.
-Classifiers are better at judging position than at judging length so grouped bar charts are easier to grasp in terms of computational perception.
-	\item \textbf{H3.2} \textbf{not yet} Any ideas?
-\end{itemize}
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{position_length_original1.png}
-	  \includegraphics[width=\linewidth]{position_length_original2.png}
-
-  \caption{\textbf{Position-Length Experiment.} (Not yet) Rasterized versions of the graphs of Cleveland and McGill's position-length experiment.
-The perceptual task involves comparing.
-the two dot-marked quantities across five different visual encodings of either grouped or divided bar charts.
-We evaluate which type of bar chart performs better with our neural networks as a combined classification and regression problem.
-The first task is to select which of the marked quantities is smaller (classification) and the second task is to specify how much smaller it is (regression).}
-	\label{fig:position_length_experiment}
-\end{figure}
-
-\section{Bars and Framed Rectangles Experiment}
-
-Visual cues can help converting graphical elements back to their real world variables.
-Cleveland and McGill introduced the bars and framed rectangles experiment which judges the elementary perceptual task of position along non-aligned scales~\cite{cleveland_mcgill}.
-
-\subsection{Hypotheses}
-
-We proposed two hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H4.1} \textbf{Classifiers can leverage additional visual cues.} The original bar and framed rectangle experiment shows how visual cues aid humans in mapping graphical elements to quantitative variables.
-This should be the same for feed-forward neural networks since they are inspired by the visual system.
-	\item \textbf{H4.2} \textbf{Weber's law can be transferred to computational perception.} Cleveland and McGill confirmed Weber's law based on the bar and framed rectangle experiment.
-For humans, the ability to perceive change within a distribution is proportional to the size of the initial distribution.
-\end{itemize}
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_overview}
-  \caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales.
-For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles (right).
-In this case, the right bar is heigher as visible with less free space when adding the frame.
-We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
-	\label{fig:bars_and_framed_rectangles_experiment}
-\end{figure}
-
-\subsection{Weber-Fechner's Law}
-
-As identified by Cleveland and McGill, the bar and framed rectangle experiment is closely related to Weber's law.
-This psychophysics law states that perceivable difference within a distribution is proportional to the initial size of the distribution.
-Weber's law goes hand-in-hand with Fechner's law.
-We conduct an additional experiment based on the original illustrations of the Weber-Fechner law to investigate whether this law can be applied to computational perception of our classifiers (Fig.~\ref{fig:webers_law}).
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{webers_law}
-  \caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution.
-The lower square contains 10 more dots than the upper one on both sides.
-However, the difference is easily perceivable on the left while the squares on the right almost look the same.
-We generate rasterized visualizations similar to this setup and evaluate our classifiers.}
-	\label{fig:webers_law}
-\end{figure}
-
-\section{Results and Discussion}
-
-\subsection{Elementary Perceptual Tasks}
-
-some are good and some are bad.. why?
-
-\begin{figure*}[h]
-	  \includegraphics[width=\linewidth]{figure1.pdf}
-  \caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure1_results}
-\end{figure*}
-
-\textbf{Computational Perception Ranking.}
-
-Cleveland McGills Ranking - can we observe something similar?
-
-\begin{enumerate}
-	\item Position along a common scale e.g. scatter plot
-	\item Position on identical but nonaligned scales e.g. multiple scatter plots
-	\item Length e.g. bar chart
-	\item Angle \& Slope (tie) e.g. pie chart
-	\item Area e.g. bubbles
-	\item Volume, density, and color saturation (tie) e.g. heatmap
-	\item Color hue e.g. newsmap
-\end{enumerate}
-
-\noindent{\textbf{Cross-classifier variability.}}
-
-Can a neural network generalize on simple perceptual tasks?
-
-\subsection{Position-Angle Experiment}
-
-Bar charts are more accurate (Fig.~\ref{fig:figure3_mlae}) and networks converge faster (Fig.~\ref{fig:figure3_val_loss}). This is great.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
-  \caption{\textbf{Classifier Efficiency of the Position-Angle experiment.} Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts.
-We report the MSE measure for both encodings of four different classifier on previously unseen validation data.}
-	\label{fig:figure3_val_loss}
-\end{figure}
-
-\begin{figure*}[t]
-	  \includegraphics[width=\linewidth]{figure3_mlae.pdf}
-  \caption{\textbf{Computational results of the Position-Angle experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure3_mlae}
-\end{figure*}
-
-\subsection{Position-Length Experiment}
-
-We have nothing yet here...
-
-\subsection{Bars and Framed Rectangles Experiment}
-
-First run indicates that framed rectangles perform better but we dont really know it yet.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_val_loss.png}
-  \caption{\textbf{Classifier Efficiency of the Bars and Framed Rectangles experiment.} Categorical Cross-Entropy loss for the \emph{bars and framed rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-The frame around the bars adds an additional visual cue enables faster network convergence.
-This is not yet reproducible!}
-	\label{fig:figure12_val_loss}
-\end{figure}
-
-\section{Conclusions}
-
-Future work: allow insights for infovis for machines
-
-%% if specified like this the section will be committed in review mode
-%\acknowledgments{
-%The authors wish to thank A, B, and C. This work was supported in part by
-%a grant from XYZ (\# 12345-67890).}
-
-%\bibliographystyle{abbrv}
-\bibliographystyle{abbrv-doi}
-%\bibliographystyle{abbrv-doi-narrow}
-%\bibliographystyle{abbrv-doi-hyperref}
-%\bibliographystyle{abbrv-doi-hyperref-narrow}
-
-\bibliography{paper.bib}
-\end{document}

### 2018-03-21 PAPER/04_position_angle_experiment.tex

--- 2018-03-16 PAPER/paper.tex
@@
-\documentclass[journal]{vgtc}                % final (journal style)
-%\documentclass[review,journal]{vgtc}         % review (journal style)
-%\documentclass[widereview]{vgtc}             % wide-spaced review
-%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
-
-%% Uncomment one of the lines above depending on where your paper is
-%% in the conference process. ``review'' and ``widereview'' are for review
-%% submission, ``preprint'' is for pre-publication, and the final version
-%% doesn't use a specific qualifier.
-
-%% Please use one of the ``review'' options in combination with the
-%% assigned online id (see below) ONLY if your paper uses a double blind
-%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
-%% in the past.
-
-%% Please note that the use of figures other than the optional teaser is not permitted on the first page
-%% of the journal version.  Figures should begin on the second page and be
-%% in CMYK or Grey scale format, otherwise, colour shifting may occur
-%% during the printing process.  Papers submitted with figures other than the optional teaser on the
-%% first page will be refused. Also, the teaser figure should only have the
-%% width of the abstract as the template enforces it.
-
-%% These few lines make a distinction between latex and pdflatex calls and they
-%% bring in essential packages for graphics and font handling.
-%% Note that due to the \DeclareGraphicsExtensions{} call it is no longer necessary
-%% to provide the the path and extension of a graphics file:
-%% \includegraphics{diamondrule} is completely sufficient.
-%%
-\ifpdf%                                % if we use pdflatex
-  \pdfoutput=1\relax                   % create PDFs from pdfLaTeX
-  \pdfcompresslevel=9                  % PDF Compression
-  \pdfoptionpdfminorversion=7          % create PDF 1.7
-  \ExecuteOptions{pdftex}
-  \usepackage{graphicx}                % allow us to embed graphics files
-  \DeclareGraphicsExtensions{.pdf,.png,.jpg,.jpeg} % for pdflatex we expect .pdf, .png, or .jpg files
-\else%                                 % else we use pure latex
-  \ExecuteOptions{dvips}
-  \usepackage{graphicx}                % allow us to embed graphics files
-  \DeclareGraphicsExtensions{.eps}     % for pure latex we expect eps files
-\fi%
-
-%% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
-\graphicspath{{gfx/}{./}} % where to search for the images
-
-\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
-\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
-\usepackage{textcomp}                  % use better special symbols
-\usepackage{mathptmx}                  % use matching math font
-\usepackage{times}                     % we use Times as the main font
-\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
-\usepackage{cite}                      % needed to automatically sort the references
-\usepackage{tabu}                      % only used for the table example
-\usepackage{booktabs}                  % only used for the table example
-\usepackage{makecell}
-\usepackage{multirow}
-
-%% We encourage the use of mathptmx for consistent usage of times font
-%% throughout the proceedings. However, if you encounter conflicts
-%% with other math-related packages, you may want to disable it.
-
-%% In preprint mode you may define your own headline.
-%\preprinttext{To appear in IEEE Transactions on Visualization and Computer Graphics.}
-
-%% If you are submitting a paper to a conference for review with a double
-%% blind reviewing process, please replace the value ``0'' below with your
-%% OnlineID. Otherwise, you may safely leave it at ``0''.
-\onlineid{0}
-
-%% declare the category of your paper, only shown in review mode
-\vgtccategory{Research}
-%% please declare the paper type of your paper to help reviewers, only shown in review mode
-%% choices:
-%% * algorithm/technique
-%% * application/design study
-%% * evaluation
-%% * system
-%% * theory/model
-\vgtcpapertype{theory/model}
-
-%% Paper title.
-\title{Computing Graphical Perception}
-
-%% This is how authors are specified in the journal style
-
-%% indicate IEEE Member or Student Member in form indicated below
-\author{Daniel Haehn, \textit{Member, IEEE}, James Tompkin, and Hanspeter Pfister}
-\authorfooter{
-%% insert punctuation at end of each item
-\item Daniel Haehn, and Hanspeter Pfister are with the Paulson School of Engineering and Applied Sciences at Harvard University. \\
-E-mail: \{haehn,pfister\}@seas.harvard.edu.
-%
-\item James Tompkin is with the Thomas J. Watson Sr. Center for Information Technology at Brown University. \\E-mail: james\_tompkin@brown.edu.
-}
-
-%other entries to be set up for journal
-\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Computing Graphical Perception}
-%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
-
-%% Abstract section.
-\abstract{TODO%
-} % end of abstract
-
-%% Keywords that describe your work. Will show as 'Index Terms' in journal
-%% please capitalize first letter and insert punctuation after last keyword
-\keywords{Machine Perception, Deep Learning}
-
-%% ACM Computing Classification System (CCS). 
-%% See <http://www.acm.org/class/1998/> for details.
-%% The ``\CCScat'' command takes four arguments.
-
-%\CCScatlist{ % not used in journal version
-% \CCScat{K.6.1}{Management of Computing and Information Systems}%
-%{Project and People Management}{Life Cycle};
-% \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
-%}
-
-%% Uncomment below to include a teaser figure.
-\teaser{
-  \centering
-  \includegraphics[width=\linewidth]{teaser.pdf}
-  \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking visual cortex inspired machine learning classifiers to assess the relationship between values encoded in pie charts and bar charts.
-Similar to the findings of Clevenland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs predict more accurately when working with bar charts (mean squared error, MSE in green).}
-	\label{fig:teaser}
-}
-
-%% Uncomment below to disable the manuscript note
-%\renewcommand{\manuscriptnotetxt}{}
-
-%% Copyright space is enabled by default as required by guidelines.
-%% It is disabled by the 'review' option or via the following command:
-% \nocopyrightspace
-
-\vgtcinsertpkg
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-%% The ``\maketitle'' command must be the first command after the
-%% ``\begin{document}'' command. It prepares and prints the title block.
-
-%% the only exception to this rule is the \firstsection command
-
-\input{00_introduction.tex}
-
-\input{01_related_work.tex}
-
-\section{Experimental Setup}
-
-The experiments shown in this paper are either supervised regression or classification tasks.
-We formulate any estimation of quantities (e.g.
-angles, positions, lengths etc.) as a regression problem between $0$ and $1$.
-The output indicates the percentage in regards to the degrees of freedom of the individual experiment.
-If the experiment involves a choice, we formulate it as a classification problem.
-
-\subsection{Measures}
-
-\textbf{Accuracy.} We use the same metric as Cleveland and McGill to measure accuracy.
-\begin{equation}
-	\log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
-\end{equation}
-\\~\\
-\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the confidence intervals.
-\\~\\
-\noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
-For regression tasks the loss is defined as mean squared error (MSE) and for classification tasks the loss is categorical cross-entropy.
-
-\subsection{Classifiers}
-
-Our classifiers are built upon a multilayer perceptron (MLP) which is a feedforward artificial neural network.
-We combine this MLP with different convolutional neural networks (CNNs) for preprocessing and feature generation.
-These include the traditional LeNet trained from scratch, as well as VGG19 and Xception trained using ImageNet.
-\\~\\
-\noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units~(Fig.~\ref{fig:classifiers}).
-We then add a dropout layer to prevent overfitting and compute linear regression or classification (softmax).
-\\~\\
-\noindent{\textbf{Convolutional Neural Networks.}} We use CNNs to generate additional features as input to the MLP.
-We train the \emph{LeNet} classifier with tune it specifically towards each visualization.
-For \emph{VGG19} and \emph{Xception}, we generate features using previously trained weights on ImageNet.
-\\~\\
-\noindent{\textbf{Optimization.}} All networks are optimized using stochastic gradient descent with Nesterov momentum using fixed parameters (Table \ref{tab:parameters}).
-We train for $1000$ epochs but stop early if the loss does not decrease for ten epochs.
-
-\begin{table}[t]
-\centering
-\caption{We use different feature generators as input to a multilayer perceptron which performs linear regression or the classification task.
-This yields different sets of trainable parameters.
-We also train the MLP directly on the visualizations without any additional feature generation.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lrl}
-%	\toprule
-%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
-%	\midrule
-%	MLP & $0$ & $2,560,513$ \\
-%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
-%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
-%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
-%	\bottomrule
-	\toprule
-	Classifier & \makecell{Trainable\\Parameters} & Optimization \\
-	\midrule
-	MLP & $2,560,513$ & SGD (Nesterov momentum)\\
-	\emph{LeNet} + MLP & $8,026,083$ & Learning rate: $0.0001$\\
-	\emph{VGG19} + MLP & $21,204,545$ & Momentum: $0.9$ \\
-	\emph{Xception} + MLP & $25,580,585$ & \makecell[tl]{Batchsize: 32\\Epochs: $1000$ (Early Stopping)}\\
-	\bottomrule
-\end{tabular}
-}
-\label{tab:parameters}
-\vspace{-4mm}
-\end{table}
-
-\begin{figure}[t]
-	\centering
-	  \includegraphics[width=.8\linewidth]{classifiers.pdf}
-  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
-We use Dropout regularization to prevent overfitting.
-We learn categorical and unordered dependent variables using the softmax function and perform linear regression for continuous variables.
-The MLP can learn the visualizations directly but we also learn features generated by LeNet (2 conv.
-layers, filter size $5\time5$), VGG19 trained on ImageNet (16 conv.
-layers, filter size $3\times3$), or Xception trained on ImageNet (36 conv.
-layers, filter size $3\times3$) to increase the number of trainable parameters.}
-	\label{fig:classifiers}
-\end{figure}
-
-\subsection{Data}
-
-We create all visualizations as parametrized rasterized images without interpolation.
-The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}.
-We add subtle random noise ($0.05$) to each pixel to introduce additional variation.
-\\~\\
-\noindent\textbf{Training/Validation/Test Splits.} We specify the size of each split set as follows: 60,000 training images, 20,000 validation images, and 20,000 test images.
-We then randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables.
-This eliminates leakage during training and evaluation.
-We also scale each set independently: images to the range of $-.5$ to $.5$ and labels to the range of $0.0$ to $1.0$.
-\\~\\
-\noindent\textbf{Cross-classifier variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.
-
-\section{Elementary Perceptual Tasks}
-
-Cleveland and McGill describe the mapping of graphical elements to quantitative variables as \emph{elementary perceptual tasks} and introduce a list of ten different encodings in their paper~\cite{cleveland_mcgill}.
-We create visualizations of these tasks as rasterized images (Fig.~\ref{fig:elementary_perceptual_tasks}).
-
-\subsection{Parametrizations}
-\label{sec:parametrizations}
-We generate multiple parameterizations for each elementary perceptual task (Fig.~\ref{fig:elementary_perceptual_tasks}) and sequentially increase the number of parameters.
-For instance, for \emph{position non-aligned scale} we first only vary the origin of the coordinate system which yields just 10 different parameters.
-We then include translation along the y-axis with a significant increase in variability.
-We then also add x-movement and a variable spot size.
-This results in more complex datasets depending on the variability setting.
-Table~\ref{tab:encoding_parameters} shows the different settings.
-It is important to consider this variability when evaluating different classifiers with individual trainable parameters (Table~\ref{tab:parameters}).
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure1_overview.pdf}
-  \caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of the elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
-We vary the parameters of each perceptual task and then assess the interpretability of feed-forward neural networks.}
-	\label{fig:elementary_perceptual_tasks}
-\end{figure}
-
-\begin{table}[t]
-\centering
-\caption{\textbf{Variability of Elementary Perceptual Tasks.} We sequentially increase the number of parameters for every visual encoding of the elementary perceptual tasks.
-This introduces variability and increasingly more complex datasets.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{llr}
-%	\toprule
-%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
-%	\midrule
-%	MLP & $0$ & $2,560,513$ \\
-%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
-%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
-%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
-%	\bottomrule
-	\toprule
-	Elementary Perceptual Task & Variability & Parameters\\
-	\midrule
-	\makecell[tl]{\emph{Position Common Scale}} & \makecell[tl]{Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $60$ \\ $3600$ \\ $21600$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Position Non-Aligned Scale}} & \makecell[tl]{Origin\\+ Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $10$ \\$600$ \\ $36000$ \\ $216000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Length}} & \makecell[tl]{Length\\+ Position Y \\+ Position X \\+ Width} & \makecell[tr]{ $60$ \\ $2400$ \\ $144000$\\$864000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Direction}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $360$ \\ $21600$ \\ $1296000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Angle}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $90$ \\ $5400$ \\ $324000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Area}} & \makecell[tl]{Radius\\+ Position Y \\+ Position X} & \makecell[tr]{ $40$ \\ $800$ \\ $16000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Volume}} & \makecell[tl]{Cube Sidelength\\+ Position Y \\+ Position X} & \makecell[tr]{ $20$ \\ $400$ \\ $8000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Curvature}} & \makecell[tl]{Midpoint Curvature\\+ Position Y \\+ Position X\\+ Width} & \makecell[tr]{ $80$ \\ $1600$ \\ $64000$ \\ $2560000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Shading}} & \makecell[tl]{Density\\+ Position Y \\+ Position X} & \makecell[tr]{ $100$ \\ $1600$ \\ $2000$ \\ $40000$}\\
-
-	\bottomrule
-\end{tabular}
-}
-\label{tab:encoding_parameters}
-\end{table}
-
-\subsection{Hypotheses}
-
-We proposed four hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H1.1} \textbf{Visual cortex inspired classifiers are able to connect graphical elements to their quantitative variables.} While much simpler models than their biological pendant, convolutional neural networks are heavily influenced by our biological knowledge of the visual system.
-Such classifiers therefor follow the same principles as human perception.
-	\item \textbf{H1.2} \textbf{Computed perceptual performance is dependent on classifier complexity.} We evaluate multiple classifiers with different numbers of trainable parameters.
-A more complex classifier (with higher number of parameters) will perform better on elementary perceptual tasks.
-	\item \textbf{H1.3} \textbf{Some visual encodings are better than others for computations.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
-We investigate whether this order is also relevant for computing graphical perception.
-	\item \textbf{H1.4} \textbf{Classifiers trained on perceptual tasks can generalize to more or less complex variations of the same task.} Recent research suggests that convolutional neural networks generalize extremely well.
-While the underlying reasons are mainly yet unknown, this property allows them to perform on variations of a similar perceptual task.
-\end{itemize}
-
 \section{Position-Angle Experiment}
@@
 \end{itemize}
-
-\section{Position-Length Experiment}
-
-This is the one where we estimate two selected bars compared to the longest one - very similar to the previous one but, not yet done.
-We basically test divided versus grouped barchart and we estimate one relation between two marked quanitities: what percent the smaller is of the larger.
-
-There are five types: type 1-3 this is a position judgement along a common scale.
-(btw all classifiers seem to do that extremely well in the elementary tasks so we assume this will work well here too).
-Types 4-5 are length judgements and we know that the classifiers struggle with that quite a bit.
-
-The setup from Cleveland McGill is first a classification task: which one is smaller? and then a regression task: how much smaller. So we have to see how to encode this.
-
-\subsection{Hypotheses}
-
-We proposed two hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H3.1} \textbf{Grouped bar charts are better computational perceivable than divided bar charts.} A grouped bar chart involves judging a position while a divided bar chart most likely (if not the bottom is looked at) requires length judgements.
-Classifiers are better at judging position than at judging length so grouped bar charts are easier to grasp in terms of computational perception.
-	\item \textbf{H3.2} \textbf{not yet} Any ideas?
-\end{itemize}
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{position_length_original1.png}
-	  \includegraphics[width=\linewidth]{position_length_original2.png}
-
-  \caption{\textbf{Position-Length Experiment.} (Not yet) Rasterized versions of the graphs of Cleveland and McGill's position-length experiment.
-The perceptual task involves comparing.
-the two dot-marked quantities across five different visual encodings of either grouped or divided bar charts.
-We evaluate which type of bar chart performs better with our neural networks as a combined classification and regression problem.
-The first task is to select which of the marked quantities is smaller (classification) and the second task is to specify how much smaller it is (regression).}
-	\label{fig:position_length_experiment}
-\end{figure}
-
-\section{Bars and Framed Rectangles Experiment}
-
-Visual cues can help converting graphical elements back to their real world variables.
-Cleveland and McGill introduced the bars and framed rectangles experiment which judges the elementary perceptual task of position along non-aligned scales~\cite{cleveland_mcgill}.
-
-\subsection{Hypotheses}
-
-We proposed two hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H4.1} \textbf{Classifiers can leverage additional visual cues.} The original bar and framed rectangle experiment shows how visual cues aid humans in mapping graphical elements to quantitative variables.
-This should be the same for feed-forward neural networks since they are inspired by the visual system.
-	\item \textbf{H4.2} \textbf{Weber's law can be transferred to computational perception.} Cleveland and McGill confirmed Weber's law based on the bar and framed rectangle experiment.
-For humans, the ability to perceive change within a distribution is proportional to the size of the initial distribution.
-\end{itemize}
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_overview}
-  \caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales.
-For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles (right).
-In this case, the right bar is heigher as visible with less free space when adding the frame.
-We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
-	\label{fig:bars_and_framed_rectangles_experiment}
-\end{figure}
-
-\subsection{Weber-Fechner's Law}
-
-As identified by Cleveland and McGill, the bar and framed rectangle experiment is closely related to Weber's law.
-This psychophysics law states that perceivable difference within a distribution is proportional to the initial size of the distribution.
-Weber's law goes hand-in-hand with Fechner's law.
-We conduct an additional experiment based on the original illustrations of the Weber-Fechner law to investigate whether this law can be applied to computational perception of our classifiers (Fig.~\ref{fig:webers_law}).
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{webers_law}
-  \caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution.
-The lower square contains 10 more dots than the upper one on both sides.
-However, the difference is easily perceivable on the left while the squares on the right almost look the same.
-We generate rasterized visualizations similar to this setup and evaluate our classifiers.}
-	\label{fig:webers_law}
-\end{figure}
-
-\section{Results and Discussion}
-
-\subsection{Elementary Perceptual Tasks}
-
-some are good and some are bad.. why?
-
-\begin{figure*}[h]
-	  \includegraphics[width=\linewidth]{figure1.pdf}
-  \caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure1_results}
-\end{figure*}
-
-\textbf{Computational Perception Ranking.}
-
-Cleveland McGills Ranking - can we observe something similar?
-
-\begin{enumerate}
-	\item Position along a common scale e.g. scatter plot
-	\item Position on identical but nonaligned scales e.g. multiple scatter plots
-	\item Length e.g. bar chart
-	\item Angle \& Slope (tie) e.g. pie chart
-	\item Area e.g. bubbles
-	\item Volume, density, and color saturation (tie) e.g. heatmap
-	\item Color hue e.g. newsmap
-\end{enumerate}
-
-\noindent{\textbf{Cross-classifier variability.}}
-
-Can a neural network generalize on simple perceptual tasks?
-
-\subsection{Position-Angle Experiment}
-
-Bar charts are more accurate (Fig.~\ref{fig:figure3_mlae}) and networks converge faster (Fig.~\ref{fig:figure3_val_loss}). This is great.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
-  \caption{\textbf{Classifier Efficiency of the Position-Angle experiment.} Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts.
-We report the MSE measure for both encodings of four different classifier on previously unseen validation data.}
-	\label{fig:figure3_val_loss}
-\end{figure}
-
-\begin{figure*}[t]
-	  \includegraphics[width=\linewidth]{figure3_mlae.pdf}
-  \caption{\textbf{Computational results of the Position-Angle experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure3_mlae}
-\end{figure*}
-
-\subsection{Position-Length Experiment}
-
-We have nothing yet here...
-
-\subsection{Bars and Framed Rectangles Experiment}
-
-First run indicates that framed rectangles perform better but we dont really know it yet.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_val_loss.png}
-  \caption{\textbf{Classifier Efficiency of the Bars and Framed Rectangles experiment.} Categorical Cross-Entropy loss for the \emph{bars and framed rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-The frame around the bars adds an additional visual cue enables faster network convergence.
-This is not yet reproducible!}
-	\label{fig:figure12_val_loss}
-\end{figure}
-
-\section{Conclusions}
-
-Future work: allow insights for infovis for machines
-
-%% if specified like this the section will be committed in review mode
-%\acknowledgments{
-%The authors wish to thank A, B, and C. This work was supported in part by
-%a grant from XYZ (\# 12345-67890).}
-
-%\bibliographystyle{abbrv}
-\bibliographystyle{abbrv-doi}
-%\bibliographystyle{abbrv-doi-narrow}
-%\bibliographystyle{abbrv-doi-hyperref}
-%\bibliographystyle{abbrv-doi-hyperref-narrow}
-
-\bibliography{paper.bib}
-\end{document}

### 2018-03-21 PAPER/06_bars_and_framed_rectangles_experiment.tex

--- 2018-03-16 PAPER/paper.tex
@@
-\documentclass[journal]{vgtc}                % final (journal style)
-%\documentclass[review,journal]{vgtc}         % review (journal style)
-%\documentclass[widereview]{vgtc}             % wide-spaced review
-%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
-
-%% Uncomment one of the lines above depending on where your paper is
-%% in the conference process. ``review'' and ``widereview'' are for review
-%% submission, ``preprint'' is for pre-publication, and the final version
-%% doesn't use a specific qualifier.
-
-%% Please use one of the ``review'' options in combination with the
-%% assigned online id (see below) ONLY if your paper uses a double blind
-%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
-%% in the past.
-
-%% Please note that the use of figures other than the optional teaser is not permitted on the first page
-%% of the journal version.  Figures should begin on the second page and be
-%% in CMYK or Grey scale format, otherwise, colour shifting may occur
-%% during the printing process.  Papers submitted with figures other than the optional teaser on the
-%% first page will be refused. Also, the teaser figure should only have the
-%% width of the abstract as the template enforces it.
-
-%% These few lines make a distinction between latex and pdflatex calls and they
-%% bring in essential packages for graphics and font handling.
-%% Note that due to the \DeclareGraphicsExtensions{} call it is no longer necessary
-%% to provide the the path and extension of a graphics file:
-%% \includegraphics{diamondrule} is completely sufficient.
-%%
-\ifpdf%                                % if we use pdflatex
-  \pdfoutput=1\relax                   % create PDFs from pdfLaTeX
-  \pdfcompresslevel=9                  % PDF Compression
-  \pdfoptionpdfminorversion=7          % create PDF 1.7
-  \ExecuteOptions{pdftex}
-  \usepackage{graphicx}                % allow us to embed graphics files
-  \DeclareGraphicsExtensions{.pdf,.png,.jpg,.jpeg} % for pdflatex we expect .pdf, .png, or .jpg files
-\else%                                 % else we use pure latex
-  \ExecuteOptions{dvips}
-  \usepackage{graphicx}                % allow us to embed graphics files
-  \DeclareGraphicsExtensions{.eps}     % for pure latex we expect eps files
-\fi%
-
-%% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
-\graphicspath{{gfx/}{./}} % where to search for the images
-
-\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
-\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
-\usepackage{textcomp}                  % use better special symbols
-\usepackage{mathptmx}                  % use matching math font
-\usepackage{times}                     % we use Times as the main font
-\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
-\usepackage{cite}                      % needed to automatically sort the references
-\usepackage{tabu}                      % only used for the table example
-\usepackage{booktabs}                  % only used for the table example
-\usepackage{makecell}
-\usepackage{multirow}
-
-%% We encourage the use of mathptmx for consistent usage of times font
-%% throughout the proceedings. However, if you encounter conflicts
-%% with other math-related packages, you may want to disable it.
-
-%% In preprint mode you may define your own headline.
-%\preprinttext{To appear in IEEE Transactions on Visualization and Computer Graphics.}
-
-%% If you are submitting a paper to a conference for review with a double
-%% blind reviewing process, please replace the value ``0'' below with your
-%% OnlineID. Otherwise, you may safely leave it at ``0''.
-\onlineid{0}
-
-%% declare the category of your paper, only shown in review mode
-\vgtccategory{Research}
-%% please declare the paper type of your paper to help reviewers, only shown in review mode
-%% choices:
-%% * algorithm/technique
-%% * application/design study
-%% * evaluation
-%% * system
-%% * theory/model
-\vgtcpapertype{theory/model}
-
-%% Paper title.
-\title{Computing Graphical Perception}
-
-%% This is how authors are specified in the journal style
-
-%% indicate IEEE Member or Student Member in form indicated below
-\author{Daniel Haehn, \textit{Member, IEEE}, James Tompkin, and Hanspeter Pfister}
-\authorfooter{
-%% insert punctuation at end of each item
-\item Daniel Haehn, and Hanspeter Pfister are with the Paulson School of Engineering and Applied Sciences at Harvard University. \\
-E-mail: \{haehn,pfister\}@seas.harvard.edu.
-%
-\item James Tompkin is with the Thomas J. Watson Sr. Center for Information Technology at Brown University. \\E-mail: james\_tompkin@brown.edu.
-}
-
-%other entries to be set up for journal
-\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Computing Graphical Perception}
-%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
-
-%% Abstract section.
-\abstract{TODO%
-} % end of abstract
-
-%% Keywords that describe your work. Will show as 'Index Terms' in journal
-%% please capitalize first letter and insert punctuation after last keyword
-\keywords{Machine Perception, Deep Learning}
-
-%% ACM Computing Classification System (CCS). 
-%% See <http://www.acm.org/class/1998/> for details.
-%% The ``\CCScat'' command takes four arguments.
-
-%\CCScatlist{ % not used in journal version
-% \CCScat{K.6.1}{Management of Computing and Information Systems}%
-%{Project and People Management}{Life Cycle};
-% \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
-%}
-
-%% Uncomment below to include a teaser figure.
-\teaser{
-  \centering
-  \includegraphics[width=\linewidth]{teaser.pdf}
-  \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking visual cortex inspired machine learning classifiers to assess the relationship between values encoded in pie charts and bar charts.
-Similar to the findings of Clevenland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs predict more accurately when working with bar charts (mean squared error, MSE in green).}
-	\label{fig:teaser}
-}
-
-%% Uncomment below to disable the manuscript note
-%\renewcommand{\manuscriptnotetxt}{}
-
-%% Copyright space is enabled by default as required by guidelines.
-%% It is disabled by the 'review' option or via the following command:
-% \nocopyrightspace
-
-\vgtcinsertpkg
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-%% The ``\maketitle'' command must be the first command after the
-%% ``\begin{document}'' command. It prepares and prints the title block.
-
-%% the only exception to this rule is the \firstsection command
-
-\input{00_introduction.tex}
-
-\input{01_related_work.tex}
-
-\section{Experimental Setup}
-
-The experiments shown in this paper are either supervised regression or classification tasks.
-We formulate any estimation of quantities (e.g.
-angles, positions, lengths etc.) as a regression problem between $0$ and $1$.
-The output indicates the percentage in regards to the degrees of freedom of the individual experiment.
-If the experiment involves a choice, we formulate it as a classification problem.
-
-\subsection{Measures}
-
-\textbf{Accuracy.} We use the same metric as Cleveland and McGill to measure accuracy.
-\begin{equation}
-	\log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
-\end{equation}
-\\~\\
-\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the confidence intervals.
-\\~\\
-\noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
-For regression tasks the loss is defined as mean squared error (MSE) and for classification tasks the loss is categorical cross-entropy.
-
-\subsection{Classifiers}
-
-Our classifiers are built upon a multilayer perceptron (MLP) which is a feedforward artificial neural network.
-We combine this MLP with different convolutional neural networks (CNNs) for preprocessing and feature generation.
-These include the traditional LeNet trained from scratch, as well as VGG19 and Xception trained using ImageNet.
-\\~\\
-\noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units~(Fig.~\ref{fig:classifiers}).
-We then add a dropout layer to prevent overfitting and compute linear regression or classification (softmax).
-\\~\\
-\noindent{\textbf{Convolutional Neural Networks.}} We use CNNs to generate additional features as input to the MLP.
-We train the \emph{LeNet} classifier with tune it specifically towards each visualization.
-For \emph{VGG19} and \emph{Xception}, we generate features using previously trained weights on ImageNet.
-\\~\\
-\noindent{\textbf{Optimization.}} All networks are optimized using stochastic gradient descent with Nesterov momentum using fixed parameters (Table \ref{tab:parameters}).
-We train for $1000$ epochs but stop early if the loss does not decrease for ten epochs.
-
-\begin{table}[t]
-\centering
-\caption{We use different feature generators as input to a multilayer perceptron which performs linear regression or the classification task.
-This yields different sets of trainable parameters.
-We also train the MLP directly on the visualizations without any additional feature generation.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lrl}
-%	\toprule
-%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
-%	\midrule
-%	MLP & $0$ & $2,560,513$ \\
-%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
-%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
-%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
-%	\bottomrule
-	\toprule
-	Classifier & \makecell{Trainable\\Parameters} & Optimization \\
-	\midrule
-	MLP & $2,560,513$ & SGD (Nesterov momentum)\\
-	\emph{LeNet} + MLP & $8,026,083$ & Learning rate: $0.0001$\\
-	\emph{VGG19} + MLP & $21,204,545$ & Momentum: $0.9$ \\
-	\emph{Xception} + MLP & $25,580,585$ & \makecell[tl]{Batchsize: 32\\Epochs: $1000$ (Early Stopping)}\\
-	\bottomrule
-\end{tabular}
-}
-\label{tab:parameters}
-\vspace{-4mm}
-\end{table}
-
-\begin{figure}[t]
-	\centering
-	  \includegraphics[width=.8\linewidth]{classifiers.pdf}
-  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
-We use Dropout regularization to prevent overfitting.
-We learn categorical and unordered dependent variables using the softmax function and perform linear regression for continuous variables.
-The MLP can learn the visualizations directly but we also learn features generated by LeNet (2 conv.
-layers, filter size $5\time5$), VGG19 trained on ImageNet (16 conv.
-layers, filter size $3\times3$), or Xception trained on ImageNet (36 conv.
-layers, filter size $3\times3$) to increase the number of trainable parameters.}
-	\label{fig:classifiers}
-\end{figure}
-
-\subsection{Data}
-
-We create all visualizations as parametrized rasterized images without interpolation.
-The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}.
-We add subtle random noise ($0.05$) to each pixel to introduce additional variation.
-\\~\\
-\noindent\textbf{Training/Validation/Test Splits.} We specify the size of each split set as follows: 60,000 training images, 20,000 validation images, and 20,000 test images.
-We then randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables.
-This eliminates leakage during training and evaluation.
-We also scale each set independently: images to the range of $-.5$ to $.5$ and labels to the range of $0.0$ to $1.0$.
-\\~\\
-\noindent\textbf{Cross-classifier variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.
-
-\section{Elementary Perceptual Tasks}
-
-Cleveland and McGill describe the mapping of graphical elements to quantitative variables as \emph{elementary perceptual tasks} and introduce a list of ten different encodings in their paper~\cite{cleveland_mcgill}.
-We create visualizations of these tasks as rasterized images (Fig.~\ref{fig:elementary_perceptual_tasks}).
-
-\subsection{Parametrizations}
-\label{sec:parametrizations}
-We generate multiple parameterizations for each elementary perceptual task (Fig.~\ref{fig:elementary_perceptual_tasks}) and sequentially increase the number of parameters.
-For instance, for \emph{position non-aligned scale} we first only vary the origin of the coordinate system which yields just 10 different parameters.
-We then include translation along the y-axis with a significant increase in variability.
-We then also add x-movement and a variable spot size.
-This results in more complex datasets depending on the variability setting.
-Table~\ref{tab:encoding_parameters} shows the different settings.
-It is important to consider this variability when evaluating different classifiers with individual trainable parameters (Table~\ref{tab:parameters}).
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure1_overview.pdf}
-  \caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of the elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
-We vary the parameters of each perceptual task and then assess the interpretability of feed-forward neural networks.}
-	\label{fig:elementary_perceptual_tasks}
-\end{figure}
-
-\begin{table}[t]
-\centering
-\caption{\textbf{Variability of Elementary Perceptual Tasks.} We sequentially increase the number of parameters for every visual encoding of the elementary perceptual tasks.
-This introduces variability and increasingly more complex datasets.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{llr}
-%	\toprule
-%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
-%	\midrule
-%	MLP & $0$ & $2,560,513$ \\
-%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
-%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
-%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
-%	\bottomrule
-	\toprule
-	Elementary Perceptual Task & Variability & Parameters\\
-	\midrule
-	\makecell[tl]{\emph{Position Common Scale}} & \makecell[tl]{Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $60$ \\ $3600$ \\ $21600$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Position Non-Aligned Scale}} & \makecell[tl]{Origin\\+ Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $10$ \\$600$ \\ $36000$ \\ $216000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Length}} & \makecell[tl]{Length\\+ Position Y \\+ Position X \\+ Width} & \makecell[tr]{ $60$ \\ $2400$ \\ $144000$\\$864000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Direction}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $360$ \\ $21600$ \\ $1296000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Angle}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $90$ \\ $5400$ \\ $324000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Area}} & \makecell[tl]{Radius\\+ Position Y \\+ Position X} & \makecell[tr]{ $40$ \\ $800$ \\ $16000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Volume}} & \makecell[tl]{Cube Sidelength\\+ Position Y \\+ Position X} & \makecell[tr]{ $20$ \\ $400$ \\ $8000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Curvature}} & \makecell[tl]{Midpoint Curvature\\+ Position Y \\+ Position X\\+ Width} & \makecell[tr]{ $80$ \\ $1600$ \\ $64000$ \\ $2560000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Shading}} & \makecell[tl]{Density\\+ Position Y \\+ Position X} & \makecell[tr]{ $100$ \\ $1600$ \\ $2000$ \\ $40000$}\\
-
-	\bottomrule
-\end{tabular}
-}
-\label{tab:encoding_parameters}
-\end{table}
-
-\subsection{Hypotheses}
-
-We proposed four hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H1.1} \textbf{Visual cortex inspired classifiers are able to connect graphical elements to their quantitative variables.} While much simpler models than their biological pendant, convolutional neural networks are heavily influenced by our biological knowledge of the visual system.
-Such classifiers therefor follow the same principles as human perception.
-	\item \textbf{H1.2} \textbf{Computed perceptual performance is dependent on classifier complexity.} We evaluate multiple classifiers with different numbers of trainable parameters.
-A more complex classifier (with higher number of parameters) will perform better on elementary perceptual tasks.
-	\item \textbf{H1.3} \textbf{Some visual encodings are better than others for computations.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
-We investigate whether this order is also relevant for computing graphical perception.
-	\item \textbf{H1.4} \textbf{Classifiers trained on perceptual tasks can generalize to more or less complex variations of the same task.} Recent research suggests that convolutional neural networks generalize extremely well.
-While the underlying reasons are mainly yet unknown, this property allows them to perform on variations of a similar perceptual task.
-\end{itemize}
-
-\section{Position-Angle Experiment}
-
-The position-angle experiment was originally performed by Cleveland and McGill to measure whether humans can better perceive quantities encoded as positions or as angles~\cite{cleveland_mcgill}.
-The actual experiment then compares pie charts versus bar charts since these map down to elementary position and angle judgement.
-We create rasterized images mimicking Cleveland and McGill's proposed encoding and investigate computational perception of our four classifiers.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_overview}
-  \caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
-The experimental task involves the judgement of different encoded values in comparison to the largest encoded values.
-The pie chart (left) and the bar chart (right) visualize the same data point.
-In their paper, Cleveland and McGill report less errors using bar charts.}
-	\label{fig:position_angle_experiment}
-\end{figure}
-
-\subsection{Hypotheses}
-
-We proposed four hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H2.1} \textbf{Computed perceptual performance is better using bar charts than pie charts.} Cleveland and McGill report that position judgements are almost twice as accurate as angle judgements.
-This renders bar charts superior to pie charts and should also be the case for convolutional neural networks.
-	\item \textbf{H2.2} \textbf{Classifiers can learn position faster than angles.} We assume that understanding bar charts is easier than understanding pie charts.
-We suspect that our classifiers learn encodings of positions faster than of angles resulting in more efficient training and faster convergence.
-\end{itemize}
-
-\section{Position-Length Experiment}
-
-This is the one where we estimate two selected bars compared to the longest one - very similar to the previous one but, not yet done.
-We basically test divided versus grouped barchart and we estimate one relation between two marked quanitities: what percent the smaller is of the larger.
-
-There are five types: type 1-3 this is a position judgement along a common scale.
-(btw all classifiers seem to do that extremely well in the elementary tasks so we assume this will work well here too).
-Types 4-5 are length judgements and we know that the classifiers struggle with that quite a bit.
-
-The setup from Cleveland McGill is first a classification task: which one is smaller? and then a regression task: how much smaller. So we have to see how to encode this.
-
-\subsection{Hypotheses}
-
-We proposed two hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H3.1} \textbf{Grouped bar charts are better computational perceivable than divided bar charts.} A grouped bar chart involves judging a position while a divided bar chart most likely (if not the bottom is looked at) requires length judgements.
-Classifiers are better at judging position than at judging length so grouped bar charts are easier to grasp in terms of computational perception.
-	\item \textbf{H3.2} \textbf{not yet} Any ideas?
-\end{itemize}
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{position_length_original1.png}
-	  \includegraphics[width=\linewidth]{position_length_original2.png}
-
-  \caption{\textbf{Position-Length Experiment.} (Not yet) Rasterized versions of the graphs of Cleveland and McGill's position-length experiment.
-The perceptual task involves comparing.
-the two dot-marked quantities across five different visual encodings of either grouped or divided bar charts.
-We evaluate which type of bar chart performs better with our neural networks as a combined classification and regression problem.
-The first task is to select which of the marked quantities is smaller (classification) and the second task is to specify how much smaller it is (regression).}
-	\label{fig:position_length_experiment}
-\end{figure}
-
 \section{Bars and Framed Rectangles Experiment}
@@
 \end{figure}
-
-\section{Results and Discussion}
-
-\subsection{Elementary Perceptual Tasks}
-
-some are good and some are bad.. why?
-
-\begin{figure*}[h]
-	  \includegraphics[width=\linewidth]{figure1.pdf}
-  \caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure1_results}
-\end{figure*}
-
-\textbf{Computational Perception Ranking.}
-
-Cleveland McGills Ranking - can we observe something similar?
-
-\begin{enumerate}
-	\item Position along a common scale e.g. scatter plot
-	\item Position on identical but nonaligned scales e.g. multiple scatter plots
-	\item Length e.g. bar chart
-	\item Angle \& Slope (tie) e.g. pie chart
-	\item Area e.g. bubbles
-	\item Volume, density, and color saturation (tie) e.g. heatmap
-	\item Color hue e.g. newsmap
-\end{enumerate}
-
-\noindent{\textbf{Cross-classifier variability.}}
-
-Can a neural network generalize on simple perceptual tasks?
-
-\subsection{Position-Angle Experiment}
-
-Bar charts are more accurate (Fig.~\ref{fig:figure3_mlae}) and networks converge faster (Fig.~\ref{fig:figure3_val_loss}). This is great.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
-  \caption{\textbf{Classifier Efficiency of the Position-Angle experiment.} Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts.
-We report the MSE measure for both encodings of four different classifier on previously unseen validation data.}
-	\label{fig:figure3_val_loss}
-\end{figure}
-
-\begin{figure*}[t]
-	  \includegraphics[width=\linewidth]{figure3_mlae.pdf}
-  \caption{\textbf{Computational results of the Position-Angle experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure3_mlae}
-\end{figure*}
-
-\subsection{Position-Length Experiment}
-
-We have nothing yet here...
-
-\subsection{Bars and Framed Rectangles Experiment}
-
-First run indicates that framed rectangles perform better but we dont really know it yet.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_val_loss.png}
-  \caption{\textbf{Classifier Efficiency of the Bars and Framed Rectangles experiment.} Categorical Cross-Entropy loss for the \emph{bars and framed rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-The frame around the bars adds an additional visual cue enables faster network convergence.
-This is not yet reproducible!}
-	\label{fig:figure12_val_loss}
-\end{figure}
-
-\section{Conclusions}
-
-Future work: allow insights for infovis for machines
-
-%% if specified like this the section will be committed in review mode
-%\acknowledgments{
-%The authors wish to thank A, B, and C. This work was supported in part by
-%a grant from XYZ (\# 12345-67890).}
-
-%\bibliographystyle{abbrv}
-\bibliographystyle{abbrv-doi}
-%\bibliographystyle{abbrv-doi-narrow}
-%\bibliographystyle{abbrv-doi-hyperref}
-%\bibliographystyle{abbrv-doi-hyperref-narrow}
-
-\bibliography{paper.bib}
-\end{document}

### 2018-03-21 PAPER/07_results.tex

--- 2018-03-16 PAPER/paper.tex
@@
-\documentclass[journal]{vgtc}                % final (journal style)
-%\documentclass[review,journal]{vgtc}         % review (journal style)
-%\documentclass[widereview]{vgtc}             % wide-spaced review
-%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
-
-%% Uncomment one of the lines above depending on where your paper is
-%% in the conference process. ``review'' and ``widereview'' are for review
-%% submission, ``preprint'' is for pre-publication, and the final version
-%% doesn't use a specific qualifier.
-
-%% Please use one of the ``review'' options in combination with the
-%% assigned online id (see below) ONLY if your paper uses a double blind
-%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
-%% in the past.
-
-%% Please note that the use of figures other than the optional teaser is not permitted on the first page
-%% of the journal version.  Figures should begin on the second page and be
-%% in CMYK or Grey scale format, otherwise, colour shifting may occur
-%% during the printing process.  Papers submitted with figures other than the optional teaser on the
-%% first page will be refused. Also, the teaser figure should only have the
-%% width of the abstract as the template enforces it.
-
-%% These few lines make a distinction between latex and pdflatex calls and they
-%% bring in essential packages for graphics and font handling.
-%% Note that due to the \DeclareGraphicsExtensions{} call it is no longer necessary
-%% to provide the the path and extension of a graphics file:
-%% \includegraphics{diamondrule} is completely sufficient.
-%%
-\ifpdf%                                % if we use pdflatex
-  \pdfoutput=1\relax                   % create PDFs from pdfLaTeX
-  \pdfcompresslevel=9                  % PDF Compression
-  \pdfoptionpdfminorversion=7          % create PDF 1.7
-  \ExecuteOptions{pdftex}
-  \usepackage{graphicx}                % allow us to embed graphics files
-  \DeclareGraphicsExtensions{.pdf,.png,.jpg,.jpeg} % for pdflatex we expect .pdf, .png, or .jpg files
-\else%                                 % else we use pure latex
-  \ExecuteOptions{dvips}
-  \usepackage{graphicx}                % allow us to embed graphics files
-  \DeclareGraphicsExtensions{.eps}     % for pure latex we expect eps files
-\fi%
-
-%% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
-\graphicspath{{gfx/}{./}} % where to search for the images
-
-\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
-\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
-\usepackage{textcomp}                  % use better special symbols
-\usepackage{mathptmx}                  % use matching math font
-\usepackage{times}                     % we use Times as the main font
-\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
-\usepackage{cite}                      % needed to automatically sort the references
-\usepackage{tabu}                      % only used for the table example
-\usepackage{booktabs}                  % only used for the table example
-\usepackage{makecell}
-\usepackage{multirow}
-
-%% We encourage the use of mathptmx for consistent usage of times font
-%% throughout the proceedings. However, if you encounter conflicts
-%% with other math-related packages, you may want to disable it.
-
-%% In preprint mode you may define your own headline.
-%\preprinttext{To appear in IEEE Transactions on Visualization and Computer Graphics.}
-
-%% If you are submitting a paper to a conference for review with a double
-%% blind reviewing process, please replace the value ``0'' below with your
-%% OnlineID. Otherwise, you may safely leave it at ``0''.
-\onlineid{0}
-
-%% declare the category of your paper, only shown in review mode
-\vgtccategory{Research}
-%% please declare the paper type of your paper to help reviewers, only shown in review mode
-%% choices:
-%% * algorithm/technique
-%% * application/design study
-%% * evaluation
-%% * system
-%% * theory/model
-\vgtcpapertype{theory/model}
-
-%% Paper title.
-\title{Computing Graphical Perception}
-
-%% This is how authors are specified in the journal style
-
-%% indicate IEEE Member or Student Member in form indicated below
-\author{Daniel Haehn, \textit{Member, IEEE}, James Tompkin, and Hanspeter Pfister}
-\authorfooter{
-%% insert punctuation at end of each item
-\item Daniel Haehn, and Hanspeter Pfister are with the Paulson School of Engineering and Applied Sciences at Harvard University. \\
-E-mail: \{haehn,pfister\}@seas.harvard.edu.
-%
-\item James Tompkin is with the Thomas J. Watson Sr. Center for Information Technology at Brown University. \\E-mail: james\_tompkin@brown.edu.
-}
-
-%other entries to be set up for journal
-\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Computing Graphical Perception}
-%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
-
-%% Abstract section.
-\abstract{TODO%
-} % end of abstract
-
-%% Keywords that describe your work. Will show as 'Index Terms' in journal
-%% please capitalize first letter and insert punctuation after last keyword
-\keywords{Machine Perception, Deep Learning}
-
-%% ACM Computing Classification System (CCS). 
-%% See <http://www.acm.org/class/1998/> for details.
-%% The ``\CCScat'' command takes four arguments.
-
-%\CCScatlist{ % not used in journal version
-% \CCScat{K.6.1}{Management of Computing and Information Systems}%
-%{Project and People Management}{Life Cycle};
-% \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
-%}
-
-%% Uncomment below to include a teaser figure.
-\teaser{
-  \centering
-  \includegraphics[width=\linewidth]{teaser.pdf}
-  \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking visual cortex inspired machine learning classifiers to assess the relationship between values encoded in pie charts and bar charts.
-Similar to the findings of Clevenland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs predict more accurately when working with bar charts (mean squared error, MSE in green).}
-	\label{fig:teaser}
-}
-
-%% Uncomment below to disable the manuscript note
-%\renewcommand{\manuscriptnotetxt}{}
-
-%% Copyright space is enabled by default as required by guidelines.
-%% It is disabled by the 'review' option or via the following command:
-% \nocopyrightspace
-
-\vgtcinsertpkg
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{document}
-
-%% The ``\maketitle'' command must be the first command after the
-%% ``\begin{document}'' command. It prepares and prints the title block.
-
-%% the only exception to this rule is the \firstsection command
-
-\input{00_introduction.tex}
-
-\input{01_related_work.tex}
-
-\section{Experimental Setup}
-
-The experiments shown in this paper are either supervised regression or classification tasks.
-We formulate any estimation of quantities (e.g.
-angles, positions, lengths etc.) as a regression problem between $0$ and $1$.
-The output indicates the percentage in regards to the degrees of freedom of the individual experiment.
-If the experiment involves a choice, we formulate it as a classification problem.
-
-\subsection{Measures}
-
-\textbf{Accuracy.} We use the same metric as Cleveland and McGill to measure accuracy.
-\begin{equation}
-	\log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
-\end{equation}
-\\~\\
-\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the confidence intervals.
-\\~\\
-\noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
-For regression tasks the loss is defined as mean squared error (MSE) and for classification tasks the loss is categorical cross-entropy.
-
-\subsection{Classifiers}
-
-Our classifiers are built upon a multilayer perceptron (MLP) which is a feedforward artificial neural network.
-We combine this MLP with different convolutional neural networks (CNNs) for preprocessing and feature generation.
-These include the traditional LeNet trained from scratch, as well as VGG19 and Xception trained using ImageNet.
-\\~\\
-\noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units~(Fig.~\ref{fig:classifiers}).
-We then add a dropout layer to prevent overfitting and compute linear regression or classification (softmax).
-\\~\\
-\noindent{\textbf{Convolutional Neural Networks.}} We use CNNs to generate additional features as input to the MLP.
-We train the \emph{LeNet} classifier with tune it specifically towards each visualization.
-For \emph{VGG19} and \emph{Xception}, we generate features using previously trained weights on ImageNet.
-\\~\\
-\noindent{\textbf{Optimization.}} All networks are optimized using stochastic gradient descent with Nesterov momentum using fixed parameters (Table \ref{tab:parameters}).
-We train for $1000$ epochs but stop early if the loss does not decrease for ten epochs.
-
-\begin{table}[t]
-\centering
-\caption{We use different feature generators as input to a multilayer perceptron which performs linear regression or the classification task.
-This yields different sets of trainable parameters.
-We also train the MLP directly on the visualizations without any additional feature generation.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lrl}
-%	\toprule
-%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
-%	\midrule
-%	MLP & $0$ & $2,560,513$ \\
-%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
-%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
-%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
-%	\bottomrule
-	\toprule
-	Classifier & \makecell{Trainable\\Parameters} & Optimization \\
-	\midrule
-	MLP & $2,560,513$ & SGD (Nesterov momentum)\\
-	\emph{LeNet} + MLP & $8,026,083$ & Learning rate: $0.0001$\\
-	\emph{VGG19} + MLP & $21,204,545$ & Momentum: $0.9$ \\
-	\emph{Xception} + MLP & $25,580,585$ & \makecell[tl]{Batchsize: 32\\Epochs: $1000$ (Early Stopping)}\\
-	\bottomrule
-\end{tabular}
-}
-\label{tab:parameters}
-\vspace{-4mm}
-\end{table}
-
-\begin{figure}[t]
-	\centering
-	  \includegraphics[width=.8\linewidth]{classifiers.pdf}
-  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
-We use Dropout regularization to prevent overfitting.
-We learn categorical and unordered dependent variables using the softmax function and perform linear regression for continuous variables.
-The MLP can learn the visualizations directly but we also learn features generated by LeNet (2 conv.
-layers, filter size $5\time5$), VGG19 trained on ImageNet (16 conv.
-layers, filter size $3\times3$), or Xception trained on ImageNet (36 conv.
-layers, filter size $3\times3$) to increase the number of trainable parameters.}
-	\label{fig:classifiers}
-\end{figure}
-
-\subsection{Data}
-
-We create all visualizations as parametrized rasterized images without interpolation.
-The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}.
-We add subtle random noise ($0.05$) to each pixel to introduce additional variation.
-\\~\\
-\noindent\textbf{Training/Validation/Test Splits.} We specify the size of each split set as follows: 60,000 training images, 20,000 validation images, and 20,000 test images.
-We then randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables.
-This eliminates leakage during training and evaluation.
-We also scale each set independently: images to the range of $-.5$ to $.5$ and labels to the range of $0.0$ to $1.0$.
-\\~\\
-\noindent\textbf{Cross-classifier variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.
-
-\section{Elementary Perceptual Tasks}
-
-Cleveland and McGill describe the mapping of graphical elements to quantitative variables as \emph{elementary perceptual tasks} and introduce a list of ten different encodings in their paper~\cite{cleveland_mcgill}.
-We create visualizations of these tasks as rasterized images (Fig.~\ref{fig:elementary_perceptual_tasks}).
-
-\subsection{Parametrizations}
-\label{sec:parametrizations}
-We generate multiple parameterizations for each elementary perceptual task (Fig.~\ref{fig:elementary_perceptual_tasks}) and sequentially increase the number of parameters.
-For instance, for \emph{position non-aligned scale} we first only vary the origin of the coordinate system which yields just 10 different parameters.
-We then include translation along the y-axis with a significant increase in variability.
-We then also add x-movement and a variable spot size.
-This results in more complex datasets depending on the variability setting.
-Table~\ref{tab:encoding_parameters} shows the different settings.
-It is important to consider this variability when evaluating different classifiers with individual trainable parameters (Table~\ref{tab:parameters}).
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure1_overview.pdf}
-  \caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of the elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
-We vary the parameters of each perceptual task and then assess the interpretability of feed-forward neural networks.}
-	\label{fig:elementary_perceptual_tasks}
-\end{figure}
-
-\begin{table}[t]
-\centering
-\caption{\textbf{Variability of Elementary Perceptual Tasks.} We sequentially increase the number of parameters for every visual encoding of the elementary perceptual tasks.
-This introduces variability and increasingly more complex datasets.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{llr}
-%	\toprule
-%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
-%	\midrule
-%	MLP & $0$ & $2,560,513$ \\
-%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
-%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
-%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
-%	\bottomrule
-	\toprule
-	Elementary Perceptual Task & Variability & Parameters\\
-	\midrule
-	\makecell[tl]{\emph{Position Common Scale}} & \makecell[tl]{Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $60$ \\ $3600$ \\ $21600$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Position Non-Aligned Scale}} & \makecell[tl]{Origin\\+ Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $10$ \\$600$ \\ $36000$ \\ $216000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Length}} & \makecell[tl]{Length\\+ Position Y \\+ Position X \\+ Width} & \makecell[tr]{ $60$ \\ $2400$ \\ $144000$\\$864000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Direction}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $360$ \\ $21600$ \\ $1296000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Angle}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $90$ \\ $5400$ \\ $324000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Area}} & \makecell[tl]{Radius\\+ Position Y \\+ Position X} & \makecell[tr]{ $40$ \\ $800$ \\ $16000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Volume}} & \makecell[tl]{Cube Sidelength\\+ Position Y \\+ Position X} & \makecell[tr]{ $20$ \\ $400$ \\ $8000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Curvature}} & \makecell[tl]{Midpoint Curvature\\+ Position Y \\+ Position X\\+ Width} & \makecell[tr]{ $80$ \\ $1600$ \\ $64000$ \\ $2560000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Shading}} & \makecell[tl]{Density\\+ Position Y \\+ Position X} & \makecell[tr]{ $100$ \\ $1600$ \\ $2000$ \\ $40000$}\\
-
-	\bottomrule
-\end{tabular}
-}
-\label{tab:encoding_parameters}
-\end{table}
-
-\subsection{Hypotheses}
-
-We proposed four hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H1.1} \textbf{Visual cortex inspired classifiers are able to connect graphical elements to their quantitative variables.} While much simpler models than their biological pendant, convolutional neural networks are heavily influenced by our biological knowledge of the visual system.
-Such classifiers therefor follow the same principles as human perception.
-	\item \textbf{H1.2} \textbf{Computed perceptual performance is dependent on classifier complexity.} We evaluate multiple classifiers with different numbers of trainable parameters.
-A more complex classifier (with higher number of parameters) will perform better on elementary perceptual tasks.
-	\item \textbf{H1.3} \textbf{Some visual encodings are better than others for computations.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
-We investigate whether this order is also relevant for computing graphical perception.
-	\item \textbf{H1.4} \textbf{Classifiers trained on perceptual tasks can generalize to more or less complex variations of the same task.} Recent research suggests that convolutional neural networks generalize extremely well.
-While the underlying reasons are mainly yet unknown, this property allows them to perform on variations of a similar perceptual task.
-\end{itemize}
-
-\section{Position-Angle Experiment}
-
-The position-angle experiment was originally performed by Cleveland and McGill to measure whether humans can better perceive quantities encoded as positions or as angles~\cite{cleveland_mcgill}.
-The actual experiment then compares pie charts versus bar charts since these map down to elementary position and angle judgement.
-We create rasterized images mimicking Cleveland and McGill's proposed encoding and investigate computational perception of our four classifiers.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_overview}
-  \caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
-The experimental task involves the judgement of different encoded values in comparison to the largest encoded values.
-The pie chart (left) and the bar chart (right) visualize the same data point.
-In their paper, Cleveland and McGill report less errors using bar charts.}
-	\label{fig:position_angle_experiment}
-\end{figure}
-
-\subsection{Hypotheses}
-
-We proposed four hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H2.1} \textbf{Computed perceptual performance is better using bar charts than pie charts.} Cleveland and McGill report that position judgements are almost twice as accurate as angle judgements.
-This renders bar charts superior to pie charts and should also be the case for convolutional neural networks.
-	\item \textbf{H2.2} \textbf{Classifiers can learn position faster than angles.} We assume that understanding bar charts is easier than understanding pie charts.
-We suspect that our classifiers learn encodings of positions faster than of angles resulting in more efficient training and faster convergence.
-\end{itemize}
-
-\section{Position-Length Experiment}
-
-This is the one where we estimate two selected bars compared to the longest one - very similar to the previous one but, not yet done.
-We basically test divided versus grouped barchart and we estimate one relation between two marked quanitities: what percent the smaller is of the larger.
-
-There are five types: type 1-3 this is a position judgement along a common scale.
-(btw all classifiers seem to do that extremely well in the elementary tasks so we assume this will work well here too).
-Types 4-5 are length judgements and we know that the classifiers struggle with that quite a bit.
-
-The setup from Cleveland McGill is first a classification task: which one is smaller? and then a regression task: how much smaller. So we have to see how to encode this.
-
-\subsection{Hypotheses}
-
-We proposed two hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H3.1} \textbf{Grouped bar charts are better computational perceivable than divided bar charts.} A grouped bar chart involves judging a position while a divided bar chart most likely (if not the bottom is looked at) requires length judgements.
-Classifiers are better at judging position than at judging length so grouped bar charts are easier to grasp in terms of computational perception.
-	\item \textbf{H3.2} \textbf{not yet} Any ideas?
-\end{itemize}
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{position_length_original1.png}
-	  \includegraphics[width=\linewidth]{position_length_original2.png}
-
-  \caption{\textbf{Position-Length Experiment.} (Not yet) Rasterized versions of the graphs of Cleveland and McGill's position-length experiment.
-The perceptual task involves comparing.
-the two dot-marked quantities across five different visual encodings of either grouped or divided bar charts.
-We evaluate which type of bar chart performs better with our neural networks as a combined classification and regression problem.
-The first task is to select which of the marked quantities is smaller (classification) and the second task is to specify how much smaller it is (regression).}
-	\label{fig:position_length_experiment}
-\end{figure}
-
-\section{Bars and Framed Rectangles Experiment}
-
-Visual cues can help converting graphical elements back to their real world variables.
-Cleveland and McGill introduced the bars and framed rectangles experiment which judges the elementary perceptual task of position along non-aligned scales~\cite{cleveland_mcgill}.
-
-\subsection{Hypotheses}
-
-We proposed two hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H4.1} \textbf{Classifiers can leverage additional visual cues.} The original bar and framed rectangle experiment shows how visual cues aid humans in mapping graphical elements to quantitative variables.
-This should be the same for feed-forward neural networks since they are inspired by the visual system.
-	\item \textbf{H4.2} \textbf{Weber's law can be transferred to computational perception.} Cleveland and McGill confirmed Weber's law based on the bar and framed rectangle experiment.
-For humans, the ability to perceive change within a distribution is proportional to the size of the initial distribution.
-\end{itemize}
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_overview}
-  \caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales.
-For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles (right).
-In this case, the right bar is heigher as visible with less free space when adding the frame.
-We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
-	\label{fig:bars_and_framed_rectangles_experiment}
-\end{figure}
-
-\subsection{Weber-Fechner's Law}
-
-As identified by Cleveland and McGill, the bar and framed rectangle experiment is closely related to Weber's law.
-This psychophysics law states that perceivable difference within a distribution is proportional to the initial size of the distribution.
-Weber's law goes hand-in-hand with Fechner's law.
-We conduct an additional experiment based on the original illustrations of the Weber-Fechner law to investigate whether this law can be applied to computational perception of our classifiers (Fig.~\ref{fig:webers_law}).
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{webers_law}
-  \caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution.
-The lower square contains 10 more dots than the upper one on both sides.
-However, the difference is easily perceivable on the left while the squares on the right almost look the same.
-We generate rasterized visualizations similar to this setup and evaluate our classifiers.}
-	\label{fig:webers_law}
-\end{figure}
-
 \section{Results and Discussion}
@@
 Can a neural network generalize on simple perceptual tasks?
+
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{crossnetwork.pdf}
+  \caption{\textbf{Cross-classifier variability for the perceptual task of measuring curvature.} We use predictions of LeNet classifiers trained on different parametrizations of the \emph{curvature} elementary perceptual task and measure the mean logistic absolute error (MLAE).
+The lower score, the better.
+Classifiers trained on curves with variable position can generalize even if the axis of translation varies.
+However, classifiers trained on fixed positions of curves are not able to measure translated curves.}
+	\label{fig:cross_network}
+\end{figure}
 
@@
 \end{figure}
-
-\section{Conclusions}
-
-Future work: allow insights for infovis for machines
-
-%% if specified like this the section will be committed in review mode
-%\acknowledgments{
-%The authors wish to thank A, B, and C. This work was supported in part by
-%a grant from XYZ (\# 12345-67890).}
-
-%\bibliographystyle{abbrv}
-\bibliographystyle{abbrv-doi}
-%\bibliographystyle{abbrv-doi-narrow}
-%\bibliographystyle{abbrv-doi-hyperref}
-%\bibliographystyle{abbrv-doi-hyperref-narrow}
-
-\bibliography{paper.bib}
-\end{document}

### 2018-03-21 PAPER/paper.tex

@@
 % - 
-\title{Computing Graphical Perception}
+\title{Reproducing 'Graphical Perception' with CNNs}
 
@@
 %% indicate IEEE Member or Student Member in form indicated below
-\author{Daniel Haehn, \textit{Member, IEEE}, James Tompkin, and Hanspeter Pfister}
+%\author{Daniel Haehn, \textit{Member, IEEE}, James Tompkin, and Hanspeter Pfister}
+\author{Daniel Haehn, James Tompkin, and Hanspeter Pfister}
 \authorfooter{
@@
 %other entries to be set up for journal
-\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Computing Graphical Perception}
+\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Computing Graphical Perception with Convolutional Neural Networks}
 %\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
@@
 We further introduce a ranking of elementary visual encodings targeted towards modern deep neural networks and derive practical evidence for properties of convolutional neural networks such as the translation invariance.
-Convolutional neural networks can successfully perform many computer vision tasks on images, and their early-layer learned representations are often said to mimic the early stages of the visual cortex.
-But can CNNs solve graphical perception tasks for visualization?
+Convolutional neural networks can successfully perform many computer vision tasks on images, and their learned representations are often said to mimic the early layers of the visual cortex.
+But can CNNs understand graphical perception for visualization?
 We investigate this question by reproducing Cleveland and McGill's seminal 1984 experiments, which measured human perception efficiency of different visual encodings and defined elementary perceptual tasks for visualization.
-We measure the graphical perceptual capabilities of four classifiers on a) elementary perceptual tasks with increasing parametric complexity, b) the position-angle experiment which compares pie and bar charts, c) the position-length experiment which compares grouped and divided bar charts, and d) the bars and framed rectangles experiment where visual cues help to perceive information.
+We measure the graphical perceptual capabilities of four classifiers on a) elementary perceptual tasks with increasing parametric complexity, b) the position-angle experiment that compares pie charts to bar charts, c) the position-length experiment that compares grouped and divided bar charts, and d) the bars and framed rectangles experiment where visual cues aid perception.
 We also study how feed-forward neural networks obey Weber's law, which defines the proportional relation between perceivable information and distribution density.
-We present the results of these experiments to foster the understanding of how such classifiers succeed and fail when applied to data visualizations.
+We present the results of these experiments to foster the understanding of how CNN classifiers succeed and fail when applied to data visualizations.
 } % end of abstract
@@
   \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking visual cortex inspired machine learning classifiers to assess the relationship between values encoded in pie charts and bar charts.
-Similar to the findings of Clevenland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs predict more accurately when working with bar charts (mean squared error, MSE in green).}
+Similar to the findings of Clevenland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs read quantities more accurately from bar charts (mean squared error, MSE in green).}
 	\label{fig:teaser}
@@
 
-\section{Experimental Setup}
+\input{02_experimental_setup.tex}
 
-The experiments shown in this paper are either supervised regression or classification tasks.
-We formulate any estimation of quantities (e.g.
-angles, positions, lengths etc.) as a regression problem between $0$ and $1$.
-The output indicates the percentage in regards to the degrees of freedom of the individual experiment.
-If the experiment involves a choice, we formulate it as a classification problem.
+\input{03_elementary_perceptual_tasks.tex}
 
-\subsection{Measures}
+\input{04_position_angle_experiment.tex}
 
-\textbf{Accuracy.} We use the same metric as Cleveland and McGill to measure accuracy.
-\begin{equation}
-	\log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
-\end{equation}
-\\~\\
-\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the confidence intervals.
-\\~\\
-\noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
-For regression tasks the loss is defined as mean squared error (MSE) and for classification tasks the loss is categorical cross-entropy.
+\input{05_position_length_experiment.tex}
 
-\subsection{Classifiers}
+\input{06_bars_and_framed_rectangles_experiment.tex}
 
-Our classifiers are built upon a multilayer perceptron (MLP) which is a feedforward artificial neural network.
-We combine this MLP with different convolutional neural networks (CNNs) for preprocessing and feature generation.
-These include the traditional LeNet trained from scratch, as well as VGG19 and Xception trained using ImageNet.
-\\~\\
-\noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units~(Fig.~\ref{fig:classifiers}).
-We then add a dropout layer to prevent overfitting and compute linear regression or classification (softmax).
-\\~\\
-\noindent{\textbf{Convolutional Neural Networks.}} We use CNNs to generate additional features as input to the MLP.
-We train the \emph{LeNet} classifier with tune it specifically towards each visualization.
-For \emph{VGG19} and \emph{Xception}, we generate features using previously trained weights on ImageNet.
-\\~\\
-\noindent{\textbf{Optimization.}} All networks are optimized using stochastic gradient descent with Nesterov momentum using fixed parameters (Table \ref{tab:parameters}).
-We train for $1000$ epochs but stop early if the loss does not decrease for ten epochs.
-\\~\\
-\noindent{\textbf{Environment.}} We run all experiments on an NVIDIA DGX1 machine with Tesla V100 graphical processing units. We use the KERAS framework with tensorflow.
+\input{07_results.tex}
 
-\begin{table}[t]
-\centering
-\caption{We use different feature generators as input to a multilayer perceptron which performs linear regression or the classification task.
-This yields different sets of trainable parameters.
-We also train the MLP directly on the visualizations without any additional feature generation.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lrl}
-%	\toprule
-%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
-%	\midrule
-%	MLP & $0$ & $2,560,513$ \\
-%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
-%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
-%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
-%	\bottomrule
-	\toprule
-	Classifier & \makecell{Trainable\\Parameters} & Optimization \\
-	\midrule
-	MLP & $2,560,513$ & SGD (Nesterov momentum)\\
-	\emph{LeNet} + MLP & $8,026,083$ & Learning rate: $0.0001$\\
-	\emph{VGG19} + MLP & $21,204,545$ & Momentum: $0.9$ \\
-	\emph{Xception} + MLP & $25,580,585$ & \makecell[tl]{Batchsize: 32\\Epochs: $1000$ (Early Stopping)}\\
-	\bottomrule
-\end{tabular}
-}
-\label{tab:parameters}
-\vspace{-4mm}
-\end{table}
-
-\begin{figure}[t]
-	\centering
-	  \includegraphics[width=.8\linewidth]{classifiers.pdf}
-  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
-We use Dropout regularization to prevent overfitting.
-We learn categorical and unordered dependent variables using the softmax function and perform linear regression for continuous variables.
-The MLP can learn the visualizations directly but we also learn features generated by LeNet (2 conv.
-layers, filter size $5\time5$), VGG19 trained on ImageNet (16 conv.
-layers, filter size $3\times3$), or Xception trained on ImageNet (36 conv.
-layers, filter size $3\times3$) to increase the number of trainable parameters.}
-	\label{fig:classifiers}
-\end{figure}
-
-\subsection{Data}
-
-We create all visualizations as parametrized rasterized images without interpolation.
-The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}.
-We add subtle random noise ($0.05$) to each pixel to introduce additional variation.
-\\~\\
-\noindent\textbf{Training/Validation/Test Splits.} We specify the size of each split set as follows: 60,000 training images, 20,000 validation images, and 20,000 test images.
-We then randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables.
-This eliminates leakage during training and evaluation.
-We also scale each set independently: images to the range of $-.5$ to $.5$ and labels to the range of $0.0$ to $1.0$.
-\\~\\
-\noindent\textbf{Cross-classifier variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.
-
-\section{Elementary Perceptual Tasks}
-
-Cleveland and McGill describe the mapping of graphical elements to quantitative variables as \emph{elementary perceptual tasks} and introduce a list of ten different encodings in their paper~\cite{cleveland_mcgill}.
-We create visualizations of these tasks as rasterized images (Fig.~\ref{fig:elementary_perceptual_tasks}).
-
-\subsection{Parametrizations}
-\label{sec:parametrizations}
-We generate multiple parameterizations for each elementary perceptual task (Fig.~\ref{fig:elementary_perceptual_tasks}) and sequentially increase the number of parameters.
-For instance, for \emph{position non-aligned scale} we first only vary the origin of the coordinate system which yields just 10 different parameters.
-We then include translation along the y-axis with a significant increase in variability.
-We then also add x-movement and a variable spot size.
-This results in more complex datasets depending on the variability setting.
-Table~\ref{tab:encoding_parameters} shows the different settings.
-It is important to consider this variability when evaluating different classifiers with individual trainable parameters (Table~\ref{tab:parameters}).
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure1_overview.pdf}
-  \caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of the elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
-We vary the parameters of each perceptual task and then assess the interpretability of feed-forward neural networks.}
-	\label{fig:elementary_perceptual_tasks}
-\end{figure}
-
-\begin{table}[t]
-\centering
-\caption{\textbf{Variability of Elementary Perceptual Tasks.} We sequentially increase the number of parameters for every visual encoding of the elementary perceptual tasks.
-This introduces variability and increasingly more complex datasets.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{llr}
-%	\toprule
-%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
-%	\midrule
-%	MLP & $0$ & $2,560,513$ \\
-%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
-%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
-%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
-%	\bottomrule
-	\toprule
-	Elementary Perceptual Task & Variability & Parameters\\
-	\midrule
-	\makecell[tl]{\emph{Position Common Scale}} & \makecell[tl]{Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $60$ \\ $3600$ \\ $21600$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Position Non-Aligned Scale}} & \makecell[tl]{Origin\\+ Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $10$ \\$600$ \\ $36000$ \\ $216000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Length}} & \makecell[tl]{Length\\+ Position Y \\+ Position X \\+ Width} & \makecell[tr]{ $60$ \\ $2400$ \\ $144000$\\$864000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Direction}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $360$ \\ $21600$ \\ $1296000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Angle}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $90$ \\ $5400$ \\ $324000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Area}} & \makecell[tl]{Radius\\+ Position Y \\+ Position X} & \makecell[tr]{ $40$ \\ $800$ \\ $16000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Volume}} & \makecell[tl]{Cube Sidelength\\+ Position Y \\+ Position X} & \makecell[tr]{ $20$ \\ $400$ \\ $8000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Curvature}} & \makecell[tl]{Midpoint Curvature\\+ Position Y \\+ Position X\\+ Width} & \makecell[tr]{ $80$ \\ $1600$ \\ $64000$ \\ $2560000$}\\
-
-	\midrule
-	\makecell[tl]{\emph{Shading}} & \makecell[tl]{Density\\+ Position Y \\+ Position X} & \makecell[tr]{ $100$ \\ $1600$ \\ $2000$ \\ $40000$}\\
-
-	\bottomrule
-\end{tabular}
-}
-\label{tab:encoding_parameters}
-\end{table}
-
-\subsection{Hypotheses}
-
-We proposed four hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H1.1} \textbf{Visual cortex inspired classifiers are able to connect graphical elements to their quantitative variables.} While much simpler models than their biological pendant, convolutional neural networks are heavily influenced by our biological knowledge of the visual system.
-Such classifiers therefor follow the same principles as human perception.
-	\item \textbf{H1.2} \textbf{Computed perceptual performance is dependent on classifier complexity.} We evaluate multiple classifiers with different numbers of trainable parameters.
-A more complex classifier (with higher number of parameters) will perform better on elementary perceptual tasks.
-	\item \textbf{H1.3} \textbf{Some visual encodings are better than others for computations.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
-We investigate whether this order is also relevant for computing graphical perception.
-	\item \textbf{H1.4} \textbf{Classifiers trained on perceptual tasks can generalize to more or less complex variations of the same task.} Recent research suggests that convolutional neural networks generalize extremely well.
-While the underlying reasons are mainly yet unknown, this property allows them to perform on variations of a similar perceptual task.
-\end{itemize}
-
-\section{Position-Angle Experiment}
-
-The position-angle experiment was originally performed by Cleveland and McGill to measure whether humans can better perceive quantities encoded as positions or as angles~\cite{cleveland_mcgill}.
-The actual experiment then compares pie charts versus bar charts since these map down to elementary position and angle judgement.
-We create rasterized images mimicking Cleveland and McGill's proposed encoding and investigate computational perception of our four classifiers.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_overview}
-  \caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
-The experimental task involves the judgement of different encoded values in comparison to the largest encoded values.
-The pie chart (left) and the bar chart (right) visualize the same data point.
-In their paper, Cleveland and McGill report less errors using bar charts.}
-	\label{fig:position_angle_experiment}
-\end{figure}
-
-\subsection{Hypotheses}
-
-We proposed four hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H2.1} \textbf{Computed perceptual performance is better using bar charts than pie charts.} Cleveland and McGill report that position judgements are almost twice as accurate as angle judgements.
-This renders bar charts superior to pie charts and should also be the case for convolutional neural networks.
-	\item \textbf{H2.2} \textbf{Classifiers can learn position faster than angles.} We assume that understanding bar charts is easier than understanding pie charts.
-We suspect that our classifiers learn encodings of positions faster than of angles resulting in more efficient training and faster convergence.
-\end{itemize}
-
-\section{Position-Length Experiment}
-
-This is the one where we estimate two selected bars compared to the longest one - very similar to the previous one but, not yet done.
-We basically test divided versus grouped barchart and we estimate one relation between two marked quanitities: what percent the smaller is of the larger.
-
-There are five types: type 1-3 this is a position judgement along a common scale.
-(btw all classifiers seem to do that extremely well in the elementary tasks so we assume this will work well here too).
-Types 4-5 are length judgements and we know that the classifiers struggle with that quite a bit.
-
-The setup from Cleveland McGill is first a classification task: which one is smaller? and then a regression task: how much smaller. So we have to see how to encode this.
-
-\subsection{Hypotheses}
-
-We proposed two hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H3.1} \textbf{Grouped bar charts are better computational perceivable than divided bar charts.} A grouped bar chart involves judging a position while a divided bar chart most likely (if not the bottom is looked at) requires length judgements.
-Classifiers are better at judging position than at judging length so grouped bar charts are easier to grasp in terms of computational perception.
-	\item \textbf{H3.2} \textbf{not yet} Any ideas?
-\end{itemize}
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{position_length_original1.png}
-	  \includegraphics[width=\linewidth]{position_length_original2.png}
-
-  \caption{\textbf{Position-Length Experiment.} (Not yet) Rasterized versions of the graphs of Cleveland and McGill's position-length experiment.
-The perceptual task involves comparing.
-the two dot-marked quantities across five different visual encodings of either grouped or divided bar charts.
-We evaluate which type of bar chart performs better with our neural networks as a combined classification and regression problem.
-The first task is to select which of the marked quantities is smaller (classification) and the second task is to specify how much smaller it is (regression).}
-	\label{fig:position_length_experiment}
-\end{figure}
-
-\section{Bars and Framed Rectangles Experiment}
-
-Visual cues can help converting graphical elements back to their real world variables.
-Cleveland and McGill introduced the bars and framed rectangles experiment which judges the elementary perceptual task of position along non-aligned scales~\cite{cleveland_mcgill}.
-
-\subsection{Hypotheses}
-
-We proposed two hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H4.1} \textbf{Classifiers can leverage additional visual cues.} The original bar and framed rectangle experiment shows how visual cues aid humans in mapping graphical elements to quantitative variables.
-This should be the same for feed-forward neural networks since they are inspired by the visual system.
-	\item \textbf{H4.2} \textbf{Weber's law can be transferred to computational perception.} Cleveland and McGill confirmed Weber's law based on the bar and framed rectangle experiment.
-For humans, the ability to perceive change within a distribution is proportional to the size of the initial distribution.
-\end{itemize}
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_overview}
-  \caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales.
-For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles (right).
-In this case, the right bar is heigher as visible with less free space when adding the frame.
-We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
-	\label{fig:bars_and_framed_rectangles_experiment}
-\end{figure}
-
-\subsection{Weber-Fechner's Law}
-
-As identified by Cleveland and McGill, the bar and framed rectangle experiment is closely related to Weber's law.
-This psychophysics law states that perceivable difference within a distribution is proportional to the initial size of the distribution.
-Weber's law goes hand-in-hand with Fechner's law.
-We conduct an additional experiment based on the original illustrations of the Weber-Fechner law to investigate whether this law can be applied to computational perception of our classifiers (Fig.~\ref{fig:webers_law}).
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{webers_law}
-  \caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution.
-The lower square contains 10 more dots than the upper one on both sides.
-However, the difference is easily perceivable on the left while the squares on the right almost look the same.
-We generate rasterized visualizations similar to this setup and evaluate our classifiers.}
-	\label{fig:webers_law}
-\end{figure}
-
-\section{Results and Discussion}
-
-\subsection{Elementary Perceptual Tasks}
-
-some are good and some are bad.. why?
-
-\begin{figure*}[h]
-	  \includegraphics[width=\linewidth]{figure1.pdf}
-  \caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure1_results}
-\end{figure*}
-
-\textbf{Computational Perception Ranking.}
-
-Cleveland McGills Ranking - can we observe something similar?
-
-\begin{enumerate}
-	\item Position along a common scale e.g. scatter plot
-	\item Position on identical but nonaligned scales e.g. multiple scatter plots
-	\item Length e.g. bar chart
-	\item Angle \& Slope (tie) e.g. pie chart
-	\item Area e.g. bubbles
-	\item Volume, density, and color saturation (tie) e.g. heatmap
-	\item Color hue e.g. newsmap
-\end{enumerate}
-
-\noindent{\textbf{Cross-classifier variability.}}
-
-Can a neural network generalize on simple perceptual tasks?
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{crossnetwork.pdf}
-  \caption{\textbf{Cross-classifier variability for the perceptual task of measuring curvature.} We use predictions of LeNet classifiers trained on different parametrizations of the \emph{curvature} elementary perceptual task and measure the mean logistic absolute error (MLAE).
-The lower score, the better.
-Classifiers trained on curves with variable position can generalize even if the axis of translation varies.
-However, classifiers trained on fixed positions of curves are not able to measure translated curves.}
-	\label{fig:cross_network}
-\end{figure}
-
-\subsection{Position-Angle Experiment}
-
-Bar charts are more accurate (Fig.~\ref{fig:figure3_mlae}) and networks converge faster (Fig.~\ref{fig:figure3_val_loss}). This is great.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
-  \caption{\textbf{Classifier Efficiency of the Position-Angle experiment.} Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts.
-We report the MSE measure for both encodings of four different classifier on previously unseen validation data.}
-	\label{fig:figure3_val_loss}
-\end{figure}
-
-\begin{figure*}[t]
-	  \includegraphics[width=\linewidth]{figure3_mlae.pdf}
-  \caption{\textbf{Computational results of the Position-Angle experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure3_mlae}
-\end{figure*}
-
-\subsection{Position-Length Experiment}
-
-We have nothing yet here...
-
-\subsection{Bars and Framed Rectangles Experiment}
-
-First run indicates that framed rectangles perform better but we dont really know it yet.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_val_loss.png}
-  \caption{\textbf{Classifier Efficiency of the Bars and Framed Rectangles experiment.} Categorical Cross-Entropy loss for the \emph{bars and framed rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-The frame around the bars adds an additional visual cue enables faster network convergence.
-This is not yet reproducible!}
-	\label{fig:figure12_val_loss}
-\end{figure}
-
-\section{Conclusions}
-
-Future work: allow insights for infovis for machines
+\input{08_conclusions.tex}
 

### 2018-03-21 EXP/ClevelandMcGill/__init__.py
Cleaned everything up!

@@ -1,7 +1,3 @@
 from figure1 import Figure1
 from figure3 import Figure3
-
-from figure1_length import Figure1_Length
-from figure1_position_nonaligned_scale import Figure1_Position_Nonaligned_Scale
-from figure1_position_common_scale import Figure1_Position_Common_Scale
 from figure12 import Figure12

### 2018-03-21 EXP/ClevelandMcGill/figure12.py

@@ -1,321 +1,0 @@
-import cPickle as pickle
-import matplotlib.pyplot as plt
-import numpy as np
-import os
-import sys
-
-sys.path.append('../')
-from util import Util
-
-class Figure12:
-
-  def __init__(self, variable_y=False, \
-                     variable_width=False, \
-                     variable_x=False, \
-                     constrain=True):
-    '''
-  	'''
-
-    # the flag for the visual aid
-    self.framed = False
-
-    self.A_x = None
-    self.A_y = None
-    self.A_value = None
-    self.A_width = None
-
-    self.B_x = None
-    self.B_y = None
-    self.B_value = None
-    self.B_width = None
-
-    self.label = None
-
-    # measured from Cleveland McGill paper via pixel ruler
-    self._size = (100,100)
-    self._max_height = 65
-    self._default_width = 20
-
-    self._delta = 5 # only if we constrain the differences between A and B
-    # this enables cases which are difficult for the human eye
-
-    #
-    # variability switches
-    #
-    #  stage 1: var. Y
-    #  stage 2: var. WIDTH
-    #  stage 3: var. X
-    #
-    self._variable_y = variable_y
-    self._variable_width = variable_width
-    self._variable_x = variable_x
-    self._constrain = constrain
-
-
-  def create(self, verbose=False):
-    '''
-    Create a valid datapoint which fulfills constrains
-    and the variability settings.
-    '''
-    self.create_(verbose)
-
-    while (self.A_value == self.B_value): #or (self._variable_y and self._constrain and (np.abs(self.A_y - self.B_y) < 10)):
-      # constrain means that the difference between the y position has to be larger than 10 pixel
-      # this only enables that the human eye can't really perceive the difference
-      # we always eliminate cases where the two bars are of the same value since we can not label these as 0 or 1
-      # (it is really not part of the experiment)
-      self.create_(verbose)
-
-    # update the label
-    if self.A_value > self.B_value:
-      self.label = 0
-    elif self.B_value > self.A_value:
-      self.label = 1
-
-
-  def parametrize_rectangle(self, X, Y, WIDTH):
-    '''
-    '''
-
-    parameters = 1
-
-    #
-    # X can be variable
-    #
-    if self._variable_x:
-      X, p = Util.parameter(X, X, delta=self._delta)
-      parameters *= p
-
-    #
[…]

### 2018-03-21 EXP/ClevelandMcGill/figure1_length.py

@@ -1,225 +1,0 @@
-import cPickle as pickle
-import matplotlib.pyplot as plt
-import numpy as np
-import os
-import sys
-
-sys.path.append('../')
-from util import Util
-
-class Figure1_Length(object):
-
-  def __init__(self, variable_x=False, \
-                     variable_y=False, \
-                     variable_width=False, \
-                     constrain = True):
-
-    '''
-    '''
-    self.X = None
-    self.Y = None
-    self.LENGTH = None
-    self.WIDTH = None
-
-    self.label = None
-
-    self._delta = 20 # we use a bit larger delta here, it needs to be bigger
-    # than the max. spot size / 2. for sure
-
-    self._size = (100,100)
-    self._widths = [1, 3, 5, 7, 9, 11]
-
-    #
-    # variability switches
-    #
-    #  stage 1: var. X
-    #  stage 2: var. Y
-    #  stage 3: var. WIDTH
-    #    
-    self._variable_x = variable_x
-    self._variable_y = variable_y
-    self._variable_width = variable_width
-    self._constrain = constrain
-
-
-  def create(self, verbose=False):
-    '''
-    Create a valid datapoint with variability settings in mind.
-    '''
-
-    parameters = 1
-
-    length, p = Util.parameter(1,41)
-    parameters *= p
-
-    Y = self._size[0] / 2
-    if self._variable_y:
-      Y, p = Util.parameter(self._delta, self._size[0]-self._delta + 1)
-      parameters *= p
-
-    X = self._size[1] / 2
-    if self._variable_x:
-      X, p = Util.parameter(self._delta, self._size[1]-self._delta + 1)
-      parameters *= p
-
-    width = 1
-    if self._variable_width:
-      width = np.random.choice(self._widths)
-      parameters *= len(self._widths)
-
-    # update the label
-    self.X = X
-    self.Y = Y
-    self.LENGTH = length
-    self.WIDTH = width
-    self.label = length
-
-    if verbose:
-      print '# Parameters', parameters
-
-
-  def to_sparse(self):
-    '''
-    Convert to sparse representation
-      [X, Y, origin, spot_size], label
-    '''
-    return [self.X, self.Y, self.LENGTH, self.WIDTH], self.label
-
-
-  @staticmethod
-  def from_sparse(sparse_representation, \
-                  variable_x=False, \
-                  variable_y=False, \
-                  variable_width=False, \
-                  constrain = True):
-    '''
-    From sparse:
[…]

### 2018-03-21 EXP/ClevelandMcGill/figure1_position_common_scale.py

@@ -1,217 +1,0 @@
-import cPickle as pickle
-import matplotlib.pyplot as plt
-import numpy as np
-import os
-import sys
-
-sys.path.append('../')
-from util import Util
-
-class Figure1_Position_Common_Scale(object):
-
-  def __init__(self, variable_x=False, \
-                     variable_spot_size=False, \
-                     constrain = True):
-
-    '''
-    '''
-    self.X = None
-    self.Y = None
-    self.ORIGIN = [0, 0]
-    self.SPOT_SIZE = None
-
-    self.label = None
-
-    self._delta = 20 # we use a bit larger delta here, it needs to be bigger
-    # than the max. spot size / 2. for sure
-
-    self._size = (100,100)
-    self._spot_sizes = [1, 3, 5, 7, 9, 11]
-
-    #
-    # variability switches
-    #
-    #  stage 1: var. Y
-    #  stage 2: var. WIDTH
-    #  stage 3: var. X
-    #    
-    self._variable_x = variable_x
-    self._variable_spot_size = variable_spot_size
-    self._constrain = constrain
-
-
-  def create(self, verbose=False):
-    '''
-    Create a valid datapoint with variability settings in mind.
-    '''
-
-    parameters = 1
-
-    #
-    # Y is always variable since it encodes our value
-    #
-    Y, p = Util.parameter(self._delta, self._size[0]-self._delta + 1)
-    parameters *= p
-
-    X = self._size[1] / 2
-    if self._variable_x:
-      X, p = Util.parameter(self._delta, self._size[1]-self._delta + 1)
-      parameters *= p
-
-    spot_size = 5
-    if self._variable_spot_size:
-      spot_size = np.random.choice(self._spot_sizes)
-      parameters *= len(self._spot_sizes)
-
-    origin = [0, 0]
-
-    # update the label
-    self.X = X
-    self.Y = Y
-    self.ORIGIN = origin
-    self.SPOT_SIZE = spot_size
-    self.label = Y - self.ORIGIN[0]# label is the same as Y
-
-    if verbose:
-      print '# Parameters', parameters
-
-
-  def to_sparse(self):
-    '''
-    Convert to sparse representation
-      [X, Y, origin, spot_size], label
-    '''
-    return [self.X, self.Y, self.ORIGIN, self.SPOT_SIZE], self.label
-
-
-  @staticmethod
-  def from_sparse(sparse_representation, \
-                  variable_x=False, \
[…]

### 2018-03-21 EXP/ClevelandMcGill/figure1_position_nonaligned_scale.py

@@ -1,183 +1,0 @@
-import cPickle as pickle
-import matplotlib.pyplot as plt
-import numpy as np
-import os
-import sys
-
-sys.path.append('../')
-from util import Util
-
-from figure1_position_common_scale import Figure1_Position_Common_Scale
-
-class Figure1_Position_Nonaligned_Scale(Figure1_Position_Common_Scale):
-
-  def __init__(self, variable_y=False, \
-                     variable_x=False, \
-                     variable_spot_size=False, \
-                     variable_origin=True, \
-                     constrain = True):
-
-    self._variable_y = variable_y
-    self._variable_origin = True
-
-    super(Figure1_Position_Nonaligned_Scale, self).__init__(variable_x=variable_x, \
-                                                            variable_spot_size=variable_spot_size, \
-                                                            constrain=constrain)
-
-
-  def create(self, verbose=False):
-    '''
-    Create a valid datapoint with variability settings in mind.
-    '''
-
-    parameters = 1
-
-    #
-    # Y is always variable since it encodes our value
-    #
-    Y = self._size[0] / 2
-    if self._variable_y:
-      Y, p = Util.parameter(self._delta, self._size[0]-self._delta + 1)
-      parameters *= p
-
-    X = self._size[1] / 2
-    if self._variable_x:
-      X, p = Util.parameter(self._delta, self._size[1]-self._delta + 1)
-      parameters *= p
-
-    spot_size = 5
-    if self._variable_spot_size:
-      spot_size = np.random.choice(self._spot_sizes)
-      parameters *= len(self._spot_sizes)
-
-    origin = [0, 0]
-    if self._variable_origin:
-      origin_y, p = Util.parameter(0, self._delta/2)
-      parameters *= p
-
-      origin[0] = origin_y
-      #
-      # NOTE: we do not support origin_x changes (not applicable here!)
-      #
-
-
-    # update the label
-    self.X = X
-    self.Y = Y
-    self.ORIGIN = origin
-    self.SPOT_SIZE = spot_size
-    self.label = Y - self.ORIGIN[0]# label is the same as Y
-
-    if verbose:
-      print '# Parameters', parameters
-
-  @staticmethod
-  def from_sparse(sparse_representation, \
-                  variable_y=False, \
-                  variable_x=False, \
-                  variable_spot_size=False, \
-                  variable_origin=True, \
-                  constrain = True):
-    '''
-    From sparse:
-      [X, Y, origin, spot_size]
[…]

### 2018-03-21 EXP/bulk_experiments.py

@@ -1,204 +1,0 @@
-from sklearn.model_selection import RepeatedStratifiedKFold 
-from keras import models
-from keras import layers
-from keras import optimizers
-import keras.applications
-from keras.utils.np_utils import to_categorical
-
-import cPickle as pickle
-import numpy as np
-import os
-import sys
-import time
-
-import ClevelandMcGill as C
-
-
-EXPERIMENT = sys.argv[1] #'Figure12'
-DATASET = sys.argv[2] # 1
-FEATUREGENERATOR = sys.argv[3] #'LeNet'
-FRAMED = sys.argv[4] # 'Framed'
-
-#
-#
-#
-print 'Running', EXPERIMENT, 'on dataset', DATASET, FRAMED, 'with', FEATUREGENERATOR
-
-
-EPOCHS = 10
-FRAMED_BOOL = (FRAMED == 'Framed')
-NO_SPLITS = 10
-NO_REPEATS = 4
-BATCH_SIZE = 32
-
-FEATUREGENERATORS = {'MLP': 'Just MLP',
-                     'LeNet': 'LeNet',
-                     'VGG19': keras.applications.VGG19,
-                     'Xception': keras.applications.Xception,
-                     'DenseNet201': keras.applications.DenseNet201}
-
-
-#
-#
-# We store now a folder structure:
-#   RESULTS/Experiment/Dataset/FeatureGenerator/Framed/EPOCHS_NOSPLITS_NOREPEATS_INDEX.p
-#
-# INDEX is the index of the current run
-#
-if os.path.abspath('~').startswith('/n/'):
-  # we are on the cluster
-  PREFIX = '/n/regal/pfister_lab/PERCEPTION/'
-else:
-  PREFIX = '/home/d/Projects/CP/'
-  
-OUTPUT_DIRECTORY = os.path.join(PREFIX, 'RESULTS_NEW', EXPERIMENT, str(DATASET), FEATUREGENERATOR, FRAMED)
-if not os.path.exists(OUTPUT_DIRECTORY):
-  os.makedirs(OUTPUT_DIRECTORY)
-
-#
-# CHECK IF THIS EXPERIMENT WAS COMPLETED ALREADY!
-#
-if len(os.listdir(OUTPUT_DIRECTORY)) >= (NO_SPLITS * NO_REPEATS * 2):
-  # this was done!
-  print 'Already done!'
-  sys.exit(0)
-
-#
-#
-# LOAD DATA
-#
-#
-if EXPERIMENT == 'Figure12':
[…]

### 2018-03-21 EXP/classifier.py

@@ -1,32 +1,0 @@
-class Classifier:
-
-  def __init__(self):
-    '''
-    '''
-
-    self.name = 'N/A'
-
-  def setup(self):
-    '''
-    '''
-    pass
-
-  def show(self):
-    '''
-    '''
-    pass
-
-  def train(self):
-    '''
-    '''
-    pass
-
-  def test(self):
-    '''
-    '''
-    pass
-
-  def store(self):
-    '''
-    '''
-    pass

### 2018-03-21 EXP/figure1_experiments.py

@@ -1,257 +1,0 @@
-from sklearn.model_selection import RepeatedStratifiedKFold 
-from sklearn.ensemble import RandomForestRegressor
-
-from keras import models
-from keras import layers
-from keras import optimizers
-import keras.applications
-from keras import backend as K
-from keras.utils.np_utils import to_categorical
-
-import cPickle as pickle
-import numpy as np
-import os
-import sys
-import time
-
-import ClevelandMcGill as C
-
-
-EXPERIMENT = sys.argv[1] # Figure1_Position_Common_Scale
-DATASET = int(sys.argv[2]) # 1
-CLASSIFIER = sys.argv[3] # 'LeNet'
-NORMALIZATION_FACTOR = float(sys.argv[4]) # e.g. 80
-
-#
-#
-#
-print 'Running', EXPERIMENT, 'on dataset', DATASET, 'with', CLASSIFIER
-
-EPOCHS = 10
-NO_SPLITS = 10
-NO_REPEATS = 4
-BATCH_SIZE = 32
-
-
-if os.path.abspath('~').startswith('/n/'):
-  # we are on the cluster
-  PREFIX = '/n/regal/pfister_lab/PERCEPTION/'
-else:
-  PREFIX = '/home/d/PERCEPTION/'
-
-
-#
-# DATA LOADING
-#
-
-DATAPATH = PREFIX + 'DATA/' + EXPERIMENT + '/'
-
-with open(DATAPATH + 'labels_'+str(DATASET)+'.p', 'r') as f:
-  y = pickle.load(f)
-
-if CLASSIFIER == 'RF':
-  '''
-  Load datapoints.
-  '''
-  with open(DATAPATH + 'datapoints_'+str(DATASET)+'.p', 'r') as f:
-    X = pickle.load(f)
-
-  # convert to numpy
-  X_np = np.zeros((len(X), 3), dtype=np.uint8)
-  for i,x in enumerate(X):
-    X_np[i] = [x[0], x[1], x[3]]
-
-  X = X_np
-
-  print 'Loaded datapoints!'
-
-elif CLASSIFIER == 'RF_Image' or CLASSIFIER == 'MLP' or CLASSIFIER == 'LeNet':
-  '''
-  Load images (1-channel). (This is already raveled!)
-  '''
-  X = np.load(DATAPATH + 'images_'+str(DATASET)+'.npy', mmap_mode='r')
-
[…]

### 2018-03-21 EXP/run_bar_framed_rectangle.py

@@ -1,0 +1,346 @@
+from keras import models
+from keras import layers
+from keras import optimizers
+import keras.applications
+import keras.callbacks
+from keras import backend as K
+from keras.utils.np_utils import to_categorical
+import sklearn.metrics
+
+import cPickle as pickle
+import numpy as np
+import os
+import sys
+import time
+
+import ClevelandMcGill as C
+
+
+EXPERIMENT = sys.argv[1] # f.e. Angle
+DATASET = int(sys.argv[2]) # How many TRUE flags to pass to data generation
+CLASSIFIER = sys.argv[3] # 'LeNet'
+NOISE = sys.argv[4] # True
+JOB_INDEX = int(sys.argv[5])
+
+#
+#
+#
+print 'Running', EXPERIMENT, 'on dataset', DATASET, 'with', CLASSIFIER, 'Noise:', NOISE, 'Job Index', JOB_INDEX
+
+#
+#
+# PROCESS SOME FLAGS
+#
+#
+SUFFIX = '.'
+if NOISE == 'True':
+  NOISE = True
+  SUFFIX = '_noise.'
+else:
+  NOISE = False
+
+DATATYPE = eval(EXPERIMENT)
+
+FLAGS = [False] * 10 # never more than 10 flags
+for f in range(DATASET):
+  FLAGS[f] = True
+
+
+if os.path.abspath('~').startswith('/n/'):
+  # we are on the cluster
+  PREFIX = '/n/regal/pfister_lab/PERCEPTION/'
+else:
+  PREFIX = '/home/d/PERCEPTION/'
+RESULTS_DIR = PREFIX + 'RESULTS/'
+
+OUTPUT_DIR = RESULTS_DIR + EXPERIMENT + '/' + str(DATASET) + '/' + CLASSIFIER + '/'
+if not os.path.exists(OUTPUT_DIR):
+  # here can be a race condition
+  try:
+    os.makedirs(OUTPUT_DIR)
+  except:
+    print 'Race condition!', os.path.exists(OUTPUT_DIR)
+
+STATSFILE = OUTPUT_DIR + str(JOB_INDEX).zfill(2) + SUFFIX + 'p'
+MODELFILE = OUTPUT_DIR + str(JOB_INDEX).zfill(2) + SUFFIX + 'h5'
+
+print 'Working in', OUTPUT_DIR
+print 'Storing', STATSFILE
+print 'Storing', MODELFILE
[…]

### 2018-03-21 EXP/run_experiments.sh

@@ -1,42 +1,0 @@
-#!/bin/bash
-
-python bulk_experiments.py Figure12 1 MLP Framed
-python bulk_experiments.py Figure12 1 MLP NotFramed
-python bulk_experiments.py Figure12 1 LeNet Framed
-python bulk_experiments.py Figure12 1 LeNet NotFramed
-python bulk_experiments.py Figure12 1 VGG19 Framed
-python bulk_experiments.py Figure12 1 VGG19 NotFramed
-python bulk_experiments.py Figure12 1 Xception Framed
-python bulk_experiments.py Figure12 1 Xception NotFramed
-#python bulk_experiments.py Figure12 1 DenseNet201 Framed
-#python bulk_experiments.py Figure12 1 DenseNet201 NotFramed
-python bulk_experiments.py Figure12 2 MLP Framed
-python bulk_experiments.py Figure12 2 MLP NotFramed
-python bulk_experiments.py Figure12 2 LeNet Framed
-python bulk_experiments.py Figure12 2 LeNet NotFramed
-python bulk_experiments.py Figure12 2 VGG19 Framed
-python bulk_experiments.py Figure12 2 VGG19 NotFramed
-python bulk_experiments.py Figure12 2 Xception Framed
-python bulk_experiments.py Figure12 2 Xception NotFramed
-#python bulk_experiments.py Figure12 2 DenseNet201 Framed
-#python bulk_experiments.py Figure12 2 DenseNet201 NotFramed
-python bulk_experiments.py Figure12 3 MLP Framed
-python bulk_experiments.py Figure12 3 MLP NotFramed
-python bulk_experiments.py Figure12 3 LeNet Framed
-python bulk_experiments.py Figure12 3 LeNet NotFramed
-python bulk_experiments.py Figure12 3 VGG19 Framed
-python bulk_experiments.py Figure12 3 VGG19 NotFramed
-python bulk_experiments.py Figure12 3 Xception Framed
-python bulk_experiments.py Figure12 3 Xception NotFramed
-#python bulk_experiments.py Figure12 3 DenseNet201 Framed
-#python bulk_experiments.py Figure12 3 DenseNet201 NotFramed
-python bulk_experiments.py Figure12 4 MLP Framed
-python bulk_experiments.py Figure12 4 MLP NotFramed
-python bulk_experiments.py Figure12 4 LeNet Framed
-python bulk_experiments.py Figure12 4 LeNet NotFramed
[…]

### 2018-03-21 EXP/vgg19.py

@@ -1,205 +1,0 @@
-from keras import models
-from keras import layers
-from keras import optimizers
-from keras.utils.np_utils import to_categorical
-import keras.applications.vgg19
-
-import os
-
-from sklearn.metrics import classification_report
-from sklearn.preprocessing import binarize
-from sklearn.model_selection import RepeatedStratifiedKFold 
-from sklearn.model_selection import cross_val_score
-from keras.wrappers.scikit_learn import KerasClassifier
-
-from classifier import Classifier
-
-class VGG19(Classifier):
-
-  def __init__(self):
-    '''
-    '''
-
-    Classifier.__init__(self)
-
-    self.name = 'VGG19 (Imagenet pre-trained) + MLP'
-
-    self._VGG19 = None
-    self._MLP = None
-
-    self._VGG19_features_for_kfold = None
-    self._VGG19_features_calculated = False
-
-  #
-  # setup MLP to fit in sklearn framework
-  #
-  @staticmethod
-  def create_mlp():
-
-    MLP = models.Sequential()
-    MLP.add(layers.Dense(256, activation='relu', input_dim=3 * 3 * 512))
-    MLP.add(layers.Dropout(0.5))
-    MLP.add(layers.Dense(2, activation='softmax'))
-
-    MLP.compile(loss='categorical_crossentropy', optimizer='adam', metrics=['acc'])    
-
-    return MLP
-
-
-  def setup(self, verbose=True):
-    '''
-    '''
-
-    Classifier.setup(self)
-
-    #
-    # setup VGG
-    #
-    self._VGG19 = keras.applications.vgg19.VGG19(include_top=False, weights='imagenet', input_shape=(100,100,3))
-
-    self._MLP = VGG19.create_mlp()
-
-    print 'Configured VGG19 (imagenet weights) and a MLP..'
-
-
-  def show(self):
-    '''
-    '''
-    
-    Classifier.show(self)
-
-    self._VGG19.summary()
-
-    if self._MLP:
-      self._MLP.summary()
-
-
-  def train(self, X_train, y_train, epochs=1, batch_size=32, validation_split=.25, verbose=True):
-    '''
-    '''
-
-    Classifier.train(self)
-
-    #
-    # We generate features using VGG19 (pre-trained)
-    #
-    if verbose:
-      print 'Generating features for X_train using VGG19..'
-    features_train = self._VGG19.predict(X_train, verbose=verbose)
-
-    #
[…]

### 2018-03-21 _LEGACY/bulk_experiments.py

@@ -1,0 +1,204 @@
+from sklearn.model_selection import RepeatedStratifiedKFold 
+from keras import models
+from keras import layers
+from keras import optimizers
+import keras.applications
+from keras.utils.np_utils import to_categorical
+
+import cPickle as pickle
+import numpy as np
+import os
+import sys
+import time
+
+import ClevelandMcGill as C
+
+
+EXPERIMENT = sys.argv[1] #'Figure12'
+DATASET = sys.argv[2] # 1
+FEATUREGENERATOR = sys.argv[3] #'LeNet'
+FRAMED = sys.argv[4] # 'Framed'
+
+#
+#
+#
+print 'Running', EXPERIMENT, 'on dataset', DATASET, FRAMED, 'with', FEATUREGENERATOR
+
+
+EPOCHS = 10
+FRAMED_BOOL = (FRAMED == 'Framed')
+NO_SPLITS = 10
+NO_REPEATS = 4
+BATCH_SIZE = 32
+
+FEATUREGENERATORS = {'MLP': 'Just MLP',
+                     'LeNet': 'LeNet',
+                     'VGG19': keras.applications.VGG19,
+                     'Xception': keras.applications.Xception,
+                     'DenseNet201': keras.applications.DenseNet201}
+
+
+#
+#
+# We store now a folder structure:
+#   RESULTS/Experiment/Dataset/FeatureGenerator/Framed/EPOCHS_NOSPLITS_NOREPEATS_INDEX.p
+#
+# INDEX is the index of the current run
+#
+if os.path.abspath('~').startswith('/n/'):
+  # we are on the cluster
+  PREFIX = '/n/regal/pfister_lab/PERCEPTION/'
+else:
+  PREFIX = '/home/d/Projects/CP/'
+  
+OUTPUT_DIRECTORY = os.path.join(PREFIX, 'RESULTS_NEW', EXPERIMENT, str(DATASET), FEATUREGENERATOR, FRAMED)
+if not os.path.exists(OUTPUT_DIRECTORY):
+  os.makedirs(OUTPUT_DIRECTORY)
+
+#
+# CHECK IF THIS EXPERIMENT WAS COMPLETED ALREADY!
+#
+if len(os.listdir(OUTPUT_DIRECTORY)) >= (NO_SPLITS * NO_REPEATS * 2):
+  # this was done!
+  print 'Already done!'
+  sys.exit(0)
+
+#
+#
+# LOAD DATA
+#
+#
+if EXPERIMENT == 'Figure12':
[…]

### 2018-03-21 _LEGACY/figure12.py

@@ -1,0 +1,321 @@
+import cPickle as pickle
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+import sys
+
+sys.path.append('../')
+from util import Util
+
+class Figure12:
+
+  def __init__(self, variable_y=False, \
+                     variable_width=False, \
+                     variable_x=False, \
+                     constrain=True):
+    '''
+  	'''
+
+    # the flag for the visual aid
+    self.framed = False
+
+    self.A_x = None
+    self.A_y = None
+    self.A_value = None
+    self.A_width = None
+
+    self.B_x = None
+    self.B_y = None
+    self.B_value = None
+    self.B_width = None
+
+    self.label = None
+
+    # measured from Cleveland McGill paper via pixel ruler
+    self._size = (100,100)
+    self._max_height = 65
+    self._default_width = 20
+
+    self._delta = 5 # only if we constrain the differences between A and B
+    # this enables cases which are difficult for the human eye
+
+    #
+    # variability switches
+    #
+    #  stage 1: var. Y
+    #  stage 2: var. WIDTH
+    #  stage 3: var. X
+    #
+    self._variable_y = variable_y
+    self._variable_width = variable_width
+    self._variable_x = variable_x
+    self._constrain = constrain
+
+
+  def create(self, verbose=False):
+    '''
+    Create a valid datapoint which fulfills constrains
+    and the variability settings.
+    '''
+    self.create_(verbose)
+
+    while (self.A_value == self.B_value): #or (self._variable_y and self._constrain and (np.abs(self.A_y - self.B_y) < 10)):
+      # constrain means that the difference between the y position has to be larger than 10 pixel
+      # this only enables that the human eye can't really perceive the difference
+      # we always eliminate cases where the two bars are of the same value since we can not label these as 0 or 1
+      # (it is really not part of the experiment)
+      self.create_(verbose)
+
+    # update the label
+    if self.A_value > self.B_value:
+      self.label = 0
+    elif self.B_value > self.A_value:
+      self.label = 1
+
+
+  def parametrize_rectangle(self, X, Y, WIDTH):
+    '''
+    '''
+
+    parameters = 1
+
+    #
+    # X can be variable
+    #
+    if self._variable_x:
+      X, p = Util.parameter(X, X, delta=self._delta)
+      parameters *= p
+
+    #
[…]

### 2018-03-21 _LEGACY/figure1_experiments.py

@@ -1,0 +1,257 @@
+from sklearn.model_selection import RepeatedStratifiedKFold 
+from sklearn.ensemble import RandomForestRegressor
+
+from keras import models
+from keras import layers
+from keras import optimizers
+import keras.applications
+from keras import backend as K
+from keras.utils.np_utils import to_categorical
+
+import cPickle as pickle
+import numpy as np
+import os
+import sys
+import time
+
+import ClevelandMcGill as C
+
+
+EXPERIMENT = sys.argv[1] # Figure1_Position_Common_Scale
+DATASET = int(sys.argv[2]) # 1
+CLASSIFIER = sys.argv[3] # 'LeNet'
+NORMALIZATION_FACTOR = float(sys.argv[4]) # e.g. 80
+
+#
+#
+#
+print 'Running', EXPERIMENT, 'on dataset', DATASET, 'with', CLASSIFIER
+
+EPOCHS = 10
+NO_SPLITS = 10
+NO_REPEATS = 4
+BATCH_SIZE = 32
+
+
+if os.path.abspath('~').startswith('/n/'):
+  # we are on the cluster
+  PREFIX = '/n/regal/pfister_lab/PERCEPTION/'
+else:
+  PREFIX = '/home/d/PERCEPTION/'
+
+
+#
+# DATA LOADING
+#
+
+DATAPATH = PREFIX + 'DATA/' + EXPERIMENT + '/'
+
+with open(DATAPATH + 'labels_'+str(DATASET)+'.p', 'r') as f:
+  y = pickle.load(f)
+
+if CLASSIFIER == 'RF':
+  '''
+  Load datapoints.
+  '''
+  with open(DATAPATH + 'datapoints_'+str(DATASET)+'.p', 'r') as f:
+    X = pickle.load(f)
+
+  # convert to numpy
+  X_np = np.zeros((len(X), 3), dtype=np.uint8)
+  for i,x in enumerate(X):
+    X_np[i] = [x[0], x[1], x[3]]
+
+  X = X_np
+
+  print 'Loaded datapoints!'
+
+elif CLASSIFIER == 'RF_Image' or CLASSIFIER == 'MLP' or CLASSIFIER == 'LeNet':
+  '''
+  Load images (1-channel). (This is already raveled!)
+  '''
+  X = np.load(DATAPATH + 'images_'+str(DATASET)+'.npy', mmap_mode='r')
+
[…]

### 2018-03-21 _LEGACY/figure1_position_common_scale.py

@@ -1,0 +1,217 @@
+import cPickle as pickle
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+import sys
+
+sys.path.append('../')
+from util import Util
+
+class Figure1_Position_Common_Scale(object):
+
+  def __init__(self, variable_x=False, \
+                     variable_spot_size=False, \
+                     constrain = True):
+
+    '''
+    '''
+    self.X = None
+    self.Y = None
+    self.ORIGIN = [0, 0]
+    self.SPOT_SIZE = None
+
+    self.label = None
+
+    self._delta = 20 # we use a bit larger delta here, it needs to be bigger
+    # than the max. spot size / 2. for sure
+
+    self._size = (100,100)
+    self._spot_sizes = [1, 3, 5, 7, 9, 11]
+
+    #
+    # variability switches
+    #
+    #  stage 1: var. Y
+    #  stage 2: var. WIDTH
+    #  stage 3: var. X
+    #    
+    self._variable_x = variable_x
+    self._variable_spot_size = variable_spot_size
+    self._constrain = constrain
+
+
+  def create(self, verbose=False):
+    '''
+    Create a valid datapoint with variability settings in mind.
+    '''
+
+    parameters = 1
+
+    #
+    # Y is always variable since it encodes our value
+    #
+    Y, p = Util.parameter(self._delta, self._size[0]-self._delta + 1)
+    parameters *= p
+
+    X = self._size[1] / 2
+    if self._variable_x:
+      X, p = Util.parameter(self._delta, self._size[1]-self._delta + 1)
+      parameters *= p
+
+    spot_size = 5
+    if self._variable_spot_size:
+      spot_size = np.random.choice(self._spot_sizes)
+      parameters *= len(self._spot_sizes)
+
+    origin = [0, 0]
+
+    # update the label
+    self.X = X
+    self.Y = Y
+    self.ORIGIN = origin
+    self.SPOT_SIZE = spot_size
+    self.label = Y - self.ORIGIN[0]# label is the same as Y
+
+    if verbose:
+      print '# Parameters', parameters
+
+
+  def to_sparse(self):
+    '''
+    Convert to sparse representation
+      [X, Y, origin, spot_size], label
+    '''
+    return [self.X, self.Y, self.ORIGIN, self.SPOT_SIZE], self.label
+
+
+  @staticmethod
+  def from_sparse(sparse_representation, \
+                  variable_x=False, \
[…]

### 2018-03-21 _LEGACY/figure1_position_nonaligned_scale.py

@@ -1,0 +1,183 @@
+import cPickle as pickle
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+import sys
+
+sys.path.append('../')
+from util import Util
+
+from figure1_position_common_scale import Figure1_Position_Common_Scale
+
+class Figure1_Position_Nonaligned_Scale(Figure1_Position_Common_Scale):
+
+  def __init__(self, variable_y=False, \
+                     variable_x=False, \
+                     variable_spot_size=False, \
+                     variable_origin=True, \
+                     constrain = True):
+
+    self._variable_y = variable_y
+    self._variable_origin = True
+
+    super(Figure1_Position_Nonaligned_Scale, self).__init__(variable_x=variable_x, \
+                                                            variable_spot_size=variable_spot_size, \
+                                                            constrain=constrain)
+
+
+  def create(self, verbose=False):
+    '''
+    Create a valid datapoint with variability settings in mind.
+    '''
+
+    parameters = 1
+
+    #
+    # Y is always variable since it encodes our value
+    #
+    Y = self._size[0] / 2
+    if self._variable_y:
+      Y, p = Util.parameter(self._delta, self._size[0]-self._delta + 1)
+      parameters *= p
+
+    X = self._size[1] / 2
+    if self._variable_x:
+      X, p = Util.parameter(self._delta, self._size[1]-self._delta + 1)
+      parameters *= p
+
+    spot_size = 5
+    if self._variable_spot_size:
+      spot_size = np.random.choice(self._spot_sizes)
+      parameters *= len(self._spot_sizes)
+
+    origin = [0, 0]
+    if self._variable_origin:
+      origin_y, p = Util.parameter(0, self._delta/2)
+      parameters *= p
+
+      origin[0] = origin_y
+      #
+      # NOTE: we do not support origin_x changes (not applicable here!)
+      #
+
+
+    # update the label
+    self.X = X
+    self.Y = Y
+    self.ORIGIN = origin
+    self.SPOT_SIZE = spot_size
+    self.label = Y - self.ORIGIN[0]# label is the same as Y
+
+    if verbose:
+      print '# Parameters', parameters
+
+  @staticmethod
+  def from_sparse(sparse_representation, \
+                  variable_y=False, \
+                  variable_x=False, \
+                  variable_spot_size=False, \
+                  variable_origin=True, \
+                  constrain = True):
+    '''
+    From sparse:
+      [X, Y, origin, spot_size]
[…]

### 2018-03-21 _LEGACY/run_experiments.sh

@@ -1,0 +1,42 @@
+#!/bin/bash
+
+python bulk_experiments.py Figure12 1 MLP Framed
+python bulk_experiments.py Figure12 1 MLP NotFramed
+python bulk_experiments.py Figure12 1 LeNet Framed
+python bulk_experiments.py Figure12 1 LeNet NotFramed
+python bulk_experiments.py Figure12 1 VGG19 Framed
+python bulk_experiments.py Figure12 1 VGG19 NotFramed
+python bulk_experiments.py Figure12 1 Xception Framed
+python bulk_experiments.py Figure12 1 Xception NotFramed
+#python bulk_experiments.py Figure12 1 DenseNet201 Framed
+#python bulk_experiments.py Figure12 1 DenseNet201 NotFramed
+python bulk_experiments.py Figure12 2 MLP Framed
+python bulk_experiments.py Figure12 2 MLP NotFramed
+python bulk_experiments.py Figure12 2 LeNet Framed
+python bulk_experiments.py Figure12 2 LeNet NotFramed
+python bulk_experiments.py Figure12 2 VGG19 Framed
+python bulk_experiments.py Figure12 2 VGG19 NotFramed
+python bulk_experiments.py Figure12 2 Xception Framed
+python bulk_experiments.py Figure12 2 Xception NotFramed
+#python bulk_experiments.py Figure12 2 DenseNet201 Framed
+#python bulk_experiments.py Figure12 2 DenseNet201 NotFramed
+python bulk_experiments.py Figure12 3 MLP Framed
+python bulk_experiments.py Figure12 3 MLP NotFramed
+python bulk_experiments.py Figure12 3 LeNet Framed
+python bulk_experiments.py Figure12 3 LeNet NotFramed
+python bulk_experiments.py Figure12 3 VGG19 Framed
+python bulk_experiments.py Figure12 3 VGG19 NotFramed
+python bulk_experiments.py Figure12 3 Xception Framed
+python bulk_experiments.py Figure12 3 Xception NotFramed
+#python bulk_experiments.py Figure12 3 DenseNet201 Framed
+#python bulk_experiments.py Figure12 3 DenseNet201 NotFramed
+python bulk_experiments.py Figure12 4 MLP Framed
+python bulk_experiments.py Figure12 4 MLP NotFramed
+python bulk_experiments.py Figure12 4 LeNet Framed
+python bulk_experiments.py Figure12 4 LeNet NotFramed
[…]

### 2018-03-21 EXP/ClevelandMcGill/figure1.py
Switch paper title and start new experiments.

@@ -1,3 +1,4 @@
+import math
 import numpy as np
 import os
 import skimage.draw
@@ -443,6 +444,8 @@ class Figure1:
     DEPTH = np.random.randint(1, DOF+1)
     parameters *= DOF
 
+    print 'DEPTH', DEPTH
+
     Y_RANGE = (Figure1.DELTA_MAX-20, Figure1.DELTA_MAX)
     X_RANGE = (0, Figure1.DELTA_MIN*2)
 
@@ -459,7 +462,11 @@ class Figure1:
     WIDTH = 60
     if var_width:
       WIDTH, p = Util.parameter(20, 60)
-      parameters *= p
+      # we only support even width
+      WIDTH = int(math.ceil(WIDTH / 2.) * 2)
+      parameters *= (p/2) # only half of the parameters 
+
+    print 'WIDTH', WIDTH
 
     start = (Y, X)
     mid = (DEPTH, X+WIDTH/2)
@@ -481,7 +488,7 @@ class Figure1:
 
     sparse = [Y, X, DEPTH, WIDTH]
 
-    label = curvature
+    label = np.round(curvature, 3)
 
     return sparse, img, label, parameters
 

### 2018-03-21 EXP/ClevelandMcGill/figure12.py

@@ -1,0 +1,27 @@
+import numpy as np
+import os
+import skimage.draw
+import sys
+
+sys.path.append('../')
+from util import Util
+
+class Figure12:
+
+  SIZE = (100, 100)
+
+
+  @staticmethod
+  def generate_datapoint():
+    '''
+    '''
+
+
+    return data, labels
+
+
+  @staticmethod
+  def data_to_image(data, framed=True):
+    '''
+    '''
+    image = np.zeros((100,100), dtype=np.bool)

### 2018-03-22 EXP/ClevelandMcGill/figure1.py
More debugging and tests.

@@ -64,7 +64,7 @@ class Figure1:
 
 
   @staticmethod
-  def position_non_aligned_scale(flags=[False, False, False]):
+  def position_non_aligned_scale(flags=[False, False, False], pre_offset=None):
 
     var_y = flags[0]
     var_x = flags[1]
@@ -75,9 +75,12 @@ class Figure1:
     label = None
     parameters = 1
 
-    OFFSET, p = Util.parameter(1, 11)
+    OFFSET, p = Util.parameter(-9, 11)
     # print 'OFFSET', OFFSET
     parameters *= p
+
+    if pre_offset:
+      OFFSET = pre_offset
 
     Y_RANGE = (Figure1.DELTA_MIN-OFFSET, Figure1.DELTA_MAX-OFFSET)
     X_RANGE = (Figure1.DELTA_MIN, Figure1.DELTA_MAX)

### 2018-03-22 EXP/run_regression.py

@@ -73,7 +73,7 @@ if os.path.exists(STATSFILE) and os.path.exists(MODELFILE):
   sys.exit(0)
 
 
-#
+
 #
 # DATA GENERATION
 #
@@ -81,6 +81,18 @@ if os.path.exists(STATSFILE) and os.path.exists(MODELFILE):
 train_target = 60000
 val_target = 20000
 test_target = 20000
+
+# get global min and max
+global_min = np.inf
+global_max = -np.inf
+for N in range(train_target+val_target+test_target):
+  
+  sparse, image, label, parameters = DATATYPE(FLAGS)
+  
+  global_min = min(label, global_min)
+  global_max = max(label, global_max)
+# end of global min max
+
 
 X_train = np.zeros((train_target, 100, 100), dtype=np.float32)
 y_train = np.zeros((train_target), dtype=np.float32)
@@ -96,6 +108,9 @@ test_counter = 0
 
 t0 = time.time()
 
+min_label = np.inf
+max_label = -np.inf
+
 all_counter = 0
 while train_counter < train_target or val_counter < val_target or test_counter < test_target:
   
@@ -109,8 +124,18 @@ while train_counter < train_target or val_counter < val_target or test_counter <
   # we need float
   image = image.astype(np.float32)
   
-  pot = np.random.choice(3, p=([.6,.2,.2]))
-
+  pot = np.random.choice(3)#, p=([.6,.2,.2]))
+  
+  #
+  #
+  # special: to allow normalizations, we make sure the global min
+  # and the global max for sure go into pot 0
+  # this biases in a very slight way towards the mean with 2/n
+  # this is ok.
+  #
+  if label == global_min or label == global_max:
+    pot = 0 # for sure training
+  
   if pot == 0 and train_counter < train_target:
     # a training candidate
     if label in y_val or label in y_test:

### 2018-03-23 EXP/ClevelandMcGill/figure1.py
Some updates after debugging!

@@ -447,7 +447,7 @@ class Figure1:
     DEPTH = np.random.randint(1, DOF+1)
     parameters *= DOF
 
-    print 'DEPTH', DEPTH
+    # print 'DEPTH', DEPTH
 
     Y_RANGE = (Figure1.DELTA_MAX-20, Figure1.DELTA_MAX)
     X_RANGE = (0, Figure1.DELTA_MIN*2)
@@ -469,7 +469,7 @@ class Figure1:
       WIDTH = int(math.ceil(WIDTH / 2.) * 2)
       parameters *= (p/2) # only half of the parameters 
 
-    print 'WIDTH', WIDTH
+    # print 'WIDTH', WIDTH
 
     start = (Y, X)
     mid = (DEPTH, X+WIDTH/2)

### 2018-03-23 EXP/ClevelandMcGill/figure12.py

@@ -10,18 +10,89 @@ class Figure12:
 
   SIZE = (100, 100)
 
-
   @staticmethod
   def generate_datapoint():
     '''
     '''
+    # ok james, what do you think of this:
+    # we stick to 100x100 images
+    # then we leave a delta of 20 to move the bars in Y
+    # the bars  then have a max value of 60
+    # this projects them to the range of 10..70 or 30..90 in the extreme cases which leaves the 
+    # 10 pixel border around which makes sure our 5x5 lenet picks up all pixels
 
+    # then, of the 60 bar value we say max 20% == 12 pixels white and max 80% black == 48 pixels.
 
-    return data, labels
+    # then the least JND is for white: 1/12 == 8% 
+    # and for black 1/48 == 20% and 1-p for the best case
+
+    # do you think thats enough?
+
+    # james: this is just bloody fantastic
+    # (not sure if he really said anything like that but i imagine he did LOL)
+
+    parameters = 1
+
+    # VALUE A
+    value_A, p = Util.parameter(48, 60)
+    parameters *= p
+
+    # VALUE B
+    value_B = value_A
+    while value_B == value_A:
+      value_B, p = Util.parameter(48, 60)
+    parameters *= p
+
+    # POS A
+    x_A = 20
+    y_A, p = Util.parameter(10, 30)
+    parameters *= p
+
+    # POS B
+    x_B = 60
+    y_B = y_A
+    while y_B == y_A:
+      y_B, p = Util.parameter(10, 30)
+    parameters *= p
+
+    data = [x_A, y_A, value_A, x_B, y_B, value_B]
+
+    labels = [value_A, value_B]
+
+    return data, labels, parameters
 
 
   @staticmethod
   def data_to_image(data, framed=True):
     '''
     '''
-    image = np.zeros((100,100), dtype=np.bool)
+    image = np.zeros((100,100), dtype=np.uint8)
+    
+    x_A, y_A, value_A, x_B, y_B, value_B = data
+
+    barwidth = 20
+    max_height = 60
+
+
+    image[y_A:y_A+max_height,x_A:x_A+1] = 2
+    image[y_A:y_A+max_height,x_A+barwidth:x_A+barwidth+1] = 2
+    image[y_A:y_A+1,x_A:x_A+barwidth] = 2
[…]

### 2018-03-23 EXP/run_position_angle.py

@@ -109,7 +109,7 @@ while train_counter < train_target or val_counter < val_target or test_counter <
 
   
   
-  pot = np.random.choice(3, p=([.6,.2,.2]))
+  pot = np.random.choice(3)
 
   if pot == 0 and train_counter < train_target:
     # a training candidate

### 2018-03-23 EXP/run_cross_network.py
Automate cross network testing for all experiments.

@@ -1,0 +1,190 @@
+import matploblib
+matplotlib.use('agg')
+from matplotlib.pyplot import imshow
+import matplotlib.pyplot as plt    
+
+import os, glob
+import cPickle as pickle
+
+from keras.models import load_model
+from keras import backend as K
+import sklearn.metrics
+import keras.applications
+
+import ClevelandMcGill as C
+
+from util import Util
+
+
+EXPERIMENT = sys.argv[1]#'C.Figure1.curvature'
+CLASSIFIER = sys.argv[2]#'LeNet'
+NETWORK_INDEX = sys.argv[3]#4
+
+
+DATATYPE = eval(EXPERIMENT)
+
+
+RESULTS_DIR = '/n/regal/pfister_lab/PERCEPTION/RESULTS/' + EXPERIMENT
+OUTPUT_DIR = '/n/regal/pfister_lab/PERCEPTION/CROSSNETWORK/' + EXPERIMENT +'/'
+if not os.path.exists(OUTPUT_DIR):
+  # here can be a race condition
+  try:
+    os.makedirs(OUTPUT_DIR)
+  except:
+    print 'Race condition!', os.path.exists(OUTPUT_DIR)
+
+
+
+
+
+all_labels = {'C.Figure1.position_common_scale': ['Position Y', '+ Position X', '+ Spotsize'], \
+              'C.Figure1.position_non_aligned_scale': ['Scale', '+ Position Y', '+ Position X', '+ Spotsize'],\
+              'C.Figure1.length': ['Length', '+ Position Y', '+ Position X', '+ Width'], \
+              'C.Figure1.direction': ['Direction', '+ Position Y', '+ Position X'], \
+              'C.Figure1.angle': ['Angle', '+ Position Y', '+ Position X'], \
+              'C.Figure1.area': ['Area', '+ Position Y', '+ Position X'], \
+              'C.Figure1.volume': ['Volume', '+ Position Y', '+ Position X'], \
+              'C.Figure1.curvature': ['Curvature', '+ Position Y', '+ Position X', '+ Width'], \
+              'C.Figure1.shading': ['Shading', '+ Position Y', '+ Position X']
+              }
+DATASETS = len(all_labels[EXPERIMENT])
+
+
+networks = []
+
+for x in range(DATASETS):
+    networks.append(load_model(RESULTS_DIR+str(x)+'/'+CLASSIFIER+'/'+str(NETWORK_INDEX).zfill(2)+'_noise.h5'))
+
+
+
+
+
+# generate images
[…]

### 2018-03-23 EXP/run_cross_network.py
Store cross network images in supplemental.

@@ -181,7 +181,7 @@ def plot_cross_network(cm, classes, title, cmap=plt.cm.Blues, c_name='', e_name=
     plt.tight_layout()
     plt.ylabel('Dataset')
     plt.xlabel(c_name+' Training')
-    plt.savefig('../PAPER/gfx/cn_'+str(e_name)+'_'+str(c_name)+'_'+str(c_index)+'.pdf', bbox_inches='tight', pad_inches=0)
+    plt.savefig('../PAPER/SUPPLEMENTAL/gfx/cn_'+str(e_name)+'_'+str(c_name)+'_'+str(c_index)+'.pdf', bbox_inches='tight', pad_inches=0)
     
 plt.figure()
 plot_cross_network(np.array(results), classes=all_labels[EXPERIMENT],

### 2018-03-23 EXP/run_cross_network.py
Fix typo.

@@ -1,4 +1,4 @@
-import matploblib
+import matplotlib
 matplotlib.use('agg')
 from matplotlib.pyplot import imshow
 import matplotlib.pyplot as plt    

### 2018-03-23 EXP/ClevelandMcGill/figure12.py
Update.

@@ -61,6 +61,18 @@ class Figure12:
 
     return data, labels, parameters
 
+  @staticmethod 
+  def data_to_bars(data):
+    '''
+    '''
+    return Figure12.data_to_image(data, framed=False)
+
+  @staticmethod 
+  def data_to_framed_rectangles(data):
+    '''
+    '''
+    return Figure12.data_to_image(data, framed=True)
+
 
   @staticmethod
   def data_to_image(data, framed=True):

### 2018-03-23 EXP/run_cross_network.py

@@ -3,7 +3,7 @@ matplotlib.use('agg')
 from matplotlib.pyplot import imshow
 import matplotlib.pyplot as plt    
 
-import os, glob
+import os, glob, sys
 import cPickle as pickle
 
 from keras.models import load_model

### 2018-03-23 EXP/run_cross_network.py
And another bugfix!

@@ -24,7 +24,7 @@ NETWORK_INDEX = sys.argv[3]#4
 DATATYPE = eval(EXPERIMENT)
 
 
-RESULTS_DIR = '/n/regal/pfister_lab/PERCEPTION/RESULTS/' + EXPERIMENT
+RESULTS_DIR = '/n/regal/pfister_lab/PERCEPTION/RESULTS/' + EXPERIMENT + '/'
 OUTPUT_DIR = '/n/regal/pfister_lab/PERCEPTION/CROSSNETWORK/' + EXPERIMENT +'/'
 if not os.path.exists(OUTPUT_DIR):
   # here can be a race condition

### 2018-03-23 EXP/ClevelandMcGill/figure12.py
Last bug, I swear!

@@ -38,9 +38,9 @@ class Figure12:
     parameters *= p
 
     # VALUE B
-    value_B = value_A
-    while value_B == value_A:
-      value_B, p = Util.parameter(48, 60)
+    # value_B = value_A
+    # while value_B == value_A:
+    value_B, p = Util.parameter(48, 60)
     parameters *= p
 
     # POS A
@@ -50,9 +50,9 @@ class Figure12:
 
     # POS B
     x_B = 60
-    y_B = y_A
-    while y_B == y_A:
-      y_B, p = Util.parameter(10, 30)
+    # y_B = y_A
+    # while y_B == y_A:
+    y_B, p = Util.parameter(10, 30)
     parameters *= p
 
     data = [x_A, y_A, value_A, x_B, y_B, value_B]

### 2018-03-23 EXP/run_cross_network.py

@@ -6,13 +6,14 @@ import matplotlib.pyplot as plt
 import os, glob, sys
 import cPickle as pickle
 
+import numpy as np
+
 from keras.models import load_model
 from keras import backend as K
 import sklearn.metrics
 import keras.applications
 
 import ClevelandMcGill as C
-
 from util import Util
 
 

### 2018-03-23 EXP/run_cross_network.py
Upppppdate.

@@ -12,6 +12,7 @@ from keras.models import load_model
 from keras import backend as K
 import sklearn.metrics
 import keras.applications
+import itertools
 
 import ClevelandMcGill as C
 from util import Util

### 2018-03-23 EXP/run_cross_network.py
Another typo.

@@ -157,7 +157,7 @@ for data in range(DATASETS):
   results[data] = network_results
 
 
-pickle_file = OUTPUT_DIR + '/' + str(CLASSIFIER)+'_'+str(NETWORK_INDEX)+'.pdf'
+pickle_file = OUTPUT_DIR + '/' + str(CLASSIFIER)+'_'+str(NETWORK_INDEX)+'.p'
 with open(pickle_file, 'w') as f:
   pickle.dump(results, f)
 print 'Stored', pickle_file

### 2018-03-23 EXP/ClevelandMcGill/figure12.py
Wow weird numpy behavior!

@@ -38,9 +38,9 @@ class Figure12:
     parameters *= p
 
     # VALUE B
-    # value_B = value_A
-    # while value_B == value_A:
-    value_B, p = Util.parameter(48, 60)
+    value_B = value_A
+    while value_B == value_A:
+      value_B, p = Util.parameter(48, 60)
     parameters *= p
 
     # POS A
@@ -50,9 +50,9 @@ class Figure12:
 
     # POS B
     x_B = 60
-    # y_B = y_A
-    # while y_B == y_A:
-    y_B, p = Util.parameter(10, 30)
+    y_B = y_A
+    while y_B == y_A:
+      y_B, p = Util.parameter(10, 30)
     parameters *= p
 
     data = [x_A, y_A, value_A, x_B, y_B, value_B]

### 2018-03-23 EXP/run_cross_network.py

@@ -63,8 +63,11 @@ for x in range(DATASETS):
 
 # generate images
 N = 20000
-X = [np.zeros((N, 100, 100), dtype=np.float32)] * DATASETS
-y = [np.zeros((N), dtype=np.float32)] * DATASETS
+X = [None] * DATASETS
+y = [None] * DATASETS
+for x in range(DATASETS):
+  X[x] = np.zeros((N, 100, 100), dtype=np.float32)
+  y[x] = np.zeros((N), dtype=np.float32)
 
 for x in range(DATASETS):
     FLAGS = [False] * 10 # never more than 10 flags

### 2018-03-23 EXP/ClevelandMcGill/figure12.py
Huge update: Now we have Figure12 and Position-Angle Experiments ready for SLURM!

@@ -34,13 +34,13 @@ class Figure12:
     parameters = 1
 
     # VALUE A
-    value_A, p = Util.parameter(48, 60)
+    value_A, p = Util.parameter(49, 61)
     parameters *= p
 
     # VALUE B
     value_B = value_A
     while value_B == value_A:
-      value_B, p = Util.parameter(48, 60)
+      value_B, p = Util.parameter(49, 61)
     parameters *= p
 
     # POS A

### 2018-03-23 EXP/run_bar_framed_rectangle.py

@@ -16,16 +16,15 @@ import time
 import ClevelandMcGill as C
 
 
-EXPERIMENT = sys.argv[1] # f.e. Angle
-DATASET = int(sys.argv[2]) # How many TRUE flags to pass to data generation
-CLASSIFIER = sys.argv[3] # 'LeNet'
-NOISE = sys.argv[4] # True
-JOB_INDEX = int(sys.argv[5])
+EXPERIMENT = sys.argv[1] # f.e. C.Figure12.data_to_framed_rectangles
+CLASSIFIER = sys.argv[2] # 'LeNet'
+NOISE = sys.argv[3] # True
+JOB_INDEX = int(sys.argv[4])
 
 #
 #
 #
-print 'Running', EXPERIMENT, 'on dataset', DATASET, 'with', CLASSIFIER, 'Noise:', NOISE, 'Job Index', JOB_INDEX
+print 'Running', EXPERIMENT, 'with', CLASSIFIER, 'Noise:', NOISE, 'Job Index', JOB_INDEX
 
 #
 #
@@ -41,9 +40,6 @@ else:
 
 DATATYPE = eval(EXPERIMENT)
 
-FLAGS = [False] * 10 # never more than 10 flags
-for f in range(DATASET):
-  FLAGS[f] = True
 
 
 if os.path.abspath('~').startswith('/n/'):
@@ -53,7 +49,7 @@ else:
   PREFIX = '/home/d/PERCEPTION/'
 RESULTS_DIR = PREFIX + 'RESULTS/'
 
-OUTPUT_DIR = RESULTS_DIR + EXPERIMENT + '/' + str(DATASET) + '/' + CLASSIFIER + '/'
+OUTPUT_DIR = RESULTS_DIR + EXPERIMENT + '/' + CLASSIFIER + '/'
 if not os.path.exists(OUTPUT_DIR):
   # here can be a race condition
   try:
@@ -72,27 +68,32 @@ if os.path.exists(STATSFILE) and os.path.exists(MODELFILE):
   print 'WAIT A MINUTE!! WE HAVE DONE THIS ONE BEFORE!'
   sys.exit(0)
 
-
 #
 #
 # DATA GENERATION
 #
 #
+
+train_counter = 0
+val_counter = 0
+test_counter = 0
 train_target = 60000
 val_target = 20000
 test_target = 20000
 
+train_labels = []
+val_labels = []
+test_labels = []
+
+
 X_train = np.zeros((train_target, 100, 100), dtype=np.float32)
[…]

### 2018-03-23 EXP/run_position_angle.py

@@ -74,27 +74,27 @@ if os.path.exists(STATSFILE) and os.path.exists(MODELFILE):
 # DATA GENERATION
 #
 #
-def in_array(haystack, needle):
-  for h in haystack:
-    if np.array_equal(needle, h):
-      return True
-  return False
 
+train_counter = 0
+val_counter = 0
+test_counter = 0
 train_target = 60000
 val_target = 20000
 test_target = 20000
 
+train_labels = []
+val_labels = []
+test_labels = []
+
+
 X_train = np.zeros((train_target, 100, 100), dtype=np.float32)
 y_train = np.zeros((train_target, 5), dtype=np.float32)
-train_counter = 0
 
 X_val = np.zeros((val_target, 100, 100), dtype=np.float32)
 y_val = np.zeros((val_target, 5), dtype=np.float32)
-val_counter = 0
 
 X_test = np.zeros((test_target, 100, 100), dtype=np.float32)
 y_test = np.zeros((test_target, 5), dtype=np.float32)
-test_counter = 0
 
 t0 = time.time()
 
@@ -103,20 +103,23 @@ while train_counter < train_target or val_counter < val_target or test_counter <
   
   all_counter += 1
   
-  data, label = C.Figure3.generate_datapoint()
-  
-  
-
-  
+  data, label, parameters = C.Figure3.generate_datapoint()
   
   pot = np.random.choice(3)
-
+  
+  # sometimes we know which pot is right
+  if label in train_labels:
+    pot = 0
+  if label in val_labels:
+    pot = 1
+  if label in test_labels:
+    pot = 2
+  
   if pot == 0 and train_counter < train_target:
-    # a training candidate
-    if in_array(y_val, label) or in_array(y_test, label):
-      # no thank you
-      continue
-      
+
+    if label not in train_labels:
+      train_labels.append(label)
+    
     #
     image = DATATYPE(data)
     image = image.astype(np.float32)
@@ -131,12 +134,10 @@ while train_counter < train_target or val_counter < val_target or test_counter <
     train_counter += 1
     
   elif pot == 1 and val_counter < val_target:
-    # a validation candidate
-    if in_array(y_test, label) or in_array(y_train, label):
-      # no thank you
-      continue
+
+    if label not in val_labels:
+      val_labels.append(label)
       
-    #
     image = DATATYPE(data)
     image = image.astype(np.float32)
       
[…]

### 2018-03-23 EXP/run_regression.py

@@ -342,7 +342,7 @@ y_pred = MLP.predict(X_test)
 #
 #
 # CLEVELAND MCGILL ERROR
-#  MIDMEANDS OF LOG ABSOLUTE ERRORS (MLAEs)
+#  MEANS OF LOG ABSOLUTE ERRORS (MLAEs)
 #
 MLAE = np.log2(sklearn.metrics.mean_absolute_error(y_pred*100, y_test*100)+.125)
 

### 2018-03-23 EXP/run_cross_network.py
Little bug when running cross networks for VGG.

@@ -125,6 +125,9 @@ for data in range(DATASETS):
 
     elif CLASSIFIER == 'VGG19':
 
+      # make 3d
+      X_new = np.stack((X[data],)*3, -1)
+
       feature_generator = keras.applications.VGG19(include_top=False, weights='imagenet', input_shape=(100,100,3))
       X_test_3D_features = feature_generator.predict(X_new, verbose=True)
       

### 2018-03-23 EXP/run_position_angle.py
Fix bug.

@@ -103,7 +103,7 @@ while train_counter < train_target or val_counter < val_target or test_counter <
   
   all_counter += 1
   
-  data, label, parameters = C.Figure3.generate_datapoint()
+  data, label = C.Figure3.generate_datapoint()
   
   pot = np.random.choice(3)
   

### 2018-03-23 EXP/ClevelandMcGill/figure3.py
Use list instead of array for performance.

@@ -79,7 +79,7 @@ class Figure3:
     #
     labels = np.roll(labels, 5-np.where(labels==1)[0])
 
-    return data, labels
+    return data, list(labels)
 
 
   @staticmethod

### 2018-03-23 EXP/run_cross_network.py
Some cross network plots need more memory.

@@ -52,6 +52,13 @@ all_labels = {'C.Figure1.position_common_scale': ['Position Y', '+ Position X', 
 DATASETS = len(all_labels[EXPERIMENT])
 
 
+pickle_file = OUTPUT_DIR + '/' + str(CLASSIFIER)+'_'+str(NETWORK_INDEX)+'.p'
+if os.path.exists(pickle_file):
+
+  print 'This one is doooone!'
+  sys.exit(0)
+
+
 networks = []
 
 for x in range(DATASETS):
@@ -163,7 +170,6 @@ for data in range(DATASETS):
   results[data] = network_results
 
 
-pickle_file = OUTPUT_DIR + '/' + str(CLASSIFIER)+'_'+str(NETWORK_INDEX)+'.p'
 with open(pickle_file, 'w') as f:
   pickle.dump(results, f)
 print 'Stored', pickle_file

### 2018-03-24 EXP/ClevelandMcGill/__init__.py
Figure 4 parametrization almost complete.

@@ -1,3 +1,4 @@
 from figure1 import Figure1
 from figure3 import Figure3
+from figure4 import Figure4
 from figure12 import Figure12

### 2018-03-24 EXP/ClevelandMcGill/figure4.py
Add constraint for Fig 4 Type 4.

@@ -1,0 +1,309 @@
+import numpy as np
+import os
+import skimage.draw
+import sys
+
+sys.path.append('../')
+from util import Util
+
+class Figure4:
+
+  SIZE = (100, 100)
+
+
+  @staticmethod
+  def generate_datapoint():
+    '''
+    '''
+
+    pairs = [10. * 10.**((i-1.)/12.) for i in range(1,11)]
+    value_A = np.random.choice(pairs)
+    value_B = value_A
+    while value_B == value_A:
+      value_B = np.random.choice(pairs)
+
+    data = [np.round(value_A), np.round(value_B)]
+
+    if (value_A < value_B):
+      ratio = np.round(value_A) / np.round(value_B)
+    else:
+      ratio = np.round(value_B) / np.round(value_A)
+      
+    label = ratio 
+
+    return data, label
+
+
+  @staticmethod
+  def data_to_type1(data):
+    '''
+    '''
+    barchart = np.zeros((100,100), dtype=np.bool)
+
+    # now we need 8 more pairs
+    all_values = [0] * 10
+    all_values[0] = np.random.randint(10,93)
+    all_values[1] = data[0] # fixed pos 1
+    all_values[2] = data[1] # fixed pos 2
+    all_values[3] = np.random.randint(10,93)
+    all_values[4] = np.random.randint(10,93)
+    all_values[5] = np.random.randint(10,93)
+    all_values[6] = np.random.randint(10,93)
+    all_values[7] = np.random.randint(10,93)
+    all_values[8] = np.random.randint(10,93)
+    all_values[9] = np.random.randint(10,93)
+
+    start = 0
+    for i,d in enumerate(all_values):
+
+      if i==0:
+        start += 2
+      elif i==5:
+        start += 8
+      else:
+        start += 0
+
+      gap = 2
+      b_width = 7
+        
+      left_bar = start+i*gap+i*b_width
+      right_bar = start+i*gap+i*b_width+b_width
+
+      # print left_bar, right_bar
+      
+      rr, cc = skimage.draw.line(99, left_bar, 99-int(d), left_bar)
+      barchart[rr, cc] = 1
+      rr, cc = skimage.draw.line(99, right_bar, 99-int(d), right_bar)
+      barchart[rr, cc] = 1
+      rr,cc = skimage.draw.line(99-int(d), left_bar, 99-int(d), right_bar)
+      barchart[rr, cc] = 1
+      
[…]

### 2018-03-24 EXP/ClevelandMcGill/figure4.py
Ready for the position-length experiment!

@@ -180,7 +180,7 @@ class Figure4:
 
     all_values[5] = data[1] # fixed pos but max. 56
     current_max = 93-all_values[5]
-    print current_max
+    # print current_max
     all_values[6] = np.random.randint(8,current_max/4.)
 
     all_values[7] = np.random.randint(8,current_max/4.)
@@ -304,6 +304,88 @@ class Figure4:
         barchart[99-current+(int(d)/2):99-current+(int(d)/2)+1, 75:76] = 1
 
 
+
+
+    return barchart
+
+
+
+
+  @staticmethod
+  def data_to_type5(data):
+    '''
+    '''
+    barchart = np.zeros((100,100), dtype=np.bool)
+
+    # we build the barchart to the top
+    all_values = [0] * 10
+
+    current_max = 93-data[0]-data[1]
+
+    if current_max <= 9:
+      # this won't work
+      raise Exception('Out of bounds')
+
+    all_values[0] = np.random.randint(3,current_max/3.)
+    
+    all_values[1] = np.random.randint(3,current_max/3.)
+
+    all_values[2] = np.random.randint(3,current_max/3.)
+
+    all_values[3] = data[0]
+
+    all_values[4] = data[1]
+
+    current_max = np.sum(all_values[0:5])
+
+    # draw left, right of the left stacked barchart
+    rr, cc = skimage.draw.line(99, 10, 99-int(current_max), 10)
+    barchart[rr, cc] = 1
+
+    rr, cc = skimage.draw.line(99, 40, 99-int(current_max), 40)
+    barchart[rr, cc] = 1
+
+    current = 0
+    for i,d in enumerate(all_values):
+
+      rr, cc = skimage.draw.line(99-(int(d)+current), 10, 99-(int(d)+current), 40)
+      barchart[rr, cc] = 1
+      current += int(d)
+
+      if i == 3 or i ==4:
+        # mark the max
+        # print current, d
+        barchart[99-current+(int(d)/2):99-current+(int(d)/2)+1, 25:26] = 1
+
+
+
+    current_max = 93
+
+    all_values[5] = np.random.randint(8,current_max/5.)
+    
+    all_values[6] = np.random.randint(8,current_max/5.)
+
+    all_values[7] = np.random.randint(8,current_max/5.)
+
+    all_values[8] = np.random.randint(8,current_max/5.)
+
+    all_values[9] = np.random.randint(8,current_max/5.)
+
[…]

### 2018-03-24 EXP/run_position_length.py

@@ -1,0 +1,383 @@
+from keras import models
+from keras import layers
+from keras import optimizers
+import keras.applications
+import keras.callbacks
+from keras import backend as K
+from keras.utils.np_utils import to_categorical
+import sklearn.metrics
+
+import cPickle as pickle
+import numpy as np
+import os
+import sys
+import time
+
+import ClevelandMcGill as C
+
+
+EXPERIMENT = sys.argv[1] # f.e. Typ1
+CLASSIFIER = sys.argv[2] # 'LeNet'
+NOISE = sys.argv[3] # True
+JOB_INDEX = int(sys.argv[4])
+
+#
+#
+#
+print 'Running', EXPERIMENT, 'with', CLASSIFIER, 'Noise:', NOISE, 'Job Index', JOB_INDEX
+
+#
+#
+# PROCESS SOME FLAGS
+#
+#
+SUFFIX = '.'
+if NOISE == 'True':
+  NOISE = True
+  SUFFIX = '_noise.'
+else:
+  NOISE = False
+
+if EXPERIMENT == 'C.Figure4.multi':
+
+  #
+  # one classifier with all different types
+  #
+  def DATATYPE():
+    '''
+    '''
+    choices = ['C.Figure4.data_to_type1', 'C.Figure4.data_to_type2',\
+              'C.Figure4.data_to_type3', 'C.Figure4.data_to_type4', 'C.Figure4.data_to_type5']
+
+    return eval(np.random.choice(choices))
+
+
+else: 
+  DATATYPE = eval(EXPERIMENT)
+
+
+
+if os.path.abspath('~').startswith('/n/'):
+  # we are on the cluster
+  PREFIX = '/n/regal/pfister_lab/PERCEPTION/'
+else:
+  PREFIX = '/home/d/PERCEPTION/'
+RESULTS_DIR = PREFIX + 'RESULTS/'
+
+OUTPUT_DIR = RESULTS_DIR + EXPERIMENT + '/' + CLASSIFIER + '/'
+if not os.path.exists(OUTPUT_DIR):
+  # here can be a race condition
+  try:
+    os.makedirs(OUTPUT_DIR)
+  except:
+    print 'Race condition!', os.path.exists(OUTPUT_DIR)
+
+STATSFILE = OUTPUT_DIR + str(JOB_INDEX).zfill(2) + SUFFIX + 'p'
[…]

### 2018-03-25 EXP/ClevelandMcGill/__init__.py
Weber's law experimental setup.

@@ -2,3 +2,4 @@ from figure1 import Figure1
 from figure3 import Figure3
 from figure4 import Figure4
 from figure12 import Figure12
+from weber import Weber

### 2018-03-25 EXP/ClevelandMcGill/weber.py

@@ -1,0 +1,70 @@
+import numpy as np
+import os
+import skimage.draw
+import sys
+
+sys.path.append('../')
+from util import Util
+
+class Weber:
+
+  SIZE = (100, 100)
+
+
+  @staticmethod
+  def base10():
+    '''
+    '''
+    randomnumber = np.random.randint(1,11)
+
+    image = Weber.generate(base=10, to_add=randomnumber)
+
+    return image, randomnumber
+
+
+  @staticmethod
+  def base100():
+    '''
+    '''
+    randomnumber = np.random.randint(1,11)
+
+    image = Weber.generate(base=100, to_add=randomnumber)
+
+    return image, randomnumber
+
+
+  @staticmethod
+  def base1000():
+    '''
+    '''
+    randomnumber = np.random.randint(1,11)
+
+    image = Weber.generate(base=1000, to_add=randomnumber)
+
+    return image, randomnumber
+
+
+  @staticmethod
+  def generate(base=10, to_add=10):
+
+    image = np.zeros((100,100), dtype=np.bool)
+
+    for p in range(base):
+      
+      image[np.random.randint(100), np.random.randint(100)] = 1
+
+
+    added = 0
+    while added < to_add:
+      
+      y, x = np.random.randint(100), np.random.randint(100)
+      
+      if image[y,x] == 1:
+        continue
+
+      image[y, x] = 1
+      added += 1
+
+    return image
+
+

### 2018-03-25 EXP/ClevelandMcGill/figure4.py
Fix endless loop in Figure4 type 4.

@@ -265,7 +265,15 @@ class Figure4:
     below_last_sum2 = below_last_sum
     above_last_sum2 = above_last_sum
 
+    # print below_last_sum, above_last_sum
+
+    ctr = 0
+
     while below_last_sum2 == below_last_sum or above_last_sum2 == above_last_sum:
+
+      if ctr == 20:
+        # this happens around 10 times in 100.000 samples.. so not really important
+        raise Exception()
 
       current_max = 93-data[1]
 
@@ -283,6 +291,8 @@ class Figure4:
 
       current_max = np.sum(all_values[5:])
       above_last_sum2 = current_max
+
+      ctr += 1 # exception counter
 
     # draw left, right of the left stacked barchart
     rr, cc = skimage.draw.line(99, 60, 99-int(current_max), 60)

### 2018-03-25 EXP/run_weber.py
Weber's law experiment ready to launch!

@@ -1,0 +1,327 @@
+from keras import models
+from keras import layers
+from keras import optimizers
+import keras.applications
+import keras.callbacks
+from keras import backend as K
+from keras.utils.np_utils import to_categorical
+import sklearn.metrics
+
+import cPickle as pickle
+import numpy as np
+import os
+import sys
+import time
+
+import ClevelandMcGill as C
+
+
+EXPERIMENT = sys.argv[1] # f.e. Angle
+CLASSIFIER = sys.argv[2] # 'LeNet'
+NOISE = sys.argv[3] # True
+JOB_INDEX = int(sys.argv[4])
+
+#
+#
+#
+print 'Running', EXPERIMENT, 'with', CLASSIFIER, 'Noise:', NOISE, 'Job Index', JOB_INDEX
+
+#
+#
+# PROCESS SOME FLAGS
+#
+#
+SUFFIX = '.'
+if NOISE == 'True':
+  NOISE = True
+  SUFFIX = '_noise.'
+else:
+  NOISE = False
+
+DATATYPE = eval(EXPERIMENT)
+
+
+
+if os.path.abspath('~').startswith('/n/'):
+  # we are on the cluster
+  PREFIX = '/n/regal/pfister_lab/PERCEPTION/'
+else:
+  PREFIX = '/home/d/PERCEPTION/'
+RESULTS_DIR = PREFIX + 'RESULTS/'
+
+OUTPUT_DIR = RESULTS_DIR + EXPERIMENT + '/' + CLASSIFIER + '/'
+if not os.path.exists(OUTPUT_DIR):
+  # here can be a race condition
+  try:
+    os.makedirs(OUTPUT_DIR)
+  except:
+    print 'Race condition!', os.path.exists(OUTPUT_DIR)
+
+STATSFILE = OUTPUT_DIR + str(JOB_INDEX).zfill(2) + SUFFIX + 'p'
+MODELFILE = OUTPUT_DIR + str(JOB_INDEX).zfill(2) + SUFFIX + 'h5'
+
+print 'Working in', OUTPUT_DIR
+print 'Storing', STATSFILE
+print 'Storing', MODELFILE
+
+if os.path.exists(STATSFILE) and os.path.exists(MODELFILE):
+  print 'WAIT A MINUTE!! WE HAVE DONE THIS ONE BEFORE!'
+  sys.exit(0)
+
+
+#
+#
+# DATA GENERATION
+#
+#
+
+train_counter = 0
+val_counter = 0
[…]

### 2018-03-25 EXP/ClevelandMcGill/figure1.py
Update the gfx.

@@ -14,7 +14,7 @@ class Figure1:
   SIZE = (100, 100)
 
   @staticmethod
-  def position_common_scale(flags=[False, False]):
+  def position_common_scale(flags=[False, False], preset=None):
     '''
     '''
     var_x = flags[0]
@@ -32,6 +32,9 @@ class Figure1:
 
     Y, p = Util.parameter(Y_RANGE[0], Y_RANGE[1])
     parameters *= p
+
+    if preset:
+      Y = preset
 
     X = Figure1.SIZE[1] / 2
     if var_x:
@@ -64,7 +67,7 @@ class Figure1:
 
 
   @staticmethod
-  def position_non_aligned_scale(flags=[False, False, False], pre_offset=None):
+  def position_non_aligned_scale(flags=[False, False, False], preset=None):
 
     var_y = flags[0]
     var_x = flags[1]
@@ -79,8 +82,8 @@ class Figure1:
     # print 'OFFSET', OFFSET
     parameters *= p
 
-    if pre_offset:
-      OFFSET = pre_offset
+    if preset:
+      OFFSET = preset
 
     Y_RANGE = (Figure1.DELTA_MIN-OFFSET, Figure1.DELTA_MAX-OFFSET)
     X_RANGE = (Figure1.DELTA_MIN, Figure1.DELTA_MAX)
@@ -123,7 +126,7 @@ class Figure1:
 
 
   @staticmethod
-  def length(flags=[False, False, False]):
+  def length(flags=[False, False, False], preset=None):
 
     var_y = flags[0]
     var_x = flags[1]
@@ -140,6 +143,9 @@ class Figure1:
 
     LENGTH, p = Util.parameter(1, Y_RANGE[1]-Y_RANGE[0]+1) # 1..60
     parameters *= p
+
+    if preset:
+      LENGTH = preset
 
     MAX_LENGTH = Y_RANGE[1]-Y_RANGE[0]
     # print 'Max length', MAX_LENGTH
@@ -179,7 +185,7 @@ class Figure1:
 
 
   @staticmethod
-  def direction(flags=[False, False]):
+  def direction(flags=[False, False], preset=None):
 
     var_y = flags[0]
     var_x = flags[1]
@@ -196,8 +202,15 @@ class Figure1:
     DOF = 360
     DIRECTION = np.random.randint(DOF)
     parameters *= DOF
+
+    if preset:
[…]

### 2018-03-25 EXP/ClevelandMcGill/weber.py

@@ -45,9 +45,12 @@ class Weber:
 
 
   @staticmethod
-  def generate(base=10, to_add=10):
+  def generate(base=10, to_add=10, label=1):
 
     image = np.zeros((100,100), dtype=np.bool)
+
+    if label != 1:
+      image = np.zeros((100,100), dtype=np.uint8)
 
     for p in range(base):
       
@@ -62,7 +65,7 @@ class Weber:
       if image[y,x] == 1:
         continue
 
-      image[y, x] = 1
+      image[y, x] = label
       added += 1
 
     return image

### 2018-03-26 PAPER/07_results.tex
Fix bug in multi position-length experiment and some paper updates.

@@
 \begin{figure*}[h]
-	  \includegraphics[width=\linewidth]{figure1.pdf}
+	  \includegraphics[width=\linewidth]{figure1_latest.pdf}
   \caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
@@
 
-We have nothing yet here...
+\begin{figure*}[t]
+	  \includegraphics[width=\linewidth]{figure4_mlae.pdf}
+  \caption{\textbf{Computational results of the Position-Length experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-length experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
+We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
+	\label{fig:figure4_mlae}
+\end{figure*}
 
@@
 First run indicates that framed rectangles perform better but we dont really know it yet.
+
+\begin{figure*}[t]
+	  \includegraphics[width=\linewidth]{figure12_mlae.pdf}
+  \caption{\textbf{Computational results of the Bars-and-Framed-Rectangles experiment.} Log absolute error means and 95\% confidence intervals for the \emph{bars-and-framed-rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
+We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
+	\label{fig:figure12_mlae}
+\end{figure*}
 

### 2018-03-26 EXP/run_position_length.py

@@ -43,13 +43,16 @@ if EXPERIMENT == 'C.Figure4.multi':
   #
   # one classifier with all different types
   #
-  def DATATYPE():
+  def DATATYPE(data):
     '''
     '''
+    
     choices = ['C.Figure4.data_to_type1', 'C.Figure4.data_to_type2',\
               'C.Figure4.data_to_type3', 'C.Figure4.data_to_type4', 'C.Figure4.data_to_type5']
 
-    return eval(np.random.choice(choices))
+    choice = np.random.choice(choices)
+    
+    return eval(choice)(data)
 
 
 else: 

### 2018-03-26 PAPER/03_elementary_perceptual_tasks.tex
Add the slim cross network graphic.

@@
 	\midrule
-	\makecell[tl]{\emph{Position Non-Aligned Scale}} & \makecell[tl]{Origin\\+ Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $10$ \\$600$ \\ $36000$ \\ $216000$}\\
+	\makecell[tl]{\emph{Position Non-Aligned Scale}} & \makecell[tl]{Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $600$ \\ $36000$ \\ $216000$}\\
 
@@
 	\midrule
-	\makecell[tl]{\emph{Curvature}} & \makecell[tl]{Midpoint Curvature\\+ Position Y \\+ Position X\\+ Width} & \makecell[tr]{ $80$ \\ $1600$ \\ $64000$ \\ $2560000$}\\
+	\makecell[tl]{\emph{Curvature}} & \makecell[tl]{Midpoint Curvature\\+ Position Y \\+ Position X} & \makecell[tr]{ $80$ \\ $1600$ \\ $64000$}\\
 
 	\midrule
-	\makecell[tl]{\emph{Shading}} & \makecell[tl]{Density\\+ Position Y \\+ Position X} & \makecell[tr]{ $100$ \\ $1600$ \\ $2000$ \\ $40000$}\\
+	\makecell[tl]{\emph{Shading}} & \makecell[tl]{Density\\+ Position Y \\+ Position X} & \makecell[tr]{ $100$ \\ $2000$ \\ $40000$}\\
 

### 2018-03-26 PAPER/07_results.tex
Add the slim cross network graphic.

@@
 \begin{figure}[t]
-	  \includegraphics[width=\linewidth]{crossnetwork.pdf}
-  \caption{\textbf{Cross-classifier variability for the perceptual task of measuring curvature.} We use predictions of LeNet classifiers trained on different parametrizations of the \emph{curvature} elementary perceptual task and measure the mean logistic absolute error (MLAE).
+	  \includegraphics[width=\linewidth]{cross_network_small_VGG19.pdf}
+  \caption{\textbf{Cross-classifier variability for the perceptual task of measuring length.} We use predictions of LeNet classifiers trained on different parametrizations of the \emph{curvature} elementary perceptual task and measure the mean logistic absolute error (MLAE).
 The lower score, the better.

### 2018-03-26 PAPER/00_introduction.tex
Update the gfx.
Update intro.

@@
 Originally inspired by neuroscientific discoveries, the recent advances in deep learning have been the direct results of engineering efforts, more specifically in convolutional neural networks (CNNs).
-While there has been significant advancement, this does not mean we understand what CNNs are doing and we are in fact treating them as blackboxes without detracting from their success \cite{goodfellow_book, deeplearning_blackbox2017}.
-Our current knowledge of biological vision suggests that modern machine learning models indeed mimic the underlying biology by abstracting the many details of biological neural networks \cite{yamins2016using, hassabis2017neuroscience}.
+Our current knowledge of biological vision suggests that modern machine learning models indeed mimic the underlying biology by abstracting the many details of biological neural networks \cite{yamins2016using, hassabis2017neuroscience, human_vs_machine_vision}.
+While there has been significant advancement, fully understanding how CNNs work is a still ongoing process which does not detract from their success~\cite{goodfellow_book, deeplearning_blackbox2017}.
+Recent works target the systematic evaluation of perceptional limits of CNNs~\cite{clevr, not_so_clevr}.
 
-Despite tremendous research efforts and generating massive datasets, we are far from fully understanding biological vision. 
-Similarly, we are constantly developing inventive features for deep artificial networks without truly understanding it in its entirety. As a result, a double discrepancy is observed. 
-Advances in neuroscience can revolutionize machine learning by reverse-engineering neural circuits, yielding new classifiers which, in turn, can help process the massive biological data. 
-There are many existing questions which must be answered in order to fill this gap of our understanding.
+We are interested in applying neural networks to graphs, charts, and visual encodings.
+A first step in this direction is to evaluate if and how convolutional neural networks perceive low-level graphical elements, the building blocks of information visualizations.
+This is a very different task than applying such networks to natural images (like the majority of computer vision research) since such scenes are usually more complex and identifiers such as edges are not as prominent.
+More related towards our goal are the \emph{graphical perception} experiments of Cleveland and McGill~\cite{cleveland_mcgill}.
+In this paper, we explore whether these experiments which were performed with humans can be reproduced with CNNs.
+Cleveland and McGill's work in the 1980s has led to many insights for modern information visualization research such as the identification of elementary perceptional tasks or that bar charts are easier to understand than pie charts.
 
-We focus on perception... experiments from cleveland mcgill..
+In order to perform this evaluation, we parametrize different visual encodings such as the elementary perceptual tasks suggested by Cleveland and McGill~\cite{cleveland_mcgill}.
+We then replicate the original experimental design by defining linear regression tasks for continuous variables.
+However, we select three modern feature generators based on convolutional neural networks and combine them with a multilayer perceptron (MLP) to include non-linearities.
+We include the LeNet-5 network~\cite{lenet}, the VGG19 network~\cite{simonyan_very_deep2014}, and the Xception classifiers~\cite{xception}.
+While we train LeNet-5 from scratch, for VGG19 and Xception we use pre-trained imagenet~\cite{imagenet} weights to further mimic the human visual system.
+By also using the MLP directly without convolutional feature detection as baseline, we test four different classifiers as part of the following scenarios: a) elementary perceptual tasks with increasing parameteric complexity, b) position-angle experiment comparing pie charts to bar charts, c) position-length experiment that compares grouped and divided bar charts, and d) the bars and framed rectangles experiments where visual cues aid perception.
+We also investigate other properties of CNNs such as whether they obey Weber's law which defines a proportional dependency between an initial distribution and perceivable change.
+Our experimental setup includes repetitions, randomizations, and regularizations to prevent any bias.
+
+Our motivation for replicating Cleveland and McGill's experiments stems from the thought that computational perception seems to be closely related to biological vision.
+If human perception yields certain results, maybe we can replicate these results with machine vision.
+As our first contribution, we study the elementary perceptual tasks by Cleveland and McGill and then systematically parametrize and evaluate them for computational perception with our four classifiers.
+This setup includes also cross-network evaluations which give insight into the generalizability of the classifiers.
+It also yields our second contribution: a ranking of elementary perceptual tasks for machine perception, similar to the one proposed by Cleveland and McGill.
+Further, we replicate the \emph{position-angle} and the \emph{position-length} experiments which contributes to the general knowledge of superior perceivability of bar charts to pie charts.
+We then reproduce the \emph{bars-and-framed-rectangles} experiment with our classifiers.
+Here, we also include an additional experiment to test the Weber-Fechner's law for CNNs.
+Finally, we discuss our findings and derive recommendations for allowing CNNs to perceive visualizations.
+
+We accompany this paper and detailed supplemental material with open source code~\footnote{Code, data, results and more are available at: \url{http://rhoana.org/perception}}, data, and results to enable a framework for the development and evaluation of new network architectures for graphical perception.
 
@@
 
-\subsection{Biological Vision}
+\subsection{From Biological Vision to Machine Learning}
 
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{biology_vs_cnn.png}
-  \caption{The Biological Vision (schematic)}
-	\label{fig:vision}
-\end{figure}
+%\begin{figure}[t]
+%	  \includegraphics[width=\linewidth]{biology_vs_cnn.png}
+%  \caption{The Biological Vision (schematic)}
+%	\label{fig:vision}
+%\end{figure}
 
@@
 
-GOALS
-\begin{itemize}
-	\item reduce the gap between neurobiology and data science to advance the understanding of visual cortex inspired machine learning
-	\item insights for infovis for machines
-\end{itemize}
-
-CONTRIBUTIONS
-\begin{itemize}
-	\item experiments of cleveland mcgill with systematic parametrization and evaluation for computational perception
-	\item ranking of elementary perceptual tasks for machine perception
-	\item validation of weber's law for feedforward neural networks
-	\item many insights for infovis for machines
-\end{itemize}
+We think that the biological inspiration of modern convolutional neural networks yields the evaluation of principles of human perception with computers.

### 2018-03-26 PAPER/07_results.tex
Update intro.

@@
 \begin{figure*}[h]
-	  \includegraphics[width=\linewidth]{figure1_latest.pdf}
+	\centering
+	  \includegraphics[width=.95\linewidth]{figure1.pdf}
   \caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
@@
 \begin{figure*}[t]
-	  \includegraphics[width=\linewidth]{figure4_mlae.pdf}
+	  \includegraphics[width=\linewidth]{figure4_mlae_with_multi.pdf}
   \caption{\textbf{Computational results of the Position-Length experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-length experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
@@
 
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_val_loss.png}
-  \caption{\textbf{Classifier Efficiency of the Bars and Framed Rectangles experiment.} Categorical Cross-Entropy loss for the \emph{bars and framed rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
+%\begin{figure}[t]
+%	  \includegraphics[width=\linewidth]{figure12_val_loss.png}
+%  \caption{\textbf{Classifier Efficiency of the Bars and Framed Rectangles experiment.} Categorical Cross-Entropy loss for the \emph{bars and framed rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
 The frame around the bars adds an additional visual cue enables faster network convergence.
 This is not yet reproducible!}
-	\label{fig:figure12_val_loss}
-\end{figure}
+%	\label{fig:figure12_val_loss}
+%\end{figure}

### 2018-03-27 PAPER/01_related_work.tex
Paper draft in the works.
Update related work (not done yet).

--- 2018-02-07 PAPER/01_related_work.tex
@@
-\section{Related Work}
-\cite{linegraph_vs_scatterplot} \cite{cleveland_mcgill}
+\section{Previous Work}
+
+\textbf{Graphical Perception.} Cleveland and McGill~\cite{cleveland_mcgill} introduce the fundamental concept of \emph{graphical perception} and investigate how different visual attributes and encodings are perceivable by humans.
+They define \emph{elementary perceptual tasks} as mental-visual stimuli to understand encodings in visualizations.
+Based on these definitions, the authors propose and perform different experiments such as the \emph{position-angle} experiment which compares bar charts and pie charts, the \emph{position-length} experiment where users judge relations between encoded values in grouped and divided bar charts, and the \emph{bars-and-framed-rectangles} experiment to evaluate Weber's law.
+Heer and Bostock later reproduced the Cleveland-McGill experiments crowd-sourced on Mechanical Turk~\cite{HeerBostock2010} which lead to follow-up work from Harrison \textit{et al.}~\cite{harrison2013influencing} who replicated the experiments while observing emotional states.
+Both papers report similar results to Cleveland and McGill which increased our motivation to mimmick their pioneering work.
+Our experimental setup replicates the original setup of Cleveland and McGill - just instead of humans, we use convolutional neural networks due to the connection with the human visual system.
+While we focus on Cleveland and McGill's work from 1984, many other excellent articles from the last decades target low-level visual encoding~\cite{bertin1967semiologie,cleveland1985graphical,treisman1988feature, wilkinson2006grammar, carpendale2003considering,widgor_perception2007,munzner2015visualization}.
+\\~\\
+\textbf{Comparing Visual Encodings.} Different visual encodings have advantages or disadvantages and the community does a great job comparing them.
+Higher-level comparisons include 2D versus 3D vector field and rendering studies~\cite{mckenzie_2d_3d,forsberg2009comparing_3d_vector,laidlaw_2d_vector,borkin2011arteries}, timeseries~\cite{herr2009timeseries} and scatterplots~\cite{tremmel1995visual,Wang_linegraph_vs_scatterplot}.
+Lower-level experiments target - besides others - open versus closed encodings~\cite{open_vs_closed_shapes}, and several evaluations of color space~\cite{ware1988color,rheingans1992color,Rogowitz2001_colormaps,kindlmann2002color}.
+While we investigate lower-level visual encodings in this work, we delay colorspace experiments for future work.
+
+Interesting are also the rankings of correlation visualization using Weber's law~\cite{harrison2014_webers_law_rank}.
+This law defines the proportional relation between the initial distribution density and perceivable change.
+In this paper, we investigate with a simple experiment whether this holds for convolutional neural networks.
+\\~\\
+\textbf{Visual Cortex Inspired Machine Learning.} Classifiers mimmicking the human visual system are a hot topic and many different architectures, models, and paradigms exist.
+Feed-forward neural networks
+
+MLP, LeNet, VGG, ImageNet, XCeption, ResNets etc and work from THomas Serre
+%
+%
+%- https://link.springer.com/chapter/10.1007/978-3-642-14600-8_46
+%
+%- https://github.com/tidyverse/ggplot2/wiki/Recommended-Reading
+%
+%- last year's VADL workshop: https://vadl2017.github.io/
+
+\textbf{Computing Perception.}
+
+Pineo et al.~\cite{Pineo2012_computational_perception} present a method to automatically evaluate and optimize visualizations using a computational model of human vision, based on a neural network simulation of the early perceptual processing in the retina and primary visual cortex.
+[JT] Copied from their abstract.
+Zoya's work

### 2018-03-28 EXP/ClevelandMcGill/figure3.py
Support antialiased drawing of pie charts.

@@ -146,3 +146,40 @@ class Figure3:
       random_direction -= current_angle
 
     return piechart
+
+  @staticmethod
+  def data_to_piechart_aa(data):
+    '''
+    '''
+    piechart = np.zeros((100,100), dtype=np.float32)
+    RADIUS = 30
+    rr,cc,val = skimage.draw.circle_perimeter_aa(50,50,RADIUS)
+    piechart[rr,cc] = val
+    random_direction = np.random.randint(360)
+    theta = -(np.pi / 180.0) * random_direction
+    END = (50 - RADIUS * np.cos(theta), 50 - RADIUS * np.sin(theta))
+    rr, cc, val = skimage.draw.line_aa(50, 50, int(np.round(END[0])), int(np.round(END[1])))
+    piechart[rr, cc] = val
+
+    for i,d in enumerate(data):
+
+      current_value = data[i]
+      current_angle = (current_value / 100.) * 360.
+      # print current_value, current_angle
+      # print 'from', random_direction, 'to', current_angle
+      theta = -(np.pi / 180.0) * (random_direction-current_angle)
+      END = (50 - RADIUS * np.cos(theta), 50 - RADIUS * np.sin(theta))
+      rr, cc,val = skimage.draw.line_aa(50, 50, int(np.round(END[0])), int(np.round(END[1])))
+      piechart[rr,cc] = val
+      
+      if d == np.max(data):
+        # this is the max spot
+        theta = -(np.pi / 180.0) * (random_direction-current_angle/2.)
+        END = (50 - RADIUS/2 * np.cos(theta), 50 - RADIUS/2 * np.sin(theta))
+        rr, cc,val = skimage.draw.line_aa(int(np.round(END[0])), int(np.round(END[1])), int(np.round(END[0])), int(np.round(END[1])))
+        piechart[rr,cc] = val
+      
+      random_direction -= current_angle
+
+    return piechart
+

### 2018-03-28 EXP/run_regression_from_scratch.py

@@ -51,7 +51,7 @@ if os.path.abspath('~').startswith('/n/'):
   PREFIX = '/n/regal/pfister_lab/PERCEPTION/'
 else:
   PREFIX = '/home/d/PERCEPTION/'
-RESULTS_DIR = PREFIX + 'RESULTS/'
+RESULTS_DIR = PREFIX + 'RESULTS_FROM_SCRATCH/'
 
 OUTPUT_DIR = RESULTS_DIR + EXPERIMENT + '/' + str(DATASET) + '/' + CLASSIFIER + '/'
 if not os.path.exists(OUTPUT_DIR):

### 2018-03-28 EXP/run_regression_from_scratch.py
Fix bug.

@@ -272,11 +272,11 @@ t0 = time.time()
 callbacks = [keras.callbacks.EarlyStopping(monitor='val_loss', min_delta=0, patience=10, verbose=0, mode='auto'), \
              keras.callbacks.ModelCheckpoint(MODELFILE, monitor='val_loss', verbose=1, save_best_only=True, mode='min')]
 
-history = model.fit(X_train, \
+history = model.fit(X_train_3D, \
                     y_train, \
                     epochs=1000, \
                     batch_size=32, \
-                    validation_data=(X_val, y_val),
+                    validation_data=(X_val_3D, y_val),
                     callbacks=callbacks,
                     verbose=True)
 

### 2018-03-28 EXP/run_position_length_from_scratch.py
Add Figure 4 from scratch!

@@ -1,0 +1,339 @@
+from keras import models
+from keras import layers
+from keras import optimizers
+import keras.applications
+import keras.callbacks
+from keras import backend as K
+from keras.utils.np_utils import to_categorical
+import sklearn.metrics
+
+import cPickle as pickle
+import numpy as np
+import os
+import sys
+import time
+
+import ClevelandMcGill as C
+
+
+EXPERIMENT = sys.argv[1] # f.e. Typ1
+CLASSIFIER = sys.argv[2] # 'LeNet'
+NOISE = sys.argv[3] # True
+JOB_INDEX = int(sys.argv[4])
+
+#
+#
+#
+print 'Running', EXPERIMENT, 'with', CLASSIFIER, 'Noise:', NOISE, 'Job Index', JOB_INDEX
+
+#
+#
+# PROCESS SOME FLAGS
+#
+#
+SUFFIX = '.'
+if NOISE == 'True':
+  NOISE = True
+  SUFFIX = '_noise.'
+else:
+  NOISE = False
+
+if EXPERIMENT == 'C.Figure4.multi':
+
+  #
+  # one classifier with all different types
+  #
+  def DATATYPE(data):
+    '''
+    '''
+    
+    choices = ['C.Figure4.data_to_type1', 'C.Figure4.data_to_type2',\
+              'C.Figure4.data_to_type3', 'C.Figure4.data_to_type4', 'C.Figure4.data_to_type5']
+
+    choice = np.random.choice(choices)
+    
+    return eval(choice)(data)
+
+
+else: 
+  DATATYPE = eval(EXPERIMENT)
+
+
+
+if os.path.abspath('~').startswith('/n/'):
+  # we are on the cluster
+  PREFIX = '/n/regal/pfister_lab/PERCEPTION/'
+else:
+  PREFIX = '/home/d/PERCEPTION/'
+RESULTS_DIR = PREFIX + 'RESULTS_FROM_SCRATCH/'
+
+OUTPUT_DIR = RESULTS_DIR + EXPERIMENT + '/' + CLASSIFIER + '/'
+if not os.path.exists(OUTPUT_DIR):
+  # here can be a race condition
+  try:
+    os.makedirs(OUTPUT_DIR)
+  except:
+    print 'Race condition!', os.path.exists(OUTPUT_DIR)
+
[…]

### 2018-03-28 EXP/run_position_length_from_scratch.py
Fix target classes.

@@ -106,13 +106,13 @@ test_labels = []
 
 
 X_train = np.zeros((train_target, 100, 100), dtype=np.float32)
-y_train = np.zeros((train_target, 1), dtype=np.float32)
+y_train = np.zeros((train_target, 5), dtype=np.float32)
 
 X_val = np.zeros((val_target, 100, 100), dtype=np.float32)
-y_val = np.zeros((val_target, 1), dtype=np.float32)
+y_val = np.zeros((val_target, 5), dtype=np.float32)
 
 X_test = np.zeros((test_target, 100, 100), dtype=np.float32)
-y_test = np.zeros((test_target, 1), dtype=np.float32)
+y_test = np.zeros((test_target, 5), dtype=np.float32)
 
 t0 = time.time()
 

### 2018-03-28 PAPER/01_related_work.tex
Update intro and rel. work.

@@
 While we focus on Cleveland and McGill's work from 1984, many other excellent articles from the last decades target low-level visual encoding~\cite{bertin1967semiologie,cleveland1985graphical,treisman1988feature, wilkinson2006grammar, carpendale2003considering,widgor_perception2007,munzner2015visualization}.
-\\~\\
-\textbf{Comparing Visual Encodings.} Different visual encodings have advantages or disadvantages and the community does a great job comparing them.
-Higher-level comparisons include 2D versus 3D vector field and rendering studies~\cite{mckenzie_2d_3d,forsberg2009comparing_3d_vector,laidlaw_2d_vector,borkin2011arteries}, timeseries~\cite{herr2009timeseries} and scatterplots~\cite{tremmel1995visual,Wang_linegraph_vs_scatterplot}.
-Lower-level experiments target - besides others - open versus closed encodings~\cite{open_vs_closed_shapes}, and several evaluations of color space~\cite{ware1988color,rheingans1992color,Rogowitz2001_colormaps,kindlmann2002color}.
-While we investigate lower-level visual encodings in this work, we delay colorspace experiments for future work.
 
@@
 \\~\\
-\textbf{Visual Cortex Inspired Machine Learning.} Classifiers mimmicking the human visual system are a hot topic and many different architectures, models, and paradigms exist.
-Feed-forward neural networks
+%\textbf{Comparing Visual Encodings.} Different visual encodings have advantages or disadvantages and the community does a great job comparing them.
+Higher-level comparisons include 2D versus 3D vector field and rendering studies~\cite{mckenzie_2d_3d,forsberg2009comparing_3d_vector,laidlaw_2d_vector,borkin2011arteries}, timeseries~\cite{herr2009timeseries} and scatterplots~\cite{tremmel1995visual,Wang_linegraph_vs_scatterplot}.
+Lower-level experiments target - besides others - open versus closed encodings~\cite{open_vs_closed_shapes}, and several evaluations of color space~\cite{ware1988color,rheingans1992color,Rogowitz2001_colormaps,kindlmann2002color}.
+While we investigate lower-level visual encodings in this work, we delay colorspace experiments for future work.
 
-MLP, LeNet, VGG, ImageNet, XCeption, ResNets etc and work from THomas Serre
+\textbf{Computational Visualization Understanding.}
+
+Pineo \textit{et al.}~\cite{Pineo2012_computational_perception} create computational model of human vision based on neural networks and their experiments show that understanding visualization triggers neural activity in high-level areas of cognition.
+The authors suspect that this level of understanding is produced by low-level neurons performing elementary perceptional tasks.
+We are further investigating this suspision.
+Other work tries to parse infographics by finding higher-level saliency models~\cite{bylinskii2016should}, or by extracting text or key visual elements~\cite{diagram_understanding,kembhavi2016diagram,zoya_text_visual_tags}.
+However, none of these works focus on computational understanding of lower-level building blocks of visualizations such as curvature, lengths, or position.
+\\~\\
+\textbf{Visual Cortex Inspired Machine Learning.} 
+%Classifiers mimmicking the human visual system are a hot topic and many different architectures, models, and paradigms exist. Feed-forward neural networks 
+The human visual cortex is an extremely powerful system which allows the ability, and seemingly without effort, to recognize an enormous amount of distinct objects in the world.
+This visual system is organized in layers and has inspired the theory of computational classifiers based on multilayer neural networks.
+Fukushima and Miyake developed the Neocognitron quantitative model~\cite{fukushima1982neocognitron} that ultimately led to the important work of Hinton, Bengio, and LeCun: \textit{deep neural networks}~\cite{lecun2015deep}, visual cortex inspired machine learning.
+Nowadays, such classifiers exist with many different architectures.
+For this paper, we select the traditional \emph{LeNet-5}~\cite{lenet} which was designed to recognize hand-written digits, the VGG19~\cite{simonyan_very_deep2014} classifier with 16 convolutional layers, and the Xception~\cite{xception} classifier with 36 convolutional layers.
+Selecting these specific networks allows us to compare archtitectures with different depths.
+%Object detection is extremely difficult and therefore is especially impressive as light intensities can change by levels of magnitude and contrast between foreground and background is so often low. 
+%In addition, the visual scene changes every time the human body or human eyes move. 
+%This visual system exhibits a very noisy structure but because it is organized by layers it has inspired the mathematical theory of multilayer neural networks.
+
+%In 1962 Hubel and Wiesel were the first to begin studying the visual cortex from the standpoint of a neuroscientist.
+Their experimental findings on cats and macaque monkeys suggested a hierarchy of cells with increasing complexity which was then later transferred to the hierarchical model of different layers.
+Twenty years later, this insight was translated to the Neocognitron quantitative model, by Fukushima and Miyake, which ultimately led to the important work of Hinton, Bengio, and LeCun in the 1980s.
+Their work on stochastic gradient descent approximation, and the availability of faster computer hardware then led to today’s deep learning networks.
+In the last decade, this field has exhibited rapid growth, constant evolution, and new applications in various domains.
+
+%The reported classification performance is superior to that of humans and the question in regards to their functionality opens an interesting research topic.
+%Visual cortex inspired machine learning classifiers exist with many different architectures.
+We select the traditional \emph{LeNet-5}~\cite{lenet} which was designed to recognize hand-written digits, the VGG19~\cite{simonyan_very_deep2014} classifier with 16 convolutional layers, and the Xception~\cite{xception} classifier with 36 convolutional layers.
+%
+%What is remarkable is that even though current machine learning models do not resemble the complexity of its biological pendant, they inherently are able to generalize but only after learning thousands of examples.
+%Neural networks trained on one specific task can be used to perform detection or segmentation of, seemingly, unrelated objects with relatively minor retraining~\cite{transfer_learning}. 
+%
+%
+%We think that the biological inspiration of modern convolutional neural networks yields the evaluation of principles of human perception with computers.
+%
+%
+%MLP, LeNet, VGG, ImageNet, XCeption, ResNets etc and work from THomas Serre
 %
@@
 %- last year's VADL workshop: https://vadl2017.github.io/
-
-\textbf{Computing Perception.}
-
-Pineo et al.~\cite{Pineo2012_computational_perception} present a method to automatically evaluate and optimize visualizations using a computational model of human vision, based on a neural network simulation of the early perceptual processing in the retina and primary visual cortex.
-[JT] Copied from their abstract.
-Zoya's work

### 2018-03-28 PAPER/00_introduction.tex

@@
 
-Artificial intelligence has taken the world of technology by storm.
+%Artificial intelligence has taken the world of technology by storm. 
 Deep multilayer neural networks are being successfully applied in a wide range of applications that are regularly outperforming humans in object recognition \cite{krizhevsky_imagenet2012, simonyan_very_deep2014,szegedy2015}.
 Originally inspired by neuroscientific discoveries, the recent advances in deep learning have been the direct results of engineering efforts, more specifically in convolutional neural networks (CNNs).
-Our current knowledge of biological vision suggests that modern machine learning models indeed mimic the underlying biology by abstracting the many details of biological neural networks \cite{yamins2016using, hassabis2017neuroscience, human_vs_machine_vision}.
-While there has been significant advancement, fully understanding how CNNs work is a still ongoing process which does not detract from their success~\cite{goodfellow_book, deeplearning_blackbox2017}.
-Recent works target the systematic evaluation of perceptional limits of CNNs~\cite{clevr, not_so_clevr}.
+Our current knowledge of biological vision suggests that CNNs indeed mimic feature learning similar to the early layers of the visual cortex but without incorporating the many details of biological neural networks \cite{yamins2016using, hassabis2017neuroscience, human_vs_machine_vision}.
+While there has been significant advancement, fully understanding how CNNs work is a still ongoing process which does not detract from their success~\cite{goodfellow_book, deeplearning_blackbox2017} and recent works target the systematic evaluation of perceptional limits of CNNs~\cite{clevr, not_so_clevr}.
 
@@
 This setup includes also cross-network evaluations which give insight into the generalizability of the classifiers.
-It also yields our second contribution: a ranking of elementary perceptual tasks for machine perception, similar to the one proposed by Cleveland and McGill.
-Further, we replicate the \emph{position-angle} and the \emph{position-length} experiments which contributes to the general knowledge of superior perceivability of bar charts to pie charts.
+It also yields our second contribution: a ranking of Cleveland and McGill's elementary perceptual tasks for our tested CNN architectures.
+Further, we replicate the \emph{position-angle} and the \emph{position-length} experiments which contributes to the general knowledge of superior perceivability of bar charts to pie charts in certain conditions.
 We then reproduce the \emph{bars-and-framed-rectangles} experiment with our classifiers.
 Here, we also include an additional experiment to test the Weber-Fechner's law for CNNs.
+Our experiments yield a TODO.
 Finally, we discuss our findings and derive recommendations for allowing CNNs to perceive visualizations.
@@
 %While the reasons for this are still unknown, specified research can most certainly advance the understanding of deep neural networks.
-
-\subsection{From Biological Vision to Machine Learning}
-
-%\begin{figure}[t]
-%	  \includegraphics[width=\linewidth]{biology_vs_cnn.png}
-%  \caption{The Biological Vision (schematic)}
-%	\label{fig:vision}
-%\end{figure}
-
-Biological vision is an extremely powerful system which allows humans the ability, and seemingly without effort, to recognize an enormous amount of distinct objects in the world. 
-Object detection is extremely difficult and therefore is especially impressive as light intensities can change by levels of magnitude and contrast between foreground and background is so often low. 
-In addition, the visual scene changes every time the human body or human eyes move. 
-This visual system exhibits a very noisy structure but because it is organized by layers it has inspired the mathematical theory of multilayer neural networks. 
-What is remarkable is that even though current machine learning models do not resemble the complexity of its biological pendant, they inherently generalize extremely well. 
-Neural networks trained on one specific task can be used to perform detection or segmentation of, seemingly, unrelated objects with relatively minor retraining. 
-The reported classification performance is superior to that of humans and the question in regards to their functionality opens an interesting research topic.
-
-In 1962 Hubel and Wiesel were the first to begin studying the visual cortex from the standpoint of a neuroscientist.
+%
+%\subsection{From Biological Vision to Machine Learning}
+%
+%%\begin{figure}[t]
+%%	  \includegraphics[width=\linewidth]{biology_vs_cnn.png}
+%%  \caption{The Biological Vision (schematic)}
+%%	\label{fig:vision}
+%%\end{figure}
+%
+%Biological vision is an extremely powerful system which allows humans the ability, and seemingly without effort, to recognize an enormous amount of distinct objects in the world. 
+%Object detection is extremely difficult and therefore is especially impressive as light intensities can change by levels of magnitude and contrast between foreground and background is so often low. 
+%In addition, the visual scene changes every time the human body or human eyes move. 
+%This visual system exhibits a very noisy structure but because it is organized by layers it has inspired the mathematical theory of multilayer neural networks. 
+%What is remarkable is that even though current machine learning models do not resemble the complexity of its biological pendant, they inherently generalize extremely well. 
+%Neural networks trained on one specific task can be used to perform detection or segmentation of, seemingly, unrelated objects with relatively minor retraining. 
+%The reported classification performance is superior to that of humans and the question in regards to their functionality opens an interesting research topic.
+%
+%In 1962 Hubel and Wiesel were the first to begin studying the visual cortex from the standpoint of a neuroscientist.
 Their experimental findings on cats and macaque monkeys suggested a hierarchy of cells with increasing complexity which was then later transferred to the hierarchical model of different layers.
@@
 In the last decade, this field has exhibited rapid growth, constant evolution, and new applications in various domains.
-
-We think that the biological inspiration of modern convolutional neural networks yields the evaluation of principles of human perception with computers.
+%
+%We think that the biological inspiration of modern convolutional neural networks yields the evaluation of principles of human perception with computers.
+%

### 2018-03-28 PAPER/03_elementary_perceptual_tasks.tex
Updated paper structure.

@@
 \end{itemize}
+
+\subsection{Results}
+
+some are good and some are bad.. why?
+
+\begin{figure*}[h]
+	\centering
+	  \includegraphics[width=.95\linewidth]{figure1.pdf}
+  \caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
+We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
+	\label{fig:figure1_results}
+\end{figure*}
+
+\textbf{Computational Perception Ranking.}
+
+Cleveland McGills Ranking - can we observe something similar?
+
+\begin{enumerate}
+	\item Position along a common scale e.g. scatter plot
+	\item Position on identical but nonaligned scales e.g. multiple scatter plots
+	\item Length e.g. bar chart
+	\item Angle \& Slope (tie) e.g. pie chart
+	\item Area e.g. bubbles
+	\item Volume, density, and color saturation (tie) e.g. heatmap
+	\item Color hue e.g. newsmap
+\end{enumerate}
+
+\noindent{\textbf{Cross-classifier variability.}}
+
+Can a neural network generalize on simple perceptual tasks?
+
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{cross_network_small_VGG19.pdf}
+  \caption{\textbf{Cross-classifier variability for the perceptual task of measuring length.} We use predictions of LeNet classifiers trained on different parametrizations of the \emph{curvature} elementary perceptual task and measure the mean logistic absolute error (MLAE).
+The lower score, the better.
+Classifiers trained on curves with variable position can generalize even if the axis of translation varies.
+However, classifiers trained on fixed positions of curves are not able to measure translated curves.}
+	\label{fig:cross_network}
+\end{figure}

### 2018-03-28 PAPER/04_position_angle_experiment.tex
Updated paper structure.

@@
 \end{itemize}
+
+\subsection{Results}
+
+Bar charts are more accurate (Fig.~\ref{fig:figure3_mlae}) and networks converge faster (Fig.~\ref{fig:figure3_val_loss}). This is great.
+
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
+  \caption{\textbf{Classifier Efficiency of the Position-Angle experiment.} Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts.
+We report the MSE measure for both encodings of four different classifier on previously unseen validation data.}
+	\label{fig:figure3_val_loss}
+\end{figure}
+
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{figure3_mlae_better.pdf}
+  \caption{\textbf{Computational results of the Position-Angle experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
+We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
+	\label{fig:figure3_mlae}
+\end{figure}

### 2018-03-28 PAPER/05_position_length_experiment.tex
Discussion point for position-length experiment
Update the gfx.
Updated paper structure.

@@
 
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{position_length_original1.png}
-	  \includegraphics[width=\linewidth]{position_length_original2.png}
+\begin{figure*}[t]
+	  \includegraphics[width=\linewidth]{figure4_overview.pdf}
 
@@
 	\label{fig:position_length_experiment}
+\end{figure*}
+
+\subsection{Discussion}
+
+JT: Look at the relative difficulty of the tasks.
+In Cleveland and McGill, types 1-5 were post-ordered by their log error such that type 1 was easiest and type 5 was hardest.
+Is this still the case with our CNNs?
+
+\subsection{Results}
+
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{figure4_mlae_with_multi_and_humans.pdf}
+  \caption{\textbf{Computational results of the Position-Length experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-length experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
+We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
+	\label{fig:figure4_mlae}
 \end{figure}

### 2018-03-28 PAPER/06_bars_and_framed_rectangles_experiment.tex
Update the gfx.
Updated paper structure.

@@
 \begin{figure}[t]
-	  \includegraphics[width=\linewidth]{webers_law}
+	  \includegraphics[width=\linewidth]{weber_overview}
   \caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution.
@@
 \end{figure}
+
+\subsection{Results}
+
+First run indicates that framed rectangles perform better but we dont really know it yet.
+
+\begin{figure*}[t]
+	  \includegraphics[width=\linewidth]{figure12_mlae.pdf}
+  \caption{\textbf{Computational results of the Bars-and-Framed-Rectangles experiment.} Log absolute error means and 95\% confidence intervals for the \emph{bars-and-framed-rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
+We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
+	\label{fig:figure12_mlae}
+\end{figure*}

### 2018-03-28 PAPER/07_results.tex
Updated paper structure.

@@
 
-\subsection{Elementary Perceptual Tasks}
+General discussion..
 
-some are good and some are bad.. why?
+\subsection{Classifiers}
 
-\begin{figure*}[h]
-	\centering
-	  \includegraphics[width=.95\linewidth]{figure1.pdf}
-  \caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure1_results}
-\end{figure*}
+\textbf{Transfer Learning using ImageNet.} Classifiers trained on imagenet are tuned towards natural images.
+While VGG19 and Xception perform better than the shallower LeNet, their full performance only develops when training from scratch.
+This shows how natural images are truely different than infographics.
+\\~\\
+\textbf{Anti-aliasing.} Does it help? Not sure yet!
 
-\textbf{Computational Perception Ranking.}
-
-Cleveland McGills Ranking - can we observe something similar?
-
-\begin{enumerate}
-	\item Position along a common scale e.g. scatter plot
-	\item Position on identical but nonaligned scales e.g. multiple scatter plots
-	\item Length e.g. bar chart
-	\item Angle \& Slope (tie) e.g. pie chart
-	\item Area e.g. bubbles
-	\item Volume, density, and color saturation (tie) e.g. heatmap
-	\item Color hue e.g. newsmap
-\end{enumerate}
-
-\noindent{\textbf{Cross-classifier variability.}}
-
-Can a neural network generalize on simple perceptual tasks?
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{cross_network_small_VGG19.pdf}
-  \caption{\textbf{Cross-classifier variability for the perceptual task of measuring length.} We use predictions of LeNet classifiers trained on different parametrizations of the \emph{curvature} elementary perceptual task and measure the mean logistic absolute error (MLAE).
-The lower score, the better.
-Classifiers trained on curves with variable position can generalize even if the axis of translation varies.
-However, classifiers trained on fixed positions of curves are not able to measure translated curves.}
-	\label{fig:cross_network}
-\end{figure}
-
-\subsection{Position-Angle Experiment}
-
-Bar charts are more accurate (Fig.~\ref{fig:figure3_mlae}) and networks converge faster (Fig.~\ref{fig:figure3_val_loss}). This is great.
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
-  \caption{\textbf{Classifier Efficiency of the Position-Angle experiment.} Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts.
-We report the MSE measure for both encodings of four different classifier on previously unseen validation data.}
-	\label{fig:figure3_val_loss}
-\end{figure}
-
-\begin{figure*}[t]
-	  \includegraphics[width=\linewidth]{figure3_mlae.pdf}
-  \caption{\textbf{Computational results of the Position-Angle experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure3_mlae}
-\end{figure*}
-
-\subsection{Position-Length Experiment}
-
-\begin{figure*}[t]
-	  \includegraphics[width=\linewidth]{figure4_mlae_with_multi.pdf}
-  \caption{\textbf{Computational results of the Position-Length experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-length experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure4_mlae}
-\end{figure*}
-
-\subsection{Bars and Framed Rectangles Experiment}
-
-First run indicates that framed rectangles perform better but we dont really know it yet.
-
-\begin{figure*}[t]
-	  \includegraphics[width=\linewidth]{figure12_mlae.pdf}
-  \caption{\textbf{Computational results of the Bars-and-Framed-Rectangles experiment.} Log absolute error means and 95\% confidence intervals for the \emph{bars-and-framed-rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure12_mlae}
-\end{figure*}
+%
+%\subsection{Elementary Perceptual Tasks}
+%
+%
+%\subsection{Position-Angle Experiment}
+%
+%
+%\subsection{Position-Length Experiment}
+%
+%\subsection{Bars and Framed Rectangles Experiment}
+%
 

### 2018-03-28 EXP/run_position_angle_from_scratch.py
Add Fig 3 from scratch.

@@ -16,7 +16,7 @@ import time
 import ClevelandMcGill as C
 
 
-EXPERIMENT = sys.argv[1] # f.e. Typ1
+EXPERIMENT = sys.argv[1] # f.e. Angle
 CLASSIFIER = sys.argv[2] # 'LeNet'
 NOISE = sys.argv[3] # True
 JOB_INDEX = int(sys.argv[4])
@@ -38,25 +38,7 @@ if NOISE == 'True':
 else:
   NOISE = False
 
-if EXPERIMENT == 'C.Figure4.multi':
-
-  #
-  # one classifier with all different types
-  #
-  def DATATYPE(data):
-    '''
-    '''
-    
-    choices = ['C.Figure4.data_to_type1', 'C.Figure4.data_to_type2',\
-              'C.Figure4.data_to_type3', 'C.Figure4.data_to_type4', 'C.Figure4.data_to_type5']
-
-    choice = np.random.choice(choices)
-    
-    return eval(choice)(data)
-
-
-else: 
-  DATATYPE = eval(EXPERIMENT)
+DATATYPE = eval(EXPERIMENT)
 
 
 
@@ -121,8 +103,8 @@ while train_counter < train_target or val_counter < val_target or test_counter <
   
   all_counter += 1
   
-  data, label = C.Figure4.generate_datapoint()
-#   print data
+  data, label = C.Figure3.generate_datapoint()
+  
   pot = np.random.choice(3)
   
   # sometimes we know which pot is right
@@ -134,16 +116,13 @@ while train_counter < train_target or val_counter < val_target or test_counter <
     pot = 2
   
   if pot == 0 and train_counter < train_target:
-    
-    #
-    try:
-      image = DATATYPE(data)
-      image = image.astype(np.float32)
-    except:
-      continue
-      
+
     if label not in train_labels:
       train_labels.append(label)
+    
+    #
+    image = DATATYPE(data)
+    image = image.astype(np.float32)
       
     # add noise?
     if NOISE:
@@ -156,15 +135,11 @@ while train_counter < train_target or val_counter < val_target or test_counter <
     
   elif pot == 1 and val_counter < val_target:
 
-    #
-    try:
-      image = DATATYPE(data)
-      image = image.astype(np.float32)
-    except:
-      continue
-    
     if label not in val_labels:
       val_labels.append(label)
+      
+    image = DATATYPE(data)
+    image = image.astype(np.float32)
       
[…]

### 2018-03-28 EXP/run_position_angle_from_scratch.py
Fix pos angle.

@@ -223,78 +223,31 @@ if CLASSIFIER == 'VGG19' or CLASSIFIER == 'XCEPTION':
   print 'memory usage', (X_train_3D.nbytes + X_val_3D.nbytes + X_test_3D.nbytes) / 1000000., 'MB'
 
   if CLASSIFIER == 'VGG19':  
-    feature_generator = keras.applications.VGG19(include_top=False, weights='imagenet', input_shape=(100,100,3))
+    feature_generator = keras.applications.VGG19(include_top=False, input_shape=(100,100,3))
   elif CLASSIFIER == 'XCEPTION':
-    feature_generator = keras.applications.Xception(include_top=False, weights='imagenet', input_shape=(100,100,3))
+    feature_generator = keras.applications.Xception(include_top=False, input_shape=(100,100,3))
   elif CLASSIFIER == 'RESNET50':
     print 'Not yet - we need some padding and so on!!!'
     sys.exit(1)
 
   t0 = time.time()
-  X_train_3D_features = feature_generator.predict(X_train_3D, verbose=True)
-  X_val_3D_features = feature_generator.predict(X_val_3D, verbose=True)
-  feature_time = time.time()-t0
 
-  X_test_3D_features = feature_generator.predict(X_test_3D, verbose=True)
-  print CLASSIFIER, 'features done after', time.time()-t0
-  print 'memory usage', (X_train_3D_features.nbytes + X_val_3D_features.nbytes + X_test_3D_features.nbytes) / 1000000., 'MB'
 
-  # update the shape
-  feature_shape = X_train_3D_features.shape[1] * X_train_3D_features.shape[2] * X_train_3D_features.shape[3]
-
+  #
+  # THE MLP
+  #
+  #
   MLP = models.Sequential()
+  MLP.add(layers.Flatten(input_shape=feature_generator.output_shape[1:]))
+  MLP.add(layers.Dense(256, activation='relu', input_dim=(100,100,3)))
+  MLP.add(layers.Dropout(0.5))
+  MLP.add(layers.Dense(5, activation='linear')) # REGRESSION
 
-  X_train = X_train_3D_features.reshape(len(X_train_3D_features), feature_shape)
-  X_val = X_val_3D_features.reshape(len(X_val_3D_features), feature_shape)
-  X_test = X_test_3D_features.reshape(len(X_test_3D_features), feature_shape)
-  ##
[…]

### 2018-03-28 EXP/run_weber_from_scratch.py
Add Weber from scratch.

@@ -1,0 +1,280 @@
+from keras import models
+from keras import layers
+from keras import optimizers
+import keras.applications
+import keras.callbacks
+from keras import backend as K
+from keras.utils.np_utils import to_categorical
+import sklearn.metrics
+
+import cPickle as pickle
+import numpy as np
+import os
+import sys
+import time
+
+import ClevelandMcGill as C
+
+
+EXPERIMENT = sys.argv[1] # f.e. Angle
+CLASSIFIER = sys.argv[2] # 'LeNet'
+NOISE = sys.argv[3] # True
+JOB_INDEX = int(sys.argv[4])
+
+#
+#
+#
+print 'Running', EXPERIMENT, 'with', CLASSIFIER, 'Noise:', NOISE, 'Job Index', JOB_INDEX
+
+#
+#
+# PROCESS SOME FLAGS
+#
+#
+SUFFIX = '.'
+if NOISE == 'True':
+  NOISE = True
+  SUFFIX = '_noise.'
+else:
+  NOISE = False
+
+DATATYPE = eval(EXPERIMENT)
+
+
+
+if os.path.abspath('~').startswith('/n/'):
+  # we are on the cluster
+  PREFIX = '/n/regal/pfister_lab/PERCEPTION/'
+else:
+  PREFIX = '/home/d/PERCEPTION/'
+RESULTS_DIR = PREFIX + 'RESULTS_FROM_SCRATCH/'
+
+OUTPUT_DIR = RESULTS_DIR + EXPERIMENT + '/' + CLASSIFIER + '/'
+if not os.path.exists(OUTPUT_DIR):
+  # here can be a race condition
+  try:
+    os.makedirs(OUTPUT_DIR)
+  except:
+    print 'Race condition!', os.path.exists(OUTPUT_DIR)
+
+STATSFILE = OUTPUT_DIR + str(JOB_INDEX).zfill(2) + SUFFIX + 'p'
+MODELFILE = OUTPUT_DIR + str(JOB_INDEX).zfill(2) + SUFFIX + 'h5'
+
+print 'Working in', OUTPUT_DIR
+print 'Storing', STATSFILE
+print 'Storing', MODELFILE
+
+if os.path.exists(STATSFILE) and os.path.exists(MODELFILE):
+  print 'WAIT A MINUTE!! WE HAVE DONE THIS ONE BEFORE!'
+  sys.exit(0)
+
+
+#
+#
+# DATA GENERATION
+#
+#
+
[…]

### 2018-03-28 PAPER/00_introduction.tex
Editing intro

@@
 
-%Artificial intelligence has taken the world of technology by storm. 
-Deep multilayer neural networks are being successfully applied in a wide range of applications that are regularly outperforming humans in object recognition \cite{krizhevsky_imagenet2012, simonyan_very_deep2014,szegedy2015}.
-Originally inspired by neuroscientific discoveries, the recent advances in deep learning have been the direct results of engineering efforts, more specifically in convolutional neural networks (CNNs).
-Our current knowledge of biological vision suggests that CNNs indeed mimic feature learning similar to the early layers of the visual cortex but without incorporating the many details of biological neural networks \cite{yamins2016using, hassabis2017neuroscience, human_vs_machine_vision}.
-While there has been significant advancement, fully understanding how CNNs work is a still ongoing process which does not detract from their success~\cite{goodfellow_book, deeplearning_blackbox2017} and recent works target the systematic evaluation of perceptional limits of CNNs~\cite{clevr, not_so_clevr}.
+Convolutional neural networks (CNNs) have been successfully applied to a wide range of visual tasks, most famously to natural image object recognition~\cite{krizhevsky_imagenet2012, simonyan_very_deep2014, szegedy2015}, for which some claim equivalent or better than human performance.
+This performance comparison is often motivated by the idea that CNNs model or reproduce the early layers of the visual cortex, even though they do not incorporate many details of biological neural networks or model higher-level abstract or symbolic reasoning~\cite{yamins2016using, hassabis2017neuroscience, human_vs_machine_vision}.
+While CNN techniques were originally inspired by neuroscientific discoveries, recent advances in processing larger datasets with deeper networks have been the direct results of engineering efforts.
+Throughout this significant advancement, researchers have aimed to understand why and how CNNs produce such high performance~\cite{goodfellow_book, deeplearning_blackbox2017}, with recent works targetting the systematic evaluation of the visual perception limits of CNNs~\cite{clevr, not_so_clevr}.
 
-We are interested in applying neural networks to graphs, charts, and visual encodings.
-A first step in this direction is to evaluate if and how convolutional neural networks perceive low-level graphical elements, the building blocks of information visualizations.
-This is a very different task than applying such networks to natural images (like the majority of computer vision research) since such scenes are usually more complex and identifiers such as edges are not as prominent.
-More related towards our goal are the \emph{graphical perception} experiments of Cleveland and McGill~\cite{cleveland_mcgill}.
-In this paper, we explore whether these experiments which were performed with humans can be reproduced with CNNs.
+One fundamental application of human vision is in understanding data visualizations.
+This is a task unlike processing natural images, where real-world objects and their effects are abstracted into data and represented with visual marks.
+As a field, visualization catalogues and evaluates human perception of these figures, such as in the seminal \emph{graphical perception} experiments of Cleveland and McGill~\cite{cleveland_mcgill}.
+This work describes nine elementary perceptual reasoning tasks, such as position relative to a scale, length, angle, area, and shading density, plus orders their reasoning difficulty.
+But, with more research interested in the machine analysis of graphs, charts, and visual encodings, it seems pertinent to question whether CNNs are able to process these basic graphical elements, and derive useful measurements from the building blocks of information visualization.
+
+As such, we explore whether these experiments which were performed with humans can be reproduced with CNNs.
 Cleveland and McGill's work in the 1980s has led to many insights for modern information visualization research such as the identification of elementary perceptional tasks or that bar charts are easier to understand than pie charts.

### 2018-03-29 PAPER/00_introduction.tex
Update paper and figures.

@@
 We include the LeNet-5 network~\cite{lenet}, the VGG19 network~\cite{simonyan_very_deep2014}, and the Xception classifiers~\cite{xception}.
-While we train LeNet-5 from scratch, for VGG19 and Xception we use pre-trained imagenet~\cite{imagenet} weights to further mimic the human visual system.
+While we train all classifiers from scratch, for VGG19 and Xception we also use pre-trained imagenet~\cite{imagenet} weights to further mimic the human visual system.
 By also using the MLP directly without convolutional feature detection as baseline, we test four different classifiers as part of the following scenarios: a) elementary perceptual tasks with increasing parameteric complexity, b) position-angle experiment comparing pie charts to bar charts, c) position-length experiment that compares grouped and divided bar charts, and d) the bars and framed rectangles experiments where visual cues aid perception.
@@
 
-We accompany this paper and detailed supplemental material with open source code~\footnote{Code, data, results and more are available at: \url{http://rhoana.org/perception}}, data, and results to enable a framework for the development and evaluation of new network architectures for graphical perception.
+We accompany this paper and detailed supplemental material with open source code\footnote{Code, data, results and more are available at: \url{http://rhoana.org/perception}}, data, and results to enable a framework for the development and evaluation of new network architectures for graphical perception.
 

### 2018-03-29 PAPER/03_elementary_perceptual_tasks.tex
Update paper and figures.

@@
+% !TeX root = paper.tex
 \section{Elementary Perceptual Tasks}
@@
 Cleveland and McGill describe the mapping of graphical elements to quantitative variables as \emph{elementary perceptual tasks} and introduce a list of ten different encodings in their paper~\cite{cleveland_mcgill}.
-We create visualizations of these tasks as rasterized images (Fig.~\ref{fig:elementary_perceptual_tasks}).
+We create visualizations of these tasks as rasterized images (Table~\ref{tab:encoding_parameters}).
 
@@
 
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure1_overview.pdf}
-  \caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of the elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
+%\begin{figure}[t]
+%	  \includegraphics[width=\linewidth]{figure1_overview.pdf}
+%  \caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of the elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
 We vary the parameters of each perceptual task and then assess the interpretability of feed-forward neural networks.}
-	\label{fig:elementary_perceptual_tasks}
-\end{figure}
-
+%	\label{fig:elementary_perceptual_tasks}
+%\end{figure}
 \begin{table}[t]
 \centering
-\caption{\textbf{Variability of Elementary Perceptual Tasks.} We sequentially increase the number of parameters for every visual encoding of the elementary perceptual tasks.
-This introduces variability and increasingly more complex datasets.}
+\caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of the elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
+We sequentially increase the number of parameters (e.g.
+by adding translation) for every task.
+This introduces variability and creates increasingly more complex datasets.}
 \resizebox{\linewidth}{!}{
-\begin{tabular}{llr}
-%	\toprule
-%	\makecell{Classifier} & \makecell{Convolutional\\Layers} & \makecell{Trainable\\Parameters} \\
-%	\midrule
-%	MLP & $0$ & $2,560,513$ \\
-%	\emph{LeNet} + MLP & $2$ & $8,026,083$ \\
-%	\emph{VGG19} + MLP & $16$ & $21,204,545$ \\
-%	\emph{Xception} + MLP & $36$ & $25,580,585$ \\
-%	\bottomrule
+\begin{tabular}{lllr}
 	\toprule
-	Elementary Perceptual Task & Variability & Parameters\\
+	\multicolumn{2}{l}{Elementary Perceptual Task} & ~ & Parameters\\
 	\midrule
-	\makecell[tl]{\emph{Position Common Scale}} & \makecell[tl]{Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $60$ \\ $3600$ \\ $21600$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{position_common_scale.pdf}} & \makecell[tl]{\emph{Position Common Scale}\\~~~Position Y\\~~~+ Position X \\~~~+ Spot Size \\} &~& \makecell[tr]{~\\ $60$ \\ $3600$ \\ $21600$}\\
 
 	\midrule
-	\makecell[tl]{\emph{Position Non-Aligned Scale}} & \makecell[tl]{Position Y\\+ Position X \\+ Spot Size \\} & \makecell[tr]{ $600$ \\ $36000$ \\ $216000$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{position_non_aligned_scale.pdf}} & \makecell[tl]{\emph{Position Non-Aligned Scale}\\~~~Position Y\\~~~+ Position X \\~~~+ Spot Size \\} &~& \makecell[tr]{~\\ $600$ \\ $36000$ \\ $216000$}\\
 
 	\midrule
-	\makecell[tl]{\emph{Length}} & \makecell[tl]{Length\\+ Position Y \\+ Position X \\+ Width} & \makecell[tr]{ $60$ \\ $2400$ \\ $144000$\\$864000$}\\
+	\raisebox{-.95\height}{\includegraphics[width=.5in]{length.pdf}} & \makecell[tl]{\emph{Length}\\~~~Length\\~~~+ Position Y \\~~~+ Position X \\~~~+ Width} &~& \makecell[tr]{ ~\\$60$ \\ $2400$ \\ $144000$\\$864000$}\\
 
 	\midrule
-	\makecell[tl]{\emph{Direction}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $360$ \\ $21600$ \\ $1296000$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{direction.pdf}} & \makecell[tl]{\emph{Direction}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$360$ \\ $21600$ \\ $1296000$}\\
 
 	\midrule
-	\makecell[tl]{\emph{Angle}} & \makecell[tl]{Angle\\+ Position Y \\+ Position X} & \makecell[tr]{ $90$ \\ $5400$ \\ $324000$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{angle.pdf}} & \makecell[tl]{\emph{Angle}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$90$ \\ $5400$ \\ $324000$}\\
 
 	\midrule
-	\makecell[tl]{\emph{Area}} & \makecell[tl]{Radius\\+ Position Y \\+ Position X} & \makecell[tr]{ $40$ \\ $800$ \\ $16000$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{area.pdf}} & \makecell[tl]{\emph{Area}\\~~~Radius\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$40$ \\ $800$ \\ $16000$}\\
 
 	\midrule
-	\makecell[tl]{\emph{Volume}} & \makecell[tl]{Cube Sidelength\\+ Position Y \\+ Position X} & \makecell[tr]{ $20$ \\ $400$ \\ $8000$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{volume.pdf}} & \makecell[tl]{\emph{Volume}\\~~~Cube Sidelength\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$20$ \\ $400$ \\ $8000$}\\
 
 	\midrule
-	\makecell[tl]{\emph{Curvature}} & \makecell[tl]{Midpoint Curvature\\+ Position Y \\+ Position X} & \makecell[tr]{ $80$ \\ $1600$ \\ $64000$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{curvature.pdf}} & \makecell[tl]{\emph{Curvature}\\~~~Midpoint Curvature\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$80$ \\ $1600$ \\ $64000$}\\
 
 	\midrule
-	\makecell[tl]{\emph{Shading}} & \makecell[tl]{Density\\+ Position Y \\+ Position X} & \makecell[tr]{ $100$ \\ $2000$ \\ $40000$}\\
-
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{shading.pdf}} & \makecell[tl]{\emph{Shading}\\~~~Density\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$100$ \\ $2000$ \\ $40000$}\\	
+%	
 	\bottomrule

### 2018-03-29 PAPER/00_introduction.tex
More intro edits.

--- 2018-03-28 PAPER/00_introduction.tex
@@
 While CNN techniques were originally inspired by neuroscientific discoveries, recent advances in processing larger datasets with deeper networks have been the direct results of engineering efforts.
-Throughout this significant advancement, researchers have aimed to understand why and how CNNs produce such high performance~\cite{goodfellow_book, deeplearning_blackbox2017}, with recent works targetting the systematic evaluation of the visual perception limits of CNNs~\cite{clevr, not_so_clevr}.
+Throughout this significant advancement, researchers have aimed to understand why and how CNNs produce such high performance~\cite{goodfellow_book, deeplearning_blackbox2017}, with recent works targeting the systematic evaluation of the visual perception limits of CNNs~\cite{clevr, not_so_clevr}.
 
-One fundamental application of human vision is in understanding data visualizations.
+One fundamental application of human vision is to understand data visualizations.
 This is a task unlike processing natural images, where real-world objects and their effects are abstracted into data and represented with visual marks.
-As a field, visualization catalogues and evaluates human perception of these figures, such as in the seminal \emph{graphical perception} experiments of Cleveland and McGill~\cite{cleveland_mcgill}.
+As a field, visualization catalogues and evaluates human perception of these marks, such as in the seminal \emph{graphical perception} experiments of Cleveland and McGill~\cite{cleveland_mcgill}.
 This work describes nine elementary perceptual reasoning tasks, such as position relative to a scale, length, angle, area, and shading density, plus orders their reasoning difficulty.
-But, with more research interested in the machine analysis of graphs, charts, and visual encodings, it seems pertinent to question whether CNNs are able to process these basic graphical elements, and derive useful measurements from the building blocks of information visualization.
+But, with increasing research interest in the machine analysis of graphs, charts, and visual encodings, it seems pertinent to question whether CNNs are able to process these basic graphical elements and derive useful measurements from the building blocks of information visualization.
 
-As such, we explore whether these experiments which were performed with humans can be reproduced with CNNs.
-Cleveland and McGill's work in the 1980s has led to many insights for modern information visualization research such as the identification of elementary perceptional tasks or that bar charts are easier to understand than pie charts.
+As such, we reproduce Cleveland and McGill's human perceptual experiments with CNNs, and discuss to what extent they have `graphical perception'.
+To perform this evaluation, we parametrize the elementary perceptual tasks and experiments suggested by Cleveland and McGill~\cite{cleveland_mcgill}, and define a set of regression tasks to estimate continuous variables.
+Against human perception, we pit four neural networks: a three-layer multilayer perceptron (MLP), the LeNet 5-layer CNN~\cite{lenet}, the VGG 19-layer CNN~\cite{simonyan_very_deep2014}, and the Xception 36-layer CNN~\cite{xception}.
+As CNNs trained on natural images are said to mimic layers of the human visual cortex, we investigate whether using weights trained on natural images (via ImageNet~\cite{imagenet}) or weights trained from scratch on elementary graphical perception tasks produces more accurate measurements and greater generalization.
 
-In order to perform this evaluation, we parametrize different visual encodings such as the elementary perceptual tasks suggested by Cleveland and McGill~\cite{cleveland_mcgill}.
-We then replicate the original experimental design by defining linear regression tasks for continuous variables.
-However, we select three modern feature generators based on convolutional neural networks and combine them with a multilayer perceptron (MLP) to include non-linearities.
-We include the LeNet-5 network~\cite{lenet}, the VGG19 network~\cite{simonyan_very_deep2014}, and the Xception classifiers~\cite{xception}.
-While we train LeNet-5 from scratch, for VGG19 and Xception we use pre-trained imagenet~\cite{imagenet} weights to further mimic the human visual system.
-By also using the MLP directly without convolutional feature detection as baseline, we test four different classifiers as part of the following scenarios: a) elementary perceptual tasks with increasing parameteric complexity, b) position-angle experiment comparing pie charts to bar charts, c) position-length experiment that compares grouped and divided bar charts, and d) the bars and framed rectangles experiments where visual cues aid perception.
-We also investigate other properties of CNNs such as whether they obey Weber's law which defines a proportional dependency between an initial distribution and perceivable change.
-Our experimental setup includes repetitions, randomizations, and regularizations to prevent any bias.
+We test these four networks across four scenarios presented by Cleveland and McGill~\cite{cleveland_mcgill}: 1) Nine elementary perceptual tasks with increasing parameteric complexity, e.g., estimating the length of a bar with fixed x, then with varying x, then with varying width, including cross-network evaluations testing the generalizability of networks to unseen parameters; 2) The position-angle experiment, which compares judgements of bar charts to pie charts, 3) The position-length experiment, which compares grouped and divided bar charts, and 4) The bars and framed rectangles experiments, where visual cues aid ratio perception.
+We also investigate whether CNNs obey Weber's law, which defines a proportional dependency between an initial distribution and perceivable change.
 
-Our motivation for replicating Cleveland and McGill's experiments stems from the thought that computational perception seems to be closely related to biological vision.
-If human perception yields certain results, maybe we can replicate these results with machine vision.
-As our first contribution, we study the elementary perceptual tasks by Cleveland and McGill and then systematically parametrize and evaluate them for computational perception with our four classifiers.
-This setup includes also cross-network evaluations which give insight into the generalizability of the classifiers.
-It also yields our second contribution: a ranking of Cleveland and McGill's elementary perceptual tasks for our tested CNN architectures.
-Further, we replicate the \emph{position-angle} and the \emph{position-length} experiments which contributes to the general knowledge of superior perceivability of bar charts to pie charts in certain conditions.
-We then reproduce the \emph{bars-and-framed-rectangles} experiment with our classifiers.
-Here, we also include an additional experiment to test the Weber-Fechner's law for CNNs.
-Our experiments yield a TODO.
-Finally, we discuss our findings and derive recommendations for allowing CNNs to perceive visualizations.
+with these experiments, we describe a ranking defining the ease with which our tested CNN architectures can estimate elementary perceptual tasks, as an equivalent to Cleveland and McGill's ranking for human perception.
+Further, we discuss the implications of our results and derive recommendations for the use of CNNs in perceiving visualizations.
+We accompany this paper with open source code and our input and results data~\footnote{Code and data are available at: \url{http://rhoana.org/perception}}, both to enable reproduction studies and to spur new machine perception systems more adept at graphical perception.
 
-We accompany this paper and detailed supplemental material with open source code~\footnote{Code, data, results and more are available at: \url{http://rhoana.org/perception}}, data, and results to enable a framework for the development and evaluation of new network architectures for graphical perception.
+%Cleveland and McGill's work in the 1980s has led to many insights for modern information visualization research such as the identification of elementary perceptional tasks or that bar charts are easier to understand than pie charts.
 

### 2018-03-29 PAPER/00_introduction.tex

@@
 
-with these experiments, we describe a ranking defining the ease with which our tested CNN architectures can estimate elementary perceptual tasks, as an equivalent to Cleveland and McGill's ranking for human perception.
+With these experiments, we describe a ranking defining the ease with which our tested CNN architectures can estimate elementary perceptual tasks, as an equivalent to Cleveland and McGill's ranking for human perception.
 Further, we discuss the implications of our results and derive recommendations for the use of CNNs in perceiving visualizations.

### 2018-03-29 PAPER/02_experimental_setup.tex

@@
 
-The experiments shown in this paper are either supervised regression or classification tasks.
+We conduct quantitative experiments to measure the ``perception'' of low-level visual encodings such as positions, angles, curvature, and lengths for different convolutional neural networks.
+We treat each classifier configuration as a subject in a between-subjects experiment evaluate supervised learning algorithms.
 We formulate any estimation of quantities (e.g.
-angles, positions, lengths etc.) as a regression problem between $0$ and $1$.
+angles, positions, lengths etc.) as a logistic regression problem with single or multiple outputs between $0$ and $1$.
 The output indicates the percentage in regards to the degrees of freedom of the individual experiment.
-If the experiment involves a choice, we formulate it as a classification problem.
+As baseline, we use a standard multilayer perceptron (MLP) without the generation of additional feature maps through convolutions (Fig.~\ref{fig:classifiers}).
 
@@
 
-\textbf{Accuracy.} We use the same metric as Cleveland and McGill to measure accuracy.
+\textbf{Accuracy.} Cleveland and McGill use the midmean logistic absolute error metric (\emph{MLAE}) to measure perception accuracy.
 \begin{equation}
-	\log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
+	\textnormal{MLAE} = log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
 \end{equation}
-\\~\\
-\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the confidence intervals.
+In addition to this metric, we also calculate standard error metrics such as the mean squared error (\emph{MSE}) and the mean absolute error (\emph{MAE}).
+This allow a more direct comparison of percent errors.
 \\~\\
 \noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
-For regression tasks the loss is defined as mean squared error (MSE) and for classification tasks the loss is categorical cross-entropy.
+For regression tasks the loss is defined as mean squared error (MSE).% and for classification tasks the loss is categorical cross-entropy.
+\\~\\
+\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the $95\%$ confidence intervals but rather than performing bootstrapping, we approximate the value of the $97.5$ percentile point of the normal distribution for simplicity.
 
@@
 \centering
-\caption{We use different feature generators as input to a multilayer perceptron which performs linear regression or the classification task.
+\caption{\textbf{Classifier Training.} We use different feature generators as input to a multilayer perceptron which performs linear regression.
 This yields different sets of trainable parameters.
@@
 \end{table}
-
-\begin{figure}[t]
-	\centering
-	  \includegraphics[width=.8\linewidth]{classifiers.pdf}
-  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
+%\begin{figure}[t]
+%	\centering
+%	  \includegraphics[width=\linewidth]{classifiers.pdf}
+%  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
 We use Dropout regularization to prevent overfitting.
@@
 layers, filter size $3\times3$) to increase the number of trainable parameters.}
+%	\label{fig:classifiers}
+%\end{figure}
+\begin{figure}[t]
+	\centering
+
+    \subfloat[Feature Generation]{
+		\includegraphics[width=4.8cm,valign=c]{classifier_left.pdf}
+		\vphantom{\includegraphics[width=3.5cm,valign=c]{classifier_right.pdf}}
+	}
+	\hfill
+    \subfloat[Multilayer Perceptron]{
+		\includegraphics[width=3.5cm,valign=c]{classifier_right.pdf}
+	}
+
+  \caption{\textbf{Classifier Architecture.} The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
+We use Dropout regularization to prevent overfitting.
+As output, we perform linear regression for continuous variables.
+The MLP can learn the visualizations directly but we also learn features generated by LeNet (2 conv.
+layers, filter size $5\times5$), VGG19 (16 conv.
+layers, filter size $3\times3$), or Xception (36 conv.
+layers, filter size $3\times3$) to test different model complexities.
+}
 	\label{fig:classifiers}
@@
 The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}.
-We add subtle random noise ($0.05$) to each pixel to introduce additional variation.
+\\~\\
+\noindent\textbf{Noise.} We add subtle random noise ($0.05$) to each pixel to introduce additional variation.
 \\~\\

### 2018-03-29 PAPER/01_related_work.tex
Update sections 2,3,4.

@@
 
-\textbf{Computational Visualization Understanding.}
-
-Pineo \textit{et al.}~\cite{Pineo2012_computational_perception} create computational model of human vision based on neural networks and their experiments show that understanding visualization triggers neural activity in high-level areas of cognition.
+\noindent{\textbf{Computational Visualization Understanding.}} Pineo \textit{et al.}~\cite{Pineo2012_computational_perception} create computational model of human vision based on neural networks and their experiments show that understanding visualization triggers neural activity in high-level areas of cognition.
 The authors suspect that this level of understanding is produced by low-level neurons performing elementary perceptional tasks.
@@
 \\~\\
-\textbf{Visual Cortex Inspired Machine Learning.} 
-%Classifiers mimmicking the human visual system are a hot topic and many different architectures, models, and paradigms exist. Feed-forward neural networks 
-The human visual cortex is an extremely powerful system which allows the ability, and seemingly without effort, to recognize an enormous amount of distinct objects in the world.
+\noindent{\textbf{Visual Cortex Inspired Machine Learning.}} The human visual cortex is an extremely powerful system which allows the ability, and seemingly without effort, to recognize an enormous amount of distinct objects in the world.
 This visual system is organized in layers and has inspired the theory of computational classifiers based on multilayer neural networks.

### 2018-03-29 PAPER/02_experimental_setup.tex
Update sections 2,3,4.

@@
 
-We conduct quantitative experiments to measure the ``perception'' of low-level visual encodings such as positions, angles, curvature, and lengths for different convolutional neural networks.
-We treat each classifier configuration as a subject in a between-subjects experiment evaluate supervised learning algorithms.
-We formulate any estimation of quantities (e.g.
-angles, positions, lengths etc.) as a logistic regression problem with single or multiple outputs between $0$ and $1$.
-The output indicates the percentage in regards to the degrees of freedom of the individual experiment.
+We conduct quantitative experiments to measure how different convolutional neural networks ``perceive'' low-level visual encodings such as positions, angles, curvature, and lengths.
+We treat each network as a subject in a between-subjects experiment to evaluate supervised learning algorithms.
+For this, we formulate any estimation of single quantities or relations between multiple quantities as a logistic regression problem with single or multiple outputs between $0$ and $1$.
+The output indicates the percentage in regards to the degrees of freedom of the individual experiment but all experiments have different sets of parameters.
 As baseline, we use a standard multilayer perceptron (MLP) without the generation of additional feature maps through convolutions (Fig.~\ref{fig:classifiers}).
+For our experiments, we use a series of single factor between-subject designs with the factor \emph{classifier} (MLP, LeNet, VGG19, Xception, VGG19+ImageNet, Xception+ImageNet).
+Each experiment includes a training and a testing phase on data without any overlap (Section~\ref{sec:data}).
+All parameters such as network hyperparameters, optimization methods, and stopping conditions are fixed (Section~\ref{sec:classifiers}).
+However, since the classifiers are of different complexity, the number of trainable parameters changes (Table~\ref{tab:parameters}).
+We define a series of hypotheses prior to each experiment.
 
-\subsection{Measures}
+\subsection{Measures and Analysis}
 
-\textbf{Accuracy.} Cleveland and McGill use the midmean logistic absolute error metric (\emph{MLAE}) to measure perception accuracy.
+\textbf{Accuracy.} In their 1984 paper, Cleveland and McGill use the midmean logistic absolute error metric (\emph{MLAE}) to measure perception accuracy.
 \begin{equation}
@@
 \noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
-For regression tasks the loss is defined as mean squared error (MSE).% and for classification tasks the loss is categorical cross-entropy.
+For regression tasks the loss is defined as MSE.% and for classification tasks the loss is categorical cross-entropy.
 \\~\\
 \noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the $95\%$ confidence intervals but rather than performing bootstrapping, we approximate the value of the $97.5$ percentile point of the normal distribution for simplicity.
+\\~\\
+\noindent{\textbf{Confirmatory Data Analysis.}} To accept or reject our hypotheses, we calculate these measures and treat them as continuous variables.
+We analyze these dependent variables using analysis of variance (ANOVA) followed by parametric tests.
 
 \subsection{Classifiers}
-
+\label{sec:classifiers}
 Our classifiers are built upon a multilayer perceptron (MLP) which is a feedforward artificial neural network.
 We combine this MLP with different convolutional neural networks (CNNs) for preprocessing and feature generation.
-These include the traditional LeNet trained from scratch, as well as VGG19 and Xception trained using ImageNet.
+These include the traditional LeNet-5 network as well as the VGG19 and Xception networks.
 \\~\\
 \noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units~(Fig.~\ref{fig:classifiers}).
-We then add a dropout layer to prevent overfitting and compute linear regression or classification (softmax).
+We then add a dropout layer to prevent overfitting and compute linear regression with MSE loss.
 \\~\\
 \noindent{\textbf{Convolutional Neural Networks.}} We use CNNs to generate additional features as input to the MLP.
-We train the \emph{LeNet} classifier with tune it specifically towards each visualization.
-For \emph{VGG19} and \emph{Xception}, we generate features using previously trained weights on ImageNet.
+We train the \emph{LeNet} classifier, the \emph{VGG19} network, and the \emph{Xception} model specifically towards each visualization.
+For \emph{VGG19} and \emph{Xception}, we additionally generate features using previously trained weights on ImageNet to naively model human vision.
 \\~\\
@@
 \\~\\
-\noindent{\textbf{Environment.}} We run all experiments on an NVIDIA DGX1 machine with Tesla V100 graphical processing units. We use the KERAS framework with tensorflow.
+\noindent{\textbf{Environment.}} We run all experiments on Tesla X and Tesla V100 graphical processing units.
+We use the KERAS framework with tensorflow and leverage the scikit-image and scikit-learn libraries.
 
@@
 \caption{\textbf{Classifier Training.} We use different feature generators as input to a multilayer perceptron which performs linear regression.
-This yields different sets of trainable parameters.
-We also train the MLP directly on the visualizations without any additional feature generation.}
+This results in different sets of trainable parameters.
+We also train the MLP directly on the visualizations without any additional feature generation as baseline.}
 \resizebox{\linewidth}{!}{
@@
 layers, filter size $3\times3$), or Xception (36 conv.
-layers, filter size $3\times3$) to test different model complexities.
-}
+layers, filter size $3\times3$) to test different model complexities.}
 	\label{fig:classifiers}
@@
 \subsection{Data}
-
-We create all visualizations as parametrized rasterized images without interpolation.
+\label{sec:data}
+We create all visualizations as parametrized rasterized images without interpolation (except for the anti-alias experiment).
 The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}.
@@
 This eliminates leakage during training and evaluation.
-We also scale each set independently: images to the range of $-.5$ to $.5$ and labels to the range of $0.0$ to $1.0$.
+We also scale each set independently: images to the range of $-0.5$ to $0.5$ and labels to the range of $0.0$ to $1.0$.
 \\~\\
-\noindent\textbf{Cross-classifier variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.
+\noindent\textbf{Cross Validation.} For reproducibility, we perform repeated random sub-sampling validation, also known as Monte Carlo cross-validation, during our experiments.
+We run every experiment seperately multiple times (at least 4, up to 12) and randomly select (without replacement) the $60\%$ of our data as training data, $20\%$ as validation, and $20\%$ as test.
+Our large data sample size of 100,000 guarantees that every single observation of our parameterizations will be selected at least once (excluding noise patterns).
+Finally, we average the results over the runs.
+\\~\\
+\noindent\textbf{Intra-classifier Variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.
+\\~\\
+\noindent\textbf{Multi Classifiers.} Some experiments compare different types of visual encodings.
+We first train and evaluate each classifier on one type of encoding but we also train the classifiers on a random selection of multiple different types.
+These scenarios afflict the optimization process to find global minima and result in networks with more flexible knowledge with the caveat of longer training times.

### 2018-03-29 PAPER/03_elementary_perceptual_tasks.tex
Update sections 2,3,4.

@@
 Cleveland and McGill describe the mapping of graphical elements to quantitative variables as \emph{elementary perceptual tasks} and introduce a list of ten different encodings in their paper~\cite{cleveland_mcgill}.
-We create visualizations of these tasks as rasterized images (Table~\ref{tab:encoding_parameters}).
+These tasks are the low-level building blocks for information visualizations and encode quantities.
+Cleveland and McGill did not explicitly test human perception of single instances of these encodings.
+However, we test how each classifier measures encoded values using the elementary perceptual tasks and create visualizations of these tasks as rasterized images with different parametrizations (Table~\ref{tab:encoding_parameters}).
 
@@
 \label{sec:parametrizations}
-We generate multiple parameterizations for each elementary perceptual task (Fig.~\ref{fig:elementary_perceptual_tasks}) and sequentially increase the number of parameters.
-For instance, for \emph{position non-aligned scale} we first only vary the origin of the coordinate system which yields just 10 different parameters.
-We then include translation along the y-axis with a significant increase in variability.
-We then also add x-movement and a variable spot size.
-This results in more complex datasets depending on the variability setting.
+We generate multiple parameterizations for each elementary perceptual task and sequentially increase the number of parameters (Table~\ref{tab:encoding_parameters}).
+For instance, for \emph{Position Common Scale} we first only vary the $y$-position which yields just $60$ different parameters.
+We then include translation along the x-axis with a significant increase in variability.
+We then also add a variable spot size.
+This results in more complex datasets depending on the increase of variability.
 Table~\ref{tab:encoding_parameters} shows the different settings.
 It is important to consider this variability when evaluating different classifiers with individual trainable parameters (Table~\ref{tab:parameters}).
+In theory, classifiers can memorize the images if the data set has a low variability.
+We also counteract such behavior by adding noise.
 
@@
 %\end{figure}
-\begin{table}[t]
+\begin{table}[h]
 \centering

### 2018-03-29 EXP/run_position_angle_from_scratch.py
Fix bug aaaahhhh

@@ -275,7 +275,7 @@ print 'Fitting done', time.time()-t0
 # PREDICTION
 #
 #
-y_pred = MLP.predict(X_test_3D)
+y_pred = model.predict(X_test_3D)
 
 
 #

### 2018-03-29 EXP/run_position_length_from_scratch.py

@@ -304,7 +304,7 @@ print 'Fitting done', time.time()-t0
 # PREDICTION
 #
 #
-y_pred = MLP.predict(X_test_3D)
+y_pred = model.predict(X_test_3D)
 
 
 #

### 2018-03-29 EXP/run_weber_from_scratch.py

@@ -245,7 +245,7 @@ print 'Fitting done', time.time()-t0
 # PREDICTION
 #
 #
-y_pred = MLP.predict(X_test_3D)
+y_pred = model.predict(X_test_3D)
 
 
 #

### 2018-03-29 PAPER/01_related_work.tex
Copy editing.

@@
 
-\textbf{Graphical Perception.} Cleveland and McGill~\cite{cleveland_mcgill} introduce the fundamental concept of \emph{graphical perception} and investigate how different visual attributes and encodings are perceivable by humans.
-They define \emph{elementary perceptual tasks} as mental-visual stimuli to understand encodings in visualizations.
-Based on these definitions, the authors propose and perform different experiments such as the \emph{position-angle} experiment which compares bar charts and pie charts, the \emph{position-length} experiment where users judge relations between encoded values in grouped and divided bar charts, and the \emph{bars-and-framed-rectangles} experiment to evaluate Weber's law.
-Heer and Bostock later reproduced the Cleveland-McGill experiments crowd-sourced on Mechanical Turk~\cite{HeerBostock2010} which lead to follow-up work from Harrison \textit{et al.}~\cite{harrison2013influencing} who replicated the experiments while observing emotional states.
-Both papers report similar results to Cleveland and McGill which increased our motivation to mimmick their pioneering work.
-Our experimental setup replicates the original setup of Cleveland and McGill - just instead of humans, we use convolutional neural networks due to the connection with the human visual system.
-While we focus on Cleveland and McGill's work from 1984, many other excellent articles from the last decades target low-level visual encoding~\cite{bertin1967semiologie,cleveland1985graphical,treisman1988feature, wilkinson2006grammar, carpendale2003considering,widgor_perception2007,munzner2015visualization}.
+\paragraph{Graphical Perception} Cleveland and McGill~\cite{cleveland_mcgill,cleveland1985graphical} coin the phrase \emph{graphical perception} to describe how different visual attributes and encodings are perceived by humans.
+They define \emph{elementary perceptual tasks} as mental-visual stimuli to understand encodings in visualizations, and declare a ranking based on their perceptual difficulty.
+From these definitions, the authors propose and perform different experiments such as the \emph{position-angle} experiment which compares bar charts and pie charts, the \emph{position-length} experiment where users judge relations between encoded values in grouped and divided bar charts, and the \emph{bars-and-framed-rectangles} experiment to evaluate Weber's law \cite{harrison2014_webers_law_rank} using the proportional relation between an initial distribution density and perceivable change.
 
-Interesting are also the rankings of correlation visualization using Weber's law~\cite{harrison2014_webers_law_rank}.
+Heer and Bostock later reproduced the Cleveland-McGill experiments via crowd-sourcing on Mechanical Turk~\cite{HeerBostock2010}, with similar results.
+Harrison \textit{et al.}~\cite{harrison2013influencing} repeated the Cleveland-McGill experiments while observing viewer emotional states, again with similar results.
+Our experimental setup again reproduces the Cleveland-McGill experiments, but instead of judging human perception, we judge machine perception using convolutional neural networks.
+While we focus on Cleveland and McGill's work from the mid 1980s due to their repeated reproduction, many other works also investigate human perception to visual encoding~\cite{bertin1967semiologie,treisman1988feature,carpendale2003considering,wilkinson2006grammar,widgor_perception2007,munzner2015visualization}.
+
+%Interesting are also the rankings of correlation visualization using Weber's law~\cite{harrison2014_webers_law_rank}.
 This law defines the proportional relation between the initial distribution density and perceivable change.
-In this paper, we investigate with a simple experiment whether this holds for convolutional neural networks.
-\\~\\
+In this paper, we investigate with a simple experiment whether this holds for convolutional neural networks
+
 %\textbf{Comparing Visual Encodings.} Different visual encodings have advantages or disadvantages and the community does a great job comparing them.
@@
 
-\noindent{\textbf{Computational Visualization Understanding.}} Pineo \textit{et al.}~\cite{Pineo2012_computational_perception} create computational model of human vision based on neural networks and their experiments show that understanding visualization triggers neural activity in high-level areas of cognition.
-The authors suspect that this level of understanding is produced by low-level neurons performing elementary perceptional tasks.
-We are further investigating this suspision.
-Other work tries to parse infographics by finding higher-level saliency models~\cite{bylinskii2016should}, or by extracting text or key visual elements~\cite{diagram_understanding,kembhavi2016diagram,zoya_text_visual_tags}.
-However, none of these works focus on computational understanding of lower-level building blocks of visualizations such as curvature, lengths, or position.
-\\~\\
-\noindent{\textbf{Visual Cortex Inspired Machine Learning.}} The human visual cortex is an extremely powerful system which allows the ability, and seemingly without effort, to recognize an enormous amount of distinct objects in the world.
-This visual system is organized in layers and has inspired the theory of computational classifiers based on multilayer neural networks.
-Fukushima and Miyake developed the Neocognitron quantitative model~\cite{fukushima1982neocognitron} that ultimately led to the important work of Hinton, Bengio, and LeCun: \textit{deep neural networks}~\cite{lecun2015deep}, visual cortex inspired machine learning.
-Nowadays, such classifiers exist with many different architectures.
-For this paper, we select the traditional \emph{LeNet-5}~\cite{lenet} which was designed to recognize hand-written digits, the VGG19~\cite{simonyan_very_deep2014} classifier with 16 convolutional layers, and the Xception~\cite{xception} classifier with 36 convolutional layers.
-Selecting these specific networks allows us to compare archtitectures with different depths.
+\paragraph{Computational Visualization Understanding} Pineo \textit{et al.}~\cite{Pineo2012_computational_perception} create a computational model of human vision based on neural networks.
+Their simulations show that understanding visualization triggers neural activity in high-level areas of cognition, with the authors suspecting that this activity is supported by low-level neurons performing elementary perceptional tasks.
+Other work tries to parse infographics by finding higher-level saliency models~\cite{bylinskii2016should}, or by extracting text or key visual elements from visualizations~\cite{diagram_understanding,kembhavi2016diagram,zoya_text_visual_tags}.
+However, these works do not investigate computational understanding of elementary perceptual tasks such as curvature, lengths, or position, which are the building blocks of visualization.
+
+%CNNs trained on natural images have low-level `neurons' which activate on similar stimuli to early neurons in the human visual cortex, e.g., Gabor filter-like stimuli \misscite.
+
+\paragraph{Visual-cortex-inspired Machine Learning} The human visual cortex allows us to recognize objects in the world seemingly without effort (though few remember their infancy).
+This visual system is organized in layers, which inspired computational classifiers based on multilayer neural networks.
+Fukushima and Miyake developed the early Neocognitron quantitative model~\cite{fukushima1982neocognitron}, which ultimately led to the work of Hinton, Bengio, and LeCun~\cite{lecun2015deep} and today's GPU-powered \emph{deep} neural networks.
+Such networks exist with many architectures.
+For this paper, we select the traditional LeNet-5 which was designed to recognize hand-written digits with 5 layers~\cite{lenet}, the VGG19 network with 16 convolutional layers~\cite{simonyan_very_deep2014}, and the Xception network with 36 convolutional layers ~\cite{xception}.
+This allows us to compare archtitectures with different depths.
 %Object detection is extremely difficult and therefore is especially impressive as light intensities can change by levels of magnitude and contrast between foreground and background is so often low. 

### 2018-03-29 EXP/run_bar_framed_rectangle_from_scratch.py
Update output path.

@@ -47,7 +47,7 @@ if os.path.abspath('~').startswith('/n/'):
   PREFIX = '/n/regal/pfister_lab/PERCEPTION/'
 else:
   PREFIX = '/home/d/PERCEPTION/'
-RESULTS_DIR = PREFIX + 'RESULTS/'
+RESULTS_DIR = PREFIX + 'RESULTS_FROM_SCRATCH/'
 
 OUTPUT_DIR = RESULTS_DIR + EXPERIMENT + '/' + CLASSIFIER + '/'
 if not os.path.exists(OUTPUT_DIR):

### 2018-03-29 PAPER/02_experimental_setup.tex
Add just lil.

@@
 
-We conduct quantitative experiments to measure how different convolutional neural networks ``perceive'' low-level visual encodings such as positions, angles, curvature, and lengths.
-We treat each network as a subject in a between-subjects experiment to evaluate supervised learning algorithms.
-For this, we formulate any estimation of single quantities or relations between multiple quantities as a logistic regression problem with single or multiple outputs between $0$ and $1$.
-The output indicates the percentage in regards to the degrees of freedom of the individual experiment but all experiments have different sets of parameters.
+We conduct quantitative experiments to measure how different convolutional neural networks perceive low-level visual encodings, such as positions, angles, curvatures, and lengths.
+We formulate measurement tasks as logistic regression problems, where the networks must estimate single quantities or ratio relations between multiple quantities.
+The output between $0$ and $1$ indicates the percent of the value to be measured in the individual experiment.
+
+% The output between $0$ and $1$ indicates the percent of the value to be measured in the individual experiment.
+
 As baseline, we use a standard multilayer perceptron (MLP) without the generation of additional feature maps through convolutions (Fig.~\ref{fig:classifiers}).
-For our experiments, we use a series of single factor between-subject designs with the factor \emph{classifier} (MLP, LeNet, VGG19, Xception, VGG19+ImageNet, Xception+ImageNet).
-Each experiment includes a training and a testing phase on data without any overlap (Section~\ref{sec:data}).
+For our experiments, we use a series of single factor between-subject designs with the factor \emph{network} (MLP, LeNet, VGG19, Xception, VGG19+ImageNet, Xception+ImageNet).
+We train each network in a supervised fashion, with each experiment including a training and a testing phase on data.
+without any overlap (Section~\ref{sec:data}).
 All parameters such as network hyperparameters, optimization methods, and stopping conditions are fixed (Section~\ref{sec:classifiers}).
-However, since the classifiers are of different complexity, the number of trainable parameters changes (Table~\ref{tab:parameters}).
+However, since the networks are of different complexity, the number of trainable parameters changes (Table~\ref{tab:parameters}).
 We define a series of hypotheses prior to each experiment.
@@
 \\~\\
-\noindent\textbf{Noise.} We add subtle random noise ($0.05$) to each pixel to introduce additional variation.
+\noindent\textbf{Noise.} We add subtle random noise ($0.05$) to each pixel to introduce additional variation. This prevents the classifiers of learning the actual pixels.
 \\~\\

### 2018-03-29 PAPER/01_related_work.tex
Update several things and add supplemental template.

@@
 Our experimental setup again reproduces the Cleveland-McGill experiments, but instead of judging human perception, we judge machine perception using convolutional neural networks.
-While we focus on Cleveland and McGill's work from the mid 1980s due to their repeated reproduction, many other works also investigate human perception to visual encoding~\cite{bertin1967semiologie,treisman1988feature,carpendale2003considering,wilkinson2006grammar,widgor_perception2007,munzner2015visualization}.
+While we focus on Cleveland and McGill's work from the mid 1980s due to their repeated reproduction, many other works also investigate human perception to visual encoding~\cite{mackinlay1988applying,bertin1967semiologie,treisman1988feature,carpendale2003considering,wilkinson2006grammar,widgor_perception2007,munzner2015visualization}.
 

### 2018-03-29 PAPER/03_elementary_perceptual_tasks.tex
Update several things and add supplemental template.

@@
 	\toprule
-	\multicolumn{2}{l}{Elementary Perceptual Task} & ~ & Parameters\\
+	\multicolumn{2}{l}{Elementary Perceptual Task} & ~ & Permutations\\
 	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{position_common_scale.pdf}} & \makecell[tl]{\emph{Position Common Scale}\\~~~Position Y\\~~~+ Position X \\~~~+ Spot Size \\} &~& \makecell[tr]{~\\ $60$ \\ $3600$ \\ $21600$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{position_common_scale.pdf}} & \makecell[tl]{\emph{Position Common Scale}\\~~~Position Y\\~~~+ Position X \\~~~+ Spot Size \\} &~& \makecell[tr]{~\\ $60$ \\ $3,600$ \\ $216,00$}\\
 
 	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{position_non_aligned_scale.pdf}} & \makecell[tl]{\emph{Position Non-Aligned Scale}\\~~~Position Y\\~~~+ Position X \\~~~+ Spot Size \\} &~& \makecell[tr]{~\\ $600$ \\ $36000$ \\ $216000$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{position_non_aligned_scale.pdf}} & \makecell[tl]{\emph{Position Non-Aligned Scale}\\~~~Position Y\\~~~+ Position X \\~~~+ Spot Size \\} &~& \makecell[tr]{~\\ $600$ \\ $36,000$ \\ $216,000$}\\
 
 	\midrule
-	\raisebox{-.95\height}{\includegraphics[width=.5in]{length.pdf}} & \makecell[tl]{\emph{Length}\\~~~Length\\~~~+ Position Y \\~~~+ Position X \\~~~+ Width} &~& \makecell[tr]{ ~\\$60$ \\ $2400$ \\ $144000$\\$864000$}\\
+	\raisebox{-.95\height}{\includegraphics[width=.5in]{length.pdf}} & \makecell[tl]{\emph{Length}\\~~~Length\\~~~+ Position Y \\~~~+ Position X \\~~~+ Width} &~& \makecell[tr]{ ~\\$60$ \\ $2,400$ \\ $144,000$\\$864,000$}\\
 
 	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{direction.pdf}} & \makecell[tl]{\emph{Direction}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$360$ \\ $21600$ \\ $1296000$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{direction.pdf}} & \makecell[tl]{\emph{Direction}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$360$ \\ $21,600$ \\ $1,296,000$}\\
 
 	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{angle.pdf}} & \makecell[tl]{\emph{Angle}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$90$ \\ $5400$ \\ $324000$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{angle.pdf}} & \makecell[tl]{\emph{Angle}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$90$ \\ $5,400$ \\ $324,000$}\\
 
 	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{area.pdf}} & \makecell[tl]{\emph{Area}\\~~~Radius\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$40$ \\ $800$ \\ $16000$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{area.pdf}} & \makecell[tl]{\emph{Area}\\~~~Radius\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$40$ \\ $800$ \\ $16,000$}\\
 
 	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{volume.pdf}} & \makecell[tl]{\emph{Volume}\\~~~Cube Sidelength\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$20$ \\ $400$ \\ $8000$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{volume.pdf}} & \makecell[tl]{\emph{Volume}\\~~~Cube Sidelength\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$20$ \\ $400$ \\ $8,000$}\\
 
 	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{curvature.pdf}} & \makecell[tl]{\emph{Curvature}\\~~~Midpoint Curvature\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$80$ \\ $1600$ \\ $64000$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{curvature.pdf}} & \makecell[tl]{\emph{Curvature}\\~~~Midpoint Curvature\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$80$ \\ $1,600$ \\ $64,000$}\\
 
 	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{shading.pdf}} & \makecell[tl]{\emph{Shading}\\~~~Density\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$100$ \\ $2000$ \\ $40000$}\\	
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{shading.pdf}} & \makecell[tl]{\emph{Shading}\\~~~Density\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$100$ \\ $2,000$ \\ $40,000$}\\	
 %	

### 2018-03-29 PAPER/04_position_angle_experiment.tex
Update several things and add supplemental template.

@@
 
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_overview}
-  \caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
+%\begin{figure}[t]
+%	  \includegraphics[width=\linewidth]{figure3_overview}
+%  \caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
 The experimental task involves the judgement of different encoded values in comparison to the largest encoded values.
@@
 In their paper, Cleveland and McGill report less errors using bar charts.}
-	\label{fig:position_angle_experiment}
-\end{figure}
+%	\label{fig:position_angle_experiment}
+%\end{figure}
+\begin{table}[h]
+\centering
+\caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
+The experimental task involves the judgement of different encoded values in comparison to the largest encoded values.
+The pie chart and the bar chart visualize the same data point.
+In their paper, Cleveland and McGill report less errors using bar charts.}
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lllr}
+	\toprule
+	\multicolumn{2}{l}{~} & ~ & Permutations\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure3_Bar_Chart.pdf}} & \makecell[tl]{Type 1: \emph{Bar Chart}\\~~~Perceptual Task: \emph{Position}\\~ \\~ \\} &~& \makecell[tr]{~\\ $244,827$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure3_Pie_Chart.pdf}} & \makecell[tl]{Type 2: \emph{Pie Chart}\\~~~Perceptual Task: \emph{Angle}\\~ \\~ \\} &~& \makecell[tr]{~\\ $244,827$}\\
+
+	\bottomrule
+\end{tabular}
+}
+\label{tab:pos_angle_parameters}
+\end{table}
 
@@
 \begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure3_mlae_better.pdf}
+	  \includegraphics[width=.8\linewidth]{figure3_mlae_better.pdf}
   \caption{\textbf{Computational results of the Position-Angle experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.

### 2018-03-29 PAPER/05_position_length_experiment.tex
Add Fig 3 from scratch.
Update several things and add supplemental template.

@@
 
-\begin{figure*}[t]
-	  \includegraphics[width=\linewidth]{figure4_overview.pdf}
-
-  \caption{\textbf{Position-Length Experiment.} (Not yet) Rasterized versions of the graphs of Cleveland and McGill's position-length experiment.
+%\begin{figure*}[t]
+%	  \includegraphics[width=\linewidth]{figure4_overview.pdf}
+%	  
+%  \caption{\textbf{Position-Length Experiment.} (Not yet) Rasterized versions of the graphs of Cleveland and McGill's position-length experiment.
 The perceptual task involves comparing.
@@
 The first task is to select which of the marked quantities is smaller (classification) and the second task is to specify how much smaller it is (regression).}
-	\label{fig:position_length_experiment}
-\end{figure*}
+%	\label{fig:position_length_experiment}
+%\end{figure*}
+\begin{table}[h]
+\centering
+\caption{\textbf{Position-Length Experiment.} Rasterized versions of the graphs of Cleveland and McGill's position-length experiment.
+The perceptual task involves comparing the two dot-marked quantities across five different visual encodings of either grouped or divided bar charts.
+We evaluate which type of bar chart performs better with our neural networks.
+The two marked values are chosen from a set of ten pairs which defines the dual regression task.
+Since the other 8 values are chosen randomly, the parameter space for images of this experiment is massive.}
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lllr}
+	\toprule
+	\multicolumn{2}{l}{~} & ~ & Permutations\\
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure4_type_1.pdf}} & \makecell[tl]{Type 1: \emph{Grouped Bar Chart}\\~~~Perceptual Task: \emph{Position}\\~ \\~ \\} &~& \makecell[tr]{~\\ $9.20E+16$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure4_type_2.pdf}} & \makecell[tl]{Type 2: \emph{Divided Bar Chart}\\~~~Perceptual Task: \emph{Position}\\~ \\~ \\} &~& \makecell[tr]{~\\ $9.20E+16$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure4_type_3.pdf}} & \makecell[tl]{Type 3: \emph{Grouped Bar Chart}\\~~~Perceptual Task: \emph{Position}\\~ \\~ \\} &~& \makecell[tr]{~\\ $9.20E+16$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure4_type_4.pdf}} & \makecell[tl]{Type 4: \emph{Divided Bar Chart}\\~~~Perceptual Task: \emph{Length}\\~ \\~ \\} &~& \makecell[tr]{~\\ $9.20E+16$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure4_type_5.pdf}} & \makecell[tl]{Type 5: \emph{Divided Bar Chart}\\~~~Perceptual Task: \emph{Length}\\~ \\~ \\} &~& \makecell[tr]{~\\ $9.20E+16$}\\
+
+	\bottomrule
+\end{tabular}
+}
+\label{tab:pos_length_parameters}
+\end{table}
 
@@
 \begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure4_mlae_with_multi_and_humans.pdf}
+	\centering
+	  \includegraphics[width=.8\linewidth]{figure4_mlae_with_multi_and_humans.pdf}
   \caption{\textbf{Computational results of the Position-Length experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-length experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.

### 2018-03-29 PAPER/SUPPLEMENTAL/supplemental.tex
Update several things and add supplemental template.

--- 2018-03-21 PAPER/paper.tex
@@
 \usepackage{multirow}
+\usepackage[export]{adjustbox}
+\usepackage{subfig}
 
@@
 % - 
-\title{Reproducing 'Graphical Perception' with CNNs}
+\title{Supplemental Material for\\Evaluating `Graphical Perception' with CNNs}
 
@@
 %other entries to be set up for journal
-\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Computing Graphical Perception with Convolutional Neural Networks}
+\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Evaluating `Graphical Perception' with CNNs}
 %\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
-
-%% Abstract section.
-\abstract{%Convolutional neural networks are being successfully used for image understanding and object recognition tasks while regularly outperforming humans.
-Despite their tremendous success, little is known perceptional capabilities.
-In there Graphical Perception paper from 1984, Cleveland and McGill define elementary perceptual tasks that let people extract quantitative information from visualizations and measure human perception of different visual encodings with in depth user studies.
-%We replicate the experimental setup of Cleveland and McGill and evaluate the perceptional capabilities of four modern classifiers based on neural networks.
-We systematically test how the classifers perform on a) elementary perceptual tasks with increasing parametric complexity, b) the position-angle experiment which compares pie and bar charts, c) the position-length experiment which compares grouped and divided bar charts, and d) the bars and framed rectangles experiment where visual cues help to perceive information.
-We present the results of these experiments to foster the understanding of how such classifiers can be successfully applied to data visualizations.
-On this journey, we also study how the feed-forward neural networks obey Weber's law which defines the proportional relation between perceivable information and distribution size.
-We further introduce a ranking of elementary visual encodings targeted towards modern deep neural networks and derive practical evidence for properties of convolutional neural networks such as the translation invariance.
-Convolutional neural networks can successfully perform many computer vision tasks on images, and their learned representations are often said to mimic the early layers of the visual cortex.
-But can CNNs understand graphical perception for visualization?
-We investigate this question by reproducing Cleveland and McGill's seminal 1984 experiments, which measured human perception efficiency of different visual encodings and defined elementary perceptual tasks for visualization.
-We measure the graphical perceptual capabilities of four classifiers on a) elementary perceptual tasks with increasing parametric complexity, b) the position-angle experiment that compares pie charts to bar charts, c) the position-length experiment that compares grouped and divided bar charts, and d) the bars and framed rectangles experiment where visual cues aid perception.
-We also study how feed-forward neural networks obey Weber's law, which defines the proportional relation between perceivable information and distribution density.
-We present the results of these experiments to foster the understanding of how CNN classifiers succeed and fail when applied to data visualizations.
-} % end of abstract
-
-%% Keywords that describe your work. Will show as 'Index Terms' in journal
-%% please capitalize first letter and insert punctuation after last keyword
-\keywords{Machine Perception, Deep Learning}
 
@@
 %% Uncomment below to include a teaser figure.
-\teaser{
-  \centering
-  \includegraphics[width=\linewidth]{teaser.pdf}
-  \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking visual cortex inspired machine learning classifiers to assess the relationship between values encoded in pie charts and bar charts.
+%\teaser{
+%  \centering
+%  \includegraphics[width=\linewidth]{teaser.pdf}
+%  \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking visual cortex inspired machine learning classifiers to assess the relationship between values encoded in pie charts and bar charts.
 Similar to the findings of Clevenland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs read quantities more accurately from bar charts (mean squared error, MSE in green).}
-	\label{fig:teaser}
-}
+%	\label{fig:teaser}
+%}
 
@@
 \begin{document}
-
+\maketitle
 %% The ``\maketitle'' command must be the first command after the
@@
 %% the only exception to this rule is the \firstsection command
-
-\input{00_introduction.tex}
-
-\input{01_related_work.tex}
-
-\input{02_experimental_setup.tex}
-
-\input{03_elementary_perceptual_tasks.tex}
-
-\input{04_position_angle_experiment.tex}
-
-\input{05_position_length_experiment.tex}
-
-\input{06_bars_and_framed_rectangles_experiment.tex}
-
-\input{07_results.tex}
-
-\input{08_conclusions.tex}
 
@@
 
-\bibliography{paper.bib}
+\bibliography{../paper.bib}
 \end{document}

### 2018-03-30 PAPER/00_introduction.tex
Changes to Section 2

@@
 To perform this evaluation, we parametrize the elementary perceptual tasks and experiments suggested by Cleveland and McGill~\cite{cleveland_mcgill}, and define a set of regression tasks to estimate continuous variables.
-Against human perception, we pit four neural networks: a three-layer multilayer perceptron (MLP), the LeNet 5-layer CNN~\cite{lenet}, the VGG 19-layer CNN~\cite{simonyan_very_deep2014}, and the Xception 36-layer CNN~\cite{xception}.
+Against human perception, we pit four neural networks: a three-layer multilayer perceptron (MLP), the LeNet 2-layer CNN~\cite{lenet}, the VGG 19-layer CNN~\cite{simonyan_very_deep2014}, and the Xception 36-layer CNN~\cite{xception}.
 As CNNs trained on natural images are said to mimic layers of the human visual cortex, we investigate whether using weights trained on natural images (via ImageNet~\cite{imagenet}) or weights trained from scratch on elementary graphical perception tasks produces more accurate measurements and greater generalization.

### 2018-03-30 PAPER/01_related_work.tex
Changes to Section 2

--- 2018-03-29 PAPER/01_related_work.tex
@@
 
-\paragraph{Graphical Perception} Cleveland and McGill~\cite{cleveland_mcgill,cleveland1985graphical} coin the phrase \emph{graphical perception} to describe how different visual attributes and encodings are perceived by humans.
+\noindent{\textbf{Graphical Perception.}} Cleveland and McGill~\cite{cleveland_mcgill,cleveland1985graphical} coin the phrase \emph{graphical perception} to describe how different visual attributes and encodings are perceived by humans.
 They define \emph{elementary perceptual tasks} as mental-visual stimuli to understand encodings in visualizations, and declare a ranking based on their perceptual difficulty.
@@
 
-\paragraph{Computational Visualization Understanding} Pineo \textit{et al.}~\cite{Pineo2012_computational_perception} create a computational model of human vision based on neural networks.
+\noindent{\textbf{Computational Visualization Understanding.}} Pineo \textit{et al.}~\cite{Pineo2012_computational_perception} create a computational model of human vision based on neural networks.
 Their simulations show that understanding visualization triggers neural activity in high-level areas of cognition, with the authors suspecting that this activity is supported by low-level neurons performing elementary perceptional tasks.
@@
 
-\paragraph{Visual-cortex-inspired Machine Learning} The human visual cortex allows us to recognize objects in the world seemingly without effort (though few remember their infancy).
+\noindent{\textbf{Visual-cortex-inspired Machine Learning.}} The human visual cortex allows us to recognize objects in the world seemingly without effort (though few remember their infancy).
 This visual system is organized in layers, which inspired computational classifiers based on multilayer neural networks.
@@
 Such networks exist with many architectures.
-For this paper, we select the traditional LeNet-5 which was designed to recognize hand-written digits with 5 layers~\cite{lenet}, the VGG19 network with 16 convolutional layers~\cite{simonyan_very_deep2014}, and the Xception network with 36 convolutional layers ~\cite{xception}.
-This allows us to compare archtitectures with different depths.
+For this paper, we compare a set of networks with different architectures and depths, plus networks with weights trained on natural images and on the elementary perceptual reasoning tasks themselves.
 %Object detection is extremely difficult and therefore is especially impressive as light intensities can change by levels of magnitude and contrast between foreground and background is so often low. 

### 2018-03-30 PAPER/02_experimental_setup.tex
Changes to Section 2

@@
 We conduct quantitative experiments to measure how different convolutional neural networks perceive low-level visual encodings, such as positions, angles, curvatures, and lengths.
-We formulate measurement tasks as logistic regression problems, where the networks must estimate single quantities or ratio relations between multiple quantities.
-The output between $0$ and $1$ indicates the percent of the value to be measured in the individual experiment.
+We formulate these measurement tasks as logistic regression problems: given a stimuli image of an elementary visualization, the networks must estimate the single quantity present or the ratio between multiple quantities present.
 
-% The output between $0$ and $1$ indicates the percent of the value to be measured in the individual experiment.
+For each experiment, we use a single factor between-subject design, with the factor being the network used.
+This lets us evaluate whether different network designs are competitive against existing human perception results.
+We train each network in a supervised fashion with a mean-squared error (MSE) loss between the ground-truth labels and the network's estimate of the measurement from observing the generated stimuli images.
+Then, we test each network's ability to generalize to new examples with a separate data, created using the same stimuli generator function but with unseen ground-truth measurements (Section~\ref{sec:data}).
 
-As baseline, we use a standard multilayer perceptron (MLP) without the generation of additional feature maps through convolutions (Fig.~\ref{fig:classifiers}).
-For our experiments, we use a series of single factor between-subject designs with the factor \emph{network} (MLP, LeNet, VGG19, Xception, VGG19+ImageNet, Xception+ImageNet).
-We train each network in a supervised fashion, with each experiment including a training and a testing phase on data.
-without any overlap (Section~\ref{sec:data}).
-All parameters such as network hyperparameters, optimization methods, and stopping conditions are fixed (Section~\ref{sec:classifiers}).
-However, since the networks are of different complexity, the number of trainable parameters changes (Table~\ref{tab:parameters}).
-We define a series of hypotheses prior to each experiment.
+% We define a series of hypotheses prior to each experiment.
 
-\subsection{Measures and Analysis}
+\subsection{Networks}
+\label{sec:networks}
 
-\textbf{Accuracy.} In their 1984 paper, Cleveland and McGill use the midmean logistic absolute error metric (\emph{MLAE}) to measure perception accuracy.
-\begin{equation}
-	\textnormal{MLAE} = log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
-\end{equation}
-In addition to this metric, we also calculate standard error metrics such as the mean squared error (\emph{MSE}) and the mean absolute error (\emph{MAE}).
-This allow a more direct comparison of percent errors.
+\noindent{\textbf{Multilayer Perceptron.}} As baseline, we use a multilayer perceptron (MLP), but without the prior convolutional layers as is typical in network designs for solving visual tasks (Fig.~\ref{fig:classifiers}).
+Our MLP contains a layer of $256$ perceptrons, which are activated as rectified linear units (ReLU)~(Fig.~\ref{fig:classifiers}).
+We train this layer with dropout (probability $= 0.5$) to prevent overfitting, and then combine these ReLU units to regress our output measurement.
 \\~\\
-\noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
-For regression tasks the loss is defined as MSE.% and for classification tasks the loss is categorical cross-entropy.
+\noindent{\textbf{Convolutional Neural Networks.}} We compare different convolutional neural networks (CNNs) with both `trained from scratch' weights and pre-trained weights on a database of natural images (1000-class ImageNet \cite{imagenet}).
+These networks are the traditional LeNet-5 with 2 layers, which was designed to recognize hand-written digits~\cite{lenet}; the VGG19 network with 19 layers, which was designed to solve the ImageNet object recognition challenge~\cite{simonyan_very_deep2014}; and the Xception network with 36 layers~\cite{xception}, which was also designed to solve the ImageNet object recognition challenge plus the 15,000-class JFT object recognition challenge~\cite{Hinton2015}.
+Each of these networks has as its last layers an MLP architecture equivalent to our baseline, and so they act as earlier image and feature processors for this final regressor.
+Since the networks are of different architectures, the number of trainable parameters changes, with some networks having more capacity than others (Table~\ref{tab:parameters}).
+
+For \emph{VGG19} and \emph{Xception}, we have two variants: the network trained from scratch on elementary perceptual tasks, plus the network using weights that were previously trained on the ImageNet object recognition challenge \emph{except} for the MLP layer.
+This is intended to produce early-layer features which mimic human vision, and then to see whether they are more or less useful than networks trained from scratch.
 \\~\\
-\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the $95\%$ confidence intervals but rather than performing bootstrapping, we approximate the value of the $97.5$ percentile point of the normal distribution for simplicity.
-\\~\\
-\noindent{\textbf{Confirmatory Data Analysis.}} To accept or reject our hypotheses, we calculate these measures and treat them as continuous variables.
-We analyze these dependent variables using analysis of variance (ANOVA) followed by parametric tests.
-
-\subsection{Classifiers}
-\label{sec:classifiers}
-Our classifiers are built upon a multilayer perceptron (MLP) which is a feedforward artificial neural network.
-We combine this MLP with different convolutional neural networks (CNNs) for preprocessing and feature generation.
-These include the traditional LeNet-5 network as well as the VGG19 and Xception networks.
-\\~\\
-\noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units~(Fig.~\ref{fig:classifiers}).
-We then add a dropout layer to prevent overfitting and compute linear regression with MSE loss.
-\\~\\
-\noindent{\textbf{Convolutional Neural Networks.}} We use CNNs to generate additional features as input to the MLP.
-We train the \emph{LeNet} classifier, the \emph{VGG19} network, and the \emph{Xception} model specifically towards each visualization.
-For \emph{VGG19} and \emph{Xception}, we additionally generate features using previously trained weights on ImageNet to naively model human vision.
-\\~\\
-\noindent{\textbf{Optimization.}} All networks are optimized using stochastic gradient descent with Nesterov momentum using fixed parameters (Table \ref{tab:parameters}).
-We train for $1000$ epochs but stop early if the loss does not decrease for ten epochs.
+\noindent{\textbf{Optimization.}} All network hyperparameters, optimization methods, and stopping conditions are fixed across networks (Table \ref{tab:parameters}).
+We train for $1000$ epochs using stochastic gradient descent with Nesterov momentum, but stop early if the loss does not decrease for ten epochs.
 \\~\\
 \noindent{\textbf{Environment.}} We run all experiments on Tesla X and Tesla V100 graphical processing units.
-We use the KERAS framework with tensorflow and leverage the scikit-image and scikit-learn libraries.
+We use the KERAS framework with a TensorFlow backend to train the networks, and use the scikit-image library to generate the stimuli.
+% and scikit-learn libraries - JT: What do we use this for?
 
@@
 \centering
-\caption{\textbf{Classifier Training.} We use different feature generators as input to a multilayer perceptron which performs linear regression.
+\caption{\textbf{Network Training.} We use different feature generators as input to a multilayer perceptron which performs linear regression.
 This results in different sets of trainable parameters.
-We also train the MLP directly on the visualizations without any additional feature generation as baseline.}
+As a baseline, we also train the MLP directly on the visualization images without any additional feature generation.}
 \resizebox{\linewidth}{!}{
@@
 	\toprule
-	Classifier & \makecell{Trainable\\Parameters} & Optimization \\
+	Network & \makecell{Trainable\\Parameters} & Optimization \\
 	\midrule
@@
 
-  \caption{\textbf{Classifier Architecture.} The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
+  \caption{\textbf{Network Architecture.} The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
 We use Dropout regularization to prevent overfitting.
 As output, we perform linear regression for continuous variables.
-The MLP can learn the visualizations directly but we also learn features generated by LeNet (2 conv.
+The MLP can learn to represent the visualizations directly, but we also learn features generated by LeNet (2 conv.
 layers, filter size $5\times5$), VGG19 (16 conv.
@@
 \label{sec:data}
-We create all visualizations as parametrized rasterized images without interpolation (except for the anti-alias experiment).
-The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}.
+We create all visualizations as parametrized rasterized images without interpolation.
+The number of possible parameter values differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and Section~\ref{sec:parametrizations}.
 \\~\\
-\noindent\textbf{Noise.} We add subtle random noise ($0.05$) to each pixel to introduce additional variation. This prevents the classifiers of learning the actual pixels.
+\noindent\textbf{Noise.} We add subtle random noise ($0.05$) to each pixel to introduce additional variation. This prevents the networks of learning the actual pixels.
 \\~\\
 \noindent\textbf{Training/Validation/Test Splits.} We specify the size of each split set as follows: 60,000 training images, 20,000 validation images, and 20,000 test images.
-We then randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables.
+Then, we randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables.
 This eliminates leakage during training and evaluation.
 We also scale each set independently: images to the range of $-0.5$ to $0.5$ and labels to the range of $0.0$ to $1.0$.
+This represents the percent of the value to be measured in the individual experiment.
+% JT: Moved; but what does this last sentence mean?
 \\~\\
@@
 These scenarios afflict the optimization process to find global minima and result in networks with more flexible knowledge with the caveat of longer training times.
+
+\subsection{Measures and Analysis}
+
+\noindent{\textbf{Accuracy.}} In their 1984 paper, Cleveland and McGill use the midmean logistic absolute error metric (\emph{MLAE}) to measure perception accuracy.
+\begin{equation}
+	\textnormal{MLAE} = log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
+\end{equation}
+In addition to this metric, we also calculate standard error metrics such as the mean squared error (\emph{MSE}) and the mean absolute error (\emph{MAE}).
+This allow a more direct comparison of percent errors.
+\\~\\
+\noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
+For regression tasks the loss is defined as MSE.% and for classification tasks the loss is categorical cross-entropy.
+\\~\\
+\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the $95\%$ confidence intervals but rather than performing bootstrapping, we approximate the value of the $97.5$ percentile point of the normal distribution for simplicity.
+\\~\\
+\noindent{\textbf{Confirmatory Data Analysis.}} To accept or reject our hypotheses, we calculate these measures and treat them as continuous variables.
+We analyze these dependent variables using analysis of variance (ANOVA) followed by parametric tests.

### 2018-03-30 PAPER/06_bars_and_framed_rectangles_experiment.tex

@@
 
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_overview}
-  \caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales.
+%\begin{figure}[t]
+%	  \includegraphics[width=\linewidth]{figure12_overview}
+%  \caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales.
 For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles (right).
@@
 We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
-	\label{fig:bars_and_framed_rectangles_experiment}
-\end{figure}
+%	\label{fig:bars_and_framed_rectangles_experiment}
+%\end{figure}
+\begin{table}[h]
+\centering
+\caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales.
+For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles.
+In this case, the right bar is heigher as visible with less free space when adding the frame.
+We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lllr}
+	\toprule
+	\multicolumn{2}{l}{~} & ~ & Permutations\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure12_0.pdf}} & \makecell[tl]{\emph{Bars}\\~~~Perceptual Task: \emph{Length}\\~~~JND: $X\%$\\~ \\~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $57,600$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure12_1.pdf}} & \makecell[tl]{\emph{Framed Rectangles}\\~~~Perceptual Task: \emph{Position}\\~~~JND: $X\%$\\~ \\~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $57,600$}\\
+
+	\bottomrule
+\end{tabular}
+}
+\label{tab:bars_and_framed_rectangles_parameters}
+\end{table}
 
@@
 
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{weber_overview}
-  \caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution.
+%\begin{figure}[t]
+%	  \includegraphics[width=\linewidth]{weber_overview}
+%  \caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution.
 The lower square contains 10 more dots than the upper one on both sides.
@@
 We generate rasterized visualizations similar to this setup and evaluate our classifiers.}
-	\label{fig:webers_law}
-\end{figure}
+%	\label{fig:webers_law}
+%\end{figure}
+\begin{table}[h]
+\centering
+\caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution.
+We create three different types of images initialized with 10, 100, and 1000 dots.
+We then mark randomly up to 10  dots in previously free pixels (here visualized in red).
+For humans, the difference is easily perceivable when the initial dot count is 10 but is hard when it is higher.
+We evaluate our classifiers on these rasterized visualizations.}
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lllr}
+	\toprule
+	\multicolumn{2}{l}{~} & ~ & Permutations\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_0.pdf}} & \makecell[tl]{\emph{Base 10}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_1.pdf}} & \makecell[tl]{\emph{Base 100}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_2.pdf}} & \makecell[tl]{\emph{Base 1000}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\
+
+	\bottomrule
+\end{tabular}
+}
+\label{tab:bars_and_framed_rectangles_parameters}
+\end{table}
 

### 2018-03-30 PAPER/paper.tex
Switch paper title and start new experiments.
Update paper and figures.
bib ref change; squash

@@
 % - 
-\title{Reproducing 'Graphical Perception' with CNNs}
+\title{Evaluating `Graphical Perception' with CNNs}
 
@@
 %
-\item James Tompkin is with the Thomas J. Watson Sr. Center for Information Technology at Brown University. \\E-mail: james\_tompkin@brown.edu.
+\item James Tompkin is with Brown University. \\E-mail: james\_tompkin@brown.edu.
 }
@@
 %other entries to be set up for journal
-\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Computing Graphical Perception with Convolutional Neural Networks}
+\shortauthortitle{Haehn \MakeLowercase{\textit{et al.}}: Evaluating `Graphical Perception' with CNNs}
 %\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
@@
 
-\bibliography{paper.bib}
+\bibliography{paper}
 \end{document}

### 2018-03-30 PAPER/00_introduction.tex
Squash, conform.

@@
 
-We test these four networks across four scenarios presented by Cleveland and McGill~\cite{cleveland_mcgill}: 1) Nine elementary perceptual tasks with increasing parameteric complexity, e.g., estimating the length of a bar with fixed x, then with varying x, then with varying width, including cross-network evaluations testing the generalizability of networks to unseen parameters; 2) The position-angle experiment, which compares judgements of bar charts to pie charts, 3) The position-length experiment, which compares grouped and divided bar charts, and 4) The bars and framed rectangles experiments, where visual cues aid ratio perception.
-We also investigate whether CNNs obey Weber's law, which defines a proportional dependency between an initial distribution and perceivable change.
+We test these four networks across four scenarios presented by Cleveland and McGill~\cite{cleveland_mcgill}: 1) Nine elementary perceptual tasks with increasing parameteric complexity, e.g., length estimation with fixed x, then with varying x, then with varying width, including cross-network evaluations testing the generalizability of networks to unseen parameters; 2) The position-angle experiment, which compares judgements of bar charts to pie charts, 3) The position-length experiment, which compares grouped and divided bar charts, and 4) The bars and framed rectangles experiments, where visual cues aid ratio perception.
+We also investigate whether our CNNs can detect a proportional change in a measurement across scales, in relation to Weber's law.
 
@@
 Further, we discuss the implications of our results and derive recommendations for the use of CNNs in perceiving visualizations.
-We accompany this paper with open source code and our input and results data~\footnote{Code and data are available at: \url{http://rhoana.org/perception}}, both to enable reproduction studies and to spur new machine perception systems more adept at graphical perception.
+We accompany this paper with open source code and our input and results data, both to enable reproduction studies and to spur new machine perception systems more adept at graphical perception: \url{http://rhoana.org/perception}
 

### 2018-03-30 PAPER/02_experimental_setup.tex
Section 2

@@
 \label{sec:data}
-We create all visualizations as parametrized rasterized images without interpolation.
-The number of possible parameter values differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and Section~\ref{sec:parametrizations}.
-\\~\\
-\noindent\textbf{Noise.} We add subtle random noise ($0.05$) to each pixel to introduce additional variation. This prevents the networks of learning the actual pixels.
-\\~\\
-\noindent\textbf{Training/Validation/Test Splits.} We specify the size of each split set as follows: 60,000 training images, 20,000 validation images, and 20,000 test images.
-Then, we randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables.
-This eliminates leakage during training and evaluation.
-We also scale each set independently: images to the range of $-0.5$ to $0.5$ and labels to the range of $0.0$ to $1.0$.
-This represents the percent of the value to be measured in the individual experiment.
-% JT: Moved; but what does this last sentence mean?
-\\~\\
-\noindent\textbf{Cross Validation.} For reproducibility, we perform repeated random sub-sampling validation, also known as Monte Carlo cross-validation, during our experiments.
-We run every experiment seperately multiple times (at least 4, up to 12) and randomly select (without replacement) the $60\%$ of our data as training data, $20\%$ as validation, and $20\%$ as test.
-Our large data sample size of 100,000 guarantees that every single observation of our parameterizations will be selected at least once (excluding noise patterns).
-Finally, we average the results over the runs.
-\\~\\
-\noindent\textbf{Intra-classifier Variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.
-\\~\\
-\noindent\textbf{Multi Classifiers.} Some experiments compare different types of visual encodings.
-We first train and evaluate each classifier on one type of encoding but we also train the classifiers on a random selection of multiple different types.
-These scenarios afflict the optimization process to find global minima and result in networks with more flexible knowledge with the caveat of longer training times.
+
+\noindent\textbf{Image Stimuli and Labels} 
+We create our stimuli visualizations as 100$\times$100 binary images, rasterized without interpolation.
+We write a parameterized stimuli generator for each elementary task.
+The number of possible parameter values differs per experiment, and we summarize these in Table~\ref{tab:encoding_parameters} and Section~\ref{sec:parametrizations}.
+Before use, we scale the generated images into an unbiased range: images to the range of $-0.5$ to $0.5$.
+Then, we add subtle random noise (between $0--0.05$) to each pixel to introduce variation which prevents the networks from simply `remembering' each different image.
+
+Each stimuli image also has an associated ground truth label representing the parameter set which generated the image, e.g., the length in pixels of a bar.
+As before, we scale these labels to the range of $0.0$ to $1.0$, which represent the maximum and minimum values that this parameter can take.
+\noindent\textbf{Training/Validation/Test Splits.} For each task, we use 60,000 training images, 20,000 validation images, and 20,000 test images.
+To create these datasets, we generate stimuli from random parameters and add them to the sets until the target number is reached, while maintaining distinct (random) parameter spaces for each set to ensure that there is no leakage between training and validation/testing.
 
@@
 
-\noindent{\textbf{Accuracy.}} In their 1984 paper, Cleveland and McGill use the midmean logistic absolute error metric (\emph{MLAE}) to measure perception accuracy.
+\noindent\textbf{Cross Validation.} For reproducibility, we perform repeated random sub-sampling validation, also known as Monte Carlo cross-validation, during our experiments.
+We run every experiment seperately twelve times, and randomly select (without replacement) the $60\%$ of our data as training data, $20\%$ as validation, and $20\%$ as test.
+%Our large data sample size of 100,000 guarantees that every single observation of our parameterizations will be selected at least once (excluding noise patterns). -> JT: No it doesn't.
+% Finally, we average the results over the runs. -> JT: Fine, but we'll do distributions anyway, so not worth saying.
+\\~\\
+\noindent{\textbf{Task Accuracy.}} In their 1984 paper, Cleveland and McGill use the midmean logistic absolute error metric (\emph{MLAE}) to measure perception accuracy.
+To allow comparison between their human results and our machine results, we also use MLAE as a presentation metric:
 \begin{equation}
@@
 This allow a more direct comparison of percent errors.
+However, please note that our networks were trained using MSE loss and not directly with MLAE.
 \\~\\
-\noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
-For regression tasks the loss is defined as MSE.% and for classification tasks the loss is categorical cross-entropy.
+\noindent{\textbf{Task Confidence Intervals.}} We follow Cleveland and McGill and present $95\%$ confidence intervals.
+However, as our networks but rather than performing bootstrapping, we approximate the value of the $97.5$ percentile point of the normal distribution for simplicity.
+JT: What?
+Why 97.5\% ?
+Why not 95\% ?
 \\~\\
-\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the $95\%$ confidence intervals but rather than performing bootstrapping, we approximate the value of the $97.5$ percentile point of the normal distribution for simplicity.
+\noindent{\textbf{Confirmatory Data Analysis.}} To accept or reject our hypotheses, we analyze dependent variables using analysis of variance (ANOVA) followed by parametric tests. JT: Which tests?
 \\~\\
-\noindent{\textbf{Confirmatory Data Analysis.}} To accept or reject our hypotheses, we calculate these measures and treat them as continuous variables.
-We analyze these dependent variables using analysis of variance (ANOVA) followed by parametric tests.
+\noindent{\textbf{Training Efficiency.}} We use the training convergence rate as a measure of how easy or hard a particular task is for the network to learn to solve.
+This is defined as the MSE loss decrease per training epoch, which is an indicator of the training efficiency of the network with respect to the visual encoding.
+\\~\\
+\noindent\textbf{Network Generalizability.} We evaluate generalizability by asking a network previously trained upon one task parameterization to answer quetsions about the same type of task stimuli but with more complex parameterization, e.g., estimating bar length without and with changes in stroke width.
+
+Further, some experiments compare different visual encoding types, e.g., bar plot vs.~stacked bar plot.
+We train and evaluate individual networks for each parameterization, plus we also train and evaluate a networks on stimuli across the different types.
+This single decision-making software better mimics the judgements that a human would be able to make.
+%This scenario affects the optimization process and result in networks with more flexible knowledge with the caveat of longer training times. -> JT: Save it for the results discussion.

### 2018-03-30 PAPER/04_position_angle_experiment.tex
Update 4.

@@
 	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure3_Bar_Chart.pdf}} & \makecell[tl]{Type 1: \emph{Bar Chart}\\~~~Perceptual Task: \emph{Position}\\~ \\~ \\} &~& \makecell[tr]{~\\ $244,827$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure3_Bar_Chart.pdf}} & \makecell[tl]{Type 1: \emph{Bar Chart}\\~~~Perceptual Task: \emph{Position}\\~ \\~ \\} &~& \makecell[tr]{~\\ $878,520$}\\
 
 	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure3_Pie_Chart.pdf}} & \makecell[tl]{Type 2: \emph{Pie Chart}\\~~~Perceptual Task: \emph{Angle}\\~ \\~ \\} &~& \makecell[tr]{~\\ $244,827$}\\
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure3_Pie_Chart.pdf}} & \makecell[tl]{Type 2: \emph{Pie Chart}\\~~~Perceptual Task: \emph{Angle}\\~ \\~ \\} &~& \makecell[tr]{~\\ $878,520$}\\
 

### 2018-03-30 PAPER/00_introduction.tex
Some typos and stuff.

@@
 One fundamental application of human vision is to understand data visualizations.
-This is a task unlike processing natural images, where real-world objects and their effects are abstracted into data and represented with visual marks.
+This is a task unlike natural image processing but includes the abstraction of real-world objects and their effects into data, represented with visual marks.
 As a field, visualization catalogues and evaluates human perception of these marks, such as in the seminal \emph{graphical perception} experiments of Cleveland and McGill~\cite{cleveland_mcgill}.
@@
 To perform this evaluation, we parametrize the elementary perceptual tasks and experiments suggested by Cleveland and McGill~\cite{cleveland_mcgill}, and define a set of regression tasks to estimate continuous variables.
-Against human perception, we pit four neural networks: a three-layer multilayer perceptron (MLP), the LeNet 2-layer CNN~\cite{lenet}, the VGG 19-layer CNN~\cite{simonyan_very_deep2014}, and the Xception 36-layer CNN~\cite{xception}.
+Against human perception, we pit four neural networks: a three-layer multilayer perceptron (MLP), the LeNet 2-layer CNN~\cite{lenet}, the VGG 16-layer CNN~\cite{simonyan_very_deep2014}, and the Xception 36-layer CNN~\cite{xception}.
 As CNNs trained on natural images are said to mimic layers of the human visual cortex, we investigate whether using weights trained on natural images (via ImageNet~\cite{imagenet}) or weights trained from scratch on elementary graphical perception tasks produces more accurate measurements and greater generalization.

### 2018-03-30 PAPER/02_experimental_setup.tex
Forgot ne thing
Some typos and stuff.

@@
 
-\noindent{\textbf{Multilayer Perceptron.}} As baseline, we use a multilayer perceptron (MLP), but without the prior convolutional layers as is typical in network designs for solving visual tasks (Fig.~\ref{fig:classifiers}).
+\noindent{\textbf{Multilayer Perceptron.}} As a baseline, we use a multilayer perceptron (MLP), but without the prior convolutional layers as is typical in network designs for solving visual tasks (Fig.~\ref{fig:classifiers}).
 Our MLP contains a layer of $256$ perceptrons, which are activated as rectified linear units (ReLU)~(Fig.~\ref{fig:classifiers}).
@@
 
-\noindent\textbf{Image Stimuli and Labels} 
+\noindent\textbf{Image Stimuli and Labels.} 
 We create our stimuli visualizations as 100$\times$100 binary images, rasterized without interpolation.
@@
 Before use, we scale the generated images into an unbiased range: images to the range of $-0.5$ to $0.5$.
-Then, we add subtle random noise (between $0--0.05$) to each pixel to introduce variation which prevents the networks from simply `remembering' each different image.
+Then, we add subtle random noise (uniformly distributed between $0-0.05$) to each pixel to introduce variation which prevents the networks from simply `remembering' each different image.
 
@@
 As before, we scale these labels to the range of $0.0$ to $1.0$, which represent the maximum and minimum values that this parameter can take.
+\\~\\
 \noindent\textbf{Training/Validation/Test Splits.} For each task, we use 60,000 training images, 20,000 validation images, and 20,000 test images.
@@
 In addition to this metric, we also calculate standard error metrics such as the mean squared error (\emph{MSE}) and the mean absolute error (\emph{MAE}).
-This allow a more direct comparison of percent errors.
-However, please note that our networks were trained using MSE loss and not directly with MLAE.
+This allows a more direct comparison of percent errors.
+Please note that our networks were trained using MSE loss and not directly with MLAE.
 \\~\\
 \noindent{\textbf{Task Confidence Intervals.}} We follow Cleveland and McGill and present $95\%$ confidence intervals.
-However, as our networks but rather than performing bootstrapping, we approximate the value of the $97.5$ percentile point of the normal distribution for simplicity.
-JT: What?
-Why 97.5\% ?
-Why not 95\% ?
+We approximate the value of the $97.5$ percentile point of the normal distribution for simplicity with $1.96$ as suggested by the central limit theorem~\cite{central_limit}.
 \\~\\
@@
 \\~\\
-\noindent\textbf{Network Generalizability.} We evaluate generalizability by asking a network previously trained upon one task parameterization to answer quetsions about the same type of task stimuli but with more complex parameterization, e.g., estimating bar length without and with changes in stroke width.
+\noindent\textbf{Network Generalizability.} We evaluate generalizability by asking a network previously trained upon one task parameterization to answer questions about the same type of task stimuli but with more complex parameterization, e.g., estimating bar length without and with changes in stroke width.
 

### 2018-03-30 PAPER/paper.tex
Some typos and stuff.

@@
   \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking visual cortex inspired machine learning classifiers to assess the relationship between values encoded in pie charts and bar charts.
-Similar to the findings of Clevenland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs read quantities more accurately from bar charts (mean squared error, MSE in green).}
+Similar to the findings of Cleveland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs read quantities more accurately from bar charts (mean squared error, MSE in green).}
 	\label{fig:teaser}

### 2018-03-30 PAPER/06_bars_and_framed_rectangles_experiment.tex
More updates.

@@
 Weber's law goes hand-in-hand with Fechner's law.
-We conduct an additional experiment based on the original illustrations of the Weber-Fechner law to investigate whether this law can be applied to computational perception of our classifiers (Fig.~\ref{fig:webers_law}).
+We conduct an additional experiment based on the original illustrations of the Weber-Fechner law to investigate whether this law can be applied to computational perception of our classifiers (Fig.~\ref{tab:weber_law}).
 
@@
 }
-\label{tab:bars_and_framed_rectangles_parameters}
+\label{tab:weber_law}
 \end{table}

### 2018-03-30 PAPER/02_experimental_setup.tex
Minor fixes

@@
 
-We conduct quantitative experiments to measure how different convolutional neural networks perceive low-level visual encodings, such as positions, angles, curvatures, and lengths.
+First, we describe the commonalities across all of our experiments.
+We measure how different convolutional neural networks perceive low-level visual encodings, such as positions, angles, curvatures, and lengths.
 We formulate these measurement tasks as logistic regression problems: given a stimuli image of an elementary visualization, the networks must estimate the single quantity present or the ratio between multiple quantities present.

### 2018-03-30 PAPER/03_elementary_perceptual_tasks.tex
Some typos and stuff.
Minor.
Minor fixes

@@
 % !TeX root = paper.tex
-\section{Elementary Perceptual Tasks}
+\section{Experiment: Elementary Perceptual Tasks}
 
-Cleveland and McGill describe the mapping of graphical elements to quantitative variables as \emph{elementary perceptual tasks} and introduce a list of ten different encodings in their paper~\cite{cleveland_mcgill}.
+Cleveland and McGill describe the mapping of graphical elements to quantitative variables as \emph{elementary perceptual tasks} and introduce a list of ten different encodings in their paper~\cite{cleveland_mcgill,cleveland1985graphical}.
 These tasks are the low-level building blocks for information visualizations and encode quantities.
@@
 %\end{figure}
-\begin{table}[h]
+\begin{table}[t]
 \centering
@@
 	\item \textbf{H1.2} \textbf{Computed perceptual performance is dependent on classifier complexity.} We evaluate multiple classifiers with different numbers of trainable parameters.
-A more complex classifier (with higher number of parameters) will perform better on elementary perceptual tasks.
+A more complex classifier (with a higher number of parameters) will perform better on elementary perceptual tasks.
 	\item \textbf{H1.3} \textbf{Some visual encodings are better than others for computations.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
@@
 
-\begin{figure*}[h]
+\begin{figure*}[t]
 	\centering
-	  \includegraphics[width=.95\linewidth]{figure1.pdf}
+	\includegraphics[width=0.48\linewidth]{figure1_slim_left.pdf}
+	\includegraphics[width=0.48\linewidth]{figure1_slim_right.pdf}
 	\caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.

### 2018-03-30 PAPER/03_elementary_perceptual_tasks.tex

@@
 %\end{figure}
-\begin{table}[t]
+\begin{table}[!ht]
 \centering
@@
 
-some are good and some are bad.. why?
-
-\begin{figure*}[t]
-	\centering
-	\includegraphics[width=0.48\linewidth]{figure1_slim_left.pdf}
-	\includegraphics[width=0.48\linewidth]{figure1_slim_right.pdf}
-	\caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure1_results}
-\end{figure*}
+The results of the elementary perceptual tasks experiment demonstrates that our networks are able to `measure' visually encoded quantities using low-level building blocks (Fig.~\cite{})figure1_results}).
+While the performance varies for the different encodings, we observe for all networks performance mostly better or at least equal compared to the estimated human performance of $9\%$~\cite{cleveland1985graphical}.
+These results confirm our initial hypotheses.
+\\~\\
+\noindent\textbf{Accuracy.} The average error across all tasks and all networks is $$.
 
@@
 
-\begin{enumerate}
-	\item Position along a common scale e.g. scatter plot
-	\item Position on identical but nonaligned scales e.g. multiple scatter plots
-	\item Length e.g. bar chart
-	\item Angle \& Slope (tie) e.g. pie chart
-	\item Area e.g. bubbles
-	\item Volume, density, and color saturation (tie) e.g. heatmap
-	\item Color hue e.g. newsmap
-\end{enumerate}
+%\begin{enumerate}
+%	\item Position along a common scale e.g. scatter plot
+%	\item Position on identical but nonaligned scales e.g. multiple scatter plots
+%	\item Length e.g. bar chart
+%	\item Angle \& Slope (tie) e.g. pie chart
+%	\item Area e.g. bubbles
+%	\item Volume, density, and color saturation (tie) e.g. heatmap
+%	\item Color hue e.g. newsmap
+%\end{enumerate}
+
+\begin{table}[!ht]
+\centering
+\caption{\textbf{Ranking of Elementary Perceptual Tasks.} We create a ranking of the elementary perceptual tasks for all of our networks.
+We report midmean logistic absolute errors (MLAE) for each network averaged across multiple runs on the most complex parametrization of the individual stimuli.
+The VGG19 network performs best compared on all tasks.
+For human performance, we report the ranking of Cleveland and McGill~\cite{cleveland_mcgill}.
+The VGG19 * and Xception * networks use ImageNet weights and yield identical rankings.}
+\resizebox{\linewidth}{!}{
+\begin{tabular}{cllllll}
+\toprule
+Human & MLP & LeNet & VGG19 * & \textbf{VGG19} & Xception * & Xception \\
+\midrule
+\multicolumn{7}{l}{\emph{Position Common Scale}} \\
+1. & 7. (3.84) & 2. (1.36) & 5. (1.02) & \textbf{3.} (-0.04) & 5. (1.65) & 1. (1.04)\\
+\multicolumn{7}{l}{\emph{Position Non aligned Scale}} \\
+2. & 6. (3.61) & 1. (1.35) & 6. (1.09) & \textbf{5.} (0.26) & 6. (1.71) & 2. (1.06)\\
+\multicolumn{7}{l}{\emph{Length}} \\
+3. & 1. (1.99) & 8. (3.19) & 4. (0.87) & \textbf{2.} (-0.14) & 4. (1.59) & 3. (1.11) \\
+\multicolumn{7}{l}{\emph{Direction}} \\
+3. & 9. (4.65) & 7. (3.07) & 9. (2.84) & \textbf{8.} (0.92) & 9. (3.46) & 6. (1.57) \\
+\multicolumn{7}{l}{\emph{Angle}} \\
+3. & 8. (4.61) & 9. (3.33) & 8. (2.31) & \textbf{9.} (0.99) & 8. (2.60) & 7. (1.72) \\
+\multicolumn{7}{l}{\emph{Area}} \\
+4. & 2. (2.01) & 5. (2.21) & 1. (0.49) & \textbf{1.} (-0.17) & 1. (0.80) & 5. (1.38) \\
+\multicolumn{7}{l}{\emph{Volume}} \\
+5. & 4. (2.38) & 4. (1.91) & 7. (1.16) & \textbf{7.} (0.87) & 7. (2.03) & 9. (2.10) \\
+\multicolumn{7}{l}{\emph{Curvature}} \\
+5. & 3. (2.34) & 3. (1.81) & 2. (0.71) & \textbf{6.} (0.28) & 2. (1.17) & 4. (1.13) \\
+\multicolumn{7}{l}{\emph{Shading}} \\
+6. & 5. (3.04) & 6. (2.23) & 3. (0.73) & \textbf{4.} (0.14) & 3. (1.57) & 8. (1.82) \\
+
+% 
+%\makecell[tl]{\emph{Position}\\~~\emph{Non-aligned Scale}} & 1 & 1 & 2 & 3 & \textbf{4} & 5 & 6 \\
+%\makecell[tl]{\emph{Length}} & 1 & 1 & 2 & 3 & \textbf{4} & 5 & 6 \\
+%\makecell[tl]{\emph{Direction}} & 1 & 1 & 2 & 3 & \textbf{4} & 5 & 6 \\
+%\makecell[tl]{\emph{Angle}} & 1 & 1 & 2 & 3 & \textbf{4} & 5 & 6 \\
+%\makecell[tl]{\emph{Area}} & 1 & 1 & 2 & 3 & \textbf{4} & 5 & 6 \\
+%\makecell[tl]{\emph{Volume}} & 1 & 1 & 2 & 3 & \textbf{4} & 5 & 6 \\
+%\makecell[tl]{\emph{Curvature}} & 1 & 1 & 2 & 3 & \textbf{4} & 5 & 6 \\
+%\makecell[tl]{\emph{Shading}} & 1 & 1 & 2 & 3 & \textbf{4} & 5 & 6 \\
+%\begin{tabular}{ll}
+%	\toprule
+%	Task & \begin{tabular}{ccccccc}
+%			Human & MLP & LeNet & VGG19 * & VGG19 & Xception * & Xception
+%			\end{tabular}\\
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{position_common_scale.pdf}} & \makecell[tl]{\emph{Position Common Scale}\\ \begin{tabular}{ccccccc}
+%
+%\end{tabular}} \\
+%
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{position_non_aligned_scale.pdf}} & \makecell[tl]{\emph{Position Non-Aligned Scale}\\~~~Position Y\\~~~+ Position X \\~~~+ Spot Size \\} &~& \makecell[tr]{~\\ $600$ \\ $36,000$ \\ $216,000$}\\
+%
+%	\midrule
+%	\raisebox{-.95\height}{\includegraphics[width=.5in]{length.pdf}} & \makecell[tl]{\emph{Length}\\~~~Length\\~~~+ Position Y \\~~~+ Position X \\~~~+ Width} &~& \makecell[tr]{ ~\\$60$ \\ $2,400$ \\ $144,000$\\$864,000$}\\
+%
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{direction.pdf}} & \makecell[tl]{\emph{Direction}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$360$ \\ $21,600$ \\ $1,296,000$}\\
+%
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{angle.pdf}} & \makecell[tl]{\emph{Angle}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$90$ \\ $5,400$ \\ $324,000$}\\
+%
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{area.pdf}} & \makecell[tl]{\emph{Area}\\~~~Radius\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$40$ \\ $800$ \\ $16,000$}\\
+%
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{volume.pdf}} & \makecell[tl]{\emph{Volume}\\~~~Cube Sidelength\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$20$ \\ $400$ \\ $8,000$}\\
+%	
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{curvature.pdf}} & \makecell[tl]{\emph{Curvature}\\~~~Midpoint Curvature\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$80$ \\ $1,600$ \\ $64,000$}\\
+%
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{shading.pdf}} & \makecell[tl]{\emph{Shading}\\~~~Density\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$100$ \\ $2,000$ \\ $40,000$}\\
+
+	\bottomrule
+\end{tabular}
+}
+\label{tab:ranking}
+\end{table}
+
+\begin{figure}[!ht]
+	\centering
+	\includegraphics[width=.8\linewidth]{figure1_slim_only_last.pdf}
+%	\includegraphics[width=0.48\linewidth]{figure1_slim_right.pdf}
+	\caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
+We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks on the most complex parametrizations.
+The * indicates networks which use ImageNet weights.
+We also compare against an approximate human baseline of $9%$ accuracy reported by Cleveland and McGill.}
+	\label{fig:figure1_results}
+\end{figure}
 

### 2018-03-30 PAPER/03_elementary_perceptual_tasks.tex
Update section 4.

@@
 %\end{figure}
+
 \begin{table}[!ht]
@@
 
+%
 \subsection{Hypotheses}
@@
 \begin{itemize}
-	\item \textbf{H1.1} \textbf{Visual cortex inspired classifiers are able to connect graphical elements to their quantitative variables.} While much simpler models than their biological pendant, convolutional neural networks are heavily influenced by our biological knowledge of the visual system.
-Such classifiers therefor follow the same principles as human perception.
-	\item \textbf{H1.2} \textbf{Computed perceptual performance is dependent on classifier complexity.} We evaluate multiple classifiers with different numbers of trainable parameters.
-A more complex classifier (with a higher number of parameters) will perform better on elementary perceptual tasks.
+	\item \textbf{H1.1} \textbf{Our convolutional neural networks are able to regress quantitative variables from graphical elements.} We parametrize different visual encodings (Table~\ref{tab:encoding_parameters}) and test whether the CNNs can measure them with an accuracy similar to humans.
+	\item \textbf{H1.2} \textbf{Computed perceptual performance is dependent on network architexture.} We evaluate multiple regressors with different numbers of trainable parameters.
+A more complex network (with a higher number of parameters) might perform better on elementary perceptual tasks.
 	\item \textbf{H1.3} \textbf{Some visual encodings are better than others for computations.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
 We investigate whether this order is also relevant for computing graphical perception.
-	\item \textbf{H1.4} \textbf{Classifiers trained on perceptual tasks can generalize to more or less complex variations of the same task.} Recent research suggests that convolutional neural networks generalize extremely well.
+	\item \textbf{H1.4} \textbf{Networks trained on perceptual tasks can generalize to more or less complex variations of the same task.} Recent research suggests that convolutional neural networks are able to generalize by interpolating between different datapoints.
 While the underlying reasons are mainly yet unknown, this property allows them to perform on variations of a similar perceptual task.
+We create visual representations of the elementary perceptual tasks with different variability (Table~\ref{tab:encoding_parameters}).
+We suspect that such variations are not a problem for successfully regressing the quantitative variables and that the networks are able to generalize.
 \end{itemize}
+
+%	
+%	We suspect that CNNs are able to 'learn` absolute quantities encoded using low-level visual 
+%	
+%	While much simpler models than their biological pendant, convolutional neural networks are heavily influenced by our biological knowledge of the visual system.
+Such classifiers therefor follow the same principles as human perception.
 
@@
 
-The results of the elementary perceptual tasks experiment demonstrates that our networks are able to `measure' visually encoded quantities using low-level building blocks (Fig.~\cite{})figure1_results}).
-While the performance varies for the different encodings, we observe for all networks performance mostly better or at least equal compared to the estimated human performance of $9\%$~\cite{cleveland1985graphical}.
+The results of the elementary perceptual tasks experiment demonstrates that our networks are able to measure visually encoded quantities using low-level building blocks (Fig.~\ref{tab:figure1_results}).
+While the performance varies for the different encodings, we observe for all networks performance similar to the measured human performance by Cleveland and McGill~\cite{cleveland1985graphical}.
 These results confirm our initial hypotheses.
 \\~\\
-\noindent\textbf{Accuracy.} The average error across all tasks and all networks is $$.
-
-\textbf{Computational Perception Ranking.}
-
-Cleveland McGills Ranking - can we observe something similar?
-
-%\begin{enumerate}
-%	\item Position along a common scale e.g. scatter plot
-%	\item Position on identical but nonaligned scales e.g. multiple scatter plots
-%	\item Length e.g. bar chart
-%	\item Angle \& Slope (tie) e.g. pie chart
-%	\item Area e.g. bubbles
-%	\item Volume, density, and color saturation (tie) e.g. heatmap
-%	\item Color hue e.g. newsmap
-%\end{enumerate}
+\noindent{\textbf{Accuracy.}} We report the average error across all tasks for our networks as \textit{MLAE}: 
+$2.943$ ($SD=0.857$) for the multilayer perceptron, 
+for LeNet $2.125$ ($SD=0.38$), 
+VGG19 on ImageNet $0.979$ ($SD=0.581$), 
+for VGG19 trained from scratch $0.404$ ($SD=0.407$), 
+Xception on ImageNet $1.627$ ($SD=0.462$), 
+and for the re-trained Xception $1.504$ ($SD=0.493$).
+The average error for all classifiers is \textit{MLAE}$=1.597$ ($SD=0.394$) and \textit{MAE}=$2.89$ ($SD=0.848$).
+VGG19 trained from scratch reaches the best performance overall.
+These values are in the same ballpark as the range of $8-14\%$ error (approx.
+MLAE=$3$) for a similar experiment by Cleveland and McGill testing human estimation of relations between elementary perceptual tasks~\cite{cleveland1985graphical}.
+We therefor carefully confirm that our networks are able to regress the graphical elements to quantities within a reasonable range of error and \textbf{accept H1.1}.
+\\~\\
+\noindent{\textbf{Regressor Complexity.}} The performance varies not only across different visual encodings but also very much across the different network architectures.
+We compare the average regression performances for our networks and report the effect of the regressor as statistical significant ($F_{5,48}=20.392,p<0.01$).
+Post hoc comparisons show that the differences between LeNet and the VGG19 network, independent of the used weights, are significant ($t_48=4.674,p<0.01$).
+VGG19 from scratch and Xception (both versions) are significant, from scratch ($t_48=4.87,p<0.01$) and for imagenet weights($t_48=5.621,p<0.01$).
+However, differences between LeNet and both Xception networks are not significant.
+In any case, the variation of results also visually inspectable in Fig.~\ref{fig:figure1_results} shows the variation of performance of the different networks.
+We therefor \textbf{partially accept H1.2}.
+\\~\\
+\noindent{\textbf{Ranking of Visual Encodings.}} Cleveland and McGill provide an ordering of elementary visual encodings based on theoretical arguments and experimental results.
+We re-create their ranking with MLAE values of our networks in Table~\ref{tab:ranking}.
+The rankings differ between the different networks and Cleveland and McGill's version.
+However, the ranking between networks using imagenet weights is identical.
+Since such a ranking see,s to heavily depend on the used regressor, the weights, and the initialization, we do not think a generalized ranking can be created and \textbf{reject H1.3}.
 
@@
 The * indicates networks which use ImageNet weights.
-We also compare against an approximate human baseline of $9%$ accuracy reported by Cleveland and McGill.}
+We also compare against an approximate human baseline of $9\%$ error reported by Cleveland and McGill.}
 	\label{fig:figure1_results}
@@
 
-\noindent{\textbf{Cross-classifier variability.}}
+\noindent{\textbf{Cross-classifier Variability and Network Generalizability.}} We measure regression performance across classifiers trained on the different parameterizations of the elementary perceptual tasks.
+Looking at the results in Fig.~\ref{fig:cross_network}, we can observe that even slight variations of an encoding such as added translation in Y or X drastically increases the error.
+We think that the networks are not really learning concepts of perception but rather learn the variation of pixel values.
 
-Can a neural network generalize on simple perceptual tasks?
-
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{cross_network_small_VGG19.pdf}
-  \caption{\textbf{Cross-classifier variability for the perceptual task of measuring length.} We use predictions of LeNet classifiers trained on different parametrizations of the \emph{curvature} elementary perceptual task and measure the mean logistic absolute error (MLAE).
-The lower score, the better.
-Classifiers trained on curves with variable position can generalize even if the axis of translation varies.
-However, classifiers trained on fixed positions of curves are not able to measure translated curves.}
+\begin{figure}[!ht]
+	\centering
+	  \includegraphics[width=.75\linewidth]{cross_network_small_VGG19_area.pdf}
+	  \includegraphics[width=.75\linewidth]{cross_network_small_VGG19_length.pdf}
+	  \includegraphics[width=.75\linewidth]{cross_network_small_VGG19_position_common_scale.pdf}
+	  \includegraphics[width=.75\linewidth]{cross_network_small_VGG19_shading.pdf}
+  \caption{\textbf{Cross-classifier variability for perceptual tasks.} We use predictions of the VGG19 network trained on different parametrizations of \emph{area}, \emph{length}, \emph{position common scale}, and \emph{shading}.
+These are the top 4 encodings in the ranking for this network.
+We measure the mean logistic absolute error (MLAE) -- the lower score, the better.
+Regressors trained on stimuli with variable position can generalize even if the axis of translation varies.
+However, if they are trained on fixed positions of the stimuli they are not able to measure translated encodings.}
 	\label{fig:cross_network}

### 2018-03-30 EXP/run_cross_network_from_scratch.py
Ready for the new cross network plots.

@@ -1,0 +1,170 @@
+import matplotlib
+matplotlib.use('agg')
+from matplotlib.pyplot import imshow
+import matplotlib.pyplot as plt    
+
+import os, glob, sys
+import cPickle as pickle
+
+import numpy as np
+
+from keras.models import load_model
+from keras import backend as K
+import sklearn.metrics
+import keras.applications
+import itertools
+
+import ClevelandMcGill as C
+from util import Util
+
+
+EXPERIMENT = sys.argv[1]#'C.Figure1.curvature'
+CLASSIFIER = sys.argv[2]#'LeNet'
+NETWORK_INDEX = sys.argv[3]#4
+
+
+DATATYPE = eval(EXPERIMENT)
+
+
+RESULTS_DIR = '/n/regal/pfister_lab/PERCEPTION/RESULTS_FROM_SCRATCH/' + EXPERIMENT + '/'
+OUTPUT_DIR = '/n/regal/pfister_lab/PERCEPTION/CROSSNETWORK_FROM_SCRATCH/' + EXPERIMENT +'/'
+if not os.path.exists(OUTPUT_DIR):
+  # here can be a race condition
+  try:
+    os.makedirs(OUTPUT_DIR)
+  except:
+    print 'Race condition!', os.path.exists(OUTPUT_DIR)
+
+
+
+
+
+all_labels = {'C.Figure1.position_common_scale': ['Position Y', '+ Position X', '+ Spotsize'], \
+              'C.Figure1.position_non_aligned_scale': ['Scale', '+ Position Y', '+ Position X', '+ Spotsize'],\
+              'C.Figure1.length': ['Length', '+ Position Y', '+ Position X', '+ Width'], \
+              'C.Figure1.direction': ['Direction', '+ Position Y', '+ Position X'], \
+              'C.Figure1.angle': ['Angle', '+ Position Y', '+ Position X'], \
+              'C.Figure1.area': ['Area', '+ Position Y', '+ Position X'], \
+              'C.Figure1.volume': ['Volume', '+ Position Y', '+ Position X'], \
+              'C.Figure1.curvature': ['Curvature', '+ Position Y', '+ Position X', '+ Width'], \
+              'C.Figure1.shading': ['Shading', '+ Position Y', '+ Position X']
+              }
+DATASETS = len(all_labels[EXPERIMENT])
+
+
[…]

### 2018-03-31 PAPER/paper.tex
Going through Section 4.

@@
 %% Abstract section.
-\abstract{%Convolutional neural networks are being successfully used for image understanding and object recognition tasks while regularly outperforming humans.
-Despite their tremendous success, little is known perceptional capabilities.
-In there Graphical Perception paper from 1984, Cleveland and McGill define elementary perceptual tasks that let people extract quantitative information from visualizations and measure human perception of different visual encodings with in depth user studies.
-%We replicate the experimental setup of Cleveland and McGill and evaluate the perceptional capabilities of four modern classifiers based on neural networks.
-We systematically test how the classifers perform on a) elementary perceptual tasks with increasing parametric complexity, b) the position-angle experiment which compares pie and bar charts, c) the position-length experiment which compares grouped and divided bar charts, and d) the bars and framed rectangles experiment where visual cues help to perceive information.
-We present the results of these experiments to foster the understanding of how such classifiers can be successfully applied to data visualizations.
-On this journey, we also study how the feed-forward neural networks obey Weber's law which defines the proportional relation between perceivable information and distribution size.
-We further introduce a ranking of elementary visual encodings targeted towards modern deep neural networks and derive practical evidence for properties of convolutional neural networks such as the translation invariance.
+\abstract{
 Convolutional neural networks can successfully perform many computer vision tasks on images, and their learned representations are often said to mimic the early layers of the visual cortex.

### 2018-03-31 PAPER/02_experimental_setup.tex

@@
 Before use, we scale the generated images into an unbiased range: images to the range of $-0.5$ to $0.5$.
-Then, we add subtle random noise (uniformly distributed between $0-0.05$) to each pixel to introduce variation which prevents the networks from simply `remembering' each different image.
+Then, we add subtle random noise (uniformly distributed between $-0.025$--$0.025$) to each pixel to introduce variation which prevents the networks from simply `remembering' each different image.
 
@@
 \subsection{Measures and Analysis}
+\label{sec:measuresandanalysis}
 
@@
 \\~\\
-\noindent\textbf{Network Generalizability.} We evaluate generalizability by asking a network previously trained upon one task parameterization to answer questions about the same type of task stimuli but with more complex parameterization, e.g., estimating bar length without and with changes in stroke width.
+\noindent\textbf{Network Generalizability.} With sufficient capacity of trainable parameters, it is often said that a network can `memorize' the images if the data set has a low variability, and so it is important to consider this variability when evaluating different networks with fixed numbers of trainable parameters (Table~\ref{tab:parameters}).
+As discussed, we add noise to each stimulus image to prevent this.
+We also evaluate generalizability by asking a network previously trained upon one task parameterization to answer questions about the same type of task stimuli but with more complex parameterization, e.g., estimating bar length without and with changes in stroke width.
 

### 2018-03-31 PAPER/03_elementary_perceptual_tasks.tex

@@
 
-Cleveland and McGill describe the mapping of graphical elements to quantitative variables as \emph{elementary perceptual tasks} and introduce a list of ten different encodings in their paper~\cite{cleveland_mcgill,cleveland1985graphical}.
-These tasks are the low-level building blocks for information visualizations and encode quantities.
-Cleveland and McGill did not explicitly test human perception of single instances of these encodings.
-However, we test how each classifier measures encoded values using the elementary perceptual tasks and create visualizations of these tasks as rasterized images with different parametrizations (Table~\ref{tab:encoding_parameters}).
+Cleveland and McGill describe a set of elementary graphical perceptual tasks across ten encodings, where each encodes a quantitative variable in a graphical element or visual mark~\cite{cleveland_mcgill,cleveland1985graphical}.
+These tasks are the low-level building blocks for information visualizations (Table~\ref{tab:encoding_parameters}): estimating position on a common scale, position on non-aligned scales, length, direction (or slope), angle, area, volume, curvature,  shading (or ink density), and color saturation.
+As human color perception is complex, and because Cleveland and McGill perform no experiments with it, for now we leave it for future work.
 
-\subsection{Parametrizations}
-\label{sec:parametrizations}
-We generate multiple parameterizations for each elementary perceptual task and sequentially increase the number of parameters (Table~\ref{tab:encoding_parameters}).
-For instance, for \emph{Position Common Scale} we first only vary the $y$-position which yields just $60$ different parameters.
-We then include translation along the x-axis with a significant increase in variability.
-We then also add a variable spot size.
-This results in more complex datasets depending on the increase of variability.
-Table~\ref{tab:encoding_parameters} shows the different settings.
-It is important to consider this variability when evaluating different classifiers with individual trainable parameters (Table~\ref{tab:parameters}).
-In theory, classifiers can memorize the images if the data set has a low variability.
-We also counteract such behavior by adding noise.
+For the remaining nine tasks, we create visualizations as 100$\times$100 raster images, and test whether each of our networks is able to regress values from the images.
+As discussed in Section~\ref{sec:measuresandanalysis}, we generate multiple parameterizations for each elementary perceptual task to allow us to increase the number of parameters that the networks must estimate, and to measure performance as we move closer towards a general representation of visual marks.
+For instance, for \emph{Position Common Scale}, first we only vary the $y$-position of the spot to estimate against the scale, then we include translation along the x-axis, and then we vary the size of the spot size.
+These parameterizations are still simple---each increase is only slightly more complex for a human to solve---but it rapidly increases the number of possible images that the network must `learn' (Table~\ref{tab:encoding_parameters}).
+
+JT: Write something about comparison to Cleveland and McGill's 1985 experimental encoding?
+
+JT: Write something more intelligent about the different kinds of parameter variations.
+
+%Cleveland and McGill did not explicitly test human perception of single instances of these encodings. -> JT: We now know this isn't true.
 
@@
 \caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of the elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
-We sequentially increase the number of parameters (e.g.
-by adding translation) for every task.
+We sequentially increase the number of parameters for every task (e.g., by adding translation).
 This introduces variability and creates increasingly more complex datasets.}
@@
 
-We proposed four hypotheses entering the elementary perceptual task experiment:
+We state four hypotheses for the elementary perceptual task experiment:
 
 \begin{itemize}
-	\item \textbf{H1.1} \textbf{Our convolutional neural networks are able to regress quantitative variables from graphical elements.} We parametrize different visual encodings (Table~\ref{tab:encoding_parameters}) and test whether the CNNs can measure them with an accuracy similar to humans.
+	\item \textbf{H1.1} \textbf{The CNNs tested will be able to regress quantitative variables from graphical elements.} We parametrize different visual encodings (Table~\ref{tab:encoding_parameters}) and test whether the CNNs can measure them, and relate the results to accuracies obtained by humans on similar tasks.
 	\item \textbf{H1.2} \textbf{Computed perceptual performance is dependent on network architexture.} We evaluate multiple regressors with different numbers of trainable parameters.
-A more complex network (with a higher number of parameters) might perform better on elementary perceptual tasks.
-	\item \textbf{H1.3} \textbf{Some visual encodings are better than others for computations.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
+We expect a more complex network (with a higher number of trainable parameters) to perform better on elementary perceptual tasks.
+	\item \textbf{H1.3} \textbf{Some visual encodings will be easier to learn than others for the CNNs tested.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
 We investigate whether this order is also relevant for computing graphical perception.
-	\item \textbf{H1.4} \textbf{Networks trained on perceptual tasks can generalize to more or less complex variations of the same task.} Recent research suggests that convolutional neural networks are able to generalize by interpolating between different datapoints.
-While the underlying reasons are mainly yet unknown, this property allows them to perform on variations of a similar perceptual task.
-We create visual representations of the elementary perceptual tasks with different variability (Table~\ref{tab:encoding_parameters}).
-We suspect that such variations are not a problem for successfully regressing the quantitative variables and that the networks are able to generalize.
+	\item \textbf{H1.4} \textbf{Networks trained on perceptual tasks can generalize to more and less complex variations of the same task.} Empirical evidence suggests that CNNs are able to generalize by interpolating between different training data points.
+This property allows them to perform on variations of a similar perceptual task.
+We create visual representations of the elementary perceptual tasks with different variability, and expect that networks will be able to generalize when presented with slight task variations.
 \end{itemize}
@@
 
-The results of the elementary perceptual tasks experiment demonstrates that our networks are able to measure visually encoded quantities using low-level building blocks (Fig.~\ref{tab:figure1_results}).
-While the performance varies for the different encodings, we observe for all networks performance similar to the measured human performance by Cleveland and McGill~\cite{cleveland1985graphical}.
+%While the performance varies for the different encodings, we observe for all networks performance similar to the measured human performance by Cleveland and McGill~\cite{cleveland1985graphical}.
 These results confirm our initial hypotheses.
+\noindent{\textbf{Overall Accuracy.}} The tested CNNs and MLP are able to regress the visually encoded quantities in most cases (Fig.~\ref{fig:figure1_results}), with average error across all classifiers and tasks as \textit{MLAE}$=1.597$ ($SD=0.394$).
+% and \textit{MAE}=$2.89$ ($SD=0.848$).
+These values are similar or better than the range of MLAE=$3.02$--$3.82$ ($8-14\%$ error) for a similar experiment by Cleveland and McGill testing human estimation of relations between elementary perceptual tasks~\cite{cleveland1985graphical}.
+From these results, we \textbf{accept H1.1}.
 \\~\\
-\noindent{\textbf{Accuracy.}} We report the average error across all tasks for our networks as \textit{MLAE}: 
-$2.943$ ($SD=0.857$) for the multilayer perceptron, 
+\noindent{\textbf{Comparing Networks.}} 
+Across network architectures and training schemes, there is considerable difference in performance. In order of decreasing error: 
+The MLP has \textit{MLAE}$=2.943$ ($SD=0.857$), 
 for LeNet $2.125$ ($SD=0.38$), 
-VGG19 on ImageNet $0.979$ ($SD=0.581$), 
-for VGG19 trained from scratch $0.404$ ($SD=0.407$), 
-Xception on ImageNet $1.627$ ($SD=0.462$), 
-and for the re-trained Xception $1.504$ ($SD=0.493$).
-The average error for all classifiers is \textit{MLAE}$=1.597$ ($SD=0.394$) and \textit{MAE}=$2.89$ ($SD=0.848$).
-VGG19 trained from scratch reaches the best performance overall.
-These values are in the same ballpark as the range of $8-14\%$ error (approx.
-MLAE=$3$) for a similar experiment by Cleveland and McGill testing human estimation of relations between elementary perceptual tasks~\cite{cleveland1985graphical}.
-We therefor carefully confirm that our networks are able to regress the graphical elements to quantities within a reasonable range of error and \textbf{accept H1.1}.
-\\~\\
-\noindent{\textbf{Regressor Complexity.}} The performance varies not only across different visual encodings but also very much across the different network architectures.
-We compare the average regression performances for our networks and report the effect of the regressor as statistical significant ($F_{5,48}=20.392,p<0.01$).
+Xception trained on ImageNet $1.627$ ($SD=0.462$), 
+Xception trained from scratch $1.504$ ($SD=0.493$),
+VGG19 trained on ImageNet $0.979$ ($SD=0.581$), 
+and VGG19 trained from scratch $0.404$ ($SD=0.407$).
+
+Across tasks, we compare the average regression performances for our networks and report the effect of the network as statistically significant ($F_{5,48}=20.392,p<0.01$).
 Post hoc comparisons show that the differences between LeNet and the VGG19 network, independent of the used weights, are significant ($t_48=4.674,p<0.01$).
-VGG19 from scratch and Xception (both versions) are significant, from scratch ($t_48=4.87,p<0.01$) and for imagenet weights($t_48=5.621,p<0.01$).
+VGG19 from scratch and Xception (both versions) perform significantly differently, with Xception from scratch ($t_48=4.87,p<0.01$) and Xception with ImageNet weights ($t_48=5.621,p<0.01$).
 However, differences between LeNet and both Xception networks are not significant.
-In any case, the variation of results also visually inspectable in Fig.~\ref{fig:figure1_results} shows the variation of performance of the different networks.
-We therefor \textbf{partially accept H1.2}.
+Taken collectively, we \textbf{partially accept H1.2}, in that more tunable parameters does not automatically infer greater performance.
 \\~\\
 \noindent{\textbf{Ranking of Visual Encodings.}} Cleveland and McGill provide an ordering of elementary visual encodings based on theoretical arguments and experimental results.
-We re-create their ranking with MLAE values of our networks in Table~\ref{tab:ranking}.
-The rankings differ between the different networks and Cleveland and McGill's version.
-However, the ranking between networks using imagenet weights is identical.
-Since such a ranking see,s to heavily depend on the used regressor, the weights, and the initialization, we do not think a generalized ranking can be created and \textbf{reject H1.3}.
+We compare their ranking with rankings of our networks in Table~\ref{tab:ranking}.
+We note that the rankings between networks using ImageNet weights is identical; however, overall, there is significant variability, and we \textbf{reject H1.3}.
 
-\begin{table}[!ht]
+%Observing Figure~\ref{fig:figure1_results}, -> JT: we should do this.
+
+%Since such a ranking see,s to heavily depend on the used regressor, the weights, and the initialization, we do not think a generalized ranking can be created and \textbf{reject H1.3}.
+
+\begin{table}[tb]
 \centering
@@
 \toprule
-Human & MLP & LeNet & VGG19 * & \textbf{VGG19} & Xception * & Xception \\
+Human (CMcG) & MLP & LeNet & VGG19 * & \textbf{VGG19} & Xception * & Xception \\
 \midrule
@@
 %	\includegraphics[width=0.48\linewidth]{figure1_slim_right.pdf}
-	\caption{\textbf{Computational results of Elementary Perceptual Tasks experiment.} Log absolute error means and 95\% confidence intervals for computed perception of different classifiers on the \emph{elementary perceptual tasks} introduced by Cleveland and McGill 1984~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks on the most complex parametrizations.
-The * indicates networks which use ImageNet weights.
-We also compare against an approximate human baseline of $9\%$ error reported by Cleveland and McGill.}
+	\caption{\textbf{Elementary perceptual tasks results for most complex task parameterization.} \emph{Left:} Example stimuli image.
+\emph{Right:} MLAE and 95\% confidence intervals for different networks.
+The * indicates networks which use ImageNet weights up until the MLP, rather than being trained from scratch.}
 	\label{fig:figure1_results}

### 2018-03-31 PAPER/03_elementary_perceptual_tasks.tex
Section 4 going through.

@@
 
-For the remaining nine tasks, we create visualizations as 100$\times$100 raster images, and test whether each of our networks is able to regress values from the images.
+For the remaining nine tasks, we create visualizations as 100$\times$100 raster images, and test whether each of our networks is able to regress absolute values from the images.
 As discussed in Section~\ref{sec:measuresandanalysis}, we generate multiple parameterizations for each elementary perceptual task to allow us to increase the number of parameters that the networks must estimate, and to measure performance as we move closer towards a general representation of visual marks.
-For instance, for \emph{Position Common Scale}, first we only vary the $y$-position of the spot to estimate against the scale, then we include translation along the x-axis, and then we vary the size of the spot size.
-These parameterizations are still simple---each increase is only slightly more complex for a human to solve---but it rapidly increases the number of possible images that the network must `learn' (Table~\ref{tab:encoding_parameters}).
-
-JT: Write something about comparison to Cleveland and McGill's 1985 experimental encoding?
-
-JT: Write something more intelligent about the different kinds of parameter variations.
+For instance, for \emph{Position Common Scale}, first we only vary the $y$-position of the spot to estimate against the scale, then we include translation along the x-axis, and then we vary the size of the spot size (Table~\ref{tab:encoding_parameters}).
+These parameterizations are still simple---each increase is only slightly more complex for a human to solve---but it increases the number of possible images for the network to `learn'.
 
@@
 
-We state four hypotheses for the elementary perceptual task experiment:
+\noindent \textbf{H1.1} \textbf{The CNNs tested will be able to regress quantitative variables from graphical elements.} We parametrize different visual encodings (Table~\ref{tab:encoding_parameters}) and test whether the CNNs can measure them, and relate the results to accuracies obtained by humans on similar tasks.
 
-\begin{itemize}
-	\item \textbf{H1.1} \textbf{The CNNs tested will be able to regress quantitative variables from graphical elements.} We parametrize different visual encodings (Table~\ref{tab:encoding_parameters}) and test whether the CNNs can measure them, and relate the results to accuracies obtained by humans on similar tasks.
-	\item \textbf{H1.2} \textbf{Computed perceptual performance is dependent on network architexture.} We evaluate multiple regressors with different numbers of trainable parameters.
-We expect a more complex network (with a higher number of trainable parameters) to perform better on elementary perceptual tasks.
-	\item \textbf{H1.3} \textbf{Some visual encodings will be easier to learn than others for the CNNs tested.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
+\noindent \textbf{H1.2} \textbf{CNN perceptual performance will depend on network architexture.} We evaluate multiple regressors with different numbers of trainable parameters.
+We expect a more complex network (with more trainable parameters) to perform better on elementary perceptual tasks.
+
+\noindent \textbf{H1.3} \textbf{Some visual encodings will be easier to learn than others for the CNNs tested.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
 We investigate whether this order is also relevant for computing graphical perception.
-	\item \textbf{H1.4} \textbf{Networks trained on perceptual tasks can generalize to more and less complex variations of the same task.} Empirical evidence suggests that CNNs are able to generalize by interpolating between different training data points.
-This property allows them to perform on variations of a similar perceptual task.
+
+\noindent \textbf{H1.4} \textbf{Networks trained on perceptual tasks will generalize to more complex variations of the same task.} Empirical evidence suggests that CNNs are able to generalize by interpolating between different training data points, and so perform on variations of a similar perceptual task.
 We create visual representations of the elementary perceptual tasks with different variability, and expect that networks will be able to generalize when presented with slight task variations.
-\end{itemize}
 
@@
 
+% These values are similar or better than the range of MLAE=$3.02$--$3.82$ ($8-14\%$ error) for a similar experiment by Cleveland and McGill testing human estimation of relations between elementary perceptual tasks~\cite{cleveland1985graphical}.
+
 %While the performance varies for the different encodings, we observe for all networks performance similar to the measured human performance by Cleveland and McGill~\cite{cleveland1985graphical}.
@@
 % and \textit{MAE}=$2.89$ ($SD=0.848$).
-These values are similar or better than the range of MLAE=$3.02$--$3.82$ ($8-14\%$ error) for a similar experiment by Cleveland and McGill testing human estimation of relations between elementary perceptual tasks~\cite{cleveland1985graphical}.
 From these results, we \textbf{accept H1.1}.
@@
 We compare their ranking with rankings of our networks in Table~\ref{tab:ranking}.
-We note that the rankings between networks using ImageNet weights is identical; however, overall, there is significant variability, and we \textbf{reject H1.3}.
+Overall, there is significant variability in the rankings between architectures (Fig.~\ref{fig:figure1_results}).
+However, area estimation is an easier task for all networks, while direction and angle estimation are more difficult.
+It is harder to distinguish differences between position, length, curvature, and shading tasks.
+Further, the volume tasks suffers high variability in performance across cross-validation splits, which suggests that the image noise affects the outcome more than for other tasks.
+In sum, we \textbf{partially accept H1.3}.
+Finally, we note that the rankings between networks using ImageNet weights is identical, which suggests that the information about elementary perceptual tasks gained from those natural images is similar (given a sufficiently-complex network).
+\\~\\
+\noindent{\textbf{Cross-network Variability and Network Generalizability.}} We measure regression performance across networks trained with different parameterizations of the elementary perceptual tasks (Fig.~\ref{fig:cross_network}).
+For our best performing network (VGG19 trained from scratch), we observe that accuracy decreases only slightly as the parameterization becomes more complex so long as training examples expressing all variability are included (diagonal entries in each matrix).
+However, VGG19 is unable to generalize to added translation or stroke width variations in the encodings, leading to increases in error.
+As such, we \textbf{reject H1.4}.
 
@@
 \centering
-\caption{\textbf{Ranking of Elementary Perceptual Tasks.} We create a ranking of the elementary perceptual tasks for all of our networks.
-We report midmean logistic absolute errors (MLAE) for each network averaged across multiple runs on the most complex parametrization of the individual stimuli.
-The VGG19 network performs best compared on all tasks.
+\caption{\textbf{Elementary Perceptual Task Ranking.} We report midmean logistic absolute errors (MLAE) for each network averaged across multiple runs on the most complex parametrization of each task.
 For human performance, we report the ranking of Cleveland and McGill~\cite{cleveland_mcgill}.
-The VGG19 * and Xception * networks use ImageNet weights and yield identical rankings.}
+VGG19 performs best overall, while VGG19 * and Xception * networks using ImageNet weights yield identical rankings.}
 \resizebox{\linewidth}{!}{
@@
 
-\begin{figure}[!ht]
+\begin{figure}[p]
 	\centering
-	\includegraphics[width=.8\linewidth]{figure1_slim_only_last.pdf}
+	\includegraphics[width=\linewidth]{figure1_slim_only_last.pdf}
 %	\includegraphics[width=0.48\linewidth]{figure1_slim_right.pdf}
@@
 	\label{fig:figure1_results}
+%	\vspace*{1in}
 \end{figure}
 
-\noindent{\textbf{Cross-classifier Variability and Network Generalizability.}} We measure regression performance across classifiers trained on the different parameterizations of the elementary perceptual tasks.
-Looking at the results in Fig.~\ref{fig:cross_network}, we can observe that even slight variations of an encoding such as added translation in Y or X drastically increases the error.
-We think that the networks are not really learning concepts of perception but rather learn the variation of pixel values.
-
-\begin{figure}[!ht]
+\begin{figure}[p]
 	\centering
-	  \includegraphics[width=.75\linewidth]{cross_network_small_VGG19_area.pdf}
-	  \includegraphics[width=.75\linewidth]{cross_network_small_VGG19_length.pdf}
-	  \includegraphics[width=.75\linewidth]{cross_network_small_VGG19_position_common_scale.pdf}
-	  \includegraphics[width=.75\linewidth]{cross_network_small_VGG19_shading.pdf}
-  \caption{\textbf{Cross-classifier variability for perceptual tasks.} We use predictions of the VGG19 network trained on different parametrizations of \emph{area}, \emph{length}, \emph{position common scale}, and \emph{shading}.
-These are the top 4 encodings in the ranking for this network.
-We measure the mean logistic absolute error (MLAE) -- the lower score, the better.
-Regressors trained on stimuli with variable position can generalize even if the axis of translation varies.
-However, if they are trained on fixed positions of the stimuli they are not able to measure translated encodings.}
+	  \includegraphics[width=.8\linewidth]{cross_network_small_VGG19_position_common_scale.pdf}
+	  \includegraphics[width=.8\linewidth]{cross_network_small_VGG19_length.pdf}
+	  \includegraphics[width=.8\linewidth]{cross_network_small_VGG19_area.pdf}
+	  \includegraphics[width=.8\linewidth]{cross_network_small_VGG19_shading.pdf}
+  \caption{\textbf{Cross-classifier variability for perceptual tasks.} VGG19 networks trained on different parametrizations (X-axis) tested across different parameterizations (Y-axis), for the top four performing encodings.
+Diagonal matrix entries represent networks trained and tested on the same parameterizations.
+Below diagonal entries are scenarios where the test data has more parameters than the training data; below diagonal entries have fewer.
+We measure the mean logistic absolute error (MLAE)---the lower the score, the better.
+VGG19 becomes only slightly less accurate as the parameterization becomes more complex; however, it is unable to generalize to unseen element translations as error increases rapidly.
+Note that all networks showed similar behavior.}
 	\label{fig:cross_network}

### 2018-03-31 PAPER/03_elementary_perceptual_tasks.tex
Updates.

@@
 These results confirm our initial hypotheses.
-\noindent{\textbf{Overall Accuracy.}} The tested CNNs and MLP are able to regress the visually encoded quantities in most cases (Fig.~\ref{fig:figure1_results}), with average error across all classifiers and tasks as \textit{MLAE}$=1.597$ ($SD=0.394$).
-% and \textit{MAE}=$2.89$ ($SD=0.848$).
+\noindent{\textbf{Overall Accuracy.}} The tested CNNs and MLP are able to regress the visually encoded quantities in most cases (Fig.~\ref{fig:figure1_results}), with average error across all classifiers and tasks as \textit{MLAE}$=1.598$ ($SD=0.392$) and \textit{MAE}=$2.903$ ($SD=0.845$).
 From these results, we \textbf{accept H1.1}.
@@
 Xception trained on ImageNet $1.627$ ($SD=0.462$), 
-Xception trained from scratch $1.504$ ($SD=0.493$),
+Xception trained from scratch $1.511$ ($SD=0.485$),
 VGG19 trained on ImageNet $0.979$ ($SD=0.581$), 
@@
 
-Across tasks, we compare the average regression performances for our networks and report the effect of the network as statistically significant ($F_{5,48}=20.392,p<0.01$).
-Post hoc comparisons show that the differences between LeNet and the VGG19 network, independent of the used weights, are significant ($t_48=4.674,p<0.01$).
-VGG19 from scratch and Xception (both versions) perform significantly differently, with Xception from scratch ($t_48=4.87,p<0.01$) and Xception with ImageNet weights ($t_48=5.621,p<0.01$).
+Across tasks, we compare the average regression performances for our networks and report the effect of the network as statistically significant ($F_{5,48}=20.470,p<0.01$).
+Post hoc comparisons show that the differences between LeNet and the VGG19 network, independent of the used weights, are significant ($t_48=4.674,p<0.01$ and $t_48=8.746,p<0.01$).
+VGG19 from scratch and Xception (both versions) perform significantly differently, with Xception from scratch ($t_48=4.944,p<0.01$) and Xception with ImageNet weights ($t_48=5.621,p<0.01$).
 However, differences between LeNet and both Xception networks are not significant.
@@
 In sum, we \textbf{partially accept H1.3}.
-Finally, we note that the rankings between networks using ImageNet weights is identical, which suggests that the information about elementary perceptual tasks gained from those natural images is similar (given a sufficiently-complex network).
+Finally, we note that the rankings between networks using ImageNet weights are identical, which suggests that the information about elementary perceptual tasks gained from those natural images is similar (given a sufficiently-complex network).
 \\~\\
@@
 \midrule
-\multicolumn{7}{l}{\emph{Position Common Scale}} \\
-1. & 7. (3.84) & 2. (1.36) & 5. (1.02) & \textbf{3.} (-0.04) & 5. (1.65) & 1. (1.04)\\
-\multicolumn{7}{l}{\emph{Position Non aligned Scale}} \\
-2. & 6. (3.61) & 1. (1.35) & 6. (1.09) & \textbf{5.} (0.26) & 6. (1.71) & 2. (1.06)\\
+\multicolumn{7}{l}{\emph{Position common scale}} \\
+1. & 7. (3.84) & 2. (1.36) & 5. (1.02) & \textbf{3 (-0.04)} & 5. (1.65) & 2. (1.04) \\
+\multicolumn{7}{l}{\emph{Position non-aligned scale}} \\
+2. & 6. (3.61) & 1. (1.35) & 6. (1.09) & \textbf{5 (0.26)} & 6. (1.71) & 1. (1.02) \\
 \multicolumn{7}{l}{\emph{Length}} \\
-3. & 1. (1.99) & 8. (3.19) & 4. (0.87) & \textbf{2.} (-0.14) & 4. (1.59) & 3. (1.11) \\
+3. & 1. (1.99) & 8. (3.19) & 4. (0.87) & \textbf{2 (-0.14)} & 4. (1.59) & 3. (1.11) \\
 \multicolumn{7}{l}{\emph{Direction}} \\
-3. & 9. (4.65) & 7. (3.07) & 9. (2.84) & \textbf{8.} (0.92) & 9. (3.46) & 6. (1.57) \\
+3. & 9. (4.65) & 7. (3.07) & 9. (2.84) & \textbf{8 (0.92)} & 9. (3.46) & 6. (1.57) \\
 \multicolumn{7}{l}{\emph{Angle}} \\
-3. & 8. (4.61) & 9. (3.33) & 8. (2.31) & \textbf{9.} (0.99) & 8. (2.60) & 7. (1.72) \\
+3. & 8. (4.61) & 9. (3.33) & 8. (2.31) & \textbf{9 (0.99)} & 8. (2.60) & 7. (1.69) \\
 \multicolumn{7}{l}{\emph{Area}} \\
-4. & 2. (2.01) & 5. (2.21) & 1. (0.49) & \textbf{1.} (-0.17) & 1. (0.80) & 5. (1.38) \\
+4. & 2. (2.01) & 5. (2.21) & 1. (0.49) & \textbf{1 (-0.17)} & 1. (0.80) & 5. (1.38) \\
 \multicolumn{7}{l}{\emph{Volume}} \\
-5. & 4. (2.38) & 4. (1.91) & 7. (1.16) & \textbf{7.} (0.87) & 7. (2.03) & 9. (2.10) \\
+5. & 4. (2.38) & 4. (1.91) & 7. (1.16) & \textbf{7 (0.87)} & 7. (2.03) & 9. (2.10) \\
 \multicolumn{7}{l}{\emph{Curvature}} \\
-5. & 3. (2.34) & 3. (1.81) & 2. (0.71) & \textbf{6.} (0.28) & 2. (1.17) & 4. (1.13) \\
+5. & 3. (2.34) & 3. (1.81) & 2. (0.71) & \textbf{6 (0.28)} & 2. (1.17) & 4. (1.13) \\
 \multicolumn{7}{l}{\emph{Shading}} \\
-6. & 5. (3.04) & 6. (2.23) & 3. (0.73) & \textbf{4.} (0.14) & 3. (1.57) & 8. (1.82) \\
+6. & 5. (3.04) & 6. (2.23) & 3. (0.73) & \textbf{4 (0.14)} & 3. (1.57) & 8. (1.82) \\
 

### 2018-03-31 PAPER/SUPPLEMENTAL/supplemental.tex
Updates.

@@
 %  \centering
-%  \includegraphics[width=\linewidth]{teaser.pdf}
-%  \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking visual cortex inspired machine learning classifiers to assess the relationship between values encoded in pie charts and bar charts.
-Similar to the findings of Clevenland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs read quantities more accurately from bar charts (mean squared error, MSE in green).}
-%	\label{fig:teaser}
+%\par{\\~\\~\\~\\}
 %}
@@
 %% Uncomment below to disable the manuscript note
-%\renewcommand{\manuscriptnotetxt}{}
+\renewcommand{\manuscriptnotetxt}{}
 
@@
 
-\vgtcinsertpkg
+%\vgtcinsertpkg
 
@@
 %% the only exception to this rule is the \firstsection command
+\begin{figure*}
+\centering
+This material contains several plots accompanying the main manuscript.
+\end{figure*}
+\clearpage
+\begin{figure*}[t]
+	\centering
+	  \includegraphics[width=.8\linewidth]{error_inspection_new.pdf}
+  \caption{\textbf{Error Distributions.} Error distributions of our networks when decoding elementary perceptual tasks.}
+	\label{fig:errors}
+\end{figure*}
+
+\begin{figure*}[t]
+	\centering
+	  \includegraphics[width=\linewidth]{cross_network_new.pdf}
+  \caption{\textbf{Cross-Network Validation.} Our networks fail when the stimuli is translated.}
+	\label{fig:cross_network}
+\end{figure*}
+
+\begin{figure*}[t]
+	\centering
+		\includegraphics[width=.48\linewidth]{../gfx/figure1_slim_left.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/figure1_slim_right.pdf}
+  \caption{\textbf{Cross-Network Validation.} Our networks fail when the stimuli is translated.}
+	\label{fig:cross_network}
+\end{figure*}
+
+\begin{figure}[t]
+	\centering
+	  \includegraphics[width=\linewidth]{../gfx/figure3_mlae_better_all_AA.pdf}
+  \caption{\textbf{Anti-Aliasing.} We test whether anti-aliasing effects the performance of our networks on pie charts by measuring MLAE.
+The difference is not statistically significant ($F(1,30)=0.341,p>0.5$).}
+	\label{fig:aa}
+\end{figure}
 

### 2018-03-31 PAPER/SUPPLEMENTAL/supplemental.tex
Updates.

@@
 	  \includegraphics[width=.8\linewidth]{error_inspection_new.pdf}
-  \caption{\textbf{Error Distributions.} Error distributions of our networks when decoding elementary perceptual tasks.}
+  \caption{\textbf{Error distributions.} Error distributions of our networks when decoding elementary perceptual tasks.}
 	\label{fig:errors}
@@
 	  \includegraphics[width=\linewidth]{cross_network_new.pdf}
-  \caption{\textbf{Cross-Network Validation.} Our networks fail when the stimuli is translated.}
+  \caption{\textbf{Cross-network variability.} Our networks fail when the stimuli changes through translation or stroke width.
+The x-labels indicate the training configuration while the y-labels indicate the stimuli variation.
+Numbers represent MLAE.}
 	\label{fig:cross_network}
@@
 	  \includegraphics[width=.48\linewidth]{../gfx/figure1_slim_right.pdf}
-  \caption{\textbf{Cross-Network Validation.} Our networks fail when the stimuli is translated.}
+  \caption{\textbf{Elementary perceptual tasks.} Midmean logistic absolute errors (MLAE) for all generated stimuli and across all networks.
+The * indicates networks which use ImageNet weights instead of bein trained from scratch.}
+	\label{fig:cross_network}
+\end{figure*}
+
+\begin{figure*}[t]
+	\centering
+		\includegraphics[width=\linewidth]{../gfx/figure1_boxplot_new.pdf}
+  \caption{\textbf{Elementary perceptual tasks.} Midmean logistic absolute errors (MLAE) visualized as box plots.}
 	\label{fig:cross_network}
@@
 	  \includegraphics[width=\linewidth]{../gfx/figure3_mlae_better_all_AA.pdf}
-  \caption{\textbf{Anti-Aliasing.} We test whether anti-aliasing effects the performance of our networks on pie charts by measuring MLAE.
+  \caption{\textbf{Anti-aliasing.} We test whether anti-aliasing effects the performance of our networks on pie charts by measuring MLAE.
 The difference is not statistically significant ($F(1,30)=0.341,p>0.5$).}

### 2018-03-31 PAPER/05_position_length_experiment.tex
Minor fixes
Newest version.
Section 5 in the works.

--- 2018-03-20 PAPER/05_position_length_experiment.tex
@@
-\section{Position-Length Experiment}
+\section{Experiment: Position-Length}
 
-This is the one where we estimate two selected bars compared to the longest one - very similar to the previous one but, not yet done.
-We basically test divided versus grouped barchart and we estimate one relation between two marked quanitities: what percent the smaller is of the larger.
+The position-length experiment by Cleveland and McGill involves the perception of five types of visualizations~\cite{cleveland_mcgill}.
+The visualizations are either grouped bar charts or divided bar charts (Fig.~\ref{fig:figure4_mlae}.
+Both types of graphs can show the same information but the elementary perceptual task is different.
+According to the theory by Cleveland and McGill, a grouped bar chart always involves the judgement of positions along a common scale.
+On the other hand, a divided bar chart might require length judgements in addition.
+Figure~\ref{fig:figure4_mlae} shows different charts of both types on the left.
+As identified by Cleveland and McGill, types 1, 2, and 3 involve the judgement of positions along a common scale while types 4 and 5 involve the measure of length.
+This means that this experiment does not just compare grouped versus divided bar charts but rather comparing position and length judgements.
 
-There are five types: type 1-3 this is a position judgement along a common scale.
-(btw all classifiers seem to do that extremely well in the elementary tasks so we assume this will work well here too).
-Types 4-5 are length judgements and we know that the classifiers struggle with that quite a bit.
+For data generation we follow the same approach as in the original experiment. We generate pairs from ten values generated using Cleveland and McGill's equation
 
-The setup from Cleveland McGill is first a classification task: which one is smaller? and then a regression task: how much smaller. So we have to see how to encode this.
+\begin{equation}
+s_i = 10\times10^{(i-1)/12},\textnormal{~~~~~}i=1,...,10,
+\end{equation}
+
+with two different values for each pair.
+We then generate 8 other random values in the range of 10 and 93.
+These boundaries were chosen because we create rasterized visualizations using $100\times100$ pixel and want to stay in frame for a convolutional filter size of $5\times5$ in LeNet.
+We then visually encode the ten values as type 1--5.
+The paired quantities get marked by a single pixel.
+These are our quantities of interest and the task is to estimate the ratio of the smaller to the larger.
+We model this task as a single value regression.
+For type 4, we follow Cleveland and McGill's constraint that neither the top or the bottom of the marked quantities match to force length estimations rather than position.
+
+We model this experiment as a single value regression task and ask networks to estimate the percentage of the smaller to the larger marked quantity in each visual encoding.
+The network first has to find the two marked quantities, then identify the smaller one, and finally estimate the ratio in comparison to the larger quantitiy.
+The 8 random quantities (which should be ignored by the network) push the number of possible permutations to a massive $9.20E+16$ and creates a very challenging problem.
+
+Finally, we also train `multi' networks which include all five types which creates an even bigger challenge.
 
@@
 \begin{itemize}
-	\item \textbf{H3.1} \textbf{Grouped bar charts are better computational perceivable than divided bar charts.} A grouped bar chart involves judging a position while a divided bar chart most likely (if not the bottom is looked at) requires length judgements.
-Classifiers are better at judging position than at judging length so grouped bar charts are easier to grasp in terms of computational perception.
-	\item \textbf{H3.2} \textbf{not yet} Any ideas?
+	\item \textbf{H3.1} \textbf{Our networks can estimate grouped bar charts better than divided bar charts.}
+
+	Grouped bar charts are better computational perceivable than divided bar charts.} A grouped bar chart involves judging a position while a divided bar chart most likely (if not type 2) requires length judgements.
+
+	\item \textbf{H3.2} \textbf{Training `multi' networks } Any ideas?
 \end{itemize}
 
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{position_length_original1.png}
-	  \includegraphics[width=\linewidth]{position_length_original2.png}
-
-  \caption{\textbf{Position-Length Experiment.} (Not yet) Rasterized versions of the graphs of Cleveland and McGill's position-length experiment.
+%\begin{figure*}[t]
+%	  \includegraphics[width=\linewidth]{figure4_overview.pdf}
+%	  
+%  \caption{\textbf{Position-Length Experiment.} (Not yet) Rasterized versions of the graphs of Cleveland and McGill's position-length experiment.
 The perceptual task involves comparing.
@@
 The first task is to select which of the marked quantities is smaller (classification) and the second task is to specify how much smaller it is (regression).}
-	\label{fig:position_length_experiment}
+%	\label{fig:position_length_experiment}
+%\end{figure*}
+%\begin{table}[h]
+%\centering
+%\caption{\textbf{Position-Length Experiment.} Rasterized versions of the graphs of Cleveland and McGill's position-length experiment.
+The perceptual task involves comparing the two dot-marked quantities across five different visual encodings of either grouped or divided bar charts.
+We evaluate which type of bar chart performs better with our neural networks.
+The two marked values are chosen from a set of ten pairs which defines the dual regression task.
+Since the other 8 values are chosen randomly, the parameter space for images of this experiment is massive.}
+%\resizebox{\linewidth}{!}{
+%\begin{tabular}{lllr}
+%	\toprule
+%	\multicolumn{2}{l}{~} & ~ & Permutations\\
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure4_type_1.pdf}} & \makecell[tl]{Type 1: \emph{Grouped Bar Chart}\\~~~Perceptual Task: \emph{Position}\\~ \\~ \\} &~& \makecell[tr]{~\\ $9.20E+16$}\\
+%
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure4_type_2.pdf}} & \makecell[tl]{Type 2: \emph{Divided Bar Chart}\\~~~Perceptual Task: \emph{Position}\\~ \\~ \\} &~& \makecell[tr]{~\\ $9.20E+16$}\\
+%	
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure4_type_3.pdf}} & \makecell[tl]{Type 3: \emph{Grouped Bar Chart}\\~~~Perceptual Task: \emph{Position}\\~ \\~ \\} &~& \makecell[tr]{~\\ $9.20E+16$}\\
+%	
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure4_type_4.pdf}} & \makecell[tl]{Type 4: \emph{Divided Bar Chart}\\~~~Perceptual Task: \emph{Length}\\~ \\~ \\} &~& \makecell[tr]{~\\ $9.20E+16$}\\
+%	
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure4_type_5.pdf}} & \makecell[tl]{Type 5: \emph{Divided Bar Chart}\\~~~Perceptual Task: \emph{Length}\\~ \\~ \\} &~& \makecell[tr]{~\\ $9.20E+16$}\\
+%
+%	\bottomrule
+%\end{tabular}
+%}
+%\label{tab:pos_length_parameters}
+%\end{table}
+
+\subsection{Discussion}
+
+JT: Look at the relative difficulty of the tasks.
+In Cleveland and McGill, types 1-5 were post-ordered by their log error such that type 1 was easiest and type 5 was hardest.
+Is this still the case with our CNNs?
+
+\subsection{Results}
+
+\begin{figure}[!ht]
+	\centering
+	  \includegraphics[width=\linewidth]{figure4_mlae_with_multi_and_humans_all.pdf}
+  \caption{\textbf{Computational results of the position-length experiment.} \textit{Left:} Rasterized visualizations of type 1--5 for divided and grouped bar charts of Cleveland and McGill's position-length experiment.
+\textit{Right:} MLAE and $95\%$ confidence intervals for different regressors estimating the value of marked quantities in the visualizations.
+The VGG19 * and Xception * networks are using ImageNet weights.
+The top 5 rows represent networks trained on a single encoding while the last row shows `multi' networks which were trained on a random stream of types 1--5.
+We visually compare against human performance from the original experiment.}
+	\label{fig:figure4_mlae}
 \end{figure}

### 2018-03-31 PAPER/05_position_length_experiment.tex
Supplemental changes.

@@
 
-	Grouped bar charts are better computational perceivable than divided bar charts.} A grouped bar chart involves judging a position while a divided bar chart most likely (if not type 2) requires length judgements.
+	Grouped bar charts are better computational perceivable than divided bar charts.
+A grouped bar chart involves judging a position while a divided bar chart most likely (if not type 2) requires length judgements.
 

### 2018-03-31 PAPER/SUPPLEMENTAL/supplemental.tex
Supplemental changes.

@@
 \maketitle
+
+We present several plots which contain complete results for the elementary perceptual task experiment, for which, given the number of experiments, the main paper presented only a selection.
+We report MLAE for all added parameters to the stimuli, rather than just the most complex parameterization as in the main paper (Figures~\ref{fig:epc_mlae} and \ref{fig:epc_mlae_boxplots}).
+From this, we see that performance is largely equal across parameters, showing that most networks have sufficient capacity for the given parameterizations.
+
+We also report complete results for the cross-network variability experiment on the elementary perceptual task experiment (Figure \ref{fig:cross_network}).
+As in the main paper, this shows that our networks are not able to generalize to additional translation or stroke width parameters without representative training data.
+
+Further, we show how the errors for each network are distributed, across elementary perceptual tasks and across different cross-validation splits (Figure~\ref{fig:errors}).
+Most errors are approximately normally distributed, though our CNN with less parameters (LeNet) and our MLP often have errors which are farther from a normal distribution, and show structure.
+
+\section{Experiment: Anti-aliasing}
+We also include an experimental result showing that, for our stimuli, adding anti-aliasing to the line generation was not important to the tested CNNs (Figure~\ref{fig:aa}).
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\linewidth]{../gfx/figure3_mlae_better_all_AA.pdf}
+\caption{\textbf{Anti-aliasing.} We test whether anti-aliasing effects the performance of our networks on pie charts by measuring MLAE.
+The difference is not statistically significant ($F(1,30)=0.341,p>0.5$).}
+\label{fig:aa}
+\end{figure}
+
 %% The ``\maketitle'' command must be the first command after the
@@
 %% the only exception to this rule is the \firstsection command
-\begin{figure*}
+
+\begin{figure*}[p]
 	\centering
-This material contains several plots accompanying the main manuscript.
-\end{figure*}
-\clearpage
-\begin{figure*}[t]
-	\centering
-	  \includegraphics[width=.8\linewidth]{error_inspection_new.pdf}
-  \caption{\textbf{Error distributions.} Error distributions of our networks when decoding elementary perceptual tasks.}
-	\label{fig:errors}
+		\includegraphics[width=.48\linewidth]{../gfx/figure1_slim_left.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/figure1_slim_right.pdf}
+  \caption{\textbf{Elementary perceptual tasks.} Midmean logistic absolute errors (MLAE) for all generated stimuli and across all networks.
+The * indicates networks which use ImageNet weights instead of bein trained from scratch.}
+	\label{fig:epc_mlae}
 \end{figure*}
 
-\begin{figure*}[t]
+\begin{figure*}[p]
+	\centering
+		\includegraphics[width=\linewidth]{../gfx/figure1_boxplot_new.pdf}
+  \caption{\textbf{Elementary perceptual tasks.} Midmean logistic absolute errors (MLAE) visualized as box plots.}
+	\label{fig:epc_mlae_boxplots}
+\end{figure*}
+
+\begin{figure*}[p]
 	\centering
@@
 
-\begin{figure*}[t]
+\begin{figure*}[p]
 	\centering
-		\includegraphics[width=.48\linewidth]{../gfx/figure1_slim_left.pdf}
-	  \includegraphics[width=.48\linewidth]{../gfx/figure1_slim_right.pdf}
-  \caption{\textbf{Elementary perceptual tasks.} Midmean logistic absolute errors (MLAE) for all generated stimuli and across all networks.
-The * indicates networks which use ImageNet weights instead of bein trained from scratch.}
-	\label{fig:cross_network}
+	  \includegraphics[width=.8\linewidth]{error_inspection_new.pdf}
+  \caption{\textbf{Error distributions.} Error distributions of our networks when decoding elementary perceptual tasks.}
+	\label{fig:errors}
 \end{figure*}
-
-\begin{figure*}[t]
-	\centering
-		\includegraphics[width=\linewidth]{../gfx/figure1_boxplot_new.pdf}
-  \caption{\textbf{Elementary perceptual tasks.} Midmean logistic absolute errors (MLAE) visualized as box plots.}
-	\label{fig:cross_network}
-\end{figure*}
-
-\begin{figure}[t]
-	\centering
-	  \includegraphics[width=\linewidth]{../gfx/figure3_mlae_better_all_AA.pdf}
-  \caption{\textbf{Anti-aliasing.} We test whether anti-aliasing effects the performance of our networks on pie charts by measuring MLAE.
-The difference is not statistically significant ($F(1,30)=0.341,p>0.5$).}
-	\label{fig:aa}
-\end{figure}
 
@@
 %\bibliographystyle{abbrv}
-\bibliographystyle{abbrv-doi}
+%\bibliographystyle{abbrv-doi}
 %\bibliographystyle{abbrv-doi-narrow}
@@
 
-\bibliography{../paper.bib}
+%\bibliography{../paper}
 \end{document}

### 2018-03-31 PAPER/03_elementary_perceptual_tasks.tex

@@
 Across tasks, we compare the average regression performances for our networks and report the effect of the network as statistically significant ($F_{5,48}=20.470,p<0.01$).
-Post hoc comparisons show that the differences between LeNet and the VGG19 network, independent of the used weights, are significant ($t_48=4.674,p<0.01$ and $t_48=8.746,p<0.01$).
-VGG19 from scratch and Xception (both versions) perform significantly differently, with Xception from scratch ($t_48=4.944,p<0.01$) and Xception with ImageNet weights ($t_48=5.621,p<0.01$).
+Post hoc comparisons show that the differences between LeNet and the VGG19 network, independent of the used weights, are significant ($t_{16}=4.674,p<0.01$ and $t_{16}=8.746,p<0.01$).
+VGG19 from scratch and Xception (both versions) perform significantly differently, with Xception from scratch ($t_{16}=4.944,p<0.01$) and Xception with ImageNet weights ($t_{16}=5.621,p<0.01$).
 However, differences between LeNet and both Xception networks are not significant.
@@
 	  \includegraphics[width=.8\linewidth]{cross_network_small_VGG19_shading.pdf}
-  \caption{\textbf{Cross-classifier variability for perceptual tasks.} VGG19 networks trained on different parametrizations (X-axis) tested across different parameterizations (Y-axis), for the top four performing encodings.
+  \caption{\textbf{Cross-network variability for perceptual tasks.} VGG19 networks trained on different parametrizations (X-axis) tested across different parameterizations (Y-axis), for the top four performing encodings.
 Diagonal matrix entries represent networks trained and tested on the same parameterizations.

### 2018-03-31 PAPER/04_position_angle_experiment.tex
Minor fixes
Newest version.
Update section 4.

@@
-\section{Position-Angle Experiment}
+\clearpage
+\section{Experiment: Position-Angle}
 
-The position-angle experiment was originally performed by Cleveland and McGill to measure whether humans can better perceive quantities encoded as positions or as angles~\cite{cleveland_mcgill}.
-The actual experiment then compares pie charts versus bar charts since these map down to elementary position and angle judgement.
-We create rasterized images mimicking Cleveland and McGill's proposed encoding and investigate computational perception of our four classifiers.
+Cleveland and McGill measure human perception of quantities encoded as positions and as angles through their position-angle experiment~\cite{cleveland_mcgill}.
+The actual experiment compares pie charts versus bar charts since these map down to elementary position and angle judgement.
+We create rasterized images mimicking Cleveland and McGill's proposed encoding and investigate computational perception of our four networks.
+
+We follow the data generation according to Cleveland and McGill and generate datasets of 5 numbers which add to 100.
+The numbers fulfill their proposed requirements of being greater than 3 smaller than 39, with differences between values being greater than $0.1$.
+Similar to Cleveland and McGill, we create pie chart and bar chart representations (Fig.~\ref{fig:figure3_mlae}, left).
+We create these visualizations as $100\times100$ pixel raster images.
+We then mark the largest quantitiy of the five in each visualization with a single pixel dot.
+The regression task, again similar to the experiment bei Cleveland and McGill, is to estimate what value each quantity is in relation to the marked largest.
+Since the position of the largest element changes, we generate the targets in such fashion that the largest element is marked with 1 and the other quantities follow counter-clockwise for the pie chart and to the right for the bar chart.
+To be successful, the networks essentially first have to find the marked quantity, have the `rolling' encoding figured out, and then estimate the quantities properly.
+We generate the pie chart and bar chart visualizations with $878,520$ possible permutations each which renders this regression task as a decent problem.
 
@@
 %\end{figure}
-\begin{table}[h]
-\centering
-\caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
+%\begin{table}[h]
+%\centering
+%\caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
 The experimental task involves the judgement of different encoded values in comparison to the largest encoded values.
@@
 In their paper, Cleveland and McGill report less errors using bar charts.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lllr}
-	\toprule
-	\multicolumn{2}{l}{~} & ~ & Permutations\\
-
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure3_Bar_Chart.pdf}} & \makecell[tl]{Type 1: \emph{Bar Chart}\\~~~Perceptual Task: \emph{Position}\\~ \\~ \\} &~& \makecell[tr]{~\\ $878,520$}\\
-
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure3_Pie_Chart.pdf}} & \makecell[tl]{Type 2: \emph{Pie Chart}\\~~~Perceptual Task: \emph{Angle}\\~ \\~ \\} &~& \makecell[tr]{~\\ $878,520$}\\
-
-	\bottomrule
-\end{tabular}
-}
-\label{tab:pos_angle_parameters}
-\end{table}
+%\resizebox{\linewidth}{!}{
+%\begin{tabular}{lllr}
+%	\toprule
+%	\multicolumn{2}{l}{~} & ~ & Permutations\\
+%
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure3_Bar_Chart.pdf}} & \makecell[tl]{Type 1: \emph{Bar Chart}\\~~~Perceptual Task: \emph{Position}\\~ \\~ \\} &~& \makecell[tr]{~\\ $878,520$}\\
+%
+%
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure3_Pie_Chart.pdf}} & \makecell[tl]{Type 2: \emph{Pie Chart}\\~~~Perceptual Task: \emph{Angle}\\~ \\~ \\} &~& \makecell[tr]{~\\ $878,520$}\\
+%
+%
+%	\bottomrule
+%\end{tabular}
+%}
+%\label{tab:pos_angle_parameters}
+%\end{table}
 
@@
 
-We proposed four hypotheses entering the elementary perceptual task experiment:
+We proposed two hypotheses entering the elementary perceptual task experiment:
 
@@
 This renders bar charts superior to pie charts and should also be the case for convolutional neural networks.
-	\item \textbf{H2.2} \textbf{Classifiers can learn position faster than angles.} We assume that understanding bar charts is easier than understanding pie charts.
-We suspect that our classifiers learn encodings of positions faster than of angles resulting in more efficient training and faster convergence.
+	\item \textbf{H2.2} \textbf{Convolutional neural networks can learn position faster than angles.} We assume that regressing bar charts is easier than understanding pie charts.
+Following our ranking of elementary perceptual tasks (Table~\ref{tab:ranking}), we suspect that our networks learn encodings of positions faster than angles resulting in more efficient training and faster convergence.
 \end{itemize}
@@
 
-Bar charts are more accurate (Fig.~\ref{fig:figure3_mlae}) and networks converge faster (Fig.~\ref{fig:figure3_val_loss}). This is great.
+\begin{figure}[t]
+	\centering
+	  \includegraphics[width=\linewidth]{figure3_mlae_better_all.pdf}
+  \caption{\textbf{Computational results of the position-angle experiment.} \textit{Left:} Our encodings of one data point as a pie chart and a bar chart.
+\textit{Right:} MLAE and 95\% confidence intervals for different networks.
+VGG19 * and Xception * are using ImageNet weights while all other networks were trained on the stimuli.
+We mimmick the original experiment of Cleveland and McGill and compare against their human results~\cite{cleveland_mcgill}.}
+	\label{fig:figure3_mlae}
+\end{figure}
+
+\noindent{\textbf{Perceptual Performance.}} Our networks are able to perform the regression task for bar charts and for pie charts (Fig.~\ref{fig:figure3_mlae}).
+We evaluate over 56 runs for each condition \textit{visual encoding} (12 runs per network, but only 4 for VGG19 and Xception due to higher training times), which yields an average $MLAE=2.176$ for bar chart ($SD=0.456$), and $3.296$ ($SD=0.77$) for pie chart.
+This difference is statistically significant ($F_{1,110}=86.061, p<0.01$) and leads us to \textbf{accept H2.1}.
+Post hoc comparisons show that this holds for most networks:
+MLP for pie charts $4.09$ ($SD=0.027$) and for bar charts $2.494$ ($0.068$) is significant ($t_{22}=72.300,p<0.01$), 
+LeNet for pie charts $ 3.556 $ ($SD= 0.022 $) and for bar charts $ 1.902 $ ($SD= 0.08 $) is significant $t_{22}=66.111, p<0.01$, 
+VGG19 *  for pie charts $ 3.561 $ ($SD= 0.047 $) and for bar charts $ 2.601 $ ($SD= 0.113 $) is significant $t_{22}=25.919,p<0.01$, 
+Xception * for pie charts  $ 3.094 $ ($SD= 0.046 $) and for bar charts $ 2.315 $ ($SD= 0.032 $) is significant $t_{22}=46.329,p<0.01$, 
+and Xception for pie charts $ 1.939 $ ($SD= 0.1 $) and for bar charts $ 1.375 $ ($SD= 0.062 $) is significant $t_{22}=8.276,p<0.01$.
+The difference for VGG19 (pie charts $ 1.297 $ ($SD= 0.129 $), bar charts $ 1.153 $ ($SD= 0.09 $)) was not significant with $p<0.05$.
+This is not surprising since VGG19 is a very powerful network which can adapt to seemingly any visual encoding as seen in our ranking for elementary perceptual tasks (Table~\ref{tab:ranking}).
+\\~\\
+\noindent{\textbf{Training Efficiency.} We measure the MSE loss for all networks on previously unseen validation data during training.
+We count a network as converged when this validation loss does not decrease after 10 sequential epochs meaning that each network and even each run stops after a varying number of training epochs.
+To measure the training efficiency, we look at the MSE validation loss during the first twenty epochs of 56 runs for each condition.
+Visually inspected, the pie chart loss decreases slower (Fig.~\ref{fig:figure3_val_loss}).
+The average loss in this period is for pie charts $0.052$ ($SD=0.015$) and for bar charts $0.037$ ($SD=0.018$).
+This difference is statistically significant ($F_{1,2238}=20.656, p<0.01$).
+We therefor conclude that networks train more efficiently and faster when learning bar charts and \textbf{accept H2.2}.
+\\~\\
+It seems that the visual encoding of bar charts is superior to pie charts in terms of performance and efficiency.
+This is interesting since Cleveland and McGill observe the same effect during their human experiment and conclude that the perceptual task of estimating position is easier for humans than the estimation of angles.
+Our ranking of elementary perceptual tasks yields a low score for angles and the related encoding of directions while position ranks in the mid to top.
 
@@
 	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
-  \caption{\textbf{Classifier Efficiency of the Position-Angle experiment.} Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts.
-We report the MSE measure for both encodings of four different classifier on previously unseen validation data.}
+  \caption{\textbf{Training efficiency of the position-angle experiment.} Mean Squared Error (MSE) loss during training of our networks computed on previously unseen validation data after each epoch.
+The regressors estimate quantities in pie charts and bar charts.
+We train all networks 12 times (4 times for VGG19 and Xception due to longer training times).
+VGG19 * and Xception * use ImageNet weights.
+All networks converge faster when learning bar charts.}
 	\label{fig:figure3_val_loss}
 \end{figure}
-
-\begin{figure}[t]
-	  \includegraphics[width=.8\linewidth]{figure3_mlae_better.pdf}
-  \caption{\textbf{Computational results of the Position-Angle experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure3_mlae}
-\end{figure}

### 2018-03-31 PAPER/05_position_length_experiment.tex
Some updates.

@@
 
-Finally, we also train `multi' networks which include all five types which creates an even bigger challenge.
+Finally, we also train `multi' networks which include all five types resulting in an even bigger challenge.
 
@@
 \begin{itemize}
-	\item \textbf{H3.1} \textbf{Our networks can estimate grouped bar charts better than divided bar charts.}
-
-	Grouped bar charts are better computational perceivable than divided bar charts.
-A grouped bar chart involves judging a position while a divided bar chart most likely (if not type 2) requires length judgements.
-
-	\item \textbf{H3.2} \textbf{Training `multi' networks } Any ideas?
+  \item \textbf{H3.1} \textbf{Our networks can estimate all types equally well.} A grouped bar chart involves judging a position while a divided bar chart most likely (if not type 2) requires length judgements.
+Our rankings of elementary perceptual tasks do not yield a strong preference for either across all networks.
+  \item \textbf{H3.2} \textbf{Trained `multi' networks work as well as individual trained networks.} Convolutional neural networks have massive numbers of trainable parameters.
+Their complexity allows them to learn different types of visual encodings in one training session.
 \end{itemize}
@@
 %\end{table}
-
-\subsection{Discussion}
-
-JT: Look at the relative difficulty of the tasks.
+%
+%\subsection{Discussion}
+%
+%JT: Look at the relative difficulty of the tasks.
 In Cleveland and McGill, types 1-5 were post-ordered by their log error such that type 1 was easiest and type 5 was hardest.
@@
 
-\begin{figure}[!ht]
+\begin{figure}[!t]
   \centering
@@
 \end{figure}
+
+\noindent\textbf{Perceptual Performance.} In the original position-length experiment, Cleveland and McGill report that types 1-5 were post-ordered by their perceptual difficulty with type 1 being the easiest and type 5 the hardest.
+We report the average MLAE for each type across our networks (and total 56 runs per type) as follows:
+for type 1 $MLAE= 3.956 $ ($SD= 0.274 $),
+for type 2 $MLAE= 3.952 $ ($SD= 0.441 $),
+for type 3  $MLAE= 4.349 $ ($SD= 0.367 $),
+for type 4 $MLAE= 3.668 $ ($SD= 0.256 $), and 
+for type 5 $MLAE= 3.902 $ ($SD= 0.253 $).
+These distributions yield significance ($F_{4,25}=2.815, p<0.05$) but post hoc comparions show that really only type 3 and type 4 differ ($t_{10}=3.406, p<0.01$).
+This means that the networks do not prefer a certain type on average which leads us to \textbf{partially accept H2.1} and we do not replicate the same pattern as Cleveland and McGill in their human studies even though our networks' performance is clearly worse than their human baseline.
+\\~\\
+\noindent\textbf{`Multi' Network Performance.} In Cleveland and McGill's original position-length experiment, humans were asked to judge visualizations of types 1-5.
+In the bottom of Figure~\ref{fig:figure4_mlae} we average the human performances of types 1-5 to create a `multi' human.
+Similarly, when training our networks, we first limit each training to one stimuli.
+However, a `multi' network which learns types 1-5 simultaenously is closer to the human in the original user study.

### 2018-03-31 PAPER/05_position_length_experiment.tex
Update supplemental.

@@
 These distributions yield significance ($F_{4,25}=2.815, p<0.05$) but post hoc comparions show that really only type 3 and type 4 differ ($t_{10}=3.406, p<0.01$).
-This means that the networks do not prefer a certain type on average which leads us to \textbf{partially accept H2.1} and we do not replicate the same pattern as Cleveland and McGill in their human studies even though our networks' performance is clearly worse than their human baseline.
+This means that the networks do not prefer a certain type on average which leads us to \textbf{partially accept H3.1} and we do not replicate the same pattern as Cleveland and McGill in their human studies even though our networks' performance is clearly worse than their human baseline.
 \\~\\
@@
 However, a `multi' network which learns types 1-5 simultaenously is closer to the human in the original user study.
+WHAT SHALL WE TEST HERE?

### 2018-03-31 PAPER/06_bars_and_framed_rectangles_experiment.tex
Minor fixes
Update supplemental.

@@
-\section{Bars and Framed Rectangles Experiment}
+\section{Experiment: Bars and Framed Rectangles}
 
@@
 %\end{figure}
-\begin{table}[h]
-\centering
-\caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales.
+%\begin{table}[h]
+%\centering
+%\caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales.
 For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles.
@@
 We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lllr}
-	\toprule
-	\multicolumn{2}{l}{~} & ~ & Permutations\\
-
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure12_0.pdf}} & \makecell[tl]{\emph{Bars}\\~~~Perceptual Task: \emph{Length}\\~~~JND: $X\%$\\~ \\~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $57,600$}\\
-
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure12_1.pdf}} & \makecell[tl]{\emph{Framed Rectangles}\\~~~Perceptual Task: \emph{Position}\\~~~JND: $X\%$\\~ \\~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $57,600$}\\
-
-	\bottomrule
-\end{tabular}
-}
-\label{tab:bars_and_framed_rectangles_parameters}
-\end{table}
+%\resizebox{\linewidth}{!}{
+%\begin{tabular}{lllr}
+%	\toprule
+%	\multicolumn{2}{l}{~} & ~ & Permutations\\
+%
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure12_0.pdf}} & \makecell[tl]{\emph{Bars}\\~~~Perceptual Task: \emph{Length}\\~~~JND: $X\%$\\~ \\~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $57,600$}\\
+%
+%
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure12_1.pdf}} & \makecell[tl]{\emph{Framed Rectangles}\\~~~Perceptual Task: \emph{Position}\\~~~JND: $X\%$\\~ \\~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $57,600$}\\
+%
+%
+%	\bottomrule
+%\end{tabular}
+%}
+%\label{tab:bars_and_framed_rectangles_parameters}
+%\end{table}
 
@@
 %\end{figure}
-\begin{table}[h]
-\centering
-\caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution.
+%\begin{table}[h]
+%\centering
+%\caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution.
 We create three different types of images initialized with 10, 100, and 1000 dots.
@@
 We evaluate our classifiers on these rasterized visualizations.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lllr}
-	\toprule
-	\multicolumn{2}{l}{~} & ~ & Permutations\\
-
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_0.pdf}} & \makecell[tl]{\emph{Base 10}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\
-
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_1.pdf}} & \makecell[tl]{\emph{Base 100}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\
-
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_2.pdf}} & \makecell[tl]{\emph{Base 1000}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\
-
-	\bottomrule
-\end{tabular}
-}
-\label{tab:weber_law}
-\end{table}
+%\resizebox{\linewidth}{!}{
+%\begin{tabular}{lllr}
+%	\toprule
+%	\multicolumn{2}{l}{~} & ~ & Permutations\\
+%
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_0.pdf}} & \makecell[tl]{\emph{Base 10}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\
+%
+%
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_1.pdf}} & \makecell[tl]{\emph{Base 100}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\
+%	
+%	\midrule
+%	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_2.pdf}} & \makecell[tl]{\emph{Base 1000}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\
+%
+%	\bottomrule
+%\end{tabular}
+%}
+%\label{tab:weber_law}
+%\end{table}
 
@@
 
-\begin{figure*}[t]
-	  \includegraphics[width=\linewidth]{figure12_mlae.pdf}
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{figure12_mlae_better_all.pdf}
   \caption{\textbf{Computational results of the Bars-and-Framed-Rectangles experiment.} Log absolute error means and 95\% confidence intervals for the \emph{bars-and-framed-rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
@@
 	\label{fig:figure12_mlae}
-\end{figure*}
+\end{figure}

### 2018-03-31 PAPER/SUPPLEMENTAL/supplemental.tex
Update supplemental.

@@
 %% insert punctuation at end of each item
-\item Daniel Haehn, and Hanspeter Pfister are with the Paulson School of Engineering and Applied Sciences at Harvard University. \\
+\item Daniel Haehn, and Hanspeter Pfister are with Harvard University. \\
 E-mail: \{haehn,pfister\}@seas.harvard.edu.
@@
 \end{figure}
+\vfill\null
+\section{Experiment: Noise}
+For all our experiments, we add subtle $5\%$ noise to every pixel to enhance variability.
+We did not observe a significant effect on regression performance when comparing the weber-fechner's law experiment with and without noise averaged over 4 runs (Figure~\ref{fig:noise}).
+However, the variability of our stimuli increases.
+
+\begin{figure}[th]
+\centering
+\includegraphics[width=.8\linewidth]{../gfx/weber_mlae_no_noise.pdf}
+\includegraphics[width=.8\linewidth]{../gfx/weber_mlae_noise.pdf}
+\caption{\textbf{Noise.} We test whether noise (top: off, bottom: on) effects the performance of our networks on the weber-fechner's law experiment by measuring MLAE.
+With noise, $MLAE=4.511$ ($SD=0.512$) and without noise $MLAE=4.491$ ($SD=0.543$).
+The difference is not statistically significant ($F(1,22)=0.008,p>0.5$).}
+\label{fig:noise}
+\end{figure}
 
@@
 
+\begin{figure}[p]
+	\centering
+	  \includegraphics[width=\linewidth]{../gfx/figure4_training_loss_with_multi.pdf}
+	  \includegraphics[width=\linewidth]{../gfx/figure4_val_loss_with_multi.pdf}
+  \caption{\textbf{Loss plots for the position-length experiment.} We visualize the MSE loss on training data and for unseen validation data after each epoch.
+There is no significant difference in convergence for either encoding but spiky outliers due to monte-carlo cross validation are visible.}
+	\label{fig:fig4_loss}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+	  \includegraphics[width=\linewidth]{../gfx/figure12_training_loss.pdf}
+	  \includegraphics[width=\linewidth]{../gfx/figure12_val_loss.pdf}
+  \caption{\textbf{Loss plots for the bars-and-framed-rectangles experiment.} We visualize the MSE loss on training data and for unseen validation data after each epoch.
+There is no significant difference in convergence for either encoding.}
+	\label{fig:fig12_loss}
+\end{figure}
+
+\begin{figure*}[t]
+	\centering
+
+    \subfloat[without noise]{
+	  \includegraphics[width=.48\linewidth]{../gfx/weber_training_loss_no_noise.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/weber_validation_loss_no_noise.pdf}
+	}
+	\hfill
+    \subfloat[with noise]{
+	  \includegraphics[width=.48\linewidth]{../gfx/weber_training_loss.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/weber_validation_loss.pdf}
+	}
+
+  \caption{\textbf{Loss plots for the weber-fechner's law experiment.} We visualize the MSE loss on training data (left) and for unseen validation data (right) after each epoch (a) without noise and (b) with subtle $5\%$ noise per pixel.
+There is no significant difference when noise is added.
+The LeNet network seems to overfit with Weber Base 100 in both cases even with dropout regularization.}
+	\label{fig:weber_loss}
+\end{figure*}
+
 %% if specified like this the section will be committed in review mode

### 2018-03-31 PAPER/01_related_work.tex
Squash, conform.
Add maneesh paper.

@@
 While we focus on Cleveland and McGill's work from the mid 1980s due to their repeated reproduction, many other works also investigate human perception to visual encoding~\cite{mackinlay1988applying,bertin1967semiologie,treisman1988feature,carpendale2003considering,wilkinson2006grammar,widgor_perception2007,munzner2015visualization}.
-
+%
 %Interesting are also the rankings of correlation visualization using Weber's law~\cite{harrison2014_webers_law_rank}.
@@
 In this paper, we investigate with a simple experiment whether this holds for convolutional neural networks
-
+%
 %\textbf{Comparing Visual Encodings.} Different visual encodings have advantages or disadvantages and the community does a great job comparing them.
@@
 While we investigate lower-level visual encodings in this work, we delay colorspace experiments for future work.
-
+\\~\\
 \noindent{\textbf{Computational Visualization Understanding.}} Pineo \textit{et al.}~\cite{Pineo2012_computational_perception} create a computational model of human vision based on neural networks.
 Their simulations show that understanding visualization triggers neural activity in high-level areas of cognition, with the authors suspecting that this activity is supported by low-level neurons performing elementary perceptional tasks.
-Other work tries to parse infographics by finding higher-level saliency models~\cite{bylinskii2016should}, or by extracting text or key visual elements from visualizations~\cite{diagram_understanding,kembhavi2016diagram,zoya_text_visual_tags}.
+Other work tries to parse infographics by finding higher-level saliency models~\cite{bylinskii2016should}, or by extracting text or key visual elements from visualizations using computer vision techniques~\cite{diagram_understanding,kembhavi2016diagram,zoya_text_visual_tags,maneesh_deconstructing_d3}.
 However, these works do not investigate computational understanding of elementary perceptual tasks such as curvature, lengths, or position, which are the building blocks of visualization.
-
+%
 %CNNs trained on natural images have low-level `neurons' which activate on similar stimuli to early neurons in the human visual cortex, e.g., Gabor filter-like stimuli \misscite.
-
+\\~\\
 \noindent{\textbf{Visual-cortex-inspired Machine Learning.}} The human visual cortex allows us to recognize objects in the world seemingly without effort (though few remember their infancy).

### 2018-03-31 PAPER/06_bars_and_framed_rectangles_experiment.tex

@@
 \end{figure}
+
+\begin{figure}[t]
+	  \includegraphics[width=\linewidth]{weber_mlae_noise_all.pdf}
+  \caption{\textbf{Computational results of the Weber-Fechner's Law experiment.} Log absolute error means and 95\% confidence intervals for the \emph{bars-and-framed-rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
+We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
+	\label{fig:weber_law}
+\end{figure}

### 2018-03-31 PAPER/SUPPLEMENTAL/supplemental.tex

@@
 
+\begin{figure}[p]
+	\centering
+	  \includegraphics[width=\linewidth]{../gfx/weber_training_loss_all.pdf}
+	  \includegraphics[width=\linewidth]{../gfx/weber_validation_loss_all.pdf}
+  \caption{\textbf{Loss plots for the weber-fechner's law experiment including VGG19 and Xception.} We visualize the MSE loss on training data and for unseen validation data after each epoch.
+This plot includes the VGG19 and Xception networks trained from scratch.}
+	\label{fig:weber_loss_all}
+\end{figure}
+
 %% if specified like this the section will be committed in review mode

### 2018-03-31 PAPER/05_position_length_experiment.tex
Update section 5.

@@
 However, a `multi' network which learns types 1-5 simultaenously is closer to the human in the original user study.
-WHAT SHALL WE TEST HERE?
+For our `multi' experiment, we record an average error across all classifiers of $MLAE=4.358$ ($SD=0.327$).
+We then compare against the average errors for all types, as reported above for perceptual performance.
+We reach significant differences ($F_{5,30}=3.454,p<0.05$).
+Post hoc comparisons yield significant differences between the `multi' experiment and type 4 ($t_{10}=3.716, p<0.01$) and also to type 5 ($t_{10}=2.467, p<0.05$).
+Since our the average MLAE is worse than all of type 1--5, and the distributions observe significant differences, we acknowledge that the `multi' task is harder for the networks than learning single types of encodings.
+This might relate to us fixing all other parameters of the experiment such as the number of early stopping epochs and the size of training data, however, we \textbf{reject H3.2}.

### 2018-03-31 PAPER/07_results.tex
Some updates (not done yet).

@@
 
-General discussion..
-
-\subsection{Classifiers}
-
+\textbf{Graphical Perception by CNNs.} In all experiments, CNNs were able to regress visual encodings to their quantitative variables reasonably with error rates comparable to humans.
+This suggests that future work can enable full-blown understanding of different chart types, e.g.
+a classifier which can identify one type of bar charts.
+\\~\\
+\textbf{Understanding Infographics by CNNs.} While elementary perceptual tasks can be learned by CNNs, it seems to be a very challenging task to have CNNs `understand' information visualizations which come in all variations.
+A simple \textit{google search} for barchart yields an incredible amount of variations.
+\\~\\
+\textbf{Stimuli Variability.} All our generated stimuli exhibit a certain variability which ranges from a very low number of permutations of 20, for the most simple volume elementary perceptual task, to millions, for the position-length experiment.
+We additionally add random noise to each stimuli to ensure that the CNNs do not just memorize images.
+We do not observe a direct correlation between variability and `perceptability' by our networks which suggests that the networks do not just memorize images.
+We also perform a direct noise and no-noise comparison (see supplemental) without any significant effect.
+\\~\\
+\textbf{Concept Learning.} Our cross-network experiments show that simple variations throw off the networks and result in high error rates.
+This suggests that the networks are not actually learning concepts but rather learn slight variations of pixel values.
+While we try to counteract this with our variability settings, it seems that this is the way the networks work.
+\\~\\
 \textbf{Transfer Learning using ImageNet.} Classifiers trained on imagenet are tuned towards natural images.
@@
 \\~\\
-\textbf{Anti-aliasing.} Does it help? Not sure yet!
-
-%
-%\subsection{Elementary Perceptual Tasks}
-%
-%
-%\subsection{Position-Angle Experiment}
-%
-%
-%\subsection{Position-Length Experiment}
-%
-%\subsection{Bars and Framed Rectangles Experiment}
-%
-
-%\begin{figure}[t]
-%	  \includegraphics[width=\linewidth]{figure12_val_loss.png}
-%  \caption{\textbf{Classifier Efficiency of the Bars and Framed Rectangles experiment.} Categorical Cross-Entropy loss for the \emph{bars and framed rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-The frame around the bars adds an additional visual cue enables faster network convergence.
-This is not yet reproducible!}
-%	\label{fig:figure12_val_loss}
-%\end{figure}
+\textbf{Anti-aliasing.} To keep things simple, we choose to create rasterized visualizations without interpolation.
+However, there might be a bias from networks trained on natural images (such as VGG19 * and Xception *, with ImageNet weights) which prefer smoother and not so prominent edges.
+We compare anti-aliased stimuli against the original ones without any significant effect (see supplemental).

### 2018-03-31 PAPER/04_position_angle_experiment.tex
Section 4!

--- 2018-03-30 PAPER/04_position_angle_experiment.tex
@@
-\section{Position-Angle Experiment}
+\clearpage
+\section{Experiment: Position-Angle}
 
-The position-angle experiment was originally performed by Cleveland and McGill to measure whether humans can better perceive quantities encoded as positions or as angles~\cite{cleveland_mcgill}.
-The actual experiment then compares pie charts versus bar charts since these map down to elementary position and angle judgement.
-We create rasterized images mimicking Cleveland and McGill's proposed encoding and investigate computational perception of our four classifiers.
+Cleveland and McGill measure human perception to the ratios of positions and angles through comparisons on bar charts and pie charts~\cite{cleveland_mcgill}.
+We create rasterized images following Cleveland and McGill's proposed encoding and investigate computational perception of our networks (Figure~\ref{fig:teaser}).
+These have five bar or pie sectors representing numbers which add to 100, where each is greater than three and smaller than 39.
+One required change is in the minimal differences between the values: Cleveland and McGill create stimuli where the differences between each number are greater than $0.1$.
+However, as our networks only take $100\times100$ pixel images as input, we can only minimally represent a difference of $1$ pixel.
+
+Cleveland and McGill ask participants to estimate the ratio of the four smaller bars or sectors to the known and marked largest bar or sector.
+As such, we mark the largest quantity of the five in each visualization with a single pixel dot, then ask our networks to perform multiple regression and produce the four ratio estimates.
+Since the position of the largest element changes, we generate the targets such that the largest element is marked with 1 and the smaller elements follow counter-clockwise for the pie chart and to the right for the bar chart.
+Each of the bar and pie chart visualizations has $878,520$ possible permutations.
+
+% To be successful, the networks essentially first have to find the marked quantity, have the `rolling' encoding figured out, and then estimate the quantities properly.
+-> JT: I don't want to prescribe how the network must solve the task.
+There are other ways to solve it, and this might not be how the network solves the problem.
 
@@
 
-We proposed four hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-	\item \textbf{H2.1} \textbf{Computed perceptual performance is better using bar charts than pie charts.} Cleveland and McGill report that position judgements are almost twice as accurate as angle judgements.
-This renders bar charts superior to pie charts and should also be the case for convolutional neural networks.
-	\item \textbf{H2.2} \textbf{Classifiers can learn position faster than angles.} We assume that understanding bar charts is easier than understanding pie charts.
-We suspect that our classifiers learn encodings of positions faster than of angles resulting in more efficient training and faster convergence.
-\end{itemize}
+\begin{hypolist}
+	\item \textbf{H2.1} \textbf{Computed perceptual accuracy will be higher for bar charts than pie charts.} Cleveland and McGill report that position judgements are almost twice as accurate (MLAE) as angle judgements in humans.
+Following our ranking of elementary perceptual tasks (Table~\ref{tab:ranking}), we see that our networks also judge position encodings more accurately than angles, and so our networks will be able to more easily judge bar charts than pie charts.
+	\item \textbf{H2.2} \textbf{Convolutional neural networks will learn to regress bar chart ratios faster than pie chart ratios in training.} This follows directly from H2.1.
+\end{hypolist}
 
@@
 
-Bar charts are more accurate (Fig.~\ref{fig:figure3_mlae}) and networks converge faster (Fig.~\ref{fig:figure3_val_loss}). This is great.
+\begin{figure}[t]
+	\centering
+	  \includegraphics[width=\linewidth]{figure3_mlae_better_all.pdf}
+  \caption{\textbf{Computational results of the position-angle experiment.} \textit{Left:} Our encodings of one data point as a pie chart and a bar chart.
+\textit{Right:} MLAE and 95\% confidence intervals for different networks.
+VGG19 * and Xception * are using ImageNet weights while all other networks were trained on the stimuli.
+We train all networks 12 times (4 times for VGG19 and Xception due to significantly longer training times).
+VGG19 * and Xception * use ImageNet weights.
+Our results align with Cleveland and McGill's human results, shown in black~\cite{cleveland_mcgill}.}
+	\label{fig:figure3_mlae}
+\end{figure}
+
+% We evaluate over 56 runs for each condition \textit{visual encoding} (12 runs per network, but only 4 for VGG19 and Xception due to higher training times),
+
+\noindent{\textbf{Perceptual Accuracy.}} Our networks are able to regress the task ratios for bar charts and pie charts (Fig.~\ref{fig:figure3_mlae}).
+Cross-validation yields an average $MLAE=2.176$ ($SD=0.456$) for bar charts, and an average $MLAE=3.296$ ($SD=0.77$) for pie charts.
+This difference is statistically significant ($F_{1,110}=86.061, p<0.01$), and so we \textbf{accept H2.1}.
+
+Post hoc comparisons show that this holds for most networks: 
+MLP for pie charts $4.09$ ($SD=0.027$) and for bar charts $2.494$ ($0.068$) is significant ($t_{22}=72.300,p<0.01$);
+LeNet for pie charts $ 3.556 $ ($SD= 0.022 $) and for bar charts $ 1.902 $ ($SD= 0.08 $) is significant $t_{22}=66.111, p<0.01$;
+VGG19 * with ImageNet weights for pie charts $ 3.561 $ ($SD= 0.047 $) and for bar charts $ 2.601 $ ($SD= 0.113 $) is significant $t_{22}=25.919,p<0.01$;
+Xception * with ImageNet weights for pie charts  $ 3.094 $ ($SD= 0.046 $) and for bar charts $ 2.315 $ ($SD= 0.032 $) is significant $t_{22}=46.329,p<0.01$;
+Xception from scratch for pie charts $ 1.939 $ ($SD= 0.1 $) and for bar charts $ 1.375 $ ($SD= 0.062 $) is significant $t_{22}=8.276,p<0.01$; but
+the difference for VGG19 from scratch (pie charts $ 1.297 $ ($SD= 0.129 $), bar charts $ 1.153 $ ($SD= 0.09 $)) was not significant with $p<0.05$. 
+This outcome is in line with the elementary perceptual task results (Table~\ref{tab:ranking}), where VGG19 was the most successful network, where networks trained from scratch were more performant, and where angle was more difficult to learn than position.
+\\~\\
+\noindent{\textbf{Training Efficiency.} We measure the MSE loss for all networks on previously-unseen validation data during training.
+We count a network as converged when this validation loss does not decrease after 10 sequential epochs.
+Figure~\ref{fig:figure3_val_loss} shows this MSE validation loss during the first twenty epochs for each condition, plotted across all cross-validation splits with overdrawn lines.
+The pie chart loss decreases more slowly, with the average loss over epochs being $0.052$ ($SD=0.015$) for pie charts and $0.037$ ($SD=0.018$) for bar charts.
+This difference is statistically significant ($F_{1,2238}=20.656, p<0.01$).
+Thus, we \textbf{accept H2.2}.
+\\~\\
+To all our networks, the bar chart is a superior visual encoding than a pie chart, in terms of accuracy and efficiency.
+Cleveland and McGill observe the same effect for accuracy during their human experiments.
+
+% and conclude that the perceptual task of estimating position is easier for humans than the estimation of angles -> JT: Unless you have a direct quote, we're not putting words in their mouths.
+I couldn't find one in my brief look, but you might have one.
+If we do, then put a page number reference.
 
@@
 	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
-  \caption{\textbf{Classifier Efficiency of the Position-Angle experiment.} Mean Square Error (MSE) loss for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill} which compares the visualization of pie charts and bar charts.
-We report the MSE measure for both encodings of four different classifier on previously unseen validation data.}
+  \caption{\textbf{Training efficiency of the position-angle experiment.} Mean Squared Error (MSE) loss after each epoch during training, computed on previously-unseen validation data.
+We train all networks 12 times (4 times for VGG19 and Xception due to significantly longer training times).
+VGG19 * and Xception * use ImageNet weights.
+All networks converge faster when learning bar charts.}
 	\label{fig:figure3_val_loss}
 \end{figure}
-
-\begin{figure}[t]
-	  \includegraphics[width=.8\linewidth]{figure3_mlae_better.pdf}
-  \caption{\textbf{Computational results of the Position-Angle experiment.} Log absolute error means and 95\% confidence intervals for the \emph{position-angle experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure3_mlae}
-\end{figure}

### 2018-03-31 PAPER/00_introduction.tex
Updates from ShareLaTeX

--- 2018-03-30 PAPER/00_introduction.tex
@@
 
-Convolutional neural networks (CNNs) have been successfully applied to a wide range of visual tasks, most famously to natural image object recognition~\cite{krizhevsky_imagenet2012, simonyan_very_deep2014, szegedy2015}, for which some claim equivalent or better than human performance.
-This performance comparison is often motivated by the idea that CNNs model or reproduce the early layers of the visual cortex, even though they do not incorporate many details of biological neural networks or model higher-level abstract or symbolic reasoning~\cite{yamins2016using, hassabis2017neuroscience, human_vs_machine_vision}.
-While CNN techniques were originally inspired by neuroscientific discoveries, recent advances in processing larger datasets with deeper networks have been the direct results of engineering efforts.
-Throughout this significant advancement, researchers have aimed to understand why and how CNNs produce such high performance~\cite{goodfellow_book, deeplearning_blackbox2017}, with recent works targeting the systematic evaluation of the visual perception limits of CNNs~\cite{clevr, not_so_clevr}.
+Artificial intelligence has taken the world of technology by storm.
+Deep multilayer neural networks are being successfully applied in a wide range of applications that are regularly outperforming humans in object recognition \cite{krizhevsky_imagenet2012, simonyan_very_deep2014,szegedy2015}.
+Originally inspired by neuroscientific discoveries, the recent advances in deep learning have been the direct results of engineering efforts, more specifically in convolutional neural networks (CNNs).
+While there has been significant advancement, this does not mean we understand what CNNs are doing and we are in fact treating them as blackboxes without detracting from their success \cite{goodfellow_book, deeplearning_blackbox2017}.
+Our current knowledge of biological vision suggests that modern machine learning models indeed mimic the underlying biology by abstracting the many details of biological neural networks \cite{yamins2016using, hassabis2017neuroscience}.
 
-One fundamental application of human vision is to understand data visualizations.
-This is a task unlike processing natural images, where real-world objects and their effects are abstracted into data and represented with visual marks.
-As a field, visualization catalogues and evaluates human perception of these marks, such as in the seminal \emph{graphical perception} experiments of Cleveland and McGill~\cite{cleveland_mcgill}.
-This work describes nine elementary perceptual reasoning tasks, such as position relative to a scale, length, angle, area, and shading density, plus orders their reasoning difficulty.
-But, with increasing research interest in the machine analysis of graphs, charts, and visual encodings, it seems pertinent to question whether CNNs are able to process these basic graphical elements and derive useful measurements from the building blocks of information visualization.
+Despite tremendous research efforts and generating massive datasets, we are far from fully understanding biological vision. 
+Similarly, we are constantly developing inventive features for deep artificial networks without truly understanding it in its entirety. As a result, a double discrepancy is observed. 
+Advances in neuroscience can revolutionize machine learning by reverse-engineering neural circuits, yielding new classifiers which, in turn, can help process the massive biological data. 
+There are many existing questions which must be answered in order to fill this gap of our understanding.
 
-As such, we reproduce Cleveland and McGill's human perceptual experiments with CNNs, and discuss to what extent they have `graphical perception'.
-To perform this evaluation, we parametrize the elementary perceptual tasks and experiments suggested by Cleveland and McGill~\cite{cleveland_mcgill}, and define a set of regression tasks to estimate continuous variables.
-Against human perception, we pit four neural networks: a three-layer multilayer perceptron (MLP), the LeNet 2-layer CNN~\cite{lenet}, the VGG 19-layer CNN~\cite{simonyan_very_deep2014}, and the Xception 36-layer CNN~\cite{xception}.
-As CNNs trained on natural images are said to mimic layers of the human visual cortex, we investigate whether using weights trained on natural images (via ImageNet~\cite{imagenet}) or weights trained from scratch on elementary graphical perception tasks produces more accurate measurements and greater generalization.
+\paragraph{Proposed Solution}
 
-We test these four networks across four scenarios presented by Cleveland and McGill~\cite{cleveland_mcgill}: 1) Nine elementary perceptual tasks with increasing parameteric complexity, e.g., length estimation with fixed x, then with varying x, then with varying width, including cross-network evaluations testing the generalizability of networks to unseen parameters; 2) The position-angle experiment, which compares judgements of bar charts to pie charts, 3) The position-length experiment, which compares grouped and divided bar charts, and 4) The bars and framed rectangles experiments, where visual cues aid ratio perception.
-We also investigate whether our CNNs can detect a proportional change in a measurement across scales, in relation to Weber's law.
+We focus on perception... experiments from cleveland mcgill..
 
-With these experiments, we describe a ranking defining the ease with which our tested CNN architectures can estimate elementary perceptual tasks, as an equivalent to Cleveland and McGill's ranking for human perception.
-Further, we discuss the implications of our results and derive recommendations for the use of CNNs in perceiving visualizations.
-We accompany this paper with open source code and our input and results data, both to enable reproduction studies and to spur new machine perception systems more adept at graphical perception: \url{http://rhoana.org/perception}
+\paragraph{Validation and Applications}
 
-%Cleveland and McGill's work in the 1980s has led to many insights for modern information visualization research such as the identification of elementary perceptional tasks or that bar charts are easier to understand than pie charts.
+\paragraph{Contributions}
 
@@
 %While the reasons for this are still unknown, specified research can most certainly advance the understanding of deep neural networks.
-%
-%\subsection{From Biological Vision to Machine Learning}
-%
-%%\begin{figure}[t]
-%%	  \includegraphics[width=\linewidth]{biology_vs_cnn.png}
-%%  \caption{The Biological Vision (schematic)}
-%%	\label{fig:vision}
-%%\end{figure}
-%
-%Biological vision is an extremely powerful system which allows humans the ability, and seemingly without effort, to recognize an enormous amount of distinct objects in the world. 
-%Object detection is extremely difficult and therefore is especially impressive as light intensities can change by levels of magnitude and contrast between foreground and background is so often low. 
-%In addition, the visual scene changes every time the human body or human eyes move. 
-%This visual system exhibits a very noisy structure but because it is organized by layers it has inspired the mathematical theory of multilayer neural networks. 
-%What is remarkable is that even though current machine learning models do not resemble the complexity of its biological pendant, they inherently generalize extremely well. 
-%Neural networks trained on one specific task can be used to perform detection or segmentation of, seemingly, unrelated objects with relatively minor retraining. 
-%The reported classification performance is superior to that of humans and the question in regards to their functionality opens an interesting research topic.
-%
-%In 1962 Hubel and Wiesel were the first to begin studying the visual cortex from the standpoint of a neuroscientist.
+
+\subsection{Biological Vision}
+
+%\begin{figure}[t]
+%	  \includegraphics[width=\linewidth]{biology_vs_cnn.png}
+%  \caption{The Biological Vision (schematic)}
+%	\label{fig:vision}
+%\end{figure}
+
+Biological vision is an extremely powerful system which allows humans the ability, and seemingly without effort, to recognize an enormous amount of distinct objects in the world. 
+Object detection is extremely difficult and therefore is especially impressive as light intensities can change by levels of magnitude and contrast between foreground and background is so often low. 
+In addition, the visual scene changes every time the human body or human eyes move. 
+This visual system exhibits a very noisy structure but because it is organized by layers it has inspired the mathematical theory of multilayer neural networks. 
+What is remarkable is that even though current machine learning models do not resemble the complexity of its biological pendant, they inherently generalize extremely well. 
+Neural networks trained on one specific task can be used to perform detection or segmentation of, seemingly, unrelated objects with relatively minor retraining. 
+The reported classification performance is superior to that of humans and the question in regards to their functionality opens an interesting research topic.
+
+In 1962 Hubel and Wiesel were the first to begin studying the visual cortex from the standpoint of a neuroscientist.
 Their experimental findings on cats and macaque monkeys suggested a hierarchy of cells with increasing complexity which was then later transferred to the hierarchical model of different layers.
@@
 In the last decade, this field has exhibited rapid growth, constant evolution, and new applications in various domains.
-%
-%We think that the biological inspiration of modern convolutional neural networks yields the evaluation of principles of human perception with computers.
-%
+
+GOALS
+\begin{itemize}
+	\item reduce the gap between neurobiology and data science to advance the understanding of visual cortex inspired machine learning
+	\item insights for infovis for machines
+\end{itemize}
+
+CONTRIBUTIONS
+\begin{itemize}
+	\item experiments of cleveland mcgill with systematic parametrization and evaluation for computational perception
+	\item ranking of elementary perceptual tasks for machine perception
+	\item validation of weber's law for feedforward neural networks
+	\item many insights for infovis for machines
+\end{itemize}

### 2018-03-31 PAPER/02_experimental_setup.tex
Updates from ShareLaTeX

--- 2018-03-30 PAPER/02_experimental_setup.tex
@@
 
-We conduct quantitative experiments to measure how different convolutional neural networks perceive low-level visual encodings, such as positions, angles, curvatures, and lengths.
-We formulate these measurement tasks as logistic regression problems: given a stimuli image of an elementary visualization, the networks must estimate the single quantity present or the ratio between multiple quantities present.
+The experiments shown in this paper are either supervised regression or classification tasks.
+We formulate any estimation of quantities (e.g.
+angles, positions, lengths etc.) as a regression problem between $0$ and $1$.
+The output indicates the percentage in regards to the degrees of freedom of the individual experiment.
+If the experiment involves a choice, we formulate it as a classification problem.
 
-For each experiment, we use a single factor between-subject design, with the factor being the network used.
-This lets us evaluate whether different network designs are competitive against existing human perception results.
-We train each network in a supervised fashion with a mean-squared error (MSE) loss between the ground-truth labels and the network's estimate of the measurement from observing the generated stimuli images.
-Then, we test each network's ability to generalize to new examples with a separate data, created using the same stimuli generator function but with unseen ground-truth measurements (Section~\ref{sec:data}).
+\subsection{Measures}
 
-% We define a series of hypotheses prior to each experiment.
+\textbf{Accuracy.} We use the same metric as Cleveland and McGill to measure accuracy.
+\begin{equation}
+	\log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
+\end{equation}
+\\~\\
+\noindent{\textbf{Confidence Intervals.}} We follow the notion of Cleveland and McGill to compute the confidence intervals.
+\\~\\
+\noindent{\textbf{Efficiency.}} We use the convergence rate based on the decrease of loss per training epoch as an indicator for the efficiency of the classifier in combination with a visual encoding.
+For regression tasks the loss is defined as mean squared error (MSE) and for classification tasks the loss is categorical cross-entropy.
 
-\subsection{Networks}
-\label{sec:networks}
+\subsection{Classifiers}
 
-\noindent{\textbf{Multilayer Perceptron.}} As a baseline, we use a multilayer perceptron (MLP), but without the prior convolutional layers as is typical in network designs for solving visual tasks (Fig.~\ref{fig:classifiers}).
-Our MLP contains a layer of $256$ perceptrons, which are activated as rectified linear units (ReLU)~(Fig.~\ref{fig:classifiers}).
-We train this layer with dropout (probability $= 0.5$) to prevent overfitting, and then combine these ReLU units to regress our output measurement.
+Our classifiers are built upon a multilayer perceptron (MLP) which is a feedforward artificial neural network.
+We combine this MLP with different convolutional neural networks (CNNs) for preprocessing and feature generation.
+These include the traditional LeNet trained from scratch, as well as VGG19 and Xception trained using ImageNet.
 \\~\\
-\noindent{\textbf{Convolutional Neural Networks.}} We compare different convolutional neural networks (CNNs) with both `trained from scratch' weights and pre-trained weights on a database of natural images (1000-class ImageNet \cite{imagenet}).
-These networks are the traditional LeNet-5 with 2 layers, which was designed to recognize hand-written digits~\cite{lenet}; the VGG19 network with 19 layers, which was designed to solve the ImageNet object recognition challenge~\cite{simonyan_very_deep2014}; and the Xception network with 36 layers~\cite{xception}, which was also designed to solve the ImageNet object recognition challenge plus the 15,000-class JFT object recognition challenge~\cite{Hinton2015}.
-Each of these networks has as its last layers an MLP architecture equivalent to our baseline, and so they act as earlier image and feature processors for this final regressor.
-Since the networks are of different architectures, the number of trainable parameters changes, with some networks having more capacity than others (Table~\ref{tab:parameters}).
-
-For \emph{VGG19} and \emph{Xception}, we have two variants: the network trained from scratch on elementary perceptual tasks, plus the network using weights that were previously trained on the ImageNet object recognition challenge \emph{except} for the MLP layer.
-This is intended to produce early-layer features which mimic human vision, and then to see whether they are more or less useful than networks trained from scratch.
+\noindent{\textbf{Multilayer Perceptron.}} The multilayer perceptron in this paper has $256$ neurons which are activated as rectified linear units~(Fig.~\ref{fig:classifiers}).
+We then add a dropout layer to prevent overfitting and compute linear regression or classification (softmax).
 \\~\\
-\noindent{\textbf{Optimization.}} All network hyperparameters, optimization methods, and stopping conditions are fixed across networks (Table \ref{tab:parameters}).
-We train for $1000$ epochs using stochastic gradient descent with Nesterov momentum, but stop early if the loss does not decrease for ten epochs.
+\noindent{\textbf{Convolutional Neural Networks.}} We use CNNs to generate additional features as input to the MLP.
+We train the \emph{LeNet} classifier with tune it specifically towards each visualization.
+For \emph{VGG19} and \emph{Xception}, we generate features using previously trained weights on ImageNet.
 \\~\\
-\noindent{\textbf{Environment.}} We run all experiments on Tesla X and Tesla V100 graphical processing units.
-We use the KERAS framework with a TensorFlow backend to train the networks, and use the scikit-image library to generate the stimuli.
-% and scikit-learn libraries - JT: What do we use this for?
+\noindent{\textbf{Optimization.}} All networks are optimized using stochastic gradient descent with Nesterov momentum using fixed parameters (Table \ref{tab:parameters}).
+We train for $1000$ epochs but stop early if the loss does not decrease for ten epochs.
+\\~\\
+\noindent{\textbf{Environment.}} We run all experiments on an NVIDIA DGX1 machine with Tesla V100 graphical processing units. We use the KERAS framework with tensorflow.
 
@@
 \centering
-\caption{\textbf{Network Training.} We use different feature generators as input to a multilayer perceptron which performs linear regression.
-This results in different sets of trainable parameters.
-As a baseline, we also train the MLP directly on the visualization images without any additional feature generation.}
+\caption{We use different feature generators as input to a multilayer perceptron which performs linear regression or the classification task.
+This yields different sets of trainable parameters.
+We also train the MLP directly on the visualizations without any additional feature generation.}
 \resizebox{\linewidth}{!}{
@@
 	\toprule
-	Network & \makecell{Trainable\\Parameters} & Optimization \\
+	Classifier & \makecell{Trainable\\Parameters} & Optimization \\
 	\midrule
@@
 \end{table}
-%\begin{figure}[t]
-%	\centering
-%	  \includegraphics[width=\linewidth]{classifiers.pdf}
-%  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
+
+\begin{figure}[t]
+	\centering
+	  \includegraphics[width=\linewidth]{classifiers.pdf}
+  \caption{The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
 We use Dropout regularization to prevent overfitting.
@@
 layers, filter size $3\times3$) to increase the number of trainable parameters.}
-%	\label{fig:classifiers}
-%\end{figure}
-\begin{figure}[t]
-	\centering
-
-    \subfloat[Feature Generation]{
-		\includegraphics[width=4.8cm,valign=c]{classifier_left.pdf}
-		\vphantom{\includegraphics[width=3.5cm,valign=c]{classifier_right.pdf}}
-	}
-	\hfill
-    \subfloat[Multilayer Perceptron]{
-		\includegraphics[width=3.5cm,valign=c]{classifier_right.pdf}
-	}
-
-  \caption{\textbf{Network Architecture.} The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
-We use Dropout regularization to prevent overfitting.
-As output, we perform linear regression for continuous variables.
-The MLP can learn to represent the visualizations directly, but we also learn features generated by LeNet (2 conv.
-layers, filter size $5\times5$), VGG19 (16 conv.
-layers, filter size $3\times3$), or Xception (36 conv.
-layers, filter size $3\times3$) to test different model complexities.}
 	\label{fig:classifiers}
@@
 \subsection{Data}
-\label{sec:data}
 
-\noindent\textbf{Image Stimuli and Labels.} 
-We create our stimuli visualizations as 100$\times$100 binary images, rasterized without interpolation.
-We write a parameterized stimuli generator for each elementary task.
-The number of possible parameter values differs per experiment, and we summarize these in Table~\ref{tab:encoding_parameters} and Section~\ref{sec:parametrizations}.
-Before use, we scale the generated images into an unbiased range: images to the range of $-0.5$ to $0.5$.
-Then, we add subtle random noise (uniformly distributed between $0-0.05$) to each pixel to introduce variation which prevents the networks from simply `remembering' each different image.
-
-Each stimuli image also has an associated ground truth label representing the parameter set which generated the image, e.g., the length in pixels of a bar.
-As before, we scale these labels to the range of $0.0$ to $1.0$, which represent the maximum and minimum values that this parameter can take.
+We create all visualizations as parametrized rasterized images without interpolation.
+The number of parameters differs per experiment as summarized in Table~\ref{tab:encoding_parameters} and section~\ref{sec:parametrizations}.
+We add subtle random noise ($0.05$) to each pixel to introduce additional variation.
 \\~\\
-\noindent\textbf{Training/Validation/Test Splits.} For each task, we use 60,000 training images, 20,000 validation images, and 20,000 test images.
-To create these datasets, we generate stimuli from random parameters and add them to the sets until the target number is reached, while maintaining distinct (random) parameter spaces for each set to ensure that there is no leakage between training and validation/testing.
-
-\subsection{Measures and Analysis}
-
-\noindent\textbf{Cross Validation.} For reproducibility, we perform repeated random sub-sampling validation, also known as Monte Carlo cross-validation, during our experiments.
-We run every experiment seperately twelve times, and randomly select (without replacement) the $60\%$ of our data as training data, $20\%$ as validation, and $20\%$ as test.
-%Our large data sample size of 100,000 guarantees that every single observation of our parameterizations will be selected at least once (excluding noise patterns). -> JT: No it doesn't.
-% Finally, we average the results over the runs. -> JT: Fine, but we'll do distributions anyway, so not worth saying.
+\noindent\textbf{Training/Validation/Test Splits.} We specify the size of each split set as follows: 60,000 training images, 20,000 validation images, and 20,000 test images.
+We then randomly add parameterized visualizations to the sets while guaranteeing that each set is disjunct from each other in terms of encoded variables.
+This eliminates leakage during training and evaluation.
+We also scale each set independently: images to the range of $-.5$ to $.5$ and labels to the range of $0.0$ to $1.0$.
 \\~\\
-\noindent{\textbf{Task Accuracy.}} In their 1984 paper, Cleveland and McGill use the midmean logistic absolute error metric (\emph{MLAE}) to measure perception accuracy.
-To allow comparison between their human results and our machine results, we also use MLAE as a presentation metric:
-\begin{equation}
-	\textnormal{MLAE} = log_2( | \textnormal{predicted percent} - \textnormal{true percent} | + .125)
-\end{equation}
-In addition to this metric, we also calculate standard error metrics such as the mean squared error (\emph{MSE}) and the mean absolute error (\emph{MAE}).
-This allows a more direct comparison of percent errors.
-Please note that our networks were trained using MSE loss and not directly with MLAE.
-\\~\\
-\noindent{\textbf{Task Confidence Intervals.}} We follow Cleveland and McGill and present $95\%$ confidence intervals.
-We approximate the value of the $97.5$ percentile point of the normal distribution for simplicity with $1.96$ as suggested by the central limit theorem~\cite{central_limit}.
-\\~\\
-\noindent{\textbf{Confirmatory Data Analysis.}} To accept or reject our hypotheses, we analyze dependent variables using analysis of variance (ANOVA) followed by parametric tests. JT: Which tests?
-\\~\\
-\noindent{\textbf{Training Efficiency.}} We use the training convergence rate as a measure of how easy or hard a particular task is for the network to learn to solve.
-This is defined as the MSE loss decrease per training epoch, which is an indicator of the training efficiency of the network with respect to the visual encoding.
-\\~\\
-\noindent\textbf{Network Generalizability.} We evaluate generalizability by asking a network previously trained upon one task parameterization to answer questions about the same type of task stimuli but with more complex parameterization, e.g., estimating bar length without and with changes in stroke width.
-
-Further, some experiments compare different visual encoding types, e.g., bar plot vs.~stacked bar plot.
-We train and evaluate individual networks for each parameterization, plus we also train and evaluate a networks on stimuli across the different types.
-This single decision-making software better mimics the judgements that a human would be able to make.
-%This scenario affects the optimization process and result in networks with more flexible knowledge with the caveat of longer training times. -> JT: Save it for the results discussion.
+\noindent\textbf{Cross-classifier variability.} We also evaluate classifiers previously trained with one visualization on the same type of visualizations with different parameters by decreasing and increasing the variability of the generated images.

### 2018-06-14 PAPER/03_elementary_perceptual_tasks.tex

@@
 
-\noindent \textbf{H1.1} \textbf{The CNNs tested will be able to regress quantitative variables from graphical elements.} We parametrize different visual encodings (Table~\ref{tab:encoding_parameters}) and test whether the CNNs can measure them, and relate the results to accuracies obtained by humans on similar tasks.
+\begin{hypolist}
+\item \textbf{H1.1} \textbf{The CNNs tested will be able to regress quantitative variables from graphical elements.} We parametrize different visual encodings (Table~\ref{tab:encoding_parameters}) and test whether the CNNs can measure them.
+%, and relate the results to accuracies obtained by humans on similar tasks.
 
-\noindent \textbf{H1.2} \textbf{CNN perceptual performance will depend on network architexture.} We evaluate multiple regressors with different numbers of trainable parameters.
+\item \textbf{H1.2} \textbf{CNN perceptual performance will depend on network architexture.} We evaluate multiple regressors with different numbers of trainable parameters.
 We expect a more complex network (with more trainable parameters) to perform better on elementary perceptual tasks.
 
-\noindent \textbf{H1.3} \textbf{Some visual encodings will be easier to learn than others for the CNNs tested.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
+\item \textbf{H1.3} \textbf{Some visual encodings will be easier to learn than others for the CNNs tested.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
 We investigate whether this order is also relevant for computing graphical perception.
 
-\noindent \textbf{H1.4} \textbf{Networks trained on perceptual tasks will generalize to more complex variations of the same task.} Empirical evidence suggests that CNNs are able to generalize by interpolating between different training data points, and so perform on variations of a similar perceptual task.
+\item \textbf{H1.4} \textbf{Networks trained on perceptual tasks will generalize to more complex variations of the same task.} Empirical evidence suggests that CNNs are able to generalize by interpolating between different training data points, and so perform on variations of a similar perceptual task.
 We create visual representations of the elementary perceptual tasks with different variability, and expect that networks will be able to generalize when presented with slight task variations.
+\end{hypolist}
 

### 2018-06-14 PAPER/06_bars_and_framed_rectangles_experiment.tex

--- 2018-03-31 PAPER/06_bars_and_framed_rectangles_experiment.tex
@@
+\begin{figure}[!t]
+	  \includegraphics[width=\linewidth]{figure12_mlae_better_all.pdf}
+  \caption{\textbf{Computational results of the bars-and-framed-rectangles experiment.} \textit{Left:} Visual encodings of two bars for a task of length judgement (bottom) following Cleveland and McGill's proposed experiment.
+Perceiving which bar is longer is much easier for humans when a frame is added (top).
+For networks (trained from scratch, or * indicates ImageNet weights), there seems no significant difference between the encodings as reported by MLAE and 95\% confidence intervals.}
+	\label{fig:figure12_mlae}
+\end{figure}
+
 \section{Experiment: Bars and Framed Rectangles}
@@
 Visual cues can help converting graphical elements back to their real world variables.
-Cleveland and McGill introduced the bars and framed rectangles experiment which judges the elementary perceptual task of position along non-aligned scales~\cite{cleveland_mcgill}.
+Cleveland and McGill introduced the bars-and-framed-rectangles experiment which compares the perceptual judgement tasks  of length and position along non-aligned scales~\cite{cleveland_mcgill}.
+Figure~\ref{fig:figure12_mlae} shows both variations on the left.
+It is not easy to judge which bar is larger in the bottom picture which involves a length judgement.
+However, when adding a little frame as a reference, this length judgement is transferred to a position judgement along non-aligned scales.
+Based on this little frame, it is easy to see that the right bar is slightly larger than the left since the whitespace in the top of the frame is smaller than the one on the left.
+
+As Cleveland and McGill explain in their theories, judging the whitespace is actually also a length judgement rather than a position~\cite{cleveland_mcgill}.
+They now relate this task to Weber's Law which states that the perceivable difference within a distribution is proportional to the initial size of the distribution~\cite{householder1940weber}.
+Here, it means that humans can easily measure the difference in the white scale since its initial size is small while estimating the small change in lengths of the black bars is not easily perceivable.
+The Just Noticable Difference (JND) is higher when the initial stimuli is smaller in size (here the white bars).
+
+We mimmick the bars-and-framed rectangles experiment as a two value regression task.
+We create rasterized visualizations of size $100\times100$ as shown in Figure~\ref{fig:figure12_mlae} and let our networks estimate the sizes of the stimuli.
 
@@
 
-\subsection{Weber-Fechner's Law}
-
-As identified by Cleveland and McGill, the bar and framed rectangle experiment is closely related to Weber's law.
-This psychophysics law states that perceivable difference within a distribution is proportional to the initial size of the distribution.
-Weber's law goes hand-in-hand with Fechner's law.
-We conduct an additional experiment based on the original illustrations of the Weber-Fechner law to investigate whether this law can be applied to computational perception of our classifiers (Fig.~\ref{tab:weber_law}).
-
 %\begin{figure}[t]
@@
 
+\subsection{Point Cloud Experiment}
+
+We conduct an additional experiment testing whether Weber's law applies to convolutional neural networks for graphical perception.
+For this, we generate three 2D point clouds as base stimuli -- each is created randomly by activating 10, 100, or 1000 pixels in a $100\times100$ raster image.
+We then additionally activate from 1 to 10 random pixels within the initial distribution but by carefully only selecting inactive pixels.
+We show examples for this in Figure~\ref{fig:weber_law} (left).
+This means that the number of additional points is harder to identify if they are added to the 1000 pixel set while the 10 pixel set allows to easily count.
+We then let our networks solve a regression task to estimate the number of added points.
+
 \subsection{Results}
 
-First run indicates that framed rectangles perform better but we dont really know it yet.
-
 \begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_mlae_better_all.pdf}
-  \caption{\textbf{Computational results of the Bars-and-Framed-Rectangles experiment.} Log absolute error means and 95\% confidence intervals for the \emph{bars-and-framed-rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
+	  \includegraphics[width=\linewidth]{weber_mlae_noise_all.pdf}
+  \caption{\textbf{Computational results of the point cloud experiment.} Log absolute error means and 95\% confidence intervals for the \emph{bars-and-framed-rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
 We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure12_mlae}
+	\label{fig:weber_law}
 \end{figure}

### 2018-06-14 PAPER/SUPPLEMENTAL/supplemental.tex

@@
 
+\begin{figure*}[t]
+	\centering
+
+    \subfloat[VGG19 *, Block 3 (Conv. Layers 2+3)]{
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/vgg_imagenet_block3_conv2.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/vgg_imagenet_block3_conv3.pdf}
+	}
+	\hfill
+    \subfloat[VGG19 *, Block 4 (Conv. Layers 2+3)]{
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/vgg_imagenet_block4_conv2.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/vgg_imagenet_block4_conv3.pdf}
+	}
+	\hfill
+    \subfloat[VGG19 *, Block 5 (Conv. Layers 2+3)]{
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/vgg_imagenet_block5_conv2.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/vgg_imagenet_block5_conv3.pdf}
+	}
+	\hfill
+    \subfloat[VGG19, Block 3 (Conv. Layers 2+3)]{
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/vgg_from_scratch_block3_conv2.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/vgg_from_scratch_block3_conv3.pdf}
+	}
+	\hfill
+    \subfloat[VGG19, Block 4 (Conv. Layers 2+3)]{
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/vgg_from_scratch_block4_conv2.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/vgg_from_scratch_block4_conv3.pdf}
+	}
+	\hfill
+    \subfloat[VGG19, Block 5 (Conv. Layers 2+3)]{
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/vgg_from_scratch_block5_conv2.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/vgg_from_scratch_block5_conv3.pdf}
+	}
+  \caption{\textbf{Convolutional Activation Maps for a Pie Chart.} (a)-(c) is VGG19 *, trained on ImageNet.
+The activation maps do not differ much which is surprising since VGG19 trained from scratch performs so much better in our experiments.}
+	\label{fig:activationmaps}
+\end{figure*}
+
+\begin{figure*}[t]
+	\centering
+
+    \subfloat[VGG19 *, Block 3 (Conv. Layers 2+3)]{
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/curvature_vgg_imagenet_block3_conv2.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/curvature_vgg_imagenet_block3_conv3.pdf}
+	}
+	\hfill
+    \subfloat[VGG19 *, Block 4 (Conv. Layers 2+3)]{
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/curvature_vgg_imagenet_block4_conv2.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/curvature_vgg_imagenet_block4_conv3.pdf}
+	}
+	\hfill
+    \subfloat[VGG19 *, Block 5 (Conv. Layers 2+3)]{
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/curvature_vgg_imagenet_block5_conv2.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/curvature_vgg_imagenet_block5_conv3.pdf}
+	}
+	\hfill
+    \subfloat[VGG19, Block 3 (Conv. Layers 2+3)]{
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/curvature_vgg_from_scratch_block3_conv2.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/curvature_vgg_from_scratch_block3_conv3.pdf}
+	}
+	\hfill
+    \subfloat[VGG19, Block 4 (Conv. Layers 2+3)]{
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/curvature_vgg_from_scratch_block4_conv2.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/curvature_vgg_from_scratch_block4_conv3.pdf}
+	}
+	\hfill
+    \subfloat[VGG19, Block 5 (Conv. Layers 2+3)]{
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/curvature_vgg_from_scratch_block5_conv2.pdf}
+	  \includegraphics[width=.48\linewidth]{../gfx/ACTIVATIONMAPS/curvature_vgg_from_scratch_block5_conv3.pdf}
+	}
+  \caption{\textbf{Convolutional Activation Maps for a Curvature stimuli.} (a)-(c) is VGG19 *, trained on ImageNet.
+The activation maps do not differ much which is surprising since VGG19 trained from scratch performs so much better in our experiments.}
+	\label{fig:activationmaps2}
+\end{figure*}
+
 %% if specified like this the section will be committed in review mode

### 2018-08-15 PAPER/00_introduction.tex
Update paper from overleaf.

--- 2018-03-30 PAPER/00_introduction.tex
@@
 \firstsection{Introduction}
+\maketitle
+\vspace{-0.1cm}
 
-\maketitle
-
-Convolutional neural networks (CNNs) have been successfully applied to a wide range of visual tasks, most famously to natural image object recognition~\cite{krizhevsky_imagenet2012, simonyan_very_deep2014, szegedy2015}, for which some claim equivalent or better than human performance.
-This performance comparison is often motivated by the idea that CNNs model or reproduce the early layers of the visual cortex, even though they do not incorporate many details of biological neural networks or model higher-level abstract or symbolic reasoning~\cite{yamins2016using, hassabis2017neuroscience, human_vs_machine_vision}.
-While CNN techniques were originally inspired by neuroscientific discoveries, recent advances in processing larger datasets with deeper networks have been the direct results of engineering efforts.
-Throughout this significant advancement, researchers have aimed to understand why and how CNNs produce such high performance~\cite{goodfellow_book, deeplearning_blackbox2017}, with recent works targeting the systematic evaluation of the visual perception limits of CNNs~\cite{clevr, not_so_clevr}.
-
-One fundamental application of human vision is to understand data visualizations.
-This is a task unlike processing natural images, where real-world objects and their effects are abstracted into data and represented with visual marks.
+%One fundamental application of human vision is to understand data visualizations.
+This is a task unlike natural image processing but which includes the abstraction of real-world objects and their effects into data, represented with visual marks.
 As a field, visualization catalogues and evaluates human perception of these marks, such as in the seminal \emph{graphical perception} experiments of Cleveland and McGill~\cite{cleveland_mcgill}.
 This work describes nine elementary perceptual reasoning tasks, such as position relative to a scale, length, angle, area, and shading density, plus orders their reasoning difficulty.
-But, with increasing research interest in the machine analysis of graphs, charts, and visual encodings, it seems pertinent to question whether CNNs are able to process these basic graphical elements and derive useful measurements from the building blocks of information visualization.
 
-As such, we reproduce Cleveland and McGill's human perceptual experiments with CNNs, and discuss to what extent they have `graphical perception'.
-To perform this evaluation, we parametrize the elementary perceptual tasks and experiments suggested by Cleveland and McGill~\cite{cleveland_mcgill}, and define a set of regression tasks to estimate continuous variables.
-Against human perception, we pit four neural networks: a three-layer multilayer perceptron (MLP), the LeNet 2-layer CNN~\cite{lenet}, the VGG 19-layer CNN~\cite{simonyan_very_deep2014}, and the Xception 36-layer CNN~\cite{xception}.
-As CNNs trained on natural images are said to mimic layers of the human visual cortex, we investigate whether using weights trained on natural images (via ImageNet~\cite{imagenet}) or weights trained from scratch on elementary graphical perception tasks produces more accurate measurements and greater generalization.
+Convolutional neural networks (CNNs) have been successfully applied to a wide range of visual tasks, most famously to natural image object recognition~\cite{simonyan_very_deep2014, szegedy2015}, for which some claim equivalent or better than human performance.
+This performance comparison is often motivated by the idea that CNNs model or reproduce the early layers of the human visual cortex, even though they do not incorporate many details of biological neural networks or model higher-level abstract or symbolic reasoning~\cite{yamins2016using, hassabis2017neuroscience, human_vs_machine_vision}.
+While CNN techniques were originally inspired by neuroscientific discoveries, recent advances in processing larger datasets with deeper networks have been the direct results of engineering efforts.
+Throughout this significant advancement, researchers have aimed to understand why and how CNNs produce such high performance~\cite{deeplearning_blackbox2017}, \change{with recent works targeting the systematic evaluation of the limits of feed-forward convolutional neural networks for both image recognition problems \cite{Azulay2018} and for visual relation problems~\cite{clevr, not_so_clevr}.}
 
-We test these four networks across four scenarios presented by Cleveland and McGill~\cite{cleveland_mcgill}: 1) Nine elementary perceptual tasks with increasing parameteric complexity, e.g., length estimation with fixed x, then with varying x, then with varying width, including cross-network evaluations testing the generalizability of networks to unseen parameters; 2) The position-angle experiment, which compares judgements of bar charts to pie charts, 3) The position-length experiment, which compares grouped and divided bar charts, and 4) The bars and framed rectangles experiments, where visual cues aid ratio perception.
-We also investigate whether our CNNs can detect a proportional change in a measurement across scales, in relation to Weber's law.
+In visualization, there is increasing research interest in the computational analysis of graphs, charts, and visual encodings~\cite{maneesh_deconstructing_d3,Pineo2012_computational_perception,kafle2018dvqa}, for applications like information extraction and classification, visual question answering (``computer, which category is greater?''), or even design analysis and generation~\cite{Viegas2016}.
+One might turn to a CNN for these tasks.
+However, computational analysis of visualizations is a more complex task than natural image classification~\cite{Kahou2018}, requiring the identification, estimation, and relation of visual marks to extract information.
+For instance, we take for granted the human ability to generalize an understanding of length to a previously unseen chart design, or to estimate the ratios between lengths, yet for a CNN these abilities are in question, with no clear mechanism for concept abstraction.
 
-With these experiments, we describe a ranking defining the ease with which our tested CNN architectures can estimate elementary perceptual tasks, as an equivalent to Cleveland and McGill's ranking for human perception.
-Further, we discuss the implications of our results and derive recommendations for the use of CNNs in perceiving visualizations.
-We accompany this paper with open source code and our input and results data, both to enable reproduction studies and to spur new machine perception systems more adept at graphical perception: \url{http://rhoana.org/perception}
+% this is a nicely readable motivation!
+Our goal is to better understand the abilities of CNNs for visualization analysis, and so we investigate the performance of current off-the-shelf CNNs on visualization tasks and show what they can and cannot accomplish.
+As computational visualization analysis is predicated upon an understanding of elementary perceptual tasks, we consider the seminal \emph{graphical perception} settings of Cleveland and McGill~\cite{cleveland_mcgill}.
+This work describes nine reasoning tasks, such as position relative to a scale, length, angle, area, and shading density, and measures human graphical perception performance on bar and pie chart quantity estimation.
+We reproduce Cleveland and McGill's settings with four different neural network designs of increasing sophistication (MLP, LeNet, VGG, and Xception), and compare their performance to human graphical perception.
+For this task, we collect new human measures for each elementary task, for the bars and frames rectangles setting, and for a Weber's law point cloud experiment.
+Further, as CNNs trained on natural images are said to mimic early human vision, we investigate whether using pre-trained natural image weights (via ImageNet~\cite{imagenet}) or weights trained from scratch on elementary graphical perception tasks produces more accurate predictions.
+
+%  and greater generalization
+
+%To perform this evaluation, we parametrize the elementary perceptual tasks and experiments suggested by Cleveland and McGill~\cite{cleveland_mcgill}, and define a set of regression tasks to estimate continuous variables.
+%We pit four neural networks against human perception: a three-layer multilayer perceptron (MLP), the LeNet 2-layer CNN~\cite{lenet}, the VGG 16-layer CNN~\cite{simonyan_very_deep2014}, and the Xception 36-layer CNN~\cite{xception}.
+
+%With these experiments, we describe a ranking defining the ease with which our tested CNN architectures can estimate elementary perceptual tasks, equivalent to Cleveland and McGill's ranking for human perception.
+
+%We test these four networks across four scenarios presented by Cleveland and McGill~\cite{cleveland_mcgill}: 1) Nine elementary perceptual tasks with increasing parameteric complexity, e.g., length estimation with fixed x, then with varying x and y, then with varying width, including cross-network evaluations testing the generalizability of networks to unseen parameters; 2) The position-angle experiment, which compares judgments of bar charts to pie charts; 3) The position-length experiment, which compares grouped and divided bar charts; and 4) The bars and framed rectangles experiments, where visual cues aid ratio perception.
+We also investigate whether our CNNs can detect a proportional change in a measurement across scales in relation to Weber's law.
+
+First, we find that CNNs can more accurately predict quantities than humans for nine elementary perceptual tasks, but only if their training data includes similar stimuli.
+Second, that our networks can estimate bar chart lengths and pie segment angles with accuracy similar to humans, and that our networks trained more easily on bar charts.
+Third, that our networks were largely unable to estimate length ratios across five bar chart designs, unlike humans, and that bar chart design type had largely no effect on CNN performance but a significant effect on human performance.
+Fourth, that framing bars make it no easier for our CNNs to estimate just noticeable differences in length, unlike for humans.
+Fifth, that some CNNs can solve a difficult Weber's law problem that is beyond human ability.
+Practically, we find that networks trained from scratch on visualizations perform better than using pre-trained natural images weights, and that current Xception networks perform worse than older VGG networks.
+
+A second goal of our work is to help frame the visualization community's discussion of CNNs as it builds towards future computational visualization analysis applications.
+For this, our findings suggest that CNNs are not currently a good model for human graphical perception, and as such their application to specific visualization problems may be possible but needs care.
+We discuss this in more detail with respect to designing CNNs for analyzing visualizations.
+Further, toward this goal, we accompany this paper with open source code and data, both to enable reproduction studies and to spur new machine perception systems more adept at graphical perception: \url{http://vcglab.org/perception}
 

### 2018-08-15 PAPER/01_related_work.tex
Update paper from overleaf.

--- 2018-03-30 PAPER/01_related_work.tex
@@
-\section{Previous Work}
+\section{Related Work}
 
-\noindent{\textbf{Graphical Perception.}} Cleveland and McGill~\cite{cleveland_mcgill,cleveland1985graphical} coin the phrase \emph{graphical perception} to describe how different visual attributes and encodings are perceived by humans.
-They define \emph{elementary perceptual tasks} as mental-visual stimuli to understand encodings in visualizations, and declare a ranking based on their perceptual difficulty.
-From these definitions, the authors propose and perform different experiments such as the \emph{position-angle} experiment which compares bar charts and pie charts, the \emph{position-length} experiment where users judge relations between encoded values in grouped and divided bar charts, and the \emph{bars-and-framed-rectangles} experiment to evaluate Weber's law \cite{harrison2014_webers_law_rank} using the proportional relation between an initial distribution density and perceivable change.
+\noindent{\textbf{Graphical Perception.}} Cleveland and McGill~\cite{cleveland_mcgill,cleveland1985graphical} coined the phrase \emph{graphical perception} \change{to describe ``the visual decoding of information encoded on graphs''.} To understand encodings in visualizations, they define \emph{elementary perceptual tasks} as mental-visual stimuli, and rank their perceptual difficulty \change{to humans}.
+From these definitions, the authors perform the \emph{position-angle} experiment to compare bar charts and pie charts, and the \emph{position-length} experiment where users judge relations between encoded quantities in grouped and divided bar charts.
+The authors then use this to redesign a statistical map via \emph{bars and framed rectangles} and Weber's law \cite{harrison2014_webers_law_rank}, using the proportional relation between an initial distribution density and perceivable change.
 
 Heer and Bostock later reproduced the Cleveland-McGill experiments via crowd-sourcing on Mechanical Turk~\cite{HeerBostock2010}, with similar results.
-Harrison \textit{et al.}~\cite{harrison2013influencing} repeated the Cleveland-McGill experiments while observing viewer emotional states, again with similar results.
-Our experimental setup again reproduces the Cleveland-McGill experiments, but instead of judging human perception, we judge machine perception using convolutional neural networks.
-While we focus on Cleveland and McGill's work from the mid 1980s due to their repeated reproduction, many other works also investigate human perception to visual encoding~\cite{bertin1967semiologie,treisman1988feature,carpendale2003considering,wilkinson2006grammar,widgor_perception2007,munzner2015visualization}.
+Harrison et al.~repeated the Cleveland-McGill experiments while observing viewer emotional states, again with similar results~\cite{harrison2013influencing}.
+\change{Talbot et al.~delve deeper into the impact of specific bar chart design variations on human prediction \cite{Talbot2014}}.
+While we focus on Cleveland and McGill's work due to its repeated reproduction, many works investigate human perception to visual encoding~\cite{mackinlay1988applying,bertin1967semiologie,treisman1988feature,carpendale2003considering,wilkinson2006grammar,widgor_perception2007,munzner2015visualization}.
 
+\change{
+Cleveland and McGill's definition of graphical perception judiciously excludes a human subject and a visual decoding method, leaving the door open for machine perception such as with CNNs.
+Their quality standard is only that the machine must decode \emph{information}, which leaves narrow and well-defined applications like our experiments approachable.
+However, machine graphical perception must meet human breadth in capability to be generally useful in a human world.
+}
+%
 %Interesting are also the rankings of correlation visualization using Weber's law~\cite{harrison2014_webers_law_rank}.
@@
 Lower-level experiments target - besides others - open versus closed encodings~\cite{open_vs_closed_shapes}, and several evaluations of color space~\cite{ware1988color,rheingans1992color,Rogowitz2001_colormaps,kindlmann2002color}.
-While we investigate lower-level visual encodings in this work, we delay colorspace experiments for future work.
-
-\noindent{\textbf{Computational Visualization Understanding.}} Pineo \textit{et al.}~\cite{Pineo2012_computational_perception} create a computational model of human vision based on neural networks.
-Their simulations show that understanding visualization triggers neural activity in high-level areas of cognition, with the authors suspecting that this activity is supported by low-level neurons performing elementary perceptional tasks.
-Other work tries to parse infographics by finding higher-level saliency models~\cite{bylinskii2016should}, or by extracting text or key visual elements from visualizations~\cite{diagram_understanding,kembhavi2016diagram,zoya_text_visual_tags}.
-However, these works do not investigate computational understanding of elementary perceptual tasks such as curvature, lengths, or position, which are the building blocks of visualization.
-
+While we investigate lower-level visual encodings in this work, we delay colorspace experiments for future work
+%
+%
 %CNNs trained on natural images have low-level `neurons' which activate on similar stimuli to early neurons in the human visual cortex, e.g., Gabor filter-like stimuli \misscite.
-
-\noindent{\textbf{Visual-cortex-inspired Machine Learning.}} The human visual cortex allows us to recognize objects in the world seemingly without effort (though few remember their infancy).
-This visual system is organized in layers, which inspired computational classifiers based on multilayer neural networks.
-Fukushima and Miyake developed the early Neocognitron quantitative model~\cite{fukushima1982neocognitron}, which ultimately led to the work of Hinton, Bengio, and LeCun~\cite{lecun2015deep} and today's GPU-powered \emph{deep} neural networks.
-Such networks exist with many architectures.
-For this paper, we compare a set of networks with different architectures and depths, plus networks with weights trained on natural images and on the elementary perceptual reasoning tasks themselves.
+\\~\\
+\noindent{\textbf{Visual-cortex-inspired Machine Learning.}} The human visual cortex allows us to recognize objects in the world seemingly without effort.
+This visual system is organized into layers, which inspired computational systems based on multilayer neural networks.
+Fukushima and Miyake developed the early Neocognitron quantitative model~\cite{fukushima1982neocognitron}, which ultimately led to the work of Hinton, Bengio, and LeCun~\cite{lecun2015deep} and today's GPU-powered deep neural networks.
+Such networks have been developed with many architectures \change{attempting to model properties useful to visual reasoning, like translation invariance or part hierarchies}.
+\change{Across these, the error behaviors of CNNs indicate that they process images in a different way to humans even though they are loosely biologically inspired~\cite{SzegedyICLR2014,Engstrom2017,Azulay2018}.}
+%
+%For this paper, we compare a set of networks with different architectures and depths, plus networks with weights trained on natural images and on generated stimuli of the elementary perceptual reasoning tasks.
+%
 %Object detection is extremely difficult and therefore is especially impressive as light intensities can change by levels of magnitude and contrast between foreground and background is so often low. 
@@
 %This visual system exhibits a very noisy structure but because it is organized by layers it has inspired the mathematical theory of multilayer neural networks. 
-
+%
 %In 1962 Hubel and Wiesel were the first to begin studying the visual cortex from the standpoint of a neuroscientist.
@@
 In the last decade, this field has exhibited rapid growth, constant evolution, and new applications in various domains.
-
+%
 %The reported classification performance is superior to that of humans and the question in regards to their functionality opens an interesting research topic.
@@
 %- last year's VADL workshop: https://vadl2017.github.io/
+%
+%
+\\~\\
+\noindent{\textbf{Computational Visualization Analysis.}} Pineo et al.~create a computational model of early human vision based on neural networks, then optimize flow visualizations for comprehension~\cite{Pineo2012_computational_perception}.
+Their simulations show that visualization perception triggers neural activity in higher-level areas of cognition, which the authors suspect is supported by low-level neurons performing elementary perceptional tasks.
+Other work tries to parse infographics by finding higher-level saliency models~\cite{bylinskii2016should,bylinskii2017understanding,bylinskii2017predimportance}, \change{or tries to parse text and key visual elements from visualizations using both classic and deep-learning-based computer vision~\cite{Savva2011-revision,kembhavi2016diagram,zoya_text_visual_tags,maneesh_deconstructing_d3,Poco2017-reverse-engineering-vis} (to name but a few)}.
+
+%Poco et al.~recover visual encodings from chart images by \cite{Poco2017-reverse-engineering-vis}
+%
+%Savva et al.~similarly recover chart elements by \cite{Savva2011-revision}.
+
+%R. A. Al-Zaidy and C. L. Giles. Automatic extraction of datafrombarcharts. InProceedings of the 8th International Conference on Knowledge Capture, page 30. ACM, 2015. 2
+
+\change{
+The field of visual question answering (VQA) has begun to combine image and text feature vectors for visualizations.
+Kafle et al.~extend stacked attention networks with dynamic encodings to support different bar chart designs, plus present a benchmark dataset~\cite{kafle2018dvqa}.
+However, the ability of CNNs to solve visual relation problems like VQA has been called into question~\cite{not_so_clevr}.
+Kahou et al.~attempt to answer visual questions across five chart types.
+However, the authors state that the task poses ``a significant machine learning challenge'' with no tested system able to meet human-level performance~\cite{Kahou2018}.
+}
+
+\change{Our work takes a step back.
+None of these works investigate the building blocks of visualization: elementary perceptual tasks such as position, length, and angle estimation.
+To produce useful visualization analyzers, we must be able to computationally predict human responses in these tasks.
+We investigate whether this is possible with CNNs.}
+
+%Feed forward neural networks - visual relations strain

### 2018-08-15 PAPER/02_experimental_setup.tex
Update paper from overleaf.

--- 2018-03-30 PAPER/02_experimental_setup.tex
@@
 
-First, we describe the commonalities across all of our experiments.
-We measure how different convolutional neural networks perceive low-level visual encodings, such as positions, angles, curvatures, and lengths.
-We formulate these measurement tasks as logistic regression problems: given a stimuli image of an elementary visualization, the networks must estimate the single quantity present or the ratio between multiple quantities present.
+\change{
+We compare CNNs to human baselines across five experiments:
+\begin{enumerate}[label=E\arabic*.,itemsep=0.5pt, topsep=1pt, parsep=0.5pt]
+    \item We use Cleveland and McGill's elementary perceptual tasks to directly estimate quantities for visual marks (position, length, direction, angle, area, volume, curvature, and shading) (Section~\ref{sec:elementary}).
+    \item We reproduce Cleveland and McGill's position-angle experiment that compares pie charts to bar charts (Section~\ref{sec:positionangle}).
+    \item We reproduce Cleveland and McGill's position-length experiment that compares grouped and divided bar charts (Section~\ref{sec:positionlength}).
+    \item We assess the bars and framed rectangles setting from Cleveland and McGill, where visual cues aid perception (Section~\ref{sec:barsframedrectangles}). 
+    \item We conduct a Weber's law point cloud experiment (Section~\ref{sec:weberslaw}).
+\end{enumerate} 
+}
+
+First, we describe the commonalities across all our experiments.
+In each, we measure whether different CNNs can predict values from low-level visual marks.
+\change{We formulate these measurement tasks as logistic regression rather than classification problems, so that we can estimate continuous variables such as directions and angles.} Given a stimuli image, the networks must estimate the single quantity present or the ratio between multiple quantities present.
 
 For each experiment, we use a single factor between-subject design, with the factor being the network used.
-This lets us evaluate whether different network designs are competitive against existing human perception results.
-We train each network in a supervised fashion with a mean-squared error (MSE) loss between the ground-truth labels and the network's estimate of the measurement from observing the generated stimuli images.
-Then, we test each network's ability to generalize to new examples with a separate data, created using the same stimuli generator function but with unseen ground-truth measurements (Section~\ref{sec:data}).
+This lets us evaluate whether different network designs are competitive against human perception results.
+We train each network in a supervised fashion with a mean squared error (MSE) loss between the ground-truth labels and the network's estimate of the measurement from observing the generated stimuli images.
+Then, we test each network's ability to generalize to new examples with separate test data, created using the same stimuli generator function but with unseen ground-truth measurement labels.
+% (Section~\ref{sec:data}).
+
+%This means that not only the stimuli are distinct for train and test sets but also the associated labels (Section~\ref{sec:data}).
 
@@
 
-\noindent{\textbf{Multilayer Perceptron.}} As a baseline, we use a multilayer perceptron (MLP), but without the prior convolutional layers as is typical in network designs for solving visual tasks (Fig.~\ref{fig:classifiers}).
-Our MLP contains a layer of $256$ perceptrons, which are activated as rectified linear units (ReLU)~(Fig.~\ref{fig:classifiers}).
+\noindent{\textbf{Multilayer Perceptron.}} As a baseline, we use a multilayer perceptron (MLP).
+\change{The MLP does not have the convolutional layers which help CNNs solve visual tasks, and so we include it to tests whether a CNN is really needed to solve our simple graphical perception tasks} (Fig.~\ref{fig:classifiers}).
+Our MLP contains a layer of $256$ perceptrons that are activated as Rectified Linear Units (ReLU).
 We train this layer with dropout (probability $= 0.5$) to prevent overfitting, and then combine these ReLU units to regress our output measurement.
 \\~\\
-\noindent{\textbf{Convolutional Neural Networks.}} We compare different convolutional neural networks (CNNs) with both `trained from scratch' weights and pre-trained weights on a database of natural images (1000-class ImageNet \cite{imagenet}).
-These networks are the traditional LeNet-5 with 2 layers, which was designed to recognize hand-written digits~\cite{lenet}; the VGG19 network with 19 layers, which was designed to solve the ImageNet object recognition challenge~\cite{simonyan_very_deep2014}; and the Xception network with 36 layers~\cite{xception}, which was also designed to solve the ImageNet object recognition challenge plus the 15,000-class JFT object recognition challenge~\cite{Hinton2015}.
-Each of these networks has as its last layers an MLP architecture equivalent to our baseline, and so they act as earlier image and feature processors for this final regressor.
-Since the networks are of different architectures, the number of trainable parameters changes, with some networks having more capacity than others (Table~\ref{tab:parameters}).
+\noindent{\textbf{Convolutional Neural Networks.}} \change{We test three CNN architectures of increasing sophistication.
+Each has more layers than the last, which is an indicator for the network's capacity to hierarchically represent information.
+Each network also has more trainable parameters than the last, which is an indicator for how much information the network is able to learn overall.
+While larger networks need more data to train, we expect them to perform better than simpler networks.}
 
-For \emph{VGG19} and \emph{Xception}, we have two variants: the network trained from scratch on elementary perceptual tasks, plus the network using weights that were previously trained on the ImageNet object recognition challenge \emph{except} for the MLP layer.
-This is intended to produce early-layer features which mimic human vision, and then to see whether they are more or less useful than networks trained from scratch.
+\change{
+Our smallest CNN is the traditional LeNet-5 with 2 layers, which was designed to recognize hand-written digits~\cite{lenet}.
+Next, we use the VGG19 network with 16 layers, which achieved 90\% top 5 performance in the 1000-class ImageNet object recognition challenge in 2014~\cite{simonyan_very_deep2014}.
+Finally, we use the Xception network with 36 layers~\cite{xception}, which achieved 95\% top 5 performance on ImageNet in 2017 and was also designed to solve the 17,000-class JFT object recognition challenge~\cite{Hinton2015}.
+Xception includes state-of-the-art architecture elements: residual blocks to allow it to be deeper, and depth-wise separable convolutions (or Inception blocks) to separate spatial from cross-channel correlations for more efficient parameter use.
+All three networks have as their last layers an MLP architecture equivalent to our baseline, and so they act as earlier image and feature processors for this final regressor.
+Table~\ref{tab:parameters} lists the number of trainable parameters per network.}
+
+For \emph{VGG19} and \emph{Xception}, we train all network parameters on elementary perceptual tasks (\emph{from scratch}); and we use network parameters previously trained on the ImageNet object recognition challenge but retrain the parameters in the final MLP layers (\emph{fine tuning}).
+\change{We know that humans are able to perform graphical perception tasks, and so maybe these pre-trained parameters that mimic early-layer human vision features are useful for the task.
+However, parameters trained from scratch are unlikely to mimic human features, as the network has only seen visualization tasks and not natural images (i.e., growing up not in Flatland \cite{abbott1885flatland}, but in Visland).}
 \\~\\
-\noindent{\textbf{Optimization.}} All network hyperparameters, optimization methods, and stopping conditions are fixed across networks (Table \ref{tab:parameters}).
-We train for $1000$ epochs using stochastic gradient descent with Nesterov momentum, but stop early if the loss does not decrease for ten epochs.
+\noindent{\textbf{Optimization.}} All hyperparameters, optimization methods, and stopping conditions are fixed across networks (Table \ref{tab:parameters}).
+We train for $1000$ epochs using stochastic gradient descent with Nesterov momentum but stop early if the validation loss does not decrease for ten epochs.
+\change{Each epoch trains the network upon every stimuli, with model updates after every mini-batch of 32 stimuli.}
 \\~\\
-\noindent{\textbf{Environment.}} We run all experiments on Tesla X and Tesla V100 graphical processing units.
-We use the KERAS framework with a TensorFlow backend to train the networks, and use the scikit-image library to generate the stimuli.
+\noindent{\textbf{Environment.}} We run all experiments on NVIDIA Titan X and Tesla V100 GPUs.
+We use Python scikit-image to generate the stimuli and use Keras with TensorFlow to train the networks.
 % and scikit-learn libraries - JT: What do we use this for?
@@
 \centering
-\caption{\textbf{Network Training.} We use different feature generators as input to a multilayer perceptron which performs linear regression.
-This results in different sets of trainable parameters.
-As a baseline, we also train the MLP directly on the visualization images without any additional feature generation.}
+\caption{\textbf{Network Training.} We use different CNN feature generators as input to a multilayer perceptron, which results in different sets of trainable parameters.
+As a baseline, we also train the MLP directly.
+Optimization conditions are fixed across networks and experiments.}
 \resizebox{\linewidth}{!}{
-\begin{tabular}{lrl}
+\begin{tabular}{lr}
 %	\toprule
@@
 	\toprule
-	Network & \makecell{Trainable\\Parameters} & Optimization \\
+	Network & \makecell{Trainable\\Parameters} \\
 	\midrule
-	MLP & $2,560,513$ & SGD (Nesterov momentum)\\
-	\emph{LeNet} + MLP & $8,026,083$ & Learning rate: $0.0001$\\
-	\emph{VGG19} + MLP & $21,204,545$ & Momentum: $0.9$ \\
-	\emph{Xception} + MLP & $25,580,585$ & \makecell[tl]{Batchsize: 32\\Epochs: $1000$ (Early Stopping)}\\
+	MLP & $2,560,513$ \\
+	\emph{LeNet} + MLP & $8,026,083$ \\
+	\emph{VGG19} + MLP & $21,204,545$ \\
+	\emph{Xception} + MLP & $25,580,585$ \\
+	\bottomrule
+\end{tabular}
+\begin{tabular}{lr}
+\toprule
+\multicolumn{2}{l}{Optimization (SGD)} \\
+\midrule
+Learning rate & $0.0001$ \\
+Momentum & Nesterov \\
+\hspace{0.2cm} Value & $0.9$ \\
+Batch size & 32 \\
+Epochs & $1000$ (Early stop) \\
 \bottomrule
@@
 
-  \caption{\textbf{Network Architecture.} The multilayer perceptron (MLP) in our experiments has 256 neurons which are activated as rectified linear units (ReLU).
-We use Dropout regularization to prevent overfitting.
-As output, we perform linear regression for continuous variables.
-The MLP can learn to represent the visualizations directly, but we also learn features generated by LeNet (2 conv.
-layers, filter size $5\times5$), VGG19 (16 conv.
-layers, filter size $3\times3$), or Xception (36 conv.
-layers, filter size $3\times3$) to test different model complexities.}
+  \caption{\textbf{Network Architecture.} We use a multilayer perceptron (MLP) to perform linear regression for continuous variable output.
+We also learn convolutional features through LeNet (2 layers, filter size $5\times5$), VGG19 (16 layers, filter size $3\times3$), or Xception (36 layers, filter size $3\times3$) to test different model complexities.}
   \label{fig:classifiers}
+  \vspace{-0.5cm}
 \end{figure}
@@
 \noindent\textbf{Image Stimuli and Labels.} 
-We create our stimuli visualizations as 100$\times$100 binary images, rasterized without interpolation.
-We write a parameterized stimuli generator for each elementary task.
-The number of possible parameter values differs per experiment, and we summarize these in Table~\ref{tab:encoding_parameters} and Section~\ref{sec:parametrizations}.
-Before use, we scale the generated images into an unbiased range: images to the range of $-0.5$ to $0.5$.
-Then, we add subtle random noise (uniformly distributed between $0-0.05$) to each pixel to introduce variation which prevents the networks from simply `remembering' each different image.
+We create our stimuli as 100$\times$100 binary images, rasterized without interpolation.
+We develop a parameterized stimuli generator for each elementary task, with the number of possible parameter values differing per experiment (we summarize these in Table 1 of the supplemental material).
+Before use, we scale the generated images to the range of $-0.5$ to $0.5$ for value balance.
+Then, we add 5\% random noise (uniformly distributed between $-0.025$--$0.025$) to each pixel to introduce variation that prevents the networks from simply `remembering' each individual image.
+\change{In supplemental material Section 2, we visually compare how our stimuli vary from Cleveland and McGill's original stimuli for E1--5, and justify any differences.}
 
-Each stimuli image also has an associated ground truth label representing the parameter set which generated the image, e.g., the length in pixels of a bar.
-As before, we scale these labels to the range of $0.0$ to $1.0$, which represent the maximum and minimum values that this parameter can take.
+Each stimuli image also has an associated ground truth label representing the parameter set that generated the image.
+%, e.g., the length in pixels of a bar. As before, 
+We scale these labels to the range of $0.0$ to $1.0$ and normalize to the maximum and minimum value range for each parameter.
+%, which represent the maximum and minimum values that this parameter can take.
 \\~\\
@@
 \subsection{Measures and Analysis}
+\label{sec:measuresandanalysis}
 
-\noindent\textbf{Cross Validation.} For reproducibility, we perform repeated random sub-sampling validation, also known as Monte Carlo cross-validation, during our experiments.
-We run every experiment seperately twelve times, and randomly select (without replacement) the $60\%$ of our data as training data, $20\%$ as validation, and $20\%$ as test.
+\noindent\textbf{Cross Validation.} For experiment reproducibility, we perform repeated random sub-sampling validation, also known as Monte Carlo cross-validation~\cite{xu2001monte}.
+We run every experiment separately twelve times (four times for the `from scratch' networks due to significantly-longer training times), and randomly select (without replacement) the $60\%$ of our data as training data, $20\%$ as validation, and $20\%$ as test.
 %Our large data sample size of 100,000 guarantees that every single observation of our parameterizations will be selected at least once (excluding noise patterns). -> JT: No it doesn't.
@@
 \noindent{\textbf{Task Accuracy.}} In their 1984 paper, Cleveland and McGill use the midmean logistic absolute error metric (\emph{MLAE}) to measure perception accuracy.
-To allow comparison between their human results and our machine results, we also use MLAE as a presentation metric:
+To allow comparison between their human results and our machine results, we also use MLAE for presentation:
 \begin{equation}
@@
 \\~\\
-\noindent{\textbf{Task Confidence Intervals.}} We follow Cleveland and McGill and present $95\%$ confidence intervals.
-We approximate the value of the $97.5$ percentile point of the normal distribution for simplicity with $1.96$ as suggested by the central limit theorem~\cite{central_limit}.
+\noindent{\textbf{Error Distributions/Confidence Intervals.}} \change{We check for normality in our error distributions using the D'Agostino-Pearson test: 1.14\% of our networks did not pass.
+These were typically from the smaller MLP or LeNet networks (see supplemental material for example error distributions).
+As such, we broadly assume normality of errors and follow Cleveland and McGill in presenting $95\%$ confidence intervals, computed via bootstrapping (with 10,000 rather than 1,000 samples for a more accurate estimate)}.
+%We approximate the value of the $97.5$ percentile point of the normal distribution for simplicity with $1.96$ as suggested by the central limit theorem~\cite{central_limit}.
 \\~\\
-\noindent{\textbf{Confirmatory Data Analysis.}} To accept or reject our hypotheses, we analyze dependent variables using analysis of variance (ANOVA) followed by parametric tests. JT: Which tests?
+\noindent{\textbf{Confirmatory Data Analysis.}} To accept or reject our hypotheses under this normality, we analyze dependent variables using analysis of variance (ANOVA) followed by parametric tests.
+% this is clear from the notation t_test...
+no need to mention it otherwise
 \\~\\
-\noindent{\textbf{Training Efficiency.}} We use the training convergence rate as a measure of how easy or hard a particular task is for the network to learn to solve.
+\noindent{\textbf{Training Efficiency.}} We use the training convergence rate as a measure of how easy or hard a particular task is for the network to solve.
 This is defined as the MSE loss decrease per training epoch, which is an indicator of the training efficiency of the network with respect to the visual encoding.
+Lower MSE values are better.
+% and show that the network learned the task.
 \\~\\
-\noindent\textbf{Network Generalizability.} We evaluate generalizability by asking a network previously trained upon one task parameterization to answer questions about the same type of task stimuli but with more complex parameterization, e.g., estimating bar length without and with changes in stroke width.
+\noindent\textbf{Network Generalizability.} With sufficient capacity of trainable parameters, it is often said that a network can `memorize' the images if the data set has a low variability.
+Therefore it is important to consider this variability when evaluating different networks with fixed numbers of trainable parameters (Table~\ref{tab:parameters}).
+As discussed, we add noise to each stimulus image to increase variability.
+We also evaluate generalizability by asking a network previously trained for one task parameterization to answer questions about the same type of task stimuli but with more variability, e.g., estimating bar length without and with changes in stroke width.
 
 Further, some experiments compare different visual encoding types, e.g., bar plot vs.~stacked bar plot.
-We train and evaluate individual networks for each parameterization, plus we also train and evaluate a networks on stimuli across the different types.
-This single decision-making software better mimics the judgements that a human would be able to make.
+We train and evaluate individual networks for each task, and we also train and evaluate networks with stimuli from across the encoding types.
+These single decision-making \change{networks} better mimic judgments that a human would be able to make.
 %This scenario affects the optimization process and result in networks with more flexible knowledge with the caveat of longer training times. -> JT: Save it for the results discussion.
+
+\subsection{Human Baselines}
+\change{
+We take human baseline measurements for the position-angle (E2) and position-length (E3) experiments from Cleveland and McGill~\cite{cleveland_mcgill}, which had 51 participants.
+For the position-length experiment, we are also able to take human baseline measurements from Heer and Bostock's crowdsourced reproduction of Cleveland and McGill's experiments~\cite{HeerBostock2010}, which had 50 participants.
+Each participant in both experiments reviewed 10 stimuli for each condition.
+
+For the three remaining experiments (E1,E4,E5), we use Amazon Mechanical Turk to crowdsource new human baselines from 25 participants.
+Each participant was shown 10 stimuli for each experiment condition (nine for E1, two for E4, and three for E5), with three stimuli per condition presented as practice stimuli.
+This totaled 182 HITs per participant, with each HIT worth \$0.06.
+Average HIT time was 27 seconds.
+85 HITs total were rejected for out of range values.
+Participants were recruited from the US, with Master Worker or better qualification.
+As in Cleveland and McGill, participants were requested to perform ``a quick visual judgment and not try to make precise measurements, either mentally or with a physical object such as a pencil or your finger.''
+}

### 2018-08-15 PAPER/03_elementary_perceptual_tasks.tex
Update paper from overleaf.

@@
 % !TeX root = paper.tex
-\section{Experiment: Elementary Perceptual Tasks}
+\section{Experiment 1: Elementary Perceptual Tasks}
+\label{sec:elementary}
 
 Cleveland and McGill describe a set of elementary graphical perceptual tasks across ten encodings, where each encodes a quantitative variable in a graphical element or visual mark~\cite{cleveland_mcgill,cleveland1985graphical}.
-These tasks are the low-level building blocks for information visualizations (Table~\ref{tab:encoding_parameters}): estimating position on a common scale, position on non-aligned scales, length, direction (or slope), angle, area, volume, curvature,  shading (or ink density), and color saturation.
-As human color perception is complex, and because Cleveland and McGill perform no experiments with it, for now we leave it for future work.
+These tasks are the low-level building blocks for information visualizations (Figure~\ref{fig:figure1_results}): %(Table~\ref{tab:encoding_parameters}):
+estimating position on a common scale, position on non-aligned scales, length, direction (or slope), angle, area, volume, curvature, and shading (or ink density).
+We leave color saturation experiments for future work.
 
-For the remaining nine tasks, we create visualizations as 100$\times$100 raster images, and test whether each of our networks is able to regress absolute values from the images.
-As discussed in Section~\ref{sec:measuresandanalysis}, we generate multiple parameterizations for each elementary perceptual task to allow us to increase the number of parameters that the networks must estimate, and to measure performance as we move closer towards a general representation of visual marks.
-For instance, for \emph{Position Common Scale}, first we only vary the $y$-position of the spot to estimate against the scale, then we include translation along the x-axis, and then we vary the size of the spot size (Table~\ref{tab:encoding_parameters}).
-These parameterizations are still simple---each increase is only slightly more complex for a human to solve---but it increases the number of possible images for the network to `learn'.
+For these tasks, we create visualizations as 100$\times$100 raster images, and test whether each of our networks is able to regress quantities from the images.
+We generate multiple versions of each elementary perceptual task, which allows us to increase task complexity.
+For instance, for \emph{Position Common Scale}, first we only vary the $y$-position of the spot to estimate against the scale, then we include translation along the $x$-axis, and then we vary the spot size.
+Each variation increases the size of the space of possible images for the network to predict (Table 1; supplemental material).
+Since empirical evidence suggests that CNNs can interpolate between different training data points, we expect the networks to perform on variations of a similar perceptual task.
 
@@
 
-\begin{table}[!ht]
-\centering
-\caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of the elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
+% \begin{table}[t]
+% \centering
+% \caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of our elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
 We sequentially increase the number of parameters for every task (e.g., by adding translation).
 This introduces variability and creates increasingly more complex datasets.}
-\resizebox{\linewidth}{!}{
-\begin{tabular}{lllr}
-	\toprule
-	\multicolumn{2}{l}{Elementary Perceptual Task} & ~ & Permutations\\
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{position_common_scale.pdf}} & \makecell[tl]{\emph{Position Common Scale}\\~~~Position Y\\~~~+ Position X \\~~~+ Spot Size \\} &~& \makecell[tr]{~\\ $60$ \\ $3,600$ \\ $216,00$}\\
+% \resizebox{0.9\linewidth}{!}{
+% \begin{tabular}{lllr}
+% 	\toprule
+% 	\multicolumn{2}{l}{Elementary Perceptual Task} & ~ & Permutations\\
+% 	\midrule
+% 	\raisebox{-.85\height}{\includegraphics[width=.5in]{position_common_scale.pdf}} & \makecell[tl]{\emph{Position Common Scale}\\~~~Position Y\\~~~+ Position X \\~~~+ Spot Size \\} &~& \makecell[tr]{~\\ $60$ \\ $3,600$ \\ $21,600$}\\
 
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{position_non_aligned_scale.pdf}} & \makecell[tl]{\emph{Position Non-Aligned Scale}\\~~~Position Y\\~~~+ Position X \\~~~+ Spot Size \\} &~& \makecell[tr]{~\\ $600$ \\ $36,000$ \\ $216,000$}\\
+% 	\midrule
+% 	\raisebox{-.85\height}{\includegraphics[width=.5in]{position_non_aligned_scale.pdf}} & \makecell[tl]{\emph{Position Non-Aligned Scale}\\~~~Position Y\\~~~+ Position X \\~~~+ Spot Size \\} &~& \makecell[tr]{~\\ $600$ \\ $36,000$ \\ $216,000$}\\
 
-	\midrule
-	\raisebox{-.95\height}{\includegraphics[width=.5in]{length.pdf}} & \makecell[tl]{\emph{Length}\\~~~Length\\~~~+ Position Y \\~~~+ Position X \\~~~+ Width} &~& \makecell[tr]{ ~\\$60$ \\ $2,400$ \\ $144,000$\\$864,000$}\\
+% 	\midrule
+% 	\raisebox{-.95\height}{\includegraphics[width=.5in]{length.pdf}} & \makecell[tl]{\emph{Length}\\~~~Length\\~~~+ Position Y \\~~~+ Position X \\~~~+ Width} &~& \makecell[tr]{ ~\\$60$ \\ $2,400$ \\ $144,000$\\$864,000$}\\
 
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{direction.pdf}} & \makecell[tl]{\emph{Direction}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$360$ \\ $21,600$ \\ $1,296,000$}\\
+% 	\midrule
+% 	\raisebox{-.85\height}{\includegraphics[width=.5in]{direction.pdf}} & \makecell[tl]{\emph{Direction}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$360$ \\ $21,600$ \\ $1,296,000$}\\
 
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{angle.pdf}} & \makecell[tl]{\emph{Angle}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$90$ \\ $5,400$ \\ $324,000$}\\
+% 	\midrule
+% 	\raisebox{-.85\height}{\includegraphics[width=.5in]{angle.pdf}} & \makecell[tl]{\emph{Angle}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$90$ \\ $5,400$ \\ $324,000$}\\
 
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{area.pdf}} & \makecell[tl]{\emph{Area}\\~~~Radius\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$40$ \\ $800$ \\ $16,000$}\\
+% 	\midrule
+% 	\raisebox{-.85\height}{\includegraphics[width=.5in]{area.pdf}} & \makecell[tl]{\emph{Area}\\~~~Radius\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$40$ \\ $800$ \\ $16,000$}\\
 
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{volume.pdf}} & \makecell[tl]{\emph{Volume}\\~~~Cube Sidelength\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$20$ \\ $400$ \\ $8,000$}\\
+% 	\midrule
+% 	\raisebox{-.85\height}{\includegraphics[width=.5in]{volume.pdf}} & \makecell[tl]{\emph{Volume}\\~~~Cube Sidelength\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$20$ \\ $400$ \\ $8,000$}\\
 
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{curvature.pdf}} & \makecell[tl]{\emph{Curvature}\\~~~Midpoint Curvature\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$80$ \\ $1,600$ \\ $64,000$}\\
+% 	\midrule
+% 	\raisebox{-.85\height}{\includegraphics[width=.5in]{curvature.pdf}} & \makecell[tl]{\emph{Curvature}\\~~~Midpoint Curvature\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$80$ \\ $1,600$ \\ $64,000$}\\
 
-	\midrule
-	\raisebox{-.85\height}{\includegraphics[width=.5in]{shading.pdf}} & \makecell[tl]{\emph{Shading}\\~~~Density\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$100$ \\ $2,000$ \\ $40,000$}\\	
-%	
-	\bottomrule
-\end{tabular}
-}
-\label{tab:encoding_parameters}
-\end{table}
+% 	\midrule
+% 	\raisebox{-.85\height}{\includegraphics[width=.5in]{shading.pdf}} & \makecell[tl]{\emph{Shading}\\~~~Density\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$100$ \\ $2,000$ \\ $40,000$}\\
+% %	
+% 	\bottomrule
+% \end{tabular}
+% }
+% \label{tab:encoding_parameters}
+% \vspace{-0.25cm}
+% \end{table}
 
@@
 \begin{hypolist}
-\item \textbf{H1.1} \textbf{The CNNs tested will be able to regress quantitative variables from graphical elements.} We parametrize different visual encodings (Table~\ref{tab:encoding_parameters}) and test whether the CNNs can measure them.
-%, and relate the results to accuracies obtained by humans on similar tasks.
+\item \textbf{H1.1} \textbf{The CNNs tested will be able to regress quantitative variables from graphical elements.} We generate different visual encodings %(Table~\ref{tab:encoding_parameters})
+and test whether the CNNs can measure them. %, and relate the results to accuracies obtained by humans on similar tasks.
 
-\item \textbf{H1.2} \textbf{CNN perceptual performance will depend on network architexture.} We evaluate multiple regressors with different numbers of trainable parameters.
-We expect a more complex network (with more trainable parameters) to perform better on elementary perceptual tasks.
+\item \textbf{H1.2} \textbf{CNN perceptual performance will depend on network architecture.} We evaluate multiple regressors with different numbers of trainable parameters.
+We expect a more complex network (with more trainable parameters) to perform better on elementary perceptual tasks than a network with less complexity.
 
 \item \textbf{H1.3} \textbf{Some visual encodings will be easier to learn than others for the CNNs tested.} Cleveland and McGill order the elementary perceptual tasks by accuracy.
-We investigate whether this order is also relevant for computing graphical perception.
+We expect this order to be relevant for computing graphical perception.
 
-\item \textbf{H1.4} \textbf{Networks trained on perceptual tasks will generalize to more complex variations of the same task.} Empirical evidence suggests that CNNs are able to generalize by interpolating between different training data points, and so perform on variations of a similar perceptual task.
-We create visual representations of the elementary perceptual tasks with different variability, and expect that networks will be able to generalize when presented with slight task variations.
+\item \textbf{H1.4} \textbf{Networks trained on perceptual tasks will generalize to more complex variations of the same task.} Empirical evidence suggests that CNNs can generalize between different training data points.
+We create visual representations of the elementary perceptual tasks with different variability and expect that networks will be able to generalize to slight task variations.
 \end{hypolist}
@@
 %	
-%	We suspect that CNNs are able to 'learn` absolute quantities encoded using low-level visual 
+%	We suspect that CNNs can 'learn` absolute quantities encoded using low-level visual 
 %	
@@
 
+% \begin{figure}[tbhp]
+% 	\centering
+% 	\includegraphics[width=\linewidth]{figure1_slim_only_last.pdf}
+% %	\includegraphics[width=0.48\linewidth]{figure1_slim_right.pdf}
+% 	\caption{\textbf{Elementary perceptual tasks results for most complex task parameterization.} \emph{Left:} Example stimuli image.
+\emph{Right:} MLAE and 95\% confidence intervals for different networks.
+Lower MLAE scores are better.
+The * indicates ImageNet networks instead of being trained from scratch.}
+% 	\label{fig:figure1_results}
+% %	\vspace*{1in}
+% \end{figure}
+\begin{figure*}[t]
+	\centering
+	\includegraphics[width=.445\linewidth]{figure1_slim_only_last_NEW_LEFT.pdf}
+	\includegraphics[width=.45\linewidth]{figure1_slim_only_last_NEW_RIGHT_with_multi.pdf}
+%	\includegraphics[width=0.48\linewidth]{figure1_slim_right.pdf}
+	\caption{\textbf{Elementary perceptual tasks results for the most complex task parameterization.} In each column: \emph{Left:} Example stimuli image.
+\emph{Right:} MLAE and \change{bootstrapped} 95\% confidence intervals for different networks.
+Lower MLAE scores are better.
+The * indicates fine-tuned ImageNet weights instead of weights trained from scratch.
+\change{Bottom right shows `multi' VGG19 and Xception networks trained on all perceptual tasks, combined with optional $9\times$ increase of training data.} }
+	\label{fig:figure1_results}
+%	\vspace*{1in}
+\end{figure*}
+
 \subsection{Results}
 
+\change{Midmean random performance for all tasks is $MLAE=4.8$, save for direction ($4.6$) where the 0--360 wrap improves random performance.} % JT check this!
 % These values are similar or better than the range of MLAE=$3.02$--$3.82$ ($8-14\%$ error) for a similar experiment by Cleveland and McGill testing human estimation of relations between elementary perceptual tasks~\cite{cleveland1985graphical}.
@@
 These results confirm our initial hypotheses.
-\noindent{\textbf{Overall Accuracy.}} The tested CNNs and MLP are able to regress the visually encoded quantities in most cases (Fig.~\ref{fig:figure1_results}), with average error across all classifiers and tasks as \textit{MLAE}$=1.598$ ($SD=0.392$) and \textit{MAE}=$2.903$ ($SD=0.845$).
-From these results, we \textbf{accept H1.1}.
+\noindent{\textbf{Overall Accuracy.}} 
+The tested CNNs and MLP can regress the visually-encoded quantities in most cases (Fig.~\ref{fig:figure1_results}), with average error across all classifiers and tasks as \textit{MLAE}$=1.598$ ($SD=0.392$) and \textit{MAE}=$2.903$ ($SD=0.845$).
+Based on these results, we \textbf{accept H1.1}.
 \\~\\
@@
 VGG19 trained on ImageNet $0.979$ ($SD=0.581$), 
-and VGG19 trained from scratch $0.404$ ($SD=0.407$).
+and VGG19 trained from scratch $0.404$ ($SD=0.407$). Overall, VGG19 performs best.
 
-Across tasks, we compare the average regression performances for our networks and report the effect of the network as statistically significant ($F_{5,48}=20.470,p<0.01$).
+Across tasks, we compare the average regression performances for our networks and report the effect as statistically significant ($F_{5,48}=20.470,p<0.01$).
 Post hoc comparisons show that the differences between LeNet and the VGG19 network, independent of the used weights, are significant ($t_{16}=4.674,p<0.01$ and $t_{16}=8.746,p<0.01$).
@@
 However, differences between LeNet and both Xception networks are not significant.
-Taken collectively, we \textbf{partially accept H1.2}, in that more tunable parameters does not automatically infer greater performance.
+Taken collectively, we \textbf{partially accept H1.2}, in that higher network complexity does not automatically infer greater performance.
 \\~\\
@@
 Overall, there is significant variability in the rankings between architectures (Fig.~\ref{fig:figure1_results}).
-However, area estimation is an easier task for all networks, while direction and angle estimation are more difficult.
+Area estimation is an easier task for all networks, while direction and angle estimation are more difficult.
 It is harder to distinguish differences between position, length, curvature, and shading tasks.
-Further, the volume tasks suffers high variability in performance across cross-validation splits, which suggests that the image noise affects the outcome more than for other tasks.
+Further, the volume task suffers high variability in performance across cross-validation splits, which suggests that the image noise affects the outcome more than for other tasks.
+
+We note that the number of permutations across tasks does not strictly relate to network performance.
+While area in its most complex parameterization has 16,000 permutations, and so should be easier to learn, length has 864,000, yet VGG19 is able to achieve similar performance for both tasks.
+Likewise, direction has 4$\times$ more permutations than angle, yet the networks achieve similar performance.
+%\JT{This is because...}
+
 In sum, we \textbf{partially accept H1.3}.
-Finally, we note that the rankings between networks using ImageNet weights are identical, which suggests that the information about elementary perceptual tasks gained from those natural images is similar (given a sufficiently-complex network).
+Further, the rankings between networks using ImageNet weights are identical, suggesting that the information about elementary perceptual tasks gained from natural images is similar given a sufficiently-complex network.
 \\~\\
 \noindent{\textbf{Cross-network Variability and Network Generalizability.}} We measure regression performance across networks trained with different parameterizations of the elementary perceptual tasks (Fig.~\ref{fig:cross_network}).
-For our best performing network (VGG19 trained from scratch), we observe that accuracy decreases only slightly as the parameterization becomes more complex so long as training examples expressing all variability are included (diagonal entries in each matrix).
+For our best performing network (VGG19 trained from scratch), we observe that accuracy decreases only slightly as the parameterization becomes more complex if training examples expressing all variability are included (diagonal entries in each matrix).
 However, VGG19 is unable to generalize to added translation or stroke width variations in the encodings, leading to increases in error.
 As such, we \textbf{reject H1.4}.
+These findings suggest that slight variations in visual encodings can confuse CNNs, making it difficult to generalize the measurement of quantities to unseen examples dissimilar from the training data.
+
+%\JT{Discuss multi result; discuss more data result; discuss human performance.}
 
@@
 \caption{\textbf{Elementary Perceptual Task Ranking.} We report midmean logistic absolute errors (MLAE) for each network averaged across multiple runs on the most complex parametrization of each task.
+The lower MLAE, the better (negative values are the best).
 For human performance, we report the ranking of Cleveland and McGill~\cite{cleveland_mcgill}.
@@
 \multicolumn{7}{l}{\emph{Direction}} \\
-3. & 9. (4.65) & 7. (3.07) & 9. (2.84) & \textbf{8 (0.92)} & 9. (3.46) & 6. (1.57) \\
+3. & 9. (4.65) & 7. (3.07) & 9. (2.84) & \textbf{9 (0.92)} & 9. (3.46) & 6. (1.57) \\
 \multicolumn{7}{l}{\emph{Angle}} \\
-3. & 8. (4.61) & 9. (3.33) & 8. (2.31) & \textbf{9 (0.99)} & 8. (2.60) & 7. (1.69) \\
+3. & 8. (4.61) & 9. (3.33) & 8. (2.31) & \textbf{7 (0.66)} & 8. (2.60) & 7. (1.69) \\
 \multicolumn{7}{l}{\emph{Area}} \\
@@
 \multicolumn{7}{l}{\emph{Volume}} \\
-5. & 4. (2.38) & 4. (1.91) & 7. (1.16) & \textbf{7 (0.87)} & 7. (2.03) & 9. (2.10) \\
+5. & 4. (2.38) & 4. (1.91) & 7. (1.16) & \textbf{8 (0.87)} & 7. (2.03) & 9. (2.10) \\
 \multicolumn{7}{l}{\emph{Curvature}} \\
@@
 \label{tab:ranking}
+\vspace{-0.25cm}
 \end{table}
 
-\begin{figure}[p]
+\begin{figure}[tbhp]
 	\centering
-	\includegraphics[width=\linewidth]{figure1_slim_only_last.pdf}
-%	\includegraphics[width=0.48\linewidth]{figure1_slim_right.pdf}
-	\caption{\textbf{Elementary perceptual tasks results for most complex task parameterization.} \emph{Left:} Example stimuli image.
-\emph{Right:} MLAE and 95\% confidence intervals for different networks.
-The * indicates networks which use ImageNet weights up until the MLP, rather than being trained from scratch.}
-	\label{fig:figure1_results}
-%	\vspace*{1in}
-\end{figure}
-
-\begin{figure}[p]
-	\centering
-	  \includegraphics[width=.8\linewidth]{cross_network_small_VGG19_position_common_scale.pdf}
-	  \includegraphics[width=.8\linewidth]{cross_network_small_VGG19_length.pdf}
-	  \includegraphics[width=.8\linewidth]{cross_network_small_VGG19_area.pdf}
-	  \includegraphics[width=.8\linewidth]{cross_network_small_VGG19_shading.pdf}
-  \caption{\textbf{Cross-network variability for perceptual tasks.} VGG19 networks trained on different parametrizations (X-axis) tested across different parameterizations (Y-axis), for the top four performing encodings.
+	  \includegraphics[width=.85\linewidth]{cross_network_small_VGG19_position_common_scale.pdf}
+	  \includegraphics[width=.85\linewidth]{cross_network_small_VGG19_length.pdf}
+	  \includegraphics[width=.85\linewidth]{cross_network_small_VGG19_area.pdf}
+	  \includegraphics[width=.85\linewidth]{cross_network_small_VGG19_shading.pdf}
+	  \vspace{.1cm}
+  \caption{\textbf{Cross-network variability for perceptual tasks.} VGG19 networks trained on one set of parametrizations (X-axis) while tested across different ones (Y-axis), for the top four performing encodings.
 Diagonal matrix entries represent networks trained and tested on the same parameterizations.
-Below diagonal entries are scenarios where the test data has more parameters than the training data; below diagonal entries have fewer.
+Below diagonal entries are scenarios where the test data has more parameters than the training data; above diagonal entries have fewer.
 We measure the mean logistic absolute error (MLAE)---the lower the score, the better.

### 2018-08-15 PAPER/04_position_angle_experiment.tex
Update paper from overleaf.

@@
-\clearpage
-\section{Experiment: Position-Angle}
+\section{Experiment 2: Position-Angle}
+\label{sec:positionangle}
 
-Cleveland and McGill measure human perception to the ratios of positions and angles through comparisons on bar charts and pie charts~\cite{cleveland_mcgill}.
-We create rasterized images following Cleveland and McGill's proposed encoding and investigate computational perception of our networks (Figure~\ref{fig:teaser}).
-These have five bar or pie sectors representing numbers which add to 100, where each is greater than three and smaller than 39.
-One required change is in the minimal differences between the values: Cleveland and McGill create stimuli where the differences between each number are greater than $0.1$.
+Cleveland and McGill measure how humans perceive the ratios of positions and angles through comparisons on bar charts and pie charts~\cite{cleveland_mcgill}.
+We create rasterized images following Cleveland and McGill's proposed encoding and investigate computational perception of our networks (Fig.~\ref{fig:teaser}).
+These have five bar or pie sectors representing numbers that add to 100, where each is greater than three and smaller than 39.
+One required change is in the minimal differences between the values: Cleveland and McGill create stimuli with minimum scale difference of $0.1$.
 However, as our networks only take $100\times100$ pixel images as input, we can only minimally represent a difference of $1$ pixel.
@@
 %  \caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
-The experimental task involves the judgement of different encoded values in comparison to the largest encoded values.
+The experimental task involves the judgment of different encoded values in comparison to the largest encoded values.
 The pie chart (left) and the bar chart (right) visualize the same data point.
@@
 %\caption{\textbf{Position-Angle Experiment.} We create rasterized visualizations of pie charts and bar charts to follow Cleveland and McGill's position-angle experiment.
-The experimental task involves the judgement of different encoded values in comparison to the largest encoded values.
+The experimental task involves the judgment of different encoded values in comparison to the largest encoded values.
 The pie chart and the bar chart visualize the same data point.
@@
 \begin{hypolist}
-	\item \textbf{H2.1} \textbf{Computed perceptual accuracy will be higher for bar charts than pie charts.} Cleveland and McGill report that position judgements are almost twice as accurate (MLAE) as angle judgements in humans.
+	\item \textbf{H2.1} \textbf{Computed perceptual accuracy will be higher for bar charts than pie charts.} Cleveland and McGill report that position judgments are almost twice as accurate (MLAE) as angle judgments in humans.
 Following our ranking of elementary perceptual tasks (Table~\ref{tab:ranking}), we see that our networks also judge position encodings more accurately than angles, and so our networks will be able to more easily judge bar charts than pie charts.
@@
 
-\begin{figure}[t]
+\begin{figure}[tb]
 	\centering
-	  \includegraphics[width=\linewidth]{figure3_mlae_better_all.pdf}
-  \caption{\textbf{Computational results of the position-angle experiment.} \textit{Left:} Our encodings of one data point as a pie chart and a bar chart.
-\textit{Right:} MLAE and 95\% confidence intervals for different networks.
-VGG19 * and Xception * are using ImageNet weights while all other networks were trained on the stimuli.
-We train all networks 12 times (4 times for VGG19 and Xception due to significantly longer training times).
-VGG19 * and Xception * use ImageNet weights.
-Our results align with Cleveland and McGill's human results, shown in black~\cite{cleveland_mcgill}.}
+	  \includegraphics[width=0.90\linewidth]{figure3_mlae_better_all.pdf}
+  \caption{\textbf{Computational results of the position-angle experiment.} \textit{Left:} \change{Example stimuli.} \textit{Right:} MLAE and \change{bootstrapped} 95\% confidence intervals (the lower, the better).
+VGG19 * and Xception * fine tune ImageNet weights, with all other networks trained from scratch.
+We show Cleveland and McGill's human results in black~\cite{cleveland_mcgill}.} % We train all networks 12 times (4 times for VGG19 and Xception due to significantly longer training times).
 	\label{fig:figure3_mlae}
@@
 
-\noindent{\textbf{Perceptual Accuracy.}} Our networks are able to regress the task ratios for bar charts and pie charts (Fig.~\ref{fig:figure3_mlae}).
+%\JT{State random performance.}
+
+\noindent{\textbf{Perceptual Performance.}} \change{Midmean random performance is at $MLAE=4.7$.} Our networks are able to regress the task ratios for bar charts and pie charts (Fig.~\ref{fig:figure3_mlae}).
 Cross-validation yields an average $MLAE=2.176$ ($SD=0.456$) for bar charts, and an average $MLAE=3.296$ ($SD=0.77$) for pie charts.
@@
 the difference for VGG19 from scratch (pie charts $ 1.297 $ ($SD= 0.129 $), bar charts $ 1.153 $ ($SD= 0.09 $)) was not significant with $p<0.05$. 
-This outcome is in line with the elementary perceptual task results (Table~\ref{tab:ranking}), where VGG19 was the most successful network, where networks trained from scratch were more performant, and where angle was more difficult to learn than position.
+This outcome is in line with the elementary perceptual task results (Table~\ref{tab:ranking}), where VGG19 was most successful, where networks trained from scratch were more performant, and where angle was more difficult to learn than position.
 \\~\\
-\noindent{\textbf{Training Efficiency.} We measure the MSE loss for all networks on previously-unseen validation data during training.
-We count a network as converged when this validation loss does not decrease after 10 sequential epochs.
-Figure~\ref{fig:figure3_val_loss} shows this MSE validation loss during the first twenty epochs for each condition, plotted across all cross-validation splits with overdrawn lines.
+\noindent{\textbf{Training Efficiency.}} We measure the MSE loss for all networks on previously-unseen validation data during training.
+Figure~\ref{fig:figure3_val_loss} shows the first twenty epochs for each condition, plotted across all cross-validation splits with overdrawn lines.
 The pie chart loss decreases more slowly, with the average loss over epochs being $0.052$ ($SD=0.015$) for pie charts and $0.037$ ($SD=0.018$) for bar charts.
@@
 \\~\\
-To all our networks, the bar chart is a superior visual encoding than a pie chart, in terms of accuracy and efficiency.
+To all our networks, the bar chart is a superior encoding than a pie chart, in terms of accuracy and efficiency.
 Cleveland and McGill observe the same effect for accuracy during their human experiments.
@@
 
-\begin{figure}[t]
+\begin{figure}[tb]
 	  \includegraphics[width=\linewidth]{figure3_val_loss.pdf}
@@
 VGG19 * and Xception * use ImageNet weights.
-All networks converge faster when learning bar charts.}
+All networks reduce MSE loss faster when learning bar charts compared to  pie charts.}
 	\label{fig:figure3_val_loss}

### 2018-08-15 PAPER/05_position_length_experiment.tex
Update paper from overleaf.

--- 2018-03-31 PAPER/05_position_length_experiment.tex
@@
-\section{Experiment: Position-Length}
+\section{Experiment 3: Position-Length}
+\label{sec:positionlength}
 
-The position-length experiment by Cleveland and McGill involves the perception of five types of visualizations~\cite{cleveland_mcgill}.
-The visualizations are either grouped bar charts or divided bar charts (Fig.~\ref{fig:figure4_mlae}.
-Both types of graphs can show the same information but the elementary perceptual task is different.
-According to the theory by Cleveland and McGill, a grouped bar chart always involves the judgement of positions along a common scale.
-On the other hand, a divided bar chart might require length judgements in addition.
-Figure~\ref{fig:figure4_mlae} shows different charts of both types on the left.
-As identified by Cleveland and McGill, types 1, 2, and 3 involve the judgement of positions along a common scale while types 4 and 5 involve the measure of length.
-This means that this experiment does not just compare grouped versus divided bar charts but rather comparing position and length judgements.
+Cleveland and McGill assess the perception of position and length across five designs of grouped and divided bar charts (Fig.~\ref{fig:figure4_mlae}).
+Both types of chart can show the same information, but the elementary perceptual task is different: a grouped bar chart always involves the judgment of positions along a common scale, while a divided bar chart also requires length judgments.
+Types 1, 2, and 3 involve relating the judgment of positions along a common scale while types 4 and 5 involve relating the measure of length.
+For each graph, two bars or bar segments were marked, and participants were asked to judge what percentage the smaller marked element was of the larger.
+Cleveland and McGill ordered the types from easiest (type 1) to hardest (type 5).
 
-For data generation we follow the same approach as in the original experiment. We generate pairs from ten values generated using Cleveland and McGill's equation
+\setlength{\belowdisplayskip}{3pt} \setlength{\belowdisplayshortskip}{3pt}
+\setlength{\abovedisplayskip}{3pt} \setlength{\abovedisplayshortskip}{3pt}
 
+For data generation, we follow the same approach as in the original experiment. We generate ten value pairs using the following equation:
 \begin{equation}
@@
 \end{equation}
+Then, we generate eight other random values in the range of 10 and 93.
+These boundaries were chosen such that the largest first-layer convolutional filter size in our networks (of $5\times5$ in LeNet) would see all content in our $100\times100$ pixel image.
+The paired quantities/elements are marked by a single pixel.
+We ask our networks to estimate the ratio of the smaller to the larger, which we model as a single value regression problem.
+For type 4, we follow Cleveland and McGill's constraint that neither the top or the bottom of the marked quantities match, forcing estimations of length rather than position.
+This task has $9.20\times10^{16}$ possible permutations---a challenging problem for the capacity of our networks, but one that Cleveland and McGill found can be solved reliably by humans with $\approx6.5\%$ error \cite{cleveland_mcgill}.
 
-with two different values for each pair.
-We then generate 8 other random values in the range of 10 and 93.
-These boundaries were chosen because we create rasterized visualizations using $100\times100$ pixel and want to stay in frame for a convolutional filter size of $5\times5$ in LeNet.
-We then visually encode the ten values as type 1--5.
-The paired quantities get marked by a single pixel.
-These are our quantities of interest and the task is to estimate the ratio of the smaller to the larger.
-We model this task as a single value regression.
-For type 4, we follow Cleveland and McGill's constraint that neither the top or the bottom of the marked quantities match to force length estimations rather than position.
-
-We model this experiment as a single value regression task and ask networks to estimate the percentage of the smaller to the larger marked quantity in each visual encoding.
-The network first has to find the two marked quantities, then identify the smaller one, and finally estimate the ratio in comparison to the larger quantitiy.
+%We model this experiment as a single value regression task and ask networks to estimate the percentage of the smaller to the larger marked quantity in each visual encoding.
+The network first has to find the two marked quantities, then identify the smaller one, and finally estimate the ratio in comparison to the larger quantity.
 The 8 random quantities (which should be ignored by the network) push the number of possible permutations to a massive $9.20E+16$ and creates a very challenging problem.
-
-Finally, we also train `multi' networks which include all five types resulting in an even bigger challenge.
 
@@
 
-We proposed two hypotheses entering the elementary perceptual task experiment:
-
-\begin{itemize}
-  \item \textbf{H3.1} \textbf{Our networks can estimate all types equally well.} A grouped bar chart involves judging a position while a divided bar chart most likely (if not type 2) requires length judgements.
+\begin{hypolist}
+  \item \textbf{H3.1} \textbf{Our networks can estimate all types equally well.} A grouped bar chart involves judging a position while a divided bar chart most likely (if not type 2) requires length judgments.
 Our rankings of elementary perceptual tasks do not yield a strong preference for either across all networks.
-  \item \textbf{H3.2} \textbf{Trained `multi' networks work as well as individual trained networks.} Convolutional neural networks have massive numbers of trainable parameters.
-Their complexity allows them to learn different types of visual encodings in one training session.
-\end{itemize}
+  \item \textbf{H3.2} \textbf{A trained multi-task network will work as well as individual trained networks.} We train a multi-task network (labeled `multi') from all five types.
+While we fix the number of trainable parameters to be the same as in the single task network, CNNs have a hierarchical structure which allows them to learn intermediate representations that are useful for multiple tasks.
+\end{hypolist}
 
@@
 
+\vspace{-0.1cm}
 \subsection{Results}
 
-\begin{figure}[!t]
+\begin{figure}[t]
   \centering
-    \includegraphics[width=\linewidth]{figure4_mlae_with_multi_and_humans_all.pdf}
-  \caption{\textbf{Computational results of the position-length experiment.} \textit{Left:} Rasterized visualizations of type 1--5 for divided and grouped bar charts of Cleveland and McGill's position-length experiment.
-\textit{Right:} MLAE and $95\%$ confidence intervals for different regressors estimating the value of marked quantities in the visualizations.
-The VGG19 * and Xception * networks are using ImageNet weights.
-The top 5 rows represent networks trained on a single encoding while the last row shows `multi' networks which were trained on a random stream of types 1--5.
-We visually compare against human performance from the original experiment.}
+  \includegraphics[width=0.90\linewidth]{figure4_mlae_with_multi_and_humans_all_NEW.pdf}
+  \caption{\textbf{Computational results of the position-length experiment.} \textit{Left:} Type 1--5 stimuli for divided and grouped bar charts (as per Cleveland and McGill).
+\textit{Right:} MLAE and \change{bootstrapped} $95\%$ confidence intervals of our networks.
+Star * denotes networks using ImageNet weights.
+The last row shows `multi' networks trained on a random stream of types 1--5 \change{(plus VGG19 with $5\times$ training data)}.
+\change{We include human performance (black) from the original experiment (ClMcG, top) and from Heer and Bostock's crowdsourced studies (HeerBos, bottom).}}
   \label{fig:figure4_mlae}
@@
 
-\noindent\textbf{Perceptual Performance.} In the original position-length experiment, Cleveland and McGill report that types 1-5 were post-ordered by their perceptual difficulty with type 1 being the easiest and type 5 the hardest.
-We report the average MLAE for each type across our networks (and total 56 runs per type) as follows:
-for type 1 $MLAE= 3.956 $ ($SD= 0.274 $),
-for type 2 $MLAE= 3.952 $ ($SD= 0.441 $),
-for type 3  $MLAE= 4.349 $ ($SD= 0.367 $),
-for type 4 $MLAE= 3.668 $ ($SD= 0.256 $), and 
-for type 5 $MLAE= 3.902 $ ($SD= 0.253 $).
-These distributions yield significance ($F_{4,25}=2.815, p<0.05$) but post hoc comparions show that really only type 3 and type 4 differ ($t_{10}=3.406, p<0.01$).
-This means that the networks do not prefer a certain type on average which leads us to \textbf{partially accept H2.1} and we do not replicate the same pattern as Cleveland and McGill in their human studies even though our networks' performance is clearly worse than their human baseline.
+%\JT{State random performance.} 
+%\JT{Mention Heer and Bostock's results in body text.}
+
+\noindent\textbf{Perceptual Performance.} \change{Midmean random performance is at $MLAE=4.7$.} Average MLAE across our networks is: 
+type 1 $MLAE= 3.956 $ ($SD= 0.274 $),
+type 2 $MLAE= 3.952 $ ($SD= 0.441 $),
+type 3 $MLAE= 4.349 $ ($SD= 0.367 $),
+type 4 $MLAE= 3.668 $ ($SD= 0.256 $), and 
+type 5 $MLAE= 3.902 $ ($SD= 0.253 $).
+This yields significance ($F_{4,25}=2.815, p<0.05$), but post-hoc comparisons show that only types 3 and 4 differ ($t_{10}=3.406, p<0.01$).
+Thus, our networks do not prefer a certain type on average, and so we \textbf{partially accept H3.1}.
+%Further, we do not replicate the same type difficulty ordering as Cleveland and McGill in their human studies.
+
+Overall, our networks' performances are clearly worse than their human baselines.
+The problem space in these visual relation tasks is much larger than in the elementary perceptual tasks, resulting in average errors of $12$--$20\%$.
+Further, the finding from the elementary perceptual tasks that position and length judgments had approximately equivalent rank is consistent with these findings.
 \\~\\
-\noindent\textbf{`Multi' Network Performance.} In Cleveland and McGill's original position-length experiment, humans were asked to judge visualizations of types 1-5.
-In the bottom of Figure~\ref{fig:figure4_mlae} we average the human performances of types 1-5 to create a `multi' human.
-Similarly, when training our networks, we first limit each training to one stimuli.
-However, a `multi' network which learns types 1-5 simultaenously is closer to the human in the original user study.
+\noindent\textbf{Multi-task Network Performance.} In the original position-length experiment, humans were asked to judge visualizations across types 1-5.
+In the last row of Fig.~\ref{fig:figure4_mlae}, we average the human performances to create an average score across tasks.
+For our multi-task networks trained on all stimuli types, we record an average error across all classifiers of $MLAE=4.358$ ($SD=0.327$).
+Then, we compare against the average errors for all types, as reported above for perceptual performance.
+We reach significant differences ($F_{5,30}=3.454,p<0.05$).
+Post-hoc comparisons yield significant differences between the multi-task network and type 4 ($t_{10}=3.716, p<0.01$) and also to type 5 ($t_{10}=2.467, p<0.05$).
+Since the average MLAE is worse than all of types 1--5, and the distributions observe significant differences, we acknowledge that the multi-type task is harder for the networks than learning single types of encodings.
+We test whether there is insufficient training data for this task by providing five times more training data; however, this decreases performance.
+Thus, we \textbf{reject H3.2}.
+One promising exception is VGG trained from scratch, though performance has a wide variance across cross-validation sets.

### 2018-08-15 PAPER/06_bars_and_framed_rectangles_experiment.tex
Update paper from overleaf.

--- 2018-03-31 PAPER/06_bars_and_framed_rectangles_experiment.tex
@@
-\section{Experiment: Bars and Framed Rectangles}
+\section{Experiment 4: Bars and Framed Rectangles}
+\label{sec:barsframedrectangles}
 
-Visual cues can help converting graphical elements back to their real world variables.
-Cleveland and McGill introduced the bars and framed rectangles experiment which judges the elementary perceptual task of position along non-aligned scales~\cite{cleveland_mcgill}.
+Visual cues can help in converting graphical elements back to their real-world variables.
+Cleveland and McGill introduced the bars-and-framed-rectangles experiment to compare the perceptual judgment of length and position along non-aligned scales.
+Fig.~\ref{fig:figure12_mlae} shows both variations on the left.
+Without framing, it is difficult to judge which bar is larger (bottom).
+However, with a frame showing maximum length, this length judgment \change{can be} converted into a position judgment along non-aligned scales, which simplifies the perceptual problem.
 
-\subsection{Hypotheses}
+Cleveland and McGill theorize that judging the framed whitespace could be considered a length rather than a position judgment.
+Given this, they relate the task to Weber's Law: the perceivable difference within a distribution is proportional to the initial size of the distribution~\cite{householder1940weber}.
+For this experiment, Weber's Law implies that humans can more easily measure the difference in the white scale since its initial size is small, whereas estimating the small change in lengths of the black bars is harder.
+The Just Noticeable Difference (JND) is higher when the initial stimulus is smaller in size.
 
-We proposed two hypotheses entering the elementary perceptual task experiment:
+We set up the experiment as a two value regression task (Fig.~\ref{fig:figure12_mlae}).
+\change{Each bar length varies randomly by between 1 and 12 of 60 pixels, with each bar undergoing an individual vertical shift of up to 20 pixels.
+There are 132 different labels over 50,160 possible stimuli.}
 
-\begin{itemize}
-	\item \textbf{H4.1} \textbf{Classifiers can leverage additional visual cues.} The original bar and framed rectangle experiment shows how visual cues aid humans in mapping graphical elements to quantitative variables.
-This should be the same for feed-forward neural networks since they are inspired by the visual system.
-	\item \textbf{H4.2} \textbf{Weber's law can be transferred to computational perception.} Cleveland and McGill confirmed Weber's law based on the bar and framed rectangle experiment.
-For humans, the ability to perceive change within a distribution is proportional to the size of the initial distribution.
-\end{itemize}
+%\subsection{Hypothesis}
+
+\begin{hypolist}
+	\item \textbf{H4.1} \textbf{Networks performance will improve with additional visual cues.} The original bar and framed rectangle setting shows how visual cues aid humans in mapping graphical elements to quantitative variables.
+This should be the same for feed-forward neural networks, as we give them more signal from which to learn.
+\end{hypolist}
 
@@
 %\end{table}
-
-\subsection{Weber-Fechner's Law}
-
-As identified by Cleveland and McGill, the bar and framed rectangle experiment is closely related to Weber's law.
-This psychophysics law states that perceivable difference within a distribution is proportional to the initial size of the distribution.
-Weber's law goes hand-in-hand with Fechner's law.
-We conduct an additional experiment based on the original illustrations of the Weber-Fechner law to investigate whether this law can be applied to computational perception of our classifiers (Fig.~\ref{tab:weber_law}).
 
@@
 
+\begin{figure}[tb]
+  \centering
+  \includegraphics[width=0.9\linewidth]{figure12_mlae_better_all_NEW.pdf}
+  \caption{\textbf{Computational results of the bars-and-framed-rectangles experiment.} \textit{Left:} Stimuli of two bars for length judgment (bottom) following Cleveland and McGill's setting.
+Perceiving which bar is longer is \change{significantly} easier for humans when a frame is added (top).
+\textit{Right:} For networks (trained from scratch, or * indicates ImageNet weights), there seems no significant difference between the encodings as reported by MLAE and \change{bootstrapped} 95\% confidence intervals.}
+	\label{fig:figure12_mlae}
+\end{figure}
+
 \subsection{Results}
 
-First run indicates that framed rectangles perform better but we dont really know it yet.
+\change{Midmean random performance in this task is equal to $MLAE=4.8$.} We observe varying performance for our networks.
+Averaged across networks: the framed rectangle encoding $MLAE=1.982$ ($SD=0.89$) and for the bars encoding $1.867$ ($SD=0.709$).
+This difference was not significant, and so we \textbf{reject H4.1}.
+VGG19 again can regress the length in both cases, for the framed rectangle encoding $MLAE=0.595$ ($SD=0.225$) and for the bar encoding $MLAE=0.735$ ($SD=0.410$), though with higher variance without the added visual cues.
+The difference in relation to the visual cue is not significant.
+%\JT{Rejection of 4.1 calls into question the memorization of networks.}
 
-\begin{figure}[t]
-	  \includegraphics[width=\linewidth]{figure12_mlae_better_all.pdf}
-  \caption{\textbf{Computational results of the Bars-and-Framed-Rectangles experiment.} Log absolute error means and 95\% confidence intervals for the \emph{bars-and-framed-rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}.
-We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
-	\label{fig:figure12_mlae}
-\end{figure}
+\change{Further, our user study provides evidence that humans can more easily measure the frame rectangles.
+Participants were able to estimate the bar length with less error when frames are present $MLAE= 3.336 $ ($SD= 0.828 $) than without this visual aid $MLAE= 3.928 $ ($SD= 0.42 $).
+This difference is significant $F_{1,48}=9.765, p<0.01$.}
+
+%\\~\\
+%Unfortunately, both hypotheses were rejected for this experiment. However, it was great to see a partial validation of our ranking of elementary perceptual tasks (Table~\ref{tab:ranking}).

### 2018-08-15 PAPER/07_webers_law_experiment.tex
Update paper from overleaf.

\section{Experiment 5: Point Cloud Experiment}
\label{sec:weberslaw}

%We conduct an additional experiment testing whether Weber's law applies to convolutional neural networks for graphical perception. For this, we generate three 2D point clouds as base stimuli -- each is created randomly by activating 10, 100, or 1000 pixels in a $100\times100$ raster image. We then additionally activate from 1 to 10 random pixels within the initial distribution but by carefully only selecting inactive pixels. We show examples for this in Fig.~\ref{fig:weber_law} (left). This means that the number of additional points is harder to identify if they are added to the 1000 pixel set while the 10 pixel set allows to easily count. We then let our networks solve a regression task to estimate the number of added points.

We also create a random 2D point cloud version of Weber's law, in which the networks must estimate the number of added dots (up to 10) over an initial number of 10, 100, or 1,000 dots (Fig.~\ref{fig:weber_law}). Each individual stimulus image has random dot placement. For 10 initial dots, \change{given a brief glance at the stimuli, a human can approximately estimate the number of the dots,} but for 100 and 1,000 initial dots, this is a difficult problem where a human is likely to randomly guess. For a CNN, this problem is also difficult: there are $\binom{100\times100}{10}=2.73\times10^{33}$ locations for the 10 initial dots, which makes memorization untenable.

%\subsection{Hypothesis}

\begin{hypolist}
	\item \textbf{H5.1} \textbf{The networks will be unable to solve the point cloud experiment.} This just-noticeable-difference problem has too many parameter variations to judge, though a human could solve the simplest version with 10 initial dots.
\end{hypolist}

\begin{figure}[tb]
  \centering
  \includegraphics[width=0.9\linewidth]{weber_mlae_noise_all_NEW.pdf}
  \caption{\textbf{Computational results of the point cloud experiment.} \textit{Left:} We create 2D point clouds with 10, 100, and 1000 initial dots. Then, we add up to 10 new dots. For humans, it is possible to estimate how many dots are added if there are initially 10 points, but it is impossible to see how many dots are added when starting with 1000 dots. \textit{Right:} We let our networks regress the number of added dots and report MLAE and \change{bootstrapped} 95\% confidence intervals.}
	\label{fig:weber_law}
\end{figure}

\subsection{Results}

VGG is the only network able to solve the 10-dots version of the problem to within 10\% error, which itself is surprising. Most networks achieve close to midmean random performance, which is at $MLAE=4.8$. \change{Humans given a quick glance are better, solving the problem to within 17\% error or $MLAE=4.00$.} Moving to 100 dots, no network succeeds, and all networks achieve random performance. \change{For 100 and 1,000 dots, human performance is actually worse than random at 42\% and 44\% error respectively ($MLAE=5.39$ and $5.46$), as the participant response distributions skew towards estimating higher numbers of dots (we include response histograms in the supplemental material)}. At 1,000 dots, the larger-capacity networks perform better, which again is surprising, with VGG able to solve this problem to a low 3\% error. \change{This is explained by the fact that, with this many dots, the problem is easier solved as a summation problem rather than as a counting problem, which is a problem better suited to the multi-layer pooling architecture of CNNs.} As such, we only \textbf{partially reject 5.1}. 
%\JT{Unsatisfying!}

% Random performance was measured in MATLAB by:
% - Generating 1,000,000 random integers via uniform distribution to represent the random number of dots added.
% - Generating 1,000,000 random integers to represent the random guesses.
% - Computing the absolute error between these two vectors
% - Converting to MLAE
%
% In average error, random guesses should give 33% error
% But our human participants seem to be getting 42% error for 100, and 44% average error for 1,000.

%\JT{Relate to max pooling.} \JT{Why is 10 somewhat possible, but 100 is definitely not, but then 1,000 is possible again?}

### 2018-08-15 PAPER/08_discussion.tex
Update paper from overleaf.

\section{Discussion}
\change{
\noindent \textbf{Performance Across Experiments.} In the elementary perceptual tasks, CNNs were able to regress at least some parameter variant to error rates of around 3--10\%. This suggests that, for well-constrained tasks, we can use CNNs to directly predict quantities from individual visual marks or shapes. Area is one such task in which the multi-layer hierarchy of receptive fields and pooling layers can aid in solving this task via summation, whereas these help less for prediction in our direction task. In general, our humans were less proficient at these direct estimation tasks, as they require precise geometric reasoning. CNNs solve this problem in a different way to humans, as they interpolate from similar training examples within the learned representation.

However, the CNNs approach is not able to solve the visual relation task of the position-length experiment where it must identify and compare multiple bars. The problem has many more permutations of stimuli than the capacity of any of our networks~\cite{not_so_clevr}. Conversely, this visual relation task is relatively simple for humans given our ability to abstract the concept of length to identify bars and compute their ratio. This suggests that new network designs are needed to solve these problems in generalizable ways and not via exhaustive training.

Finally, some tasks that we did not expect to be solved, such as the 1000-point cloud JND task, were solvable by at least one network (VGG19). In principle, estimating the number of points added to the point cloud can be computed exactly by summing over the stimuli and subtracting the base number of points (10, 100, 1,000). This is a task that is virtually impossible at a glance for a human, but one that is made possible due to the CNNs layer aggregation methodology.
%
%
%All our generated stimuli exhibit parameter variability on top of the added noise, which ranged from $\approx20$ permutations up to tens of orders of magnitudes of variability. We found no clear correlation between task parameterization and network capacity, and it appears that some tasks are easier for networks to learn. 
%
%Our networks were able to solve the constrained position-angle experiment as well as a human. 
%Estimating vs.~Relating Visual Mark Quantities
%- elementary perceptual task -> single variable, able. Could use CNNs
%- position-length -> variable relations, unable. Need something else
%
\\~\\
\noindent \textbf{Architecture Surprises and Generalizability.} We see high variance in performance across our networks. The MLP is usually least able to solve the tasks as it contains no explicit visual processing convolution layers. Next, while it has convolutional layers, the LeNet often does not have the network capacity to solve the task. Next, we find that VGG19 is the best architecture for solving graphical perception tasks, regularly outperforming the newer Xception. This is surprising because Xception has more parameters and is deeper, giving it stronger ImageNet natural image performance. Further, in our elementary perceptual task (E1) multi networks (Figure~\ref{fig:figure1_results}), even with this extra capacity, Xception was less able to exploit the 9$\times$ increase in training data than VGG.

So why is VGG19 consistently better? Recent works have discovered a similar effect in comparisons with the more complex ResNet50 and InceptionV3 networks for classification under simple geometric transformations~\cite{Engstrom2017}. VGG19 is better able to generalize to image translations because it better anti-aliases the input and feature map signals (in the Nyquist–-Shannon sense) though the pooling and stride subsampling operations across its layers~\cite{Azulay2018}. %, which is a requirement for convolution to be translation invariant.

%VGG better preserves shiftability better avoids aliasing in feature maps by combining pooling with matching kernel stride operations) .

This property helps in our E1 tasks as we vary the X and Y location of the visual marks and is a useful invariance or generalizability for visualization tasks in general. Networks that do not preserve shiftability as well, like Xception, must have the capacity to learn translation invariance through the data~\cite{kauderer2017quantifying}. For instance, through data augmentations which translates the stimuli artificially. This is a less efficient use of each trainable parameter, which also helps to explain the less efficient training on the `multi' network with nine times the data.

%We posit that this is because Xception dedicates more of its parameters to inception blocks, which attempt to provide the network with some scale invariance (a property useful for natural image object detection). However, for our 2D graphical perception tasks, this property is rarely useful.
%
%, and it relates to the generalizability of the network. VGG better preserves information (in the Shannon sense) through its layers, and is more invariant to translation (which is normally a property assumed by networks).
%- More parameters doesn't necessarily mean more training data, but VGG benefitted
%
%One takeaway
%- Clear that, going forward, visualization tasks needs their own network designs to cope with specific problem
\\~\\
\noindent \textbf{Cross-parameterization Generalizability.} That said, our cross-network and cross-parameterization experiments (Fig.~\ref{fig:cross_network}) show that, even for VGG, this ability only goes so far to interpolate between training stimuli to new test stimuli. Without seeing any examples of X or Y translation, VGG performance deteriorates rapidly (for the other networks, too; see supplemental). This applies even to simple design variations like stroke width changes. Imagine scraping visualizations from the Web---a simple web search for `bar chart' yields high design variation. We can hardly expect to standardize visualization designs for computational analysis, and so this means that CNN training data for general applications must include representation from across the design space~\cite{kafle2018dvqa}. Thankfully, generating visualization designs with parametric models is relatively simple, unlike for natural images.

Further, we note that task performance can drop if the number of parameters is higher than the network capacity (in a loose sense), meaning that CNNs for understanding graphical data across visual designs must be large. In these respects, applying CNNs to understanding real-world visualizations remains a challenge~\cite{Kahou2018}.
%
%Many of our parameter variations introduce translation to the mark in X or Y. CNNs are often said to be `translation invariant', but this relies on the translation being present in the training data to begin with~\cite{kauderer2017quantifying}. In effect, this asks the network to learn many versions of the elementary prediction problem across image positions. Many of our networks have the capacity to accomplish this task.
%
% Discussion titles
% Generalizability to Designs
% - Not particularly present
% - Use other paragraph
% - Suggest that they could be applied when the output is standardized, though we're not expecting this to happen across science.
% - Our added translations are a form of data augmentation, which can help, but this is a learned generalizability rather than a specific invariance to a geometric or style transformation like stroke width
% \\~\\
% \noindent \textbf{Training and Data Strategies.} Our VGG and XCeption networks did benefit from being trained on multiple tasks (E1), so long as they could solve the original task individually---our multi network for position length (E3) was not able to learn anything useful across tasks to help it complete the task. Likewise, our networks did benefit from seeing more training data, even for the simple elementary tasks.
% - Xception network architecture could not as easily; underlying data variability may still be too sparse as engineered for ImageNet. Inception blocks try to explicitly separate spatial and cross-channel combination, which should in principle apply to our problem (e.g., allowing line translation to be separated from line stroke). Needs further investigation.
% - Going forward, easy to generate large variations of style - required for even the simple tasks (cite Kahle)
%
\\~\\
\noindent \textbf{From Scratch vs.~Fine Tuned.} Our networks fine-tuned on ImageNet were not better than those trained from scratch on the perceptual tasks, performing worse overall. This may be unsurprising given that our from-scratch networks are `specialized'. This confounds prior comparisons made between networks trained on natural images and the visual cortex, given that humans are also able to solve graphical perception tasks with the same visual system that sees the world. %This suggests that graphical perception is a higher-level task involving more than low-level vision.
%
%Our experiments suggest that researchers looking to apply existing network architectures to elementary visualization problems should do so without using existing weights.
%
%From Scratch vs. Pretrained
%- From scratch better; makes sense given limited number of systems seen.
%- No specific reason why to use natural image weights, unless need a combined natural image/visualization network (e.g., robot looking at the world and your powerpoint slides in a meeting). 
%
%
\\~\\
\noindent \textbf{Practical Insight Summary:}
Across our experiments, we gained insights with practical application for other researchers or practitioners investigating machine graphical perception:
\begin{enumerate}[label=\arabic*.,itemsep=0.5pt, topsep=1pt, parsep=0.5pt]
\item CNN architecture performance on natural images is not a good predictor for performance on graphical perception tasks. 
\item While the general network capacity trend of `more is better' is borne out by our experiments, other factors like generalization power through invariances are important for visualization tasks.
\item VGG19 or similar architectures are reasonable starting points.
\item Training weights from scratch is a better strategy for visualization task performance than fine tuning natural image weights.
\item Be aware when using natural image augmentations like zoom, rotation, or skew, as these have specific meanings for visualizations.
\item For training sets, we can use parametric models to explore large visualization design spaces, where data augmentation is replaced by parameter and design variations (e.g., stroke width).
%\item Take care drawing conclusions - way of solving problem is different
\end{enumerate}%
}% end \change{

### 2018-08-15 PAPER/09_conclusions.tex
Update paper from overleaf.

\section{Conclusions}

%Convolutional neural networks can regress quantities from visual encodings with reasonable error rates around 10\% in most cases. However, it seems that the networks do not learn perceptional concepts but rather learn variations of pixels. This can be seen in large margins of error when there is slight variability in stimuli. Also, severely complex tasks such as the position-length experiment are not solvable by modern CNNs.

%Future work shall address these issues and maybe find certain properties of network architectures which overcome these challenges. In our experiments, VGG19 -- even though not the most complex architecture -- performed better than others. Indeed there are plenty of directions to explore: How do CNNs react to other visual encodings?, Can we leverage adversarial networks to generate more machine perceivable visualizations?, Can CNNs correlate two-dimensional encodings properly?

%A more practical application could be the replacement of one-hot encoding in CNNs with the elementary shading task. VGG19 can quantify the shading with low error (Tab.~ \ref{tab:ranking}). This would reduce data dimensionality when one needs to encode unordered quantities.

%We hope that our experiments allow initial claims of what is easily possible and what is not. We certainly wish that we stimulate further research in this direction. Understanding infographics with full variability is an ambitious undertaking and we are excited to see what the future brings.

% Are results from CNN analysis of visualizations meaningful perceptually, and at what level?

% Can we build on CNNs and construct useful insights into human perception for visualization?

% How can we show a positive? If all we show are negatives, what is the next design where we show a positive?

%%%%%%%%%%%%%%%%%%%%%%%%%%
% Repeat the introduction

\change{
We set out to investigate how current CNNs perform on graphical perception tasks, and our findings are mixed. In the constrained settings of the elementary perceptual tasks of experiment 1, CNNs perform better than humans and were able to more accurately estimate quantities directly from visual marks on images. For the CNN, these tasks require learning to predict from training stimuli with relatively minor differences (which, for it, is easy), whereas for the human these tasks require making precise geometric estimates (which, for us, is hard). In other experiments, such as the position-length experiment, CNNs cannot complete the task. These tasks require identifying the bars of interest and then measuring the ratios of their lengths. Such visual relations are much harder for CNNs to predict as the space of outcomes is much larger, requiring exhaustive training~\cite{not_so_clevr}. In contrast, humans can generalize their concept of length to the new task from few examples. Finally, in the point cloud task, which is largely impossible for humans for 1,000 points, we see that VGG19 can solve this task to a high accuracy through its aggregation over layers.

Overall, the variation in our findings suggest that CNNs are not currently a good model for human graphical perception, in that they do not predict human performance and in that their successes and failures are at odds with human ability. This implicitly shows that the respective mechanisms for graphical perception are not comparable. As such, researchers and practitioners should be careful when applying and drawing conclusions from these models in their work. 
%
%Through these endeavors, our experiments lead to the practical insights that VGG19 is a better starting point for visualization analysis than more modern networks because it better generalizes information through its layers~\cite{Azulay2018}, and that even VGG with these properties must be fed a comprehensive training set which covers the visualization design space.
}
%
%Given our findings, it would seem a mistake to apply CNNs to graphical perception tasks
%
%
%However, performance alone is not sufficient to establish the CNNs as good models. Once similar performance was achieved, one would like to see the models demonstrate similar failure modes or biases that humans did. Seems clear that they work differently, but might be interesting to investigate correlation between human and CNN task performance for this effort. ???
%
\\~\\
\noindent \textbf{Future Work.} \change{In general, we are optimistic about machine graphical perception. Our findings suggest that this requires approaches that are different from the recent architectural developments like residual and inception blocks included in Xception to aid natural image classification. Networks that explicitly preserve geometric invariances, like scale, rotation, or translation, are potentially useful for graphical perception tasks~\cite{cohen2016group,worrall2017harmonic}. For instance, in the elementary angle task, rotation invariance would factor out the overall rotation and leave only the angle estimation problem; however, this must be controlled as, in the direction task, rotation invariance would remove the signal we wish to estimate.} New capsule networks also hold promise as they include specific architectural mechanisms to compartmentalize the learning of visual attributes like position, size, and orientation~\cite{capsules}. \change{Likewise, given that most visualization designs are easily described procedurally, there is promise in investigating generative approaches for learning probabilistic programs from visual stimuli~\cite{lake2015human}. Again, these would explicitly represent visual attributes. In general, approaches that attempt to represent higher-level abstractions are a key requirement for machine graphical perception to develop beyond a memorization and interpolation task. We release our open source code and data to help spur new machine perception systems more adept at graphical perception: \url{http://vcglab.org/perception}}

% Understanding Hinton's capsule networks
% https://medium.com/ai%C2%B3-theory-practice-business/understanding-hintons-capsule-networks-part-i-intuition-b4b559d1159b

### 2018-08-15 PAPER/SUPPLEMENTAL/supplemental.tex
Update paper from overleaf.

@@
 \usepackage{subfig}
+\usepackage{enumitem}
+
+\usepackage{xcolor}
+\newcommand{\tnote}[3]{{\color{#2}#1: #3}}
+%\newcommand{\change}[1]{\textcolor{red}{#1}}
+%\newcommand{\removed}[1]{\textcolor{red}{\sout{#1}}}
+% \newcommand{\change}[1]{{\color{red}#1}}
+\newcommand{\change}[1]{{#1}}
+
+\newcommand{\removed}[1]{}
+\newcommand{\DaH}[1]{\tnote{DaH}{red}{#1}}
+\newcommand{\HP}[1]{\tnote{HP}{green}{#1}}
+\newcommand{\JT}[1]{\tnote{JT}{blue}{#1}}
+\newcommand{\IGNORE}[1]{}
 
@@
 \maketitle
+%% The ``\maketitle'' command must be the first command after the
+%% ``\begin{document}'' command. It prepares and prints the title block.
 
-We present several plots which contain complete results for the elementary perceptual task experiment, for which, given the number of experiments, the main paper presented only a selection.
+%% the only exception to this rule is the \firstsection command
+
+\section{Additional Experiments}
+
+\subsection{Anti-aliasing}
+One way to increase the fidelity of the information given to the network for angled or curved lines is to anti-alias the line drawing.
+This provides grayscale intensity values to smooth out the jagged line edges and may be more suitable to networks trained on natural images (such as VGG19~* and Xception~*, with ImageNet weights).
+
+We conduct an experiment to measure continuous pie chart angles from stimuli with anti-aliased (AA) lines (Figure~\ref{fig:aa}).
+There is a slight increase in performance for the AA case for VGG19 *, and an even slighter increase for Xception *.
+However, overall, adding anti-aliasing to the line generation was not statistically significant to the tested CNNs when comparing the performance across all networks ($F(1,30)=0.341, p>0.5$).
+This is not surprising since smoothing only changes very little information at the resolution of our stimuli.
+Smooth edges might add more value on higher-resolution stimuli.
+% which we plan to evaluate in the future.
+
+%\JT{Where is the comparison to the `from scratch' stimuli?
+Can we say anything about this here?} \JT{What statistical test is this?
+Is it across all conditions, or just for VGG19?
+YES this is across all conditions...
+whether the average result changes}
+
+% In our experiments, we chose to create visualizations without interpolation, and to estimate integer-valued measurements.
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=.825\linewidth]{../gfx/figure3_mlae_better_all_AA.pdf}
+\caption{\textbf{Anti-aliasing.} We test whether anti-aliasing affects the performance of our networks on pie charts by measuring MLAE.
+Starred networks are fine-tuned from ImageNet weights.
+The difference is not statistically significant ($F(1,30)=0.341,p>0.5$).}
+\label{fig:aa}
+\end{figure}
+
+\subsection{Noise}
+
+For all our experiments, we add subtle $5\%$ noise to every pixel to enhance variability.
+We did not observe a significant effect on regression performance when comparing the Weber-Fechner's law experiment with and without noise, averaged over 4 runs (Figure~\ref{fig:noise}).
+However, the variability of LeNet increased with the additional complexity.
+
+\begin{figure}[p]
+\centering
+\includegraphics[width=.825\linewidth]{../gfx/weber_mlae_no_noise.pdf}\\
+\vspace{0.4cm}
+\includegraphics[width=.825\linewidth]{../gfx/weber_mlae_noise.pdf}
+\caption{\textbf{Noise.} We test whether noise affects the performance of our networks on the Weber-Fechner's law experiment by measuring MLAE.
+\emph{Top:} With noise, $MLAE=4.511$ ($SD=0.512$).
+\emph{Bottom:} Without noise $MLAE=4.491$ ($SD=0.543$).
+The difference is not statistically significant ($F(1,22)=0.008,p>0.5$).}
+\label{fig:noise}
+\end{figure}
+
+\section{Comparisons to Cleveland and McGill's Stimuli}
+
+\change{
+We visually compare our stimuli to those of Cleveland and McGill~\cite{cleveland_mcgill}, and explain our choices for any differences.
+}
+
+\subsection{Resolution Differences}
+
+\change{
+Cleveland and McGill's stimuli were created on 8.5$\times$11 inch paper and presented to participants in a binder.
+While it is unclear from their paper how the stimuli were created (e.g., printed from a digital file, or drawn with pen), it is safe to assume that there is a resolution difference from our stimuli.
+Our stimuli are 100$\times$100 digital images, with their resolution chosen for computational reasons: We needed to train more than 2,500 models and, at this resolution, our most complex Xception model took 6 hours to train with 60,000 stimuli.
+
+This limited resolution affects network performance in making precise estimates, especially in tasks that require estimating lines drawn at angles or curved lines.
+As we draw binary stimuli, there are only so many angles of lines that can be drawn within 100$\times$100 pixels that produce different visual output.
+This leads to a performance floor or constant error.
+For instance, on average, we can only draw binary angled lines with 1.45 degrees of accuracy (0.67 standard deviation).
+
+% Point about input size, layers, receptive fields, etc.?
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Pre-declaring some figures for good placement %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{figure*}[tb]
+    \centering
+    \includegraphics[width=0.49\linewidth]{./gfx/CMcG_Comparison/CMcG_ElementaryPerceptualTasks.png}
+    \vrule
+    \hspace{0.2cm}
+    \includegraphics[width=0.465\linewidth,trim=0 -2.00cm 0 0, clip]{./gfx/CMcG_Comparison/Ours_ElementaryPerceptualTasks.pdf}
+    \caption{\emph{Left:} Cleveland and McGill's example figure for their elementary perceptual tasks, with each task showing variation through two sample examples (Figure 1 from \cite{cleveland_mcgill}).
+\emph{Right:} Two example stimuli from our elementary perceptual task experiment.
+Note that the border around each of our stimuli only exist in this figure to visually separate the stimuli; no border exists in our experiments.
+Our stimuli also include minor variations in style, such as dot and stroke width in position and length.
+Finally, as per Cleveland and McGill, we defer color saturation as a perceptual problem for future work.
+(Though given the success of CNNs in estimating elementary graphical perception task values in comparison to humans, we would also expect CNNs to be more successful than humans at estimating color saturation.)}
+    \label{fig:cmcg_elementary_comparison}
+\end{figure*}
+
+\subsection{Elementary Perceptual Tasks}
+
+\change{
+Cleveland and McGill provide an explanatory figure for their defined elementary perceptual tasks, with each example showing the expected variation in parameter through two samples.
+We compare visually to two samples of our stimuli (Figure \ref{fig:cmcg_elementary_comparison}).
+Our stimuli are similar in appearance, with the addition that our stimuli also include minor variations in style such as dot and stroke width in position and length.
+
+One difference is that we exclude the numerical values written as text along the scale in the `Position Common Scale' and `Position Non-aligned Scales' examples.
+While training a CNN to recognize text on graphs is possible given enough examples~\cite{kafle2018dvqa}, in our case these numbers would not change from stimuli to stimuli and so the network would not gain any information between stimuli to help solve the position estimation problem.
+As such, we simplify the task by removing the redundant numbers, with the scale values implicitly encoded by the size of the image and the location of the line start and end pixels.
+
+By the same reasoning, the scale itself is redundant in the `Position Common Scale' task as it does not move---the network must map the vertical pixel positions of the dots to the quantity range of numbers.
+We keep the scale to allow comparison with the `Position Non-aligned Scale' task.
+Further, when we train networks that look at \emph{all} elementary perceptual tasks, keeping the scale allows the network to exploit the consistency of representation across the two position estimation tasks.
+}
+
+Finally, we list the permutations of each stimuli in Table \ref{tab:encoding_parameters}.
+
+\begin{table}[t]
+\centering
+\caption{\textbf{Elementary Perceptual Tasks.} Rasterized visualizations of our elementary perceptual tasks as defined by Cleveland and McGill~\cite{cleveland_mcgill} (color saturation excluded).
+We sequentially increase the number of parameters for every task (e.g., by adding translation).
+This introduces variability and creates increasingly more complex datasets.}
+\resizebox{0.9\linewidth}{!}{
+\begin{tabular}{lllr}
+	\toprule
+	\multicolumn{2}{l}{Elementary Perceptual Task} & ~ & Permutations\\
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{../gfx/position_common_scale.pdf}} & \makecell[tl]{\emph{Position Common Scale}\\~~~Position Y\\~~~+ Position X \\~~~+ Spot Size \\} &~& \makecell[tr]{~\\ $60$ \\ $3,600$ \\ $21,600$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{../gfx/position_non_aligned_scale.pdf}} & \makecell[tl]{\emph{Position Non-Aligned Scale}\\~~~Position Y\\~~~+ Position X \\~~~+ Spot Size \\} &~& \makecell[tr]{~\\ $600$ \\ $36,000$ \\ $216,000$}\\
+
+	\midrule
+	\raisebox{-.95\height}{\includegraphics[width=.5in]{../gfx/length.pdf}} & \makecell[tl]{\emph{Length}\\~~~Length\\~~~+ Position Y \\~~~+ Position X \\~~~+ Width} &~& \makecell[tr]{ ~\\$60$ \\ $2,400$ \\ $144,000$\\$864,000$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{../gfx/direction.pdf}} & \makecell[tl]{\emph{Direction}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$360$ \\ $21,600$ \\ $1,296,000$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{../gfx/angle.pdf}} & \makecell[tl]{\emph{Angle}\\~~~Angle\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$90$ \\ $5,400$ \\ $324,000$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{../gfx/area.pdf}} & \makecell[tl]{\emph{Area}\\~~~Radius\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$40$ \\ $800$ \\ $16,000$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{../gfx/volume.pdf}} & \makecell[tl]{\emph{Volume}\\~~~Cube Sidelength\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$20$ \\ $400$ \\ $8,000$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{../gfx/curvature.pdf}} & \makecell[tl]{\emph{Curvature}\\~~~Midpoint Curvature\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$80$ \\ $1,600$ \\ $64,000$}\\
+
+	\midrule
+	\raisebox{-.85\height}{\includegraphics[width=.5in]{../gfx/shading.pdf}} & \makecell[tl]{\emph{Shading}\\~~~Density\\~~~+ Position Y \\~~~+ Position X} &~& \makecell[tr]{ ~\\$100$ \\ $2,000$ \\ $40,000$}\\
+%	
+	\bottomrule
+\end{tabular}
+}
+\label{tab:encoding_parameters}
+\end{table}
+
+% \subsubsection{Stimuli Generation}
+% \JT{Not sure of the value of this.}
+% Our code release acts as documentation for our stimuli generation process.
+Additionally, we describe it below.
+In all cases, we make sure that the drawn visual marks remain fully on the image, with a border of empty space equal to at least 5$\times$5 so that a first-layer convolution kernel will see all image content.
+% \begin{description}
+% \item[Position Common Scale] Given a 100$\times$100 image, the task is to estimate a number from 0 to 60.
+With (0,0) as bottom left, a scale one pixel thick is drawn from (10,20) to (10,80).
+The position dot is drawn offset from the scale horizontally by 40 pixels, i.e., at (50,.), with variation in Y between 20--80.
+To add complexity, the horizontal position of the dot then also varies in X (20--80).
+In the most complex variation, the dot size varies in pixel width/height from the set $\{1, 3, 5, 7, 9, 11\}$, i.e., the dot is maximally 11$\times$11 pixels large.
+% \item[Position Non-aligned Scale] This is the same as for `Position Common Scale', with the task being to estimate a number from 0 to 60.
+Now, the vertical position of the scale now also varied by offsetting by between -10 and 10 pixels, i.e., any 80 pixels within the range 10--90.
+The dot is similarly offset so that it follows the scale.
+% \item[Length] The task is to estimate the bar length, which is randomly generated between 1 and 60.
+In the simplest case, this is drawn at x-coordinate 50 pixels.
+Then, we vary the horizontal position between positions 20--80.
+Finally, we vary the stroke width from the set $\{1, 3, 5, 7, 9, 11\}$.
+% \item{Direction}
+% \item{Angle}
+% \item{Area}
+% \item{Volume}
+% \item{Curvature}
+% \item{Shading}
+% \end{description}
+
+\subsection{Position-length Experiment}
+
+\change{
+Again, we attempted to replicate Cleveland and McGill's stimuli as closely as possible, while removing the text on the scale and the A/B category labels (Figure~\ref{fig:cmcg_positionlength_comparison}).
+These values are implicitly encoded by the vertical height within the image, and by the left/right position of the bars.
+Like Cleveland and McGill, we place a dot within the regions that require ratio estimation.
+In our case, the network must learn the association of dot placement to required length estimation.
+}
+
+\begin{figure*}[tb]
+    \centering
+    %\includegraphics[width=0.8\linewidth]{./gfx/CMcG_Comparison/CMcG_PositionLength.png}\\
+    \includegraphics[width=0.8\linewidth]{./gfx/CMcG_Comparison/Ours_PositionLength.pdf}
+    \caption{\emph{Top:} Cleveland and McGill's example stimuli for their position-length experiment (Figure 4 from \cite{cleveland_mcgill}).
+\emph{Bottom:} Example stimuli from our position-length experiment.
+Note that the border around each of our stimuli only exist in this figure; no border exists in our experiments.}
+    \label{fig:cmcg_positionlength_comparison}
+\end{figure*}
+
+\subsection{Position-angle Experiment}
+
+\change{
+As before, our figures make implicit the scales and labels within these stimuli (Figure~\ref{fig:cmcg_positionangle_comparison}).
+One difference here is that Cleveland and McGill explicitly tell the participants on a supplemental sheet which pie segment or bar is largest; we replace this signal with a visual dot.
+}
+
+\begin{figure}[tb]
+    \centering
+    \includegraphics[width=0.75\linewidth,trim=-1cm 0 0 0, clip]{./gfx/CMcG_Comparison/CMcG_PositionAngle.png}\\
+    \includegraphics[width=0.72\linewidth]{./gfx/CMcG_Comparison/Ours_PositionAngle.pdf}
+    \caption{\emph{Top:} Cleveland and McGill's example stimuli for their position-angle experiment (Figure 3 from \cite{cleveland_mcgill}).
+\emph{Right:} Example stimuli from our position-length experiment.
+Note that the border around each of our stimuli only exist in this figure; no border exists in our experiments.
+Further, we have resized our stimuli on the page to more clearly show the similarity; each are 100$\times$100 pixels.}
+    \label{fig:cmcg_positionangle_comparison}
+\end{figure}
+
+\subsection{Bars/Rectangles Experiment}
+
+\change{
+In this case, the only difference is that we remove the A/B category labels, and let the comparison judgment be implicitly encoded by the left/right bars (Figure~\ref{fig:cmcg_barsrectangles_comparison}).
+}
+
+\begin{figure}[tb]
+    \centering
+    \includegraphics[width=0.8\linewidth]{./gfx/CMcG_Comparison/CMcG_BarsRectangles.png}\\
+    \includegraphics[width=0.8\linewidth]{./gfx/CMcG_Comparison/Ours_BarsRectangles.pdf}
+    \caption{\emph{Top:} Cleveland and McGill's example bars and rectangles (Figure 12 from \cite{cleveland_mcgill}).
+\emph{Right:} Example stimuli from our bars and rectangles experiment.
+Note that the border around each of our stimuli only exist in this figure; no border exists in our experiments.}
+    \label{fig:cmcg_barsrectangles_comparison}
+\end{figure}
+
+\subsection{Point Cloud Experiment}
+
+\change{
+Cleveland and McGill have no equivalent to this Weber-Fechner's law experiment, and so we show no stimuli comparison.
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{\change{Comparing the Human vs.~Machine\\Judgment Processes}}
+
+\change{
+There are many general differences in how humans and CNNs perceive: 
+\begin{itemize}[itemsep=0.5pt, topsep=1pt, parsep=0.5pt]
+    \item Humans have foveal and peripheral vision. 
+    \item Humans have explicit attention mechanisms (V1), with neural feedback to drive attention.
+    \item Humans have working memory.
+    \item Humans can make predictions about what they might see given a hypothesis. 
+\end{itemize}    
+It can be argued that a CNN can see and integrate all image information at once, as it has no fovea. This partially overcomes the need for attention mechanisms, feedback, and working memory.
+
+Having spoken to three computational human vision scientists about graphical perception, it is our understanding that we still know very little about how humans see data visualizations beyond the very early visual system.
+In some cases, through works like Cleveland and McGill, we know how \emph{well} humans can judge elementary graphical perception tasks, and this is the level at which we make comparisons about the suitability of CNNs as models for human graphical perception---implicitly, rather than explicitly.
+
+Given these limitations, we will try and describe at a high level what we think happens when a human is asked to predict from the `position common scale' stimuli, and what happens when a CNN is asked to predict given the stimuli as input.
+
+\subsection{Human Judgment}
+For training, a human being learns how to see through experience gained by existing in the world, such as our ability to interpret visual edges as lines and to read numbers.
+Given this experience, upon observing the `position common scale' stimuli at prediction time:
+\begin{enumerate}[label=\arabic*.,itemsep=0.5pt, topsep=1pt, parsep=0.5pt]
+    \item The human looks at the scale and reads the demarcation ticks and text labels. This lets the human know the numerical values associated with the visual marks and defines the scale bounds.
+    \item The human looks at the dot and mentally projects a horizontal line over to intersect the scale.
+    \item The human estimates the distance between the two closest demarcation ticks and the intersection line. The human may make additional mental ticks within the span to help in the estimation.
+    \item The human converts this into the scale using knowledge of the tick values, producing a prediction.
+\end{enumerate}
+The first two steps may be interchanged, as the output of each is stored in human memory in either case.
+
+\subsection{Brief CNN Introduction}
+For a CNN, the process is harder to describe in intuitive terms, and so we provide a brief introduction to how CNNs work.
+The network is defined to have a set of layers, with each layer feeding into the next, and each typically performing some subset of three operations: convolution with kernels, non-linear transformation with activation functions, and pooling.
+Following these, a densely-connected set of neurons known as a multi-layer perceptron (MLP) learns how to combine the convolution layer outputs to make a prediction.
+
+\paragraph{Trainable Convolution.} The network has a number of kernels within each layer, e.g., in layer one of LeNet, we have 20 trainable 5$\times$5 kernels.
+We convolve an input image with our kernels to produce a `feature map' measuring the kernel response to the image.
+Once trained, these kernels identify important visual elements in a scene, similar to a template matching process.
+In a loose biological comparison, kernels in early layers are often compared to functions in the early human visual system that identify contrast or edges, which is an important task within graphical perception.
+Through hierarchical combination, kernels in later layers are said to localize and identify objects (if the network is trained on natural images)~\cite{Zeiler2014}.
+In our graphical perception setting, this would potentially let a network identify and compare larger structures like the cube in our volume task, or glyphs.
+
+\paragraph{Activation Functions.} With the feature maps computed by the convolution with trained kernels, we apply the activation function, which defines how much we wish our network to respond to a given feature.
+LeNet uses the Rectified Linear Unit (ReLU) activation function, which sets all values less than 0 to 0, and is loosely said to be biologically inspired by one-sided human neuron activations.
+ReLU helps us train deeper neural networks by propagating large positive signals~\cite{pmlr-v15-glorot11a}.
+
+\paragraph{Pooling.} A pooling layer acts to summarize a response over a spatial region of the feature map, either by averaging or by selecting the highest response from the region (`max.~pooling').
+LeNet uses 2$\times$2 max pooling, which decreases the size of the feature map by two in each dimension.
+Across layers, pooling lets the network accumulate or select activations as part of its prediction process.
+
+\paragraph{MLP.} Finally, after a series of layers (two for LeNet: Conv$\rightarrow$ReLU$\rightarrow$MaxPool$\rightarrow$Conv$\rightarrow$ReLU$\rightarrow$MaxPool), we feed the resulting downsampled feature map into a multi-layer perceptron, which is connected to every pixel in the feature map.
+Each perceptron in the MLP is a linear classifier that is trained to weight the feature maps to produce the correct prediction.
+The MLP itself has two layers separated by a ReLU to allow for complex combination.
+
+\subsection{CNN Judgment.} Now, let us predict for the `position common scale' task.
+Through training, our network has learned a set of convolutional kernels across its layers that, in combination, help the network to predict estimates for unseen test data.
+\begin{enumerate}[label=\arabic*.,itemsep=0.5pt, topsep=1pt, parsep=0.5pt]
+    \item The kernels produce strong responses in parts of the test image that were helpful to predict the training data.
+These are typically edges or visualization parts such as bars.
+For `position common scale', early-layer kernels produce a strong response for dots.
+    \item Pooling selects the most important features from across the image.
+Over a hierarchy of layers, these are trained to be combined by kernels into more useful features.
+For `position common scale', this dot response would be propagated through the layers, but the exact position of the dot would be made imprecise.
+    \item The MLP has been trained to weight (or combine) the pooled feature maps into a prediction.
+For `position common scale', the MLP will weight the feature map response at different locations to map to a value between 0--60.
+\end{enumerate}
+
+\subsection{Human/CNN Comparison}
+While parts of this process are similar, such as the early visual system identifying edges, the high-level picture is that the human and CNN prediction processes are substantially different.
+The human prediction involves implicit geometric understanding of a space, with projection and line length comparisons.
+The CNN prediction has no explicit way to represent these high-level operations; instead, they must be implicitly represented though a series of low-level weighted sums of pixel responses, clamping activation function transformations, and spatially-local max operations.
+
+However, for some perception tasks like area estimation or our point cloud experiment in the main paper, this series of operations is beneficial.
+For instance, estimating the number of points added to the point cloud can be computed exactly by summing over the stimuli and subtracting the base number of points (10, 100, 1,000).
+This is a task that is extremely tedious at best for a human, but easier for the CNN given its layer aggregation methodology.
+%Given the training data, the CNN is able to interpolate its trained response to new examples of sufficient similarity.
+But, as we show in the main paper, this generalization is limited for many of our tasks.
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Supplemental Results}
+
+We present several plots that contain complete results for the elementary perceptual task experiment, for which, given the number of experiments, the main paper presented only a selection.
 We report MLAE for all added parameters to the stimuli, rather than just the most complex parameterization as in the main paper (Figures~\ref{fig:epc_mlae} and \ref{fig:epc_mlae_boxplots}).
@@
 Further, we show how the errors for each network are distributed, across elementary perceptual tasks and across different cross-validation splits (Figure~\ref{fig:errors}).
-Most errors are approximately normally distributed, though our CNN with less parameters (LeNet) and our MLP often have errors which are farther from a normal distribution, and show structure.
+Most errors are approximately normally distributed, though our CNN with less parameters (LeNet) and our MLP often have errors that are farther from a normal distribution and show structure.
 
-\section{Experiment: Anti-aliasing}
-We also include an experimental result showing that, for our stimuli, adding anti-aliasing to the line generation was not important to the tested CNNs (Figure~\ref{fig:aa}).
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\bibliographystyle{abbrv}
+\bibliographystyle{abbrv-doi}
+%\bibliographystyle{abbrv-doi-narrow}
+%\bibliographystyle{abbrv-doi-hyperref}
+%\bibliographystyle{abbrv-doi-hyperref-narrow}
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=\linewidth]{../gfx/figure3_mlae_better_all_AA.pdf}
-\caption{\textbf{Anti-aliasing.} We test whether anti-aliasing effects the performance of our networks on pie charts by measuring MLAE.
-The difference is not statistically significant ($F(1,30)=0.341,p>0.5$).}
-\label{fig:aa}
-\end{figure}
-\vfill\null
-\section{Experiment: Noise}
-For all our experiments, we add subtle $5\%$ noise to every pixel to enhance variability.
-We did not observe a significant effect on regression performance when comparing the weber-fechner's law experiment with and without noise averaged over 4 runs (Figure~\ref{fig:noise}).
-However, the variability of our stimuli increases.
+\bibliography{paper}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\begin{figure}[th]
-\centering
-\includegraphics[width=.8\linewidth]{../gfx/weber_mlae_no_noise.pdf}
-\includegraphics[width=.8\linewidth]{../gfx/weber_mlae_noise.pdf}
-\caption{\textbf{Noise.} We test whether noise (top: off, bottom: on) effects the performance of our networks on the weber-fechner's law experiment by measuring MLAE.
-With noise, $MLAE=4.511$ ($SD=0.512$) and without noise $MLAE=4.491$ ($SD=0.543$).
-The difference is not statistically significant ($F(1,22)=0.008,p>0.5$).}
-\label{fig:noise}
-\end{figure}
-
-%% The ``\maketitle'' command must be the first command after the
-%% ``\begin{document}'' command. It prepares and prints the title block.
-
-%% the only exception to this rule is the \firstsection command
-
-\begin{figure*}[p]
+\begin{figure*}[tbhp]
 	\centering
@@
   \caption{\textbf{Elementary perceptual tasks.} Midmean logistic absolute errors (MLAE) for all generated stimuli and across all networks.
-The * indicates networks which use ImageNet weights instead of bein trained from scratch.}
+The * indicates networks that use ImageNet weights instead of being trained from scratch.}
 	\label{fig:epc_mlae}
@@
 
-\begin{figure*}[p]
+\begin{figure*}[tbhp]
 	\centering
@@
 
-\begin{figure}[p]
+\begin{figure}[tbhp]
 	\centering
@@
   \caption{\textbf{Loss plots for the position-length experiment.} We visualize the MSE loss on training data and for unseen validation data after each epoch.
-There is no significant difference in convergence for either encoding but spiky outliers due to monte-carlo cross validation are visible.}
+There is no significant difference in convergence for either encoding.}
 	\label{fig:fig4_loss}
@@
 
-\begin{figure*}[t]
+\begin{figure*}[tbhp]
 	\centering
@@
 
-  \caption{\textbf{Loss plots for the weber-fechner's law experiment.} We visualize the MSE loss on training data (left) and for unseen validation data (right) after each epoch (a) without noise and (b) with subtle $5\%$ noise per pixel.
+  \caption{\textbf{Loss plots for the Weber-Fechner's law experiment.} We visualize the MSE loss on training data (left) and for unseen validation data (right) after each epoch (a) without noise and (b) with subtle $5\%$ noise per pixel.
 There is no significant difference when noise is added.
@@
 	  \includegraphics[width=\linewidth]{../gfx/weber_validation_loss_all.pdf}
-  \caption{\textbf{Loss plots for the weber-fechner's law experiment including VGG19 and Xception.} We visualize the MSE loss on training data and for unseen validation data after each epoch.
+  \caption{\textbf{Loss plots for the Weber-Fechner's law experiment including VGG19 and Xception.} We visualize the MSE loss on training data and for unseen validation data after each epoch.
 This plot includes the VGG19 and Xception networks trained from scratch.}
@@
   \caption{\textbf{Convolutional Activation Maps for a Pie Chart.} (a)-(c) is VGG19 *, trained on ImageNet.
-The activation maps do not differ much which is surprising since VGG19 trained from scratch performs so much better in our experiments.}
+The activation maps do not differ much, which is surprising since VGG19 trained from scratch performs so much better in our experiments.}
 	\label{fig:activationmaps}
@@
   \caption{\textbf{Convolutional Activation Maps for a Curvature stimuli.} (a)-(c) is VGG19 *, trained on ImageNet.
-The activation maps do not differ much which is surprising since VGG19 trained from scratch performs so much better in our experiments.}
+The activation maps do not differ much, which is surprising since VGG19 trained from scratch performs so much better in our experiments.}
 	\label{fig:activationmaps2}
 \end{figure*}
+
+\begin{figure*}[tbhp]
+    \centering
+	\includegraphics[width=\linewidth]{PAPER_ReviewResponse/SUPPLEMENTAL/gfx/figure12_hist.pdf}
+    \caption{\change{\textbf{Bars and Framed Rectangles experiment (E4): User responses.} The responses show that users were able to precisely measure differences with the framed rectangles stimuli.
+For bars, the estimations were often guessed as 10 ($N=55$).}}
+\end{figure*}
+
+\begin{figure*}[tbhp]
+	\includegraphics[width=\linewidth]{PAPER_ReviewResponse/SUPPLEMENTAL/gfx/weber_hist.pdf}
+    \caption{\change{\textbf{Weber's law point cloud experiment (E5): User responses.} The human results for 10, 100, 1,000 points exhibit skewed distributions.
+At 1,000 points, this task is virtually impossible, and so humans resort to simply guessing.
+However, they seem to `guess high', leading to worse than random performance on the task.}}
+\end{figure*}
+
+\clearpage
+\begin{table*}[tbhp]
+
+\caption{\change{Mean squared error (MSE) for the elementary perceptual tasks experiment (E1), to accompany the midmean logistic absolute error metric (MLAE) measures in the main paper.}}
+
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lrrrrrr}
+\toprule
+  & MLP & LeNet & VGG19 * & VGG19 & Xception * & Xception \\
+\midrule
+Position Common Scale & $0.03794 \pm 0.01436 $ & $0.00106 \pm 0.00028 $ & $0.00079 \pm 0.00027 $ & $0.00014 \pm 0.00009 $ & $0.00170 \pm 0.00036 $ & $0.00074 \pm 0.00034 $   \\ 
+Position Non-aligned Scale & $0.02974 \pm 0.00912 $ & $0.00097 \pm 0.00043 $ & $0.00081 \pm 0.00016 $ & $0.00021 \pm 0.00013 $ & $0.00183 \pm 0.00047 $ & $0.00060 \pm 0.00006 $   \\ 
+Length & $0.00267 \pm 0.00053 $ & $0.01044 \pm 0.00301 $ & $0.00046 \pm 0.00050 $ & $0.00010 \pm 0.00002 $ & $0.00142 \pm 0.00021 $ & $0.00066 \pm 0.00010 $   \\ 
+Direction & $0.08359 \pm 0.00645 $ & $0.01409 \pm 0.00575 $ & $0.01014 \pm 0.00212 $ & $0.00121 \pm 0.00090 $ & $0.02342 \pm 0.00556 $ & $0.00247 \pm 0.00159 $   \\ 
+Angle & $0.08092 \pm 0.00953 $ & $0.01859 \pm 0.00611 $ & $0.00415 \pm 0.00087 $ & $0.00053 \pm 0.00006 $ & $0.00607 \pm 0.00064 $ & $0.00158 \pm 0.00019 $   \\ 
+Area & $0.00256 \pm 0.00074 $ & $0.00383 \pm 0.00171 $ & $0.00031 \pm 0.00016 $ & $0.00010 \pm 0.00004 $ & $0.00047 \pm 0.00013 $ & $0.13955 \pm 0.22880 $   \\ 
+Volume & $0.00593 \pm 0.00382 $ & $0.00405 \pm 0.00567 $ & $0.00100 \pm 0.00067 $ & $0.00084 \pm 0.00065 $ & $0.00339 \pm 0.00218 $ & $0.00346 \pm 0.00050 $   \\ 
+Curvature & $0.00343 \pm 0.00076 $ & $0.00065 \pm 0.00015 $ & $0.00024 \pm 0.00003 $ & $0.00006 \pm 0.00001 $ & $0.00047 \pm 0.00010 $ & $0.00254 \pm 0.00108 $   \\ 
+Shading & $0.01861 \pm 0.00811 $ & $0.00448 \pm 0.00205 $ & $0.00078 \pm 0.00067 $ & $0.00032 \pm 0.00021 $ & $0.00303 \pm 0.00138 $ & $0.00657 \pm 0.00473 $   \\ 
+\bottomrule
+\end{tabular}
+}
+\end{table*}
+
+\begin{table*}[tbhp]
+
+\caption{\change{Mean squared error (MSE) for the position-angle experiment (E2), to accompany the midmean logistic absolute error metric (MLAE) measures in the main paper}}
+
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lrrrrrr}
+\toprule
+  & MLP & LeNet & VGG19 * & VGG19 & Xception * & Xception \\
+\midrule
+Pie Chart & $0.04725 \pm 0.00177 $ & $0.02429 \pm 0.00070 $ & $0.02492 \pm 0.00143 $ & $0.00095 \pm 0.00018 $ & $0.01371 \pm 0.00078 $ & $0.00260 \pm 0.00037 $  \\ 
+Bar Chart & $0.00629 \pm 0.00051 $ & $0.00322 \pm 0.00029 $ & $0.00723 \pm 0.00107 $ & $0.00076 \pm 0.00012 $ & $0.00465 \pm 0.00020 $ & $0.00109 \pm 0.0001 $  \\ 
+\bottomrule
+\end{tabular}
+}
+\end{table*}
+
+\begin{table*}[tbhp]
+
+\caption{\change{Mean squared error (MSE) for the position-length experiment (E3), to accompany the midmean logistic absolute error metric (MLAE) measures in the main paper.}}
+
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lrrrrrr}
+\toprule
+  & MLP & LeNet & VGG19 * & VGG19 & Xception * & Xception \\
+\midrule
+Type 1 & $0.03561 \pm 0.01403 $ & $0.13262 \pm 0.07927 $ & $0.04563 \pm 0.02225 $ & $0.04259 \pm 0.01621 $ & $0.07448 \pm 0.04100 $ & $0.10515 \pm 0.04370 $   \\ 
+Type 2 & $0.04270 \pm 0.04115 $ & $0.07342 \pm 0.04200 $ & $0.12256 \pm 0.06932 $ & $0.03976 \pm 0.02488 $ & $0.04624 \pm 0.02606 $ & $0.08235 \pm 0.02587 $   \\ 
+Type 3 & $0.06055 \pm 0.02853 $ & $0.04998 \pm 0.08381 $ & $0.09711 \pm 0.04050 $ & $0.02746 \pm 0.01733 $ & $0.03701 \pm 0.01777 $ & $0.08443 \pm 0.04828 $   \\ 
+Type 4 & $0.07275 \pm 0.03090 $ & $0.05433 \pm 0.04709 $ & $0.03922 \pm 0.02143 $ & $0.03065 \pm 0.02038 $ & $0.05543 \pm 0.06636 $ & $0.06119 \pm 0.06254 $   \\ 
+Type 5 & $0.06728 \pm 0.02543 $ & $0.03583 \pm 0.02154 $ & $0.12193 \pm 0.05223 $ & $0.04314 \pm 0.03622 $ & $0.04574 \pm 0.01683 $ & $0.07167 \pm 0.02072 $   \\ 
+Multi & $0.02624 \pm 0.01729 $ & $0.04419 \pm 0.02724 $ & $0.07631 \pm 0.02532 $ & $0.01756 \pm 0.00931 $ & $0.03850 \pm 0.02710 $ & $0.08365 \pm 0.02061 $   \\ 
+\bottomrule
+\end{tabular}
+}
+\end{table*}
+
+\begin{table*}[tbhp]
+
+\caption{\change{Mean squared error (MSE) for the bars and framed rectangles experiment (E4), to accompany the midmean logistic absolute error metric (MLAE) measures in the main paper}}
+
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lrrrrrr}
+\toprule
+  & MLP & LeNet & VGG19 * & VGG19 & Xception * & Xception \\
+\midrule
+Framed Rectangles & $0.00065 \pm 0.00012 $ & $0.01343 \pm 0.00484 $ & $0.00156 \pm 0.00029 $ & $0.00032 \pm 0.00011 $ & $0.00437 \pm 0.00069 $ & $0.00155 \pm 0.00062 $   \\ 
+Bars & $0.00060 \pm 0.00012 $ & $0.00553 \pm 0.00196 $ & $0.00173 \pm 0.00037 $ & $0.00043 \pm 0.00024 $ & $0.00531 \pm 0.00073 $ & $0.00145 \pm 0.00053 $   \\ 
+\bottomrule
+\end{tabular}
+}
+\end{table*}
+
+\begin{table*}[tbhp]
+
+\caption{\change{Mean squared error (MSE) for the Weber's law point cloud experiment (E5), to accompany the midmean logistic absolute error metric (MLAE) measures in the main paper.}}
+
+\resizebox{\linewidth}{!}{
+\begin{tabular}{lrrrrrr}
+\toprule
+  & MLP & LeNet & VGG19 * & VGG19 & Xception * & Xception \\
+\midrule
+10 points  & $0.10126 \pm 0.00062 $ & $0.13759 \pm 0.03183 $ & $0.09031 \pm 0.00130 $ & $0.00899 \pm 0.00150 $ & $0.09354 \pm 0.00078 $ & $0.06830 \pm 0.00207 $   \\ 
+100 points & $0.10250 \pm 0.00057 $ & $0.10259 \pm 0.00065 $ & $0.10202 \pm 0.00038 $ & $0.09835 \pm 0.00267 $ & $0.10174 \pm 0.00034 $ & $0.10818 \pm 0.01003 $   \\ 
+1,000 points & $0.09946 \pm 0.00095 $ & $0.03436 \pm 0.01137 $ & $0.01477 \pm 0.00059 $ & $0.00076 \pm 0.00008 $ & $0.02610 \pm 0.00026 $ & $0.01041 \pm 0.00258 $   \\
+\bottomrule
+\end{tabular}
+}
+\end{table*}
 
@@
 
-%\bibliographystyle{abbrv}
-%\bibliographystyle{abbrv-doi}
-%\bibliographystyle{abbrv-doi-narrow}
-%\bibliographystyle{abbrv-doi-hyperref}
-%\bibliographystyle{abbrv-doi-hyperref-narrow}
-
-%\bibliography{../paper}
 \end{document}

### 2018-08-15 PAPER/paper.tex
Section 4!
Update paper from overleaf.

@@
 \usepackage{subfig}
+\usepackage{amsmath}
+\usepackage{enumitem}
+\usepackage{flushend}
+
+\newenvironment{hypolist}
+{
+  \begin{itemize}
+    \renewcommand\labelitemi{}
+    \setlength{\itemsep}{1pt}
+    \setlength{\parskip}{1pt}
+    \setlength{\parsep}{1pt}
+    \setlength{\leftmargin}{0em}
+    %\setlength{\itemindent}{0em} -> JT: Doesn't work...
+    \addtolength{\itemindent}{-2.3em}
+
+}
+{
+  \end{itemize}
+}
+
+\usepackage{xcolor}
+\newcommand{\tnote}[3]{{\color{#2}#1: #3}}
+%\newcommand{\change}[1]{\textcolor{red}{#1}}
+%\newcommand{\removed}[1]{\textcolor{red}{\sout{#1}}}
+% \newcommand{\change}[1]{{\color{red}#1}}
+\newcommand{\change}[1]{{#1}}
+\newcommand{\chek}[1]{{\color{red}#1}}
+
+\newcommand{\removed}[1]{}
+\newcommand{\DaH}[1]{\tnote{DaH}{red}{#1}}
+\newcommand{\HP}[1]{\tnote{HP}{green}{#1}}
+\newcommand{\JT}[1]{\tnote{JT}{blue}{#1}}
+\newcommand{\IGNORE}[1]{}
 
@@
 %% OnlineID. Otherwise, you may safely leave it at ``0''.
-\onlineid{0}
+\onlineid{1171}
 
@@
 \abstract{
-Convolutional neural networks can successfully perform many computer vision tasks on images, and their learned representations are often said to mimic the early layers of the visual cortex.
-But can CNNs understand graphical perception for visualization?
-We investigate this question by reproducing Cleveland and McGill's seminal 1984 experiments, which measured human perception efficiency of different visual encodings and defined elementary perceptual tasks for visualization.
-We measure the graphical perceptual capabilities of four classifiers on a) elementary perceptual tasks with increasing parametric complexity, b) the position-angle experiment that compares pie charts to bar charts, c) the position-length experiment that compares grouped and divided bar charts, and d) the bars and framed rectangles experiment where visual cues aid perception.
-We also study how feed-forward neural networks obey Weber's law, which defines the proportional relation between perceivable information and distribution density.
-We present the results of these experiments to foster the understanding of how CNN classifiers succeed and fail when applied to data visualizations.
+Convolutional neural networks can successfully perform many computer vision tasks on images.
+\change{For visualization, how do CNNs perform when applied to graphical perception tasks?} We investigate this question by reproducing Cleveland and McGill's seminal 1984 experiments, which measured human perception efficiency of different visual encodings and defined elementary perceptual tasks for visualization.
+We measure the graphical perceptual capabilities of four network architectures on five different visualization tasks and compare to existing and new human performance baselines.
+\change{While under limited circumstances CNNs are able to meet or outperform human task performance, we find that CNNs are not currently a good model for human graphical perception.} We present the results of these experiments to foster the understanding of how CNNs succeed and fail when applied to data visualizations.
 } % end of abstract
@@
 %% please capitalize first letter and insert punctuation after last keyword
-\keywords{Machine Perception, Deep Learning}
+\keywords{Machine Perception, Graphical Perception, Deep Learning, Convolutional Neural Networks.}
 
@@
   \centering
-  \includegraphics[width=\linewidth]{teaser.pdf}
-  \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking visual cortex inspired machine learning classifiers to assess the relationship between values encoded in pie charts and bar charts.
-Similar to the findings of Cleveland and McGill~\cite{cleveland_mcgill}, our experiments show that CNNs read quantities more accurately from bar charts (mean squared error, MSE in green).}
+  \includegraphics[width=\linewidth]{teaser_v2.pdf}
+  \caption{\textbf{Computing Cleveland and McGill's Position-Angle Experiment using Convolutional Neural Networks.} We replicate the original experiment by asking CNNs to assess the relationship between values encoded in pie charts and bar charts.
+We find that CNNs can predict quantities more accurately from bar charts (mean squared error (MSE) in green).}
 	\label{fig:teaser}
@@
 
-\input{07_results.tex}
+\input{07_webers_law_experiment.tex}
 
-\input{08_conclusions.tex}
+\input{08_discussion.tex}
+
+\input{09_conclusions.tex}
 
 %% if specified like this the section will be committed in review mode
-%\acknowledgments{
-%The authors wish to thank A, B, and C. This work was supported in part by
-%a grant from XYZ (\# 12345-67890).}
+\acknowledgments{
+\change{We thank Junkyung Kim, Matthew Ricci, Daniel Pineo, Thomas Serre, John Tsotsos, Li Zhaoping, and our reviewers for helpful discussions and feedback.
+We thank the staff of the Harvard Human Research Protection Program for fast IRB approval and the Harvard Research Computing crew for their substantial support.
+We also thank the Boston Fire Department for \href{https://www.universalhub.com/2018/shorting-power-supplies-spark-fire-downtown}{extinguishing flames in our data center}, which we promise was not caused by us training over 2,500 models for this paper.
+This research is supported in part by grants NSF IIS-1447344 and NIH 5U01CA198935.}}
+
+\vfill
 
